\subsection{June}







\subsubsection{June 1st}
Today I learned about the Mahler measure, from \href{https://web.archive.org/web/20210112151948/https://www.uni-due.de/~adg350u/Skripte/DioApp.pdf}{here}. Roughly speaking, the Mahler measure is supposed to measure the ``size'' of an algebraic number. Here's our definition.
\begin{definition}
    Fix $\alpha\in\overline\QQ$ an algebraic number with minimal polynomial $f(x)=\sum_{k=0}^da_kx^k\in\ZZ[x]$ with $\gcd(a_0,\ldots,a_d)=1.$ Then we define the Mahler measure as
    \[M(\alpha):=|a_d|\prod_{\ell=1}^d\max\{1,\alpha_k\},\]
    where $\{\alpha_k\}_{\ell=1}^d$ are the Galois conjugates of $\alpha.$
\end{definition}
Note that the polynomial $f(x)=a_d\prod_{\ell=1}^d(x-\alpha_\ell)$ is uniquely determined modulo multiplication by a unit: we merely need $a_d$ to be a least common multiple of the denominators of $\prod_{\ell=1}^d(x-\alpha_\ell),$ and no further is allowed to keep $f(x)$ primitive. Thus, $|a_d|$ and therefore $M(\alpha)$ is well-defined.

There is little immediate reason to think that $M$ is the correct way to measure size of $\alpha\in\overline\QQ,$ but $M$ does have some nice properties. For example, because the polynomial is shared between all Galois conjugates, all Galois conjugates have the same Mahler measure. This is nice because we would like the Galois group to preserve $M.$

Additionally, $M$ satisfies the basic and necessary property of a size: finite bounds.
\begin{proposition}
    For a fixed bounds $D$ and $N,$ there are only finitely many $\alpha\in\overline\QQ$ of degree less than or equal to $D$ which satisfy $M(\alpha)\le N.$
\end{proposition}
The basic idea is to bound the possible polynomials which an $\alpha$ could appear from; then the polynomials need to be in $\ZZ[x],$ so there are only finitely many of those. To bound the coefficients of a minimal polynomial $f(x)$ of some $\alpha\in\overline\QQ$ satisfying $M(\alpha)\le N,$ we use Vieta's formula. For clarity, let
\[f(x):=\sum_{k=0}^da_kx^k=a_d\prod_{\ell=1}^d(x-\alpha_\ell)\]
for $\{a_k\}_{k=0}^d\subseteq\ZZ$ coprime and $\{\alpha_\ell\}_{\ell=1}^d\subseteq\overline\QQ$ the Galois conjugates of $\alpha.$ Now, Vieta's formula asserts that we make a coefficient $a_k$ by choosing $k$ of the $x$s in the product and then $d-k$ of the $\alpha_\bullet$s. In other words, fixing some $k,$
\[a_k=a_d\sum_{\substack{S\subseteq[1,d]\cap\ZZ\\|S|=d-k}}\left(\prod_{\ell\in S}-\alpha_\ell\right).\]
Then the triangle inequality plus some rearranging asserts that
\[|a_k|\le\sum_{\substack{S\subseteq[1,d]\cap\ZZ\\|S|=d-k}}\left(|a_d|\prod_{\ell\in S}|\alpha_\ell|\right)\le\sum_{S\subseteq[1,d]\cap\ZZ}M(\alpha)=2^dM(\alpha).\]
This gives us a global bound for the size of the coefficients of polynomials $f(x)$ giving $M(\alpha)\le N,$ by $|a_k|\le 2^dM(\alpha)\le 2^DN.$

To finish, we note that each $\alpha\in\overline\QQ$ with $M(\alpha)\le N$ can be associated with one of its minimal polynomials $f(x).$ Because each of these polynomials has degree bounded by $D$ and coefficients bounded (in absolute value) by $2^DN,$ we have at most
\[\left(2^DN\right)^D\]
such polynomials. (This is a very crude bound, but it will do.) Then each polynomial can be associated backwards with at most its $D$ roots, so there are at most
\[D\left(2^DN\right)^D\]
such $\alpha\in\overline\QQ.$ This finishes. $\blacksquare$

We remark that the degree requirement is necessary: fix any $n\in\NN,$ and the root of unity $\zeta_n:=e^{2\pi i/n}$ is an algebraic integer (its minimial polynomial is the cyclotomic $\Phi_n(x)$), and its conjugates are all of the form $\zeta_n^k$ for some integer $k.$ Thus, all the Galois conjugates of $\zeta_n$ are of absolute value $1,$ implying $M(\zeta_n)=1.$

As a final addendum, we note that one of the results from a few months ago has its proof nicely abbreviated by the introduction of the Mahler measure.
\begin{proposition}
    Roots of unity are the only nonzero algebraic integers with all Galois conjugates of absolute value less than or equal $1.$
\end{proposition}
Fix one such nonzero $\alpha\in\overline\ZZ\setminus\{0\}$ of degree $d$ with Galois conjugates $\{\alpha_\ell\}_{\ell=1}^d.$ What makes the Mahler measure tractable here is that $\alpha\overline\ZZ$ makes the leading coefficient of $\alpha$'s minimal polynomial equal to $1.$ Thus,
\[M(\alpha)=\prod_{\ell=1}^d\max\{1,|\alpha_\ell|\}.\]
However, we are told that $|\alpha_\ell|=1$ for each $\alpha_\ell,$ so in fact $M(\alpha)=1.$

As before, the key step is to consider what happens to $\alpha^k$ for powers $k\in\NN.$ Well, $\overline\ZZ$ is a ring, so $\alpha^k$ still has leading coefficient $1.$ And considering the orbit of $\alpha^k$ under the Galois group of the Galois closure of $\QQ(\alpha)$ over $\QQ,$ the Galois conjugates of $\alpha^k$ are $\{\alpha_\ell^k\}_{\ell=1}^d.$ Thus,
\[M\left(\alpha^k\right)=\prod_{\ell=1}^d\max\left\{1,|\alpha_\ell|^k\right\},\]
which comes out to $1$ again because $|\alpha_\ell|^k\le1$ for each $\alpha_\ell.$

To deal the killing blow, look at the sequence $\{\alpha^k\}_{k=1}^\infty.$ Every element lives in $\QQ(\alpha),$ so their degrees are bounded by the degree of $\alpha,$ and we have just established that their Mahler measures are uniformly bounded by $1.$ So there can only be finitely many distinct elements in this sequence, so the Pigeonhole principle requires repeats. So extract distinct $k_1$ and $k_2$ yielding
\[\alpha^{k_1}=\alpha^{k_2}.\]
Without loss of generality, we let $k_1>k_2.$ We required $\alpha\ne0$ in the hypothesis, so we get $\alpha^{k_1-k_2}=1,$ meaning that $\alpha$ is a root of unity. This finishes. $\blacksquare$

As some commentary, the above proof establishes some of what's nice about the Mahler measure: it at least somewhat behaves in powers because the Galois conjugates behave in powers, and the finite bound property allows some kind of ``well-ordering''-type thinking. Regardless, I am unconvinced this is the best way to measure heights.

\subsubsection{June 2nd}
Today I learned a proof that $e^q$ is irrational for any $q\in\QQ\setminus\{0\},$ following Hermite. Note that we can write $q=n/m$ for $n$ a positive integer and $m$ an integer, and $e^{n/m}$ rational implies $\left(e^{n/m}\right)^m=e^n$ rational, so it suffices to show (by contraposition) that $e^n$ is irrational for positive integers $n.$
\begin{theorem}
    The value $e^n$ is irrational for positive integers $n.$
\end{theorem}
This proof is truly magical. Our result is going to come from bounding: we are going to create an arbitrarily small integral that the rationality of $e$ forces to be an integer, and that will yield contradiction. Of course, we have to use something about $e,$ so we are going to use the fact that
\[\frac d{dx}e^{nx}=ne^{nx}.\]
Note this is a defining property of $e,$ so we can be sure we are using its full power. We would like to use this to make a (small) integral, so we have to introduce a (small) auxiliary differentiable function $F:\RR\to\RR$ here: the product rule tells us that
\[\frac d{dx}F(x)e^{nx}=F'(x)e^{nx}+nF(x)e^{nx}.\]
We promised an integral, and indeed, our integral is going to be
\[\int_0^1\left(F'(x)+nF(x)\right)e^{nx}\,dx=F(x)e^{nx}\bigg|_0^1=F(1)e^n-F(0).\]
When we suppose that $e^n$ is rational (which we will postpone as long as possible), we might hope that $F(0)$ and $F(1)$ are integers so that the above integral can be made into an integer when multiplied by the supposed denominator of $e^n.$

To proceed, we recognize we will need better control of $F'(x)+nF(x)$---after all, we have to make its integral arbitrarily small. So we define, for now, $f(x):=F'(x)+nF(x)$ and attempt to solve for $F$ in terms of $f.$ Well, we see
\[F(x)=\frac1nf(x)-\frac1nF'(x).\]
Of course, we don't know what $F'$ is directly, but we can solve for it by differentiating the above into $F'(x)=\frac1nf'(x)-\frac1nF''(x),$ so we see
\[F(x)=\frac1nf(x)-\frac1{n^2}f'(x)+\frac1{n^2}F''(x).\]
We can continue this process inductively and just write
\[F(x)=\sum_{k=0}^\infty\frac{(-1)^k}{n^{k+1}}f^{(k)}(x),\]
assuming everything converges properly. We can check that, assuming nice enough convergence (which we will have), this does indeed give $f(x)=F'(x)+nF(x).$

We said above that it would be nice for $F(0)$ and $F(1)$ to be integers, which makes choosing an $f(x)$ to define $F(x)$ a bit annoying: after all, there's that infinite sum to deal with. Our way out is to make $f(x)$ a polynomial, thus making the sum finite, and then multiplying $f$ with a very large power of $n$ to accommodate for the given denominators.

Now we introduce the star of our story. Choose $N$ very large, to be fixed later, making
\[f_0(x):=\frac{x^N(1-x)^N}{N!}\]
very small. Then we define $f(x):=n^{2N+1}f_0(x)$ to deal with the denominators so that
\[F(x):=\sum_{k=0}^\infty\frac{(-1)^k}{n^{k+1}}f^{(k)}(x)=\sum_{k=0}^{2n}n^{2N-k}f_0^{(k)}(x).\]
It might be totally obvious that $F(0)$ and $F(1)$ are integers, but they are: we can write
\[f_0(x)=\sum_{k=N}^{2N}\binom N{2N-k}(-x)^k.\]
Reading this off like a Taylor expansion, we see $k<N$ gives $f_0^{(k)}(0)=0,$ and for $k\ge N,$ we get $f^{(k)}(0)=(-1)^kk!\binom N{2N-k},$ and $k!$ does enough to cover the denominator of $N!.$ This shows the $f_0^{(k)}(0)$ are integers, so $F(0)$ is an integer. Further,
\[f_0(x)=f_0(1-x),\]
so $f_0^{(k)}(1)=(-1)^kf_0^{(k)}(0)$ is also an integer for any $k,$ meaning that $F(1)$ is an integer.

Now surely $f(x)=F'(x)+nF(x)$ still holds from our work above, for we have the nicest converge, from a finite sum. We recall our integral as
\[\int_0^1f(x)e^{nx}\,dx=F(x)e^{nx}\bigg|_0^1=F(1)e^n-F(0).\]
However, we can bound the integral because $f(x)=n^{2N+1}f_0(x)<n^{2N+1}/N!$ for $x\in(0,1).$ (Our lower-bound is that the integral is positive.) Thus,
\[0<\int_0^1f(x)e^{nx}\,dx<\frac{n^{2N+1}}{N!}\cdot e^n.\]
This can be made arbitrarily small because $N!$ is faster than exponential, which is a problem because $F(1)e^n-F(0)$ has its denominator bounded by the denominator of $e^n.$

To formalize, we finally suppose for the sake of contradiction that $e^n=p/q.$ Then
\[q\int_0^1f(x)e^{nx}\,dx=pF(1)-qF(0)\]
is an integer, but
\[0<q\int_0^1f(x)e^{nx}\,dx<\frac{pn^{2N+1}}{N!}.\]
Taking $N$ large enough, we can force $N!>pn^{2N+1},$ making the upper-bound less than $1.$ This gives us an integer between $0$ and $1,$ which is our contradiction. This finishes. $\blacksquare$

To close off loose ends, we show that we can in fact make $\frac{n^{2N+1}}{N!}$ arbitrarily small as $N\to\infty.$
\begin{lemma}
    Given a fixed integer $n,$ we can choose $N$ large enough so that
    \[\frac{n^{2N+1}}{N!}\]
    can be made arbitrarily small.
\end{lemma}
For all $N,$ note that we can say
\[\log N!=\sum_{k=1}^N\log k>\sum_{k=\ceil{N/2}}^N\log k>\frac N2\log\frac N2.\]
There is a technical detail here that there are at least $N/2$ integers between $\ceil{N/2}$ and $N$ inclusive, but this is true. (If $N$ is even, there are $N/2+1$ integers; if $N$ is odd, there are $N/2+1/2$ integers.) Anyways, this implies that
\[N!>(N/2)^{N/2}.\]
If we let $N=2n^6k$ for some large positive integer $k,$ we see
\[\frac{n^{2N+1}}{N!}\le\frac{n^{3N}}{(N/2)^{N/2}}=\frac{n^{6n^6k}}{n^{6n^3k}\cdot k^{n^6k}}=\frac1{k^{n^6k}}.\]
This is less than $\frac1k,$ for example, so we can make the original quantity as small as we please, finishing. $\blacksquare$

As commentary, I have tried my best to motivate this proof, but it is truly magical. There is a pretty large jump to the introduction of the $f_0,$ and from there everything comes together. I really don't know where this proof comes from.

In other news, today was senior salute. It was a good day.

\subsubsection{June 3rd}
Today I learned an example of spectral graph theory doing something of number-theoretic interest, from THE BOOK. Here is the statement we are building towards.
\begin{theorem}
    Fix a prime $p.$ There are $p^2$ solutions $(x,y,z)\in\FF_p^3$ to $x^2+y^2+z^2\equiv0\pmod p.$
\end{theorem}
Note that $\FF_p^\times$ acts on the solutions $(x,y,z)$ by multiplication because $k\cdot(x,y,z)=(kx,ky,kz)$ still satisfies $(kx)^2+(ky)^2+(kz)^2\equiv k^2\left(x^2+y^2+z^2\right)$ will still $0$ out when $x^2+y^2+z^2\equiv0.$ We say because the orbit of $(0,0,0)$ has size $1$ ($(0k,0k,0k)=(0,0,0)$) the nonzero points $(x,y,z)$ have size $p-1$ (the nonzero coordinate has full image).

So because the zero point is poorly behaved, we show that there are $p^2-1$ nonzero points satisfying $x^2+y^2+z^2\equiv0.$ In fact, we will want to leverage the linear algebra structure of $\FF_p^3$ later in life, so we will show that
\[\frac{\{(x,y,z)\in\FF_p^3\setminus\{0\}:x^2+y^2+z^2\equiv0\}}{\FF_p^\times}\stackrel?=\frac{p^2-1}{p-1}=p+1.\]
That is, we need to show there are $p+1$ orbits of the nonzero points on $x^2+y^2+z^2\equiv0.$ Noticing the linear algebra structure on $\FF_p^3,$ these orbits are really one-dimensional subspaces generated by a nonzero vector, say $v=(x,y,z),$ which looks
\[\{kv:k\in\FF_p\}=\FF_p^\times v\cup\{0\}.\]
So instead we are now counting the number of one-dimensional subspaces $L$ of $\FF_p^3$ whose representatives $(x,y,z)\in L$ satisfy $x^2+y^2+z^2\equiv0,$ and we need there to be $\frac{p^2-1}{p-1}=p+1$ of them because there are $p-1$ elements in each orbit.

Now we bring in the spectral graph theory, and for this we need a graph $G_p.$ Our vertices will be the one-dimensional subspaces as represented by
\[V(G_p)=\frac{\FF_p^3\setminus\{0\}}{\FF_p^\times}.\]
We quickly remark that there are $\frac{p^3-1}{p-1}=p^2+p+1$ total vertices. We would like to rig our graph to detect $x^2+y^2+z^2\equiv0$ in some nice way, and the clever way to do this is to write
\[E(G_p)=\big\{\{L_1,L_2\}\subseteq V(G_p):\langle v_1,v_2\rangle=0\text{ for all }v_1\in L_1,v_2\in L_2.\big\}.\]
Here, $\langle(x_1,y_1,z_1),(x_2,y_2,z_2)\rangle:=x_1x_2+y_1y_2+z_1z_2$ is the usual dot product. (Despite using the notation, we are not calling $\langle,\rangle$ an inner product because it is not positive-definite.) We note that the edge connection is well-defined after picking up any representative $v\in L\setminus\{0\}.$ Indeed, if $\langle v_1,v_2\rangle=0$ for $v_1\in L_1\setminus\{0\}$ and $v_2\in L_2\setminus\{0\}$ and then we choose different representatives $k_1v_1\in L_1$ and $k_2v_2\in L_2,$ then still
\[\langle k_1v_1,k_2v_2\rangle=k_1k_2\langle v_1,v_2\rangle=0.\]
Anyways, the point of using the dot product for edge connections is to say
\[x^2+y^2+z^2=\langle(x,y,z),(x,y,z)\rangle\]
vanishes if and only if $(x,y,z)$ has a loop to itself. Letting $A_p$ be the adjacency matrix of $G_p,$ it suffices for the following.
\begin{lemma}
    Let $A_p$ be the adjacency matrix of the above graph $G_p.$ Then $\op{trace}A_p=p+1.$
\end{lemma}
We have now moved theoretically far away from the original statement, but we are only a single trick away from being done. Anyways, we are going to compute $\op{trace}A_p$ by computing its eigenvalues, cementing the statement we are doing spectral graph theory.

The main idea, now, is that $A_p$ might be difficult to analyze, but $A_p^2$ is not. Indeed, given two one-dimensional subspaces $L_1,L_2\in V(G_p),$ the entry $\left(A_p^2\right)_{v_1,v_2}$ counts the number of the paths from $L_1$ to $L_2$ of length two. Fix any $v_1\in L_1\setminus\{0\}$ and $v_2\in L_2\setminus\{0\}.$

Well by definition of our graph, we are counting the number of one-dimensional subpsaces $L$ satisfying
\[\langle v,v_1\rangle=\langle v,v_2\rangle=0\]
for some nonzero representative $v\in L\setminus\{0\}.$ (Recall that edge-connection is well-defined, so our choice of representative is irrelevant.) But now we might as well be asking for the number of $v$ satisfying
\[\underbrace{\begin{bmatrix}
    - & v_1 & - \\
    - & v_2 & -
\end{bmatrix}}_Mv=\begin{bmatrix}0 \\ 0\end{bmatrix}.\]
So, roughly speaking, we want the dimension of $\ker M.$ We have two cases. Now, the set $L_1^\perp$ of vectors $v$ such that $\langle v,v_1\rangle=0$ is two-dimensional, for the map $v\mapsto\langle v,v_2\rangle$ has one-dimensoinal image. But now
\[\dim\ker M=\dim(L_1^\perp\cap L_2^\perp)=\dim L_1^\perp+\dim L_2^\perp-\dim(L_1^\perp+L_2^\perp)=4-\dim(L_1^\perp+L_2^\perp).\]
We have two cases.
\begin{itemize}
    \item If $L_1\ne L_2,$ then $L_1^\perp\ne L_2^\perp$ (recall $L^{\perp\perp}=L$ by, say, counting dimensions and building a basis). Thus, $L_2^\perp$ has a vector not in $L_1^\perp$ because they have the same size, so
    \[L_1^\perp+L_2^\perp\supsetneq L_1^\perp.\]
    Thus, $\dim(L_1^\perp+L_2^\perp)>\dim L_1^\perp=2,$ meaning that it must have dimension $3.$ So $\dim\ker M=1,$ and there is a single one-dimensional subspace to go around.
    \item If $L_1=L_2,$ then $L_1^\perp=L_2^\perp,$ so $\dim(L_1^\perp+L_2^\perp)=2.$ Thus, $\dim\ker M=2,$ implying that it has $p^2-1$ nonzero points and $\frac{p^2-1}{p-1}=p+1$ one-dimensional subspaces.
\end{itemize}
We remark briefly that we are working with one-dimensional subspaces instead of points because the condition $L_1=L_2$ creates fewer complications than $\FF_p^\times v_1=\FF_p^\times v_2.$

Anyways, it follows that the adjacency matrix looks like
\[A_p^2=\begin{bmatrix}
    p+1 & 1 & \cdots & 1 & 1 \\
    1 & p+1 & \cdots & 1 & 1 \\
    \vdots & \vdots & \ddots & \vdots & \vdots \\
    1 & 1 & \cdots & p+1 & 1 \\
    1 & 1 & \cdots & 1 & p+1
\end{bmatrix}.\]
All that remains is computation. We are interested in the eigenvalues of $A_p$ (to sum them), so we now ask for the eigenvalues of $A_p^2$ as above. Subtracting out $p$ times the identity, which will merely add $p$ to all eigenvalues, we want the eigenvalues of
\[A_p^2-pI=\begin{bmatrix}
    1 & \cdots & 1 \\
    \vdots & \ddots & \vdots \\
    1 & \cdots & 1
\end{bmatrix}.\]
Recalling that there are $\frac{p^3-1}{p-1}=p^2+p+1$ vertices, we need to find $p^2+p+1$ eigenvalues. Well, the image of this matrix lives in the one-dimensional subspace defined by $\langle1,\ldots,1\rangle$ and has a nonzero element (throw $\langle1,0,\ldots0\rangle$ in there), so the image does in fact have size $1.$

On one hand, this means that the kernel has dimension $p^2+p+1-1=p^2+p,$ so there are $p^2+p$ eigenvectors with eigenvalue $0.$ On the other hand, the vector $\langle1,\ldots,1\rangle$ maps to
\[\begin{bmatrix}
    1 & \cdots & 1 \\
    \vdots & \ddots & \vdots \\
    1 & \cdots & 1
\end{bmatrix}\begin{bmatrix}1 \\ \vdots \\ 1\end{bmatrix}=\begin{bmatrix}p^2+p+1 \\ \vdots \\ p^2+p+1\end{bmatrix}\]
and so has eigenvalue $p^2+p+1.$ Because we can only have $p^2+p+1$ total eigenvectors, this must be our last eigenvector.

Bringing this together, we see that $A_p^2$ has $p^2+p$ eigenvectors of eigenvalue $0+p=p$ and one eigenvector of eigenvalue $p^2+p+1+p=(p+1)^2.$ Taking the square-root ($A_p$ is diagonalizable because it has real entries and is symmetric because the graph is undirected), $A_p$ has one eigenvector of eigenvalue $\pm(p+1)$ and (say) $a$ eigenvectors of eigenvalue $\sqrt p$ and $b$ eigenvectors of eigenvalue $-p$ with $a+b=p^2+p.$ Thus,
\[\op{trace}A_p=\pm(p+1)+a\sqrt p-b\sqrt p.\]
Surely, $\op{trace}A$ is an integer, and $\sqrt p$ is irrational, so we must have $a=b$ to cover for this. So $\op{trace}A_p=\pm(p+1).$ But also surely $\op{trace}A_p$ is positive, so we must have $\op{trace}A_p=p+1.$ (In fact, $\langle1,\ldots,1\rangle$ has eigenvalue $p+1.$) This finishes. $\blacksquare$

It is proofs like these that make me certain I will never be a mathematician. This is pure magic. If I had to guess, the number-theoretic statement came out of other graph-theory studies, but I have no idea where the graph construction came from. Anyways, I am under the impression that this method can be generalized.

For example, we probably have the following.
\begin{theorem}
    Fix $p$ a prime and $d$ a positive odd integer. The number of $v\in\FF_p^d$ for which $\langle v,v\rangle=0$ is equal to $p^{d-1}.$
\end{theorem}
The structure for even dimension seems more difficult. For example, even with $d=2,$ are computing the number of $(x,y)\in\FF_p^2$ for which $x^2+y^2\equiv0.$ In particular, the existence of nonzero solutions is equivalent to the existence of a $\sqrt{-1}\in\FF_p$ after rearranging.

\subsubsection{June 5th}
Today I learned about the difficulty generalizing the result from yesterday to arbitrary dimensions. We are going to go done through the proof in an arbitrary dimension $d$ and see what problems we have; however, we will be less detailed. We are interested in computing
\[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}\]
for prime-power $q.$ As before, we note that, given $k\in\FF_q^\times,$ we can take $v\mapsto kv$ and preserve the condition we're counting, and as long as $v\ne0,$ there are $q-1$ nonzero vectors in $\FF_qv.$ So we can instead count
\[1+(q-1)\#\left\{L\in(\FF_q^d\setminus\{0\})/\FF_q^\times:\langle v,v\rangle=0\text{ for all }v\in L\right\}.\]
Note that if any $v\in L$ has $\langle v,v\rangle,$ then all $v'\in L$ do because $v'=kv$ for some $k\in\FF_q,$ and $\langle v',v'\rangle=k^2\langle v,v\rangle=0.$ This is to say that we don't have to keep very careful track of the condition. Very quickly, we note that the analysis above says that there are
\[\#\left((\FF_q^d\setminus\{0\})/\FF_q^\times\right)=\frac{q^d-1}{q-1}\]
total such one-dimensional subspaces of $\FF_q^d.$

As before, we count this by introducing an an auxiliary graph. Define $G_q$ with vertices labeled by each one-dimensional subspace in $(\FF_q^d\setminus\{0\})/\FF_q^\times.$ Then for $L_1,L_2\in V(G_p),$ we place an edge between $L_1$ and $L_2$ if and only if
\[\langle v_1,v_2\rangle=0\text{ for all }v_1\in L_1,v_2\in L_2.\]
Again, note that it suffices for $\langle v_1,v_2\rangle$ for a single $v_1\in L_1\setminus\{0\}$ and $v_2\in L_2\setminus\{0\}.$ Indeed, then for any other $v_1'\in L_1$ and $v_2'\in L_2,$ we can write $v_1'=k_1v_1$ and $v_2'=k_2v_2$ because $L_1,L_2$ are one-dimensional. Then we have
\[\langle v_1',v_2'\rangle=k_1k_2\langle v_1,v_2\rangle=0,\]
so indeed $L_1$ and $L_2$ are adjacent.

Now, we are interested in counting the number of one-dimensional subspaces $L\in(\FF_q^d\setminus\{0\})/\FF_q^\times$ such that $L$ is adjacent to itself: we want there to be a nonzero vector $v\in L\setminus\{0\}$ with $\langle v,v\rangle=0.$ So because we want to compute the number of loops in $G_q,$ we fix $A_q$ the adjacency matrix of $G_q$ so that we want to compute
\[\op{trace}A_q=\#\left\{L\in V(G_q):\langle v,v\rangle=0\text{ for all }v\in L\right\}.\]
But as before, the matrix $A_q$ is significantly more complicated than $A_q^2.$ We have the following.
\begin{lemma}
    The matrix $A_q^2$ is equal to $\frac{q^{d-2}-1}{q-1}J+q^{d-2}I,$ where $J$ is the all-ones matrix.
\end{lemma}
We compute the matrix coefficients of $A_q^2$ manually. By properties of the adjacency matrix, the coefficient of $A_q^2$ corresponding to the vertices $L_1$ and $L_2$ is the number of length-two paths between them. In other words, it is the number of vertices $L$ such that
\[\langle v_1,v\rangle=\langle v,v_2\rangle=0\text{ for all }v_1\in L_1,v\in L,v_2\in L_2.\]
To compute the number of one-dimensional subspaces $L,$ we'll compute the number of vectors $v$ for which $\langle v_1,v\rangle=\langle v,v_2\rangle=0$ for all $v_1\in L_1$ and $v_2\in L_2.$ This will be enough because the set of such vectors forms a subspace (it's $L_1^\perp\cap L_2^\perp$), so we can safely mod out by $\FF_q^\times$ later. So we are, roughly speaking, computing
\[\dim\left(L_1^\perp\cap L_2^\perp\right)=\dim\left(L_1^\perp\right)+\dim\left(L_2^\perp\right)-\dim\left(L_1^\perp+L_2^\perp\right),\]
where the dimensions are taken over $\FF_q.$ Note that $\dim L_\bullet^\perp=d-\dim L_\bullet=d-1$ by duality or something, so the hard part is computing $\dim\left(L_1^\perp+L_2^\perp\right).$ Now we have two cases.
\begin{itemize}
    \item If $L_1\ne L_2,$ then $L_1^\perp\ne L_2^\perp.$ But the dimension of both $L_\bullet^\perp$ has dimension $d-1,$ and $L_1^\perp+L_2^\perp$ must be strictly larger than $L_1^\perp$ or $L_2^\perp$ (one has a vector not in the other), meaning that $L_1^\perp+L_2^\perp$ has to have the largest possible dimension, $d.$
    
    Thus, $\dim\left(L_1^\perp\cap L_2^\perp\right)=d-2.$
    \item If $L_1=L_2,$ then $L_1^\perp=L_2^\perp,$ so $L_1^\perp+L_2^\perp=L_1^\perp$ has dimension $d-1.$ Thus, $\dim\left(L_1^\perp\cap L_2^\perp\right)=d-1.$
\end{itemize}
Returning to both cases simultaneously, we let $d':=\dim\left(L_1^\perp\cap L_2^\perp\right)$ as computed above. Then the number of points in $L_1^\perp\cap L_2^\perp$ is $q^{d'}.$ For each vector $v,$ it generates a one-dimensional subspace $\FF_qv,$ and these orbits are disjoint except for the origin because intersecting orbits $k_1v_1=k_2v_2$ have $v_1=k_1^{-1}k_2v_2$ and $v_2=k_2^{-1}k_1v_1$ create the same orbit.

Thus, the number of one-dimensional subspaces adjacent to both $L_1$ and $L_2$ is
\[\frac{q^{d'}-1}{q-1}.\]
This lets write out $A_q^2$ explicitly as
\[A_q^2=\frac1{q-1}\begin{bmatrix}
    q^{\color{red}d-1}-1 & q^{d-2}-1 & \cdots & q^{d-2} - 1 & q^{d-2}-1 \\
    q^{d-2}-1 & q^{\color{red}d-1}-1 & \cdots & q^{d-2} - 1 & q^{d-2}-1 \\
    \vdots & \vdots & \ddots & \vdots & \vdots \\
    q^{d-2}-1 & q^{d-2}-1 & \cdots & q^{\color{red}d-1} - 1 & q^{d-2}-1 \\
    q^{d-2}-1 & q^{d-2}-1 & \cdots & q^{d-2} - 1 & q^{\color{red}d-1}-1
\end{bmatrix}.\]
With $J$ as the all-ones matrix, we can collapse this into
\[A_q^2=\frac{q^{d-2}-1}{q-1}J+\frac{q^{d-1}-q^{d-2}}{q-1}I.\]
This rearranges into the desired, so we call it quits. $\blacksquare$

And now we can somewhat safely compute the eigenvalues of $A_q.$
\begin{lemma}
    The matrix $A_q$ has one eigenvalue of $\frac{q^{d-1}-1}{q-1}$ and $\frac{q^d-1}{q-1}-1$ eigenvalues of $\pm\sqrt{q^{d-2}}.$
\end{lemma}
Observe that $A_q,$ as the adjacency matrix of a undirected graph, is real and symmetric and therefore diagonalizable. Thus, we are guaranteed all $\frac{q^d-1}{q-1}$ eigenvalues.

We leverage what we know about $A_q^2$ because its eigenvalues are the square of the eigenvalues of $A_q.$ The matrix $J$ has one-dimensional image (all of its columns are $\langle1,\ldots,1\rangle,$ after all), so it has kernel of dimension $\frac{q^d-1}{q-1}-1.$ In other words, $J$ has $\frac{q^d-1}{q-1}-1$ eigenvalues of $0.$ It follows that $A_q^2$ has $\frac{q^d-1}{q-1}-1$ eigenvalues of
\[0+q^{d-2}.\]
Thus, $A_q$ has $\frac{q^d-1}{q-1}-1$ eigenvalues of $\pm\sqrt{q^{d-2}}.$

It remains to find our last eigenvalue of $A_q.$ Well, it should come from the eigenvector of nonzero eigenvalue from $J,$ which is $\langle1,\ldots,1\rangle.$ Note that we immediately know this eigenvector is different from our others because the other eigenvectors are from the kernel of $J.$ Anyways, we need to compute
\[A_q\begin{bmatrix}1 \\ \vdots \\ 1\end{bmatrix}.\]
Well, each component will end up being the number of adjacencies to a given vertex of $G_q,$ by properties of he adjacency matrix. (Just look at the multiplication work.) In other words, fixing a one-dimensional subspace $L\in V(G_p),$ we want the number of one-dimensional subspaces in $L^\perp.$ Well, we actually computed this in the previous lemma as
\[\frac{q^{d-1}-1}{q-1},\]
mostly because $\dim L^\perp=d-1.$ Because this is invariant of our choise of $L,$ we note
\[A_q\begin{bmatrix}1 \\ \vdots \\ 1\end{bmatrix}=\frac{q^{d-1}-1}{q-1}\begin{bmatrix}1 \\ \vdots \\ 1\end{bmatrix},\]
and so we have found our last eigenvalue. $\blacksquare$

We quickly remark that we could have found the last eigenvalue, up to sign, by noting the eigenvalue of $\langle1,\ldots,1\rangle$ is $\frac{q^d-1}{q-1}$ in $J$ (this is $J$'s width), adding in the $q^{d-2}$ to get the corresponding eigenvalue for $A_q^2,$ and finish by taking the square-root. However, we'd have to do the above computation anyways to get the sign.

Anyways, we, as before, let $a$ be the number of eigenvectors with eigenvalue $+\sqrt{q^{d-2}}$ and $b$ be the number of eigenvectors with eigenvalue $-\sqrt{q^{d-2}}.$ Then we know
\[\op{trace}A_q=\frac{q^{d-1}-1}{q-1}+(a-b)\sqrt{q^{d-2}}\]
by summing the eigenvalues. Returning to the original problem, we see that
\[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}=1+(q-1)\left(\frac{q^{d-1}-1}{q-1}+(a-b)\sqrt{q^{d-2}}\right),\]
which simplifies to
\[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}=q^{d-1}+(q-1)(a-b)\sqrt{q^{d-2}}.\]
As an example of what this can do, we immediately have the following.
\begin{proposition}
    Fix $q$ a non-square prime-power and $d$ an odd dimension. Then
    \[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}=q^{d-1}.\]
\end{proposition}
Starting from our last equation, we note that $q^{d-2}$ is a non-square raised to an odd power and is therefore odd. It follows $\sqrt{q^{d-2}}$ is irrational, but
\[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}=q^{d-1}+(q-1)(a-b)\sqrt{q^{d-2}}\]
needs to be an integer, so we had better have $a-b=0.$ This finishes. $\blacksquare$

Or we can take a slightly different approach to say the following.
\begin{proposition}
    Fix $q$ a prime-power and $d$ a dimension. If either $q$ is square or $d$ even, then
    \[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}\equiv0\pmod{\sqrt{q^{d-2}}}.\]
\end{proposition}
Note that the condition $q$ square or $d$ even is meant to ensure $q^{d-2}$ is itself a square so that $\sqrt{q^{d-2}}$ is integral. Then we note the equation
\[\#\left\{v\in\FF_q^d:\langle v,v\rangle=0\right\}=q^{d-1}+(q-1)(a-b)\sqrt{q^{d-2}}\]
immediately gives the desired upon taking$\pmod{\sqrt{q^{d-2}}}.$ There is perhaps a technicality to note $q^{(d-2)/2}\mid q^{d-1},$ but this is equivalent to $2(d-1)\ge d-2,$ or $d\ge0,$ so we don't have anything to worry about. $\blacksquare$

Very quickly, we remark that the above result does guarantee that there are a nonzero number o nonzero solutions, for being divisible by $\sqrt{q^{d-2}}$ means that the number of nonzero solutions is $-1\pmod{q^{d-2}}.$ However, this divisibility condition isn't great for size because we expect there to be about $q^{d-1}$ total solutions anyways (as in the other case).

This last result is not really very impressive, but it is cute. As alluded, the problem with square $q$ or even $d$ is that $\sqrt{q^{d-2}}$ is actually an integer, so we can't blatantly ignore it as in the other case. And indeed, the distinction is not a mere limitation to the technique but actually representative of a deeper problem.
\begin{example}
    Fix $d=2$ and $q=p$ an odd prime. Then note $\langle v,v\rangle=0$ for $v=(x,y)\in\FF_p^2$ is equivalent to
    \[x^2\equiv-y^2\pmod p.\]
    If $x=0,$ we had better have $y=0,$ so we always at least have that solution. Otherwise, nonzero solutions require $(x/y)^2\equiv-1\pmod p$ and so require a square-root of $-1.$ So the number of desired $v\in\FF_p62$ is dependent on $\left(\frac{-1}p\right),$ or equivalently on $p\pmod4.$
\end{example}
We can actually complete this example. We have two cases
\begin{itemize}
    \item If $p\equiv3\pmod4$ so that $\left(\frac{-1}p\right)=-1,$ then there will be no nonzero solutions because, as discussed, nonzero $v=(x,y)$ satisfying gives $(x/y)\equiv-1.$ So we have only $\boxed1$ solution here.
    \item Else, $p\equiv1\pmod4$ so that $\left(\frac{-1}p\right)=1$; fix $k:=\sqrt{-1}$ to be a square-root of $-1.$ As above, any nonzero solution $v=(x,y)$ gives
    \[(x/y)^2\equiv-1\pmod p.\]
    In particular, $x/y$ is equal to $k$ or $-k,$ so $x\equiv\pm ky.$ Note $k$ is nonzero, so as $y$ varies across its $p-1$ nonzero options, we get two different nonzero values of $x$ for each. So the total number of solutions is $1+(p-1)\cdot2=\boxed{2p-1}.$
\end{itemize}
In all cases, we do see that the number of solutions is (sadly) divisible by $\sqrt{p^{2-2}}=1.$

\subsubsection{June 6th}
Today I learned that a number of geometric concepts can be defined purely formally from an arbitrary distance metric, from the ARML Power Round. As a fair warning, I am under the impression that this is highly nonstandard. For starters, we will permit distances that need not be positive reals and take the following definition.
\begin{definition}
    Fix a field $F$ of characteristic not $2.$ Then for $a,b,c\in F$ with $bc\ne0,$ we define
    \[\cos([a,b,c]):=\frac{b^2+c^2-a^2}{2bc}.\]
    This is to be thought of as the cosine of the angle opposite the side of length $a$ in a triangle of side lengths $a,b,c.$
\end{definition}
Here, the elements of $F$ are acting like the distances in a triangle. Notably, we are forcing some Euclidean structure onto our geometry, even though it might not fit nicely. For example, even with just $a,b,c\in\RR,$ we can have
\[\cos[10,1,1]=\frac{1^2+1^2-10^2}{2\cdot1\cdot1}=-49,\]
which is quite strange. The reason this went bad is, of course, that our $[a,b,c]$ is not required to satisfy the triangle inequality. I think there is a way to recover the triangle inequality if we add in a valuation, but it seems painful.

Even so, we might expect some of the definitions to interact in a purely formal way, if we just define them purely formally as well. So here is squared area, using Heron's formula.
\begin{definition}
    Again fix a field $F$ of characteristic not $2.$ Then for $a,b,c\in F,$ we define the squared area of $[a,b,c]$ as
    \[\op{sqarea}([a,b,c]):=s(s-a)(s-b)(s-c),\]
    where $s=\frac{a+b+c}2$ is the (formal) semiperimeter.
\end{definition}
Note that this definition is symmetric, so we don't have to worry about the order of $[a,b,c]$ here. Also of note is that we would like to take the square-root of the squared area to get normal area, but we are not guaranteed that square-roots exist or even are well-defined in $F.$

As an example of these definitions interacting in a purely formal way, we recall the area formula
\[K\stackrel?=\frac12bc\sin\angle A,\]
where $\angle A=[a,b,c]$ is the angle. However, we don't have a $\sin,$ nor do we have a non-squared area, so we solve both problems simultaneously by squaring both sides and then using the Pythagorean identity. This looks like
\[K^2\stackrel?=\frac14b^2c^2\sin^2([a,b,c])=\frac14b^2c^2\left(1-\cos^2([a,b,c])\right).\]
Of course, we do not define the area to behave in this way, so this is something that needs proof.
\begin{proposition}
    Fix $F$ a field of characteristic not $2.$ Given $a,b,c\in F$ with $bc\ne0,$ we have
    \[\op{sqarea}([a,b,c])=\frac14b^2c^2\left(1-\cos^2([a,b,c])\right).\]
\end{proposition}
This manipulation is roughly the end of the proof of Heron's formula. Starting from the right-hand side, we note that
\[1-\cos^2([a,b,c])=1-\left(\frac{b^2+c^2-a^2}{2bc}\right)^2,\]
so
\[\frac14b^2c^2\left(1-\cos^2([a,b,c])\right)=\frac1{16}\left((2bc)^2-\left(b^2+c^2-a^2\right)^2\right).\]
Factoring by difference of squares, this looks like
\[\frac1{16}\left(2bc+\left(b^2+c^2-a^2\right)\right)\left(2bc-\left(b^2+c^2-a^2\right)\right)=\frac1{16}\left((b+c)^2-a^2\right)\left(a^2-(b-c)^2\right).\]
Using difference of squares again, this is
\[\frac1{16}(a+b+c)(-a+b+c)(a-b+c)(a+b-c).\]
Factoring in the $16,$ we can write this all as
\[\frac{a+b+c}2\left(\frac{a+b+c}2-a\right)\left(\frac{a+b+c}2-b\right)\left(\frac{a+b+c}2-c\right).\]
This is what we wanted. $\blacksquare$

As a bonus, we get the law of sines for free now. (The law of cosines is built into our definition of cosine.) We would like to write
\[\frac{\sin([a,b,c])}a\stackrel?=\frac{\sin([b,c,a])}b,\]
but again $\sin$ is somewhat meaningless. Squaring and using the Pythagorean identity, we want the following.
\begin{proposition}
    Fix $F$ a field of characteristic not $2.$ Given $a,b,c\in F$ with $abc\ne0,$ we have
    \[\frac{1-\cos^2([a,b,c])}{a^2}=\frac{1-\cos^2([b,c,a])}{b^2}.\]
\end{proposition}
We are allowed to not care about division by $0$ with $a,b,c$ because they are nonzero. So we multiply both sides by $\frac14a^2b^2c^2$ so that it suffices to show
\[\frac14b^2c^2\left(1-\cos^2([a,b,c])\right)\stackrel?=\frac14c^2a^2\left(1-\cos^2([b,c,a])\right).\]
However, the left-hand side is $\op{sqarea}([a,b,c]),$ and the right-hand side is $\op{sqarea}([b,c,a]).$ But we already remarked that the definition of $\op{sqarea}$ is symmetric in its arguments, so these are equal, as desired. $\blacksquare$

Of course, the above proof is just rewriting the proof of the law of sines from writing
\[\frac12bc\sin\angle A=\frac12ca\sin\angle B\]
directly implying $\sin\angle A/a=\sin\angle B/b.$ What is remarkable here is that we managed to codify this intuition into a purely formal manipulation of our algebraic definitions.

\subsubsection{June 6th}
Today I learned the classification of primes in polynomial rings over principal ideal domains, from Undergraduate Commutative Algebra. Before continuing, we collect some facts that should be written down somewhere. We begin with the equivalence of prime and irreducible in unique factorization domains.
\begin{lemma}
    Suppose $p$ is a non-unit prime in the integral domain $\mathcal O_K.$ Then $p$ is irreducible.
\end{lemma}
Indeed, suppose $p=ab$ for some $a,b\in\mathcal O_K.$ We need to show that one of $a$ or $b$ is a unit. Well, $p=ab$ means that $p\mid ab,$ so $p\mid a$ or $p\mid b$ by definition of prime. Without loss of generality, we say $p\mid a,$ so let $a=pk.$ Thus we see that
\[p=(pk)b.\]
Because $p$ is prime, it is nonzero, so the fact $\mathcal O_K$ is an integral domain lets us cancel this to $bk=1.$ Thus, $b\mid1,$ so $b\in\mathcal O_K^\times$ is a unit.

Technically, we also need to know that $a$ is not a unit for $p$ to be irreducible. Well, if $a$ is also a unit, then $p=ab$ is a unit as well, and we supposed that $p$ is not a unit as a hypothesis. $\blacksquare$

\begin{lemma}
    Suppose $q$ is an irreducible in the unique factorization domain $\mathcal O_K.$ Then $q$ is a prime.
\end{lemma}
Suppose $q\mid ab$ with $a,b\in\mathcal O_K.$ To use the unique factorization condition, we note that
\[\frac{ab}q=u\prod_{p\text{ irred.}}p^{\nu_p(ab/q)},\]
where all but finitely many of the $\nu_p(ab)$ are zero. In particular, we note that
\[ab=u\left(q\prod_{p\text{ irred.}}p_k\right)\]
is a perfectly valid factorization into irreducibles of $ab.$ However, we can also write
\[a=u_a\prod_{p\text{ irred.}}p^{\nu_p(a)}\quad\text{and}\quad b=u_b\prod_{p\text{ irred.}}p^{\nu_p(b)}\]
where again all but finitely many of the $\nu_p(a)$ and $\nu_p(b)$ are zero. This means that
\[ab=u_au_b\prod_{p\text{ irred.}}p^{\nu_p(a)+\nu_p(b)}\]
is another valid factorization into irreducibles of $ab.$ Comparing our two factorizations, we see that $q$ must be conjugate to some irreducible $p$ in the second factorization, and that irreducible $p$ needs to have $\nu_p(a)+\nu_p(b)$ positive because it is positive in the first factorization. But if
\[\nu_p(a)+\nu_p(b)>0,\]
then at least one of $\nu_p(a)$ or $\nu_p(b)$ must be positive. Without loss of generality, we assert $\nu_p(a)>0,$ but then $p\mid a,$ and because $q$ is conjugate to $p,$ we also have $q\mid a.$ This completes the proof. $\blacksquare$

We will also need to know that primes are maximal in principal ideal domains. Note that maximal ideals are integral domains because the corresponding quotient is a field and hence an integral domain.
\begin{lemma}
    Suppose $(p)$ is a prime ideal in the principal ideal domain $\mathcal O_K.$ Then $(p)$ is also maximal.
\end{lemma}
We show this directly. Find an ideal between $(p)$ and $\mathcal O_K,$ and name it $(a)$ because it must be principal. Namely, we know that $(p)\subseteq(a)\subseteq\mathcal O_K$ and would like to show $(a)=(p)$ or $(a)=\mathcal O_K.$

Well, $p\in(a)$ implies that $p=ab$ for some $b\in\mathcal O_K.$ However, $p$ prime implies $p$ irreducible, so $p=ab$ requires either $a$ or $b$ to be a unit. We take these cases one at a time.
\begin{itemize}
    \item If $a$ is a unit, then $(a)=\mathcal O_K.$ Using more words, $a\mid1$ requires some $k$ with $1=ak,$ so $1\in(a),$ so $\mathcal O_K=(1)\subseteq(a).$
    \item If $b$ is a unit, then $(a)=(p).$ Indeed, $a=pb^{-1},$ so $a\in(p),$ so $(a)\subseteq(p),$ but we supposed that $(p)\subseteq(a)$ already.
\end{itemize}
Having dealt with all cases, we are done here. $\blacksquare$

Now, in what follows, we fix $\mathcal O_K$ a principal ideal domain and $K$ its field of fractions. We know $\mathcal O_K$ is a unique factorization domain, so $\mathcal O_K[x]$ is a unique factorization domain as well. We are interested in classifying the primes/irreducibles in $\mathcal O_K[x].$ So fix $\mf p$ a prime ideal.

Getting rid of the easy cases, we begin by taking $\mf p=(\pi)$ principal with $\pi\in\mathcal O_K[x].$ We can always have $\mf p=(0),$ but if $\pi$ is a nonzero non-unit prime of the unique factorization domain $\mathcal O_K[x],$ then this is equivalent to $\pi$ being irreducible. So
\[(\pi)\text{ prime}\iff\pi=0\text{ or }\pi\text{ prime/irreducible}.\]
It can be checked manually that $\mathcal O_K[x]/(\pi)$ is an integral domain, but we don't have to.

It remains to deal with the case where $\mf p$ is a prime ideal but not principal. The intuition for this case is to take $\mathcal O_K=\CC[y]$ so that the maximal ideals in $\mathcal O_K[x]=\CC[x,y]$ all look like $(x-\alpha,y-\beta)$ with $\alpha,\beta\in\CC.$ What does this mean? Well,
\[(x-\alpha,y-\beta)\]
looks like the expansion of the maximal ideal $(y-\beta)\subseteq\mathcal O_K,$ so we would like to show that $\mf p$ looks like this: some kind of expansion of a maximal ideal in $\mathcal O_K.$

For this, we pick up the following lemma.
\begin{lemma}
    Fix $\mathcal O_K$ a principal ideal domain. Suppose that $\mf p$ is a non-principal ideal of $\mathcal O_K[x].$ Then $\mf p\cap(\mathcal O_K\setminus\{0\})$ is nonempty.
\end{lemma}
Fix $K$ the field of fractions of $\mathcal O_K[x].$ As a general approach, what we want to do is say that $\mf p$ non-principal means that it has elements $f_1$ and $f_2$ which are coprime in a $\mathcal O_K[x]$ sort of way, so we can reduce them to elements coprime in a $K[x]$ sort of way, and then Bezout's lemma says there are $c_1,c_2\in K[x]$ such that
\[c_1f_1+c_2f_2=1\]
because $K[x]$ is a principal ideal domain. Clearing fractions to push everything into $\mathcal O_K[x]$ gives us our element of $\mathcal O_K[x]\setminus\{0\}$ in $\mf p.$

We begin by looking for those elements $f_1$ and $f_2$ of $\mf p$ which share no non-unit common factor of $\mathcal O_K[x].$ This is surprisingly annoying, so if the reader would like to skip this, we denote where the proof of this begins and ends.
\begin{lemma}
    Using everything as defined above, we can find $f_1$ and $f_2$ in $\mf p$ which share no non-unit common factors.
\end{lemma}
Pick up any nonzero $f_1\in\mf p$ and look at its decomposition into irreducibles. (Note $\mf p\ne\{0\}$ because $\mf p$ is not principal.) One of those irreducibles needs to be in $\mf p$ because $\mf p$ is a prime (after all, the entire product $f_1\in\mf p$), so say $p_1\in\mf p$ is one such irreducible.

Now, $\mf p$ is not principal, so $\mf p\setminus(p_1)$ must be nonempty. So find some $f_2\in\mf p\setminus(\pi_1)$ and play the same trick: surely $f_2$ is nonzero because $0\in(p_1),$ so we can look at the factorization of $f_2$ into irreducibles. This factorization is just a glorified product, so one of the irreducible factors---say, $p_2$---lives in $\mf p.$

Thus, we have two irreducibles $p_1$ and $p_2$ in $\mf p,$ where $p_2\in\mf p\setminus(p_1).$ We claim that $p_1$ and $p_2$ have no non-unit common factors. Indeed, if we find $g\in\mathcal O_K[x]$ satisfying $g\mid p_1$ and $g\mid p_2,$ then dissolving everything into its prime factorization means that
\[\nu_p(g)\le\min\{\nu_p(p_1),\nu_p(p_2)\}\]
for each irreducible $p$ of $\mathcal O_K[x].$ However, $p_1=p_1$ is a valid factorization into irreducibles for $p_1,$ so $\nu_p(g)=\nu_p(p_1)=0$ for any $p$ not conjugate to $p_1.$ Similarly, $\nu_p(g)=0$ for any prime not conjugate to $p_2.$ But if $p$ is conjugate to both $p_1$ and $p_2,$ then $p_1=u_1p$ and $p_2=u_2p$ with $u_1,u_2\in\mathcal O_K[x]^\times,$ implying
\[p_2=u_2p=u_2u_1^{-1}p_1\in(p_1),\]
which is false by construction. So no prime is conjugate to both $p_1$ and $p_2,$ implying $\nu_p(g)=0$ for all irreducibles $p.$ It follows the factorization of $g$ into irreducibles features no irreducibles, so $g$ is a unit. $\blacksquare$

As our promised next step, we lift the condition that $f_1$ and $f_2$ share no non-unit common factors in $\mathcal O_K[x]$ to sharing no non-unit common factors in $K[x].$ This lifting sounds like a job for Gauss's lemma, so indeed we will use Gauss's lemma.

Suppose $g\in K[x]$ is a common factor of $f_1$ and $f_2$ in $K[x],$ and we would like to show $\deg g=0$ so that $g$ is a unit of $K.$ We need to move this argument into $\mathcal O_K[x]$ to use our hypothesis on $f_1$ and $f_2,$ so we write $f_1=gg_1$ and $f_2=gg_2$ and use our lemma from a few days ago to say
\[g=ag',\quad g_1=a_1g_1',\quad g_2=a_2g_2'\]
where $a,a_1,a_2\in K^\times$ and $g',g_1',g_2'\in\mathcal O_K[x]$ are primitive. We claim that $g'$ is a common factor of $f_1$ and $f_2$ in $\mathcal O_K[x],$ which will finish because it requires that $g'$ is a unit of $\mathcal O_K[x],$ so $g=ag'\in K$ is a unit of $K[x].$ Indeed, we'll show $g'\mid f_1,$ from which $g'\mid f_2$ will follow similarly. We write
\[f_1=gg_1=(aa_1)(g'g_1').\]
Now, $g'g_1'$ is primitive by Gauss's lemma (here is here we used it). To finish here, we need to know $aa_1\in\mathcal O_K.$ A quick way to see this is that, for any prime $p$ of $\mathcal O_K,$ we know that the minimum of $\nu_p$ across the coefficients of $f_1$ and $(aa_1)(g'g_1')$ are equal. Well, $f_1\in\mathcal O_K[x],$ so the minimum is nonnegative, and $g'g_1'$ is primitive, so the minimum is $0.$ Thus, $\nu_p(aa_1)\ge0$ for each $p,$ so $aa_1\in\mathcal O_K.$

And as our promised final step, we note that $f_1$ and $f_2$ sharing no non-unit common factors in $K[x]$ means that the ideal $(f_1,f_2)=K[x]$ in $K[x].$ In particular, we can find $c_1,c_2\in K[x]$ such that
\[c_1f_1+c_2f_2=1.\]
Multiplying together all the denominators of $c_1$ and $c_2$ gives us a giant nonzero constant $C\in\mathcal O_K$ such that $Cc_1\in\mathcal O_K[x]$ and $Cc_2\in\mathcal O_K[x],$ so
\[(Cc_1)f_1+(Cc_2)f_2=C\]
lives in $(f_1,f_2)$ in $\mathcal O_K[x]$ and therefore lives in $\mf p.$ This finishes the proof. $\blacksquare$

We can now do some analysis on what $\mf p$ looks like. We can get to the following with some small coercion.
\begin{lemma}
    Fix $\mathcal O_K$ a principal ideal domain and $\mf p$ a non-principal prime ideal. We can find some prime element $p\in\mathcal O_K\cap\mf p$ and polynomial $\pi\in\mathcal O_K[x]$ such that $\overline\pi\in\mathcal O_K/(p)[x]$ is irreducible.
\end{lemma}
It is nonzero by the lemma we worked so hard for, and it is also a prime $\mathcal O_K$-ideal. That it is an ideal follows because $\mf p$ is an $\mathcal O_K[x]$-ideal and $\mathcal O_K$ a ring. It is not all of $\mathcal O_K$ because then $1\in\mf p,$ which is problematic. And it is prime is also inherited: if $a,b\in\mathcal O_K$ have
\[ab\in\mf p\cap\mathcal O_K,\]
then $ab\in\mf p$ requires $a\in\mf p$ or $b\in\mf p.$ We already know $a,b\in\mathcal O_K,$ so $a\in\mf p\cap\mathcal O_K$ or $b\in\mf p\cap\mathcal O_K.$ Anyways, the point of all of this is to say that
\[\mf p\cap\mathcal O_K=(p)\]
for some prime $p$ of $\mathcal O_K$ because $\mathcal O_K$ is a principal ideal domain. This gives the first part of the desired.

Continuing, we note $\mf p$ is not itself principal, so we can find some $f\in\mf p\setminus(p).$ We might be tempted to pick up an irreducible $\pi$ from the factorization of $f\in\mathcal O_K[x]$ and analyze $(p,\pi).$ However, $(p,\pi)$ need not even be prime in this construction: choosing $\mathcal O_K$ with $p=2$ and $\pi=x^2+1,$ we see
\[(x+1)^2\in\left(2,x^2+1\right)\subseteq\ZZ[x],\]
but of course $x+1\not\in\left(2,x^2+1\right).$

The correct thing to do is to note that $(p)$ is prime in a principal ideal domain $\mathcal O_K$ and hence maximal, so $\mathcal O_K/(p)[x]$ is a unique factorization domain (a polynomial ring of a field). Instead of picking up any irreducible dividing $f,$ we will pick up an irreducible dividing its equivalence class $\overline f\in\mathcal O_K/(p)[x].$ We quickly establish $\deg\overline f>0$: write
\[f(x)=\sum_{k=0}^da_kx^k\in\mathcal O_K[x],\]
If all $a_k$ for $k>0$ live in $(p),$ then $a_0=f(x)-p\sum_{k=1}^d(a_k/p)x^k\in(p)$ as well, meaning $f\in(p).$ We chose $f\not\in(p),$ so there is some $k>0$ with $a_k\not\in(p),$ meaning $\deg\overline f\ge k>0.$ So we are allowed to write
\[\overline f=\prod_{k=1}^n\overline\pi_k\]
for a positive number of irreducible polynomials $\{\overline\pi_k\}_{k=1}^n\subseteq\mathcal O_K/(p)[x].$ We would like one of the $\pi_\bullet$ to live into $\mf p.$ Well, the above equality lets us find some $q\in(p)$ with
\[\prod_{k=1}^n\pi_k=f-pq\in\mf p,\]
so $\mf p$ prime lets us find some $\pi$ in the product such that $\pi\in\mf p.$ This gives the second part of the desired. $\blacksquare$

It turns out that the above lemma the $p$ and $\pi$ of the above lemma actually generate $\mf p.$ We have the following.
\begin{lemma}
    Using everything as defined in the previous lemma, we have $\mf p=(p,\pi).$
\end{lemma}
Indeed, we claim that $(p,\pi)$ is maximal, so $(p,\pi)\subseteq\mf p\subsetneq\mathcal O_K$ ($\mathcal O_K$ is not a prime ideal), from what it follows $\mf p=(p,\pi).$ To show the claim, we simply note
\[\frac{\mathcal O_K[x]}{(p,\pi)}\cong\frac{\mathcal O_K/(p)[x]}{(\overline\pi)}\cong\mathcal O_K/(p)[\alpha],\]
where $\alpha$ is a root of $\overline\pi\in\mathcal O_K/(p)[x].$ With more words, $p$ prime in $\mathcal O_K$ means $(p)$ is maximal, so $\mathcal O_K/(p)$ is a field. Adjoining the transcendental $x$ and then modding out by $\overline\pi$ is equivalent to adjoining a root $\overline\pi$ to the field $\mathcal O_K/(p),$ remaining a field. Thus, $(p,\pi)$ is maximal. $\blacksquare$

Putting this all together, we have the following theorem.
\begin{theorem}
    Fix $\mathcal O_K$ a principal ideal domain. All prime ideals $\mf p$ of $\mathcal O_K[x]$ fall into exactly one of the following categories.
    \begin{itemize}
        \item $\mf p$ is principal and looks like $(0)$ or $(f)$ for some prime/irreducible $f\in\mathcal O_K[x].$
        \item $\mf p$ is maximal and looks like $(p,\pi)$ where $p\in\mathcal O_K$ is prime, and $\pi$ reduces$\pmod p$ to some irreducible $\overline\pi\in\mathcal O_K/(p)[x].$
    \end{itemize}
\end{theorem}
This follows from the above discussion, by doing casework on if $\mf p$ is principal or not. The first part of the theorem comes from our discussion of the principal case, and the above lemma covers the second part of the theorem for when $\mf p$ is non-principal.

Further, note that all the principal primes are not maximal: if $\mf p=(p)$ with $\deg p=0,$ then $p\in\mathcal O_K,$ so $p$ is prime, and then $(p,x)$ is a strictly larger proper ideal. Indeed, $x\in\mathcal O_K[x]\setminus(p)$ while $1+p\in\mathcal O_K[x]\setminus(p,x).$

Similarly, if $\mf p=(f)$ with $\deg f>0,$ then we can pick up a prime $p\in\mathcal O_K,$ and $(p,f)$ is a strictly larger proper ideal. Indeed, $(f)$ has no nonzero elements in $\mathcal O_K,$ so $p\in\mathcal O_K[x]\setminus(f).$ And further, $1+p$ is not in $(p,f)$ either: note $1+p=ap+bf$ needs $b=0$ because $\deg(1+p)=0,$ but $1+p=ap$ means $p$ is a unit.

The point of the above is to say that our cases really are disjoint: being non-principal and prime is really the same as being maximal. The previous discussion gives the forward direction, and the above two paragraphs give the reverse direction. $\blacksquare$

What's nice about the above abstraction is that we get to leverage both geometric intuition from $\mathcal O_K=\CC[y]$ and more number-theoretic intuition from $\mathcal O_K=\ZZ.$ To be more explicit, we get a nice picture of $\op{Spec}\CC[y,x]$ as consisting of the maximal ideals
\[(y-\beta,x-\alpha)\]
because $\CC$ is algebraically closed. These can be associated with the points $(\beta,\alpha)\in\CC^2,$ so we can picture our maximal ideals as points. Then the principal ideals $(f)$ with $f\in\CC[y,x]$ prime/irreducible more or consist of ``curves'' in the plane that cannot be reduced into components.

Meanwhile we also get a nice number-theoretic statement about what the primes of $\ZZ[x]$ or $\ZZ[i][x]$ look like, which is nice. I'm not sure how much I care about these primes, and it disturbs me a bit that this much work was needed to classify them, but it is nice to have it and in this generality.

\subsubsection{June 7th}
Today I learned a truly clever proof to the Spectal theorem, from THE BOOK of course. As a brief outline, we will translate diagonalization by orthogonal matrices into the minimization of some function and then minimize that function by a compactness argument. Here are the definitions of our main characters.
\begin{definition}
    A matrix $M\in F^{n\times n}$ is said to be symmetric if and only if its coefficients satisfy $M_{k,\ell}=M_{\ell,k}$ for each valid $k,\ell.$
\end{definition}
Our main claim is that we can diagonalize real symmetric matrices by orthogonal ones, so we need to pick up the definition of orthogonal.
\begin{definition}
    A matrix $Q\in F^{n\times n}$ is said to be orthogonal if and only if $Q^\intercal=Q^{-1}.$ We will denote the set of orthogonal matrices in $\RR^{n\times n}$ by $O(n).$
\end{definition}
We remark that $Q\in O(n)$ is equivalent to $Q^\intercal Q=I.$ If we write out the matrix multiplication using column vectors, we see
\[Q^\intercal Q=\begin{bmatrix}
    - & v_1 & - \\
    & \vdots & \\
    - & v_n & -
\end{bmatrix}\begin{bmatrix}
    | & & | \\
    v_1 & \cdots & v_n \\
    | & & |
\end{bmatrix}=\begin{bmatrix}
    \langle v_1,v_1\rangle & \cdots & \langle v_n,v_1\rangle \\
    \vdots & \ddots & \vdots \\
    \langle v_1,v_n\rangle & \cdots & \langle v_n,v_n\rangle
\end{bmatrix},\]
where $\langle,\rangle$ is the usual inner product. This expansion gives the following statement.
\begin{proposition}
    A matrix $Q\in\RR^{n\times n}$ is orthogonal if and only if its column vectors form an orthonormal basis.
\end{proposition}
Indeed, from the above expansion of $Q^\intercal Q,$ we see that $Q^\intercal Q=I$ if and only if $\langle v_k,v_\ell\rangle=1_{k=\ell},$ which is exactly what is required for the column vectors to be orthonormal.

That they form a basis follows from the fact that mutually orthogonal nonzero vectors are all linearly independent: dot any relation of mutually orthogonal nonzero vectors $\{v_1\}_{k=1}^m$
\[a_1v_1+\cdots+a_mv_m=0\]
by $v_k$ to get all terms to vanish except $a_k\langle v_k,v_k\rangle=0,$ yielding $a_k=0,$ for any $k.$ Further, the columns of $Q$ have a full $n$ linearly independent vectors, so they must be a basis. $\blacksquare$

Indeed, given a real symmetric matrix $A\in\RR^{n\times n},$ we define
\[\op{od}(A):=\sum_{1\le k\ne\ell\le n}A_{k,\ell}^2\]
to be the sum of the squares of the off-diagonal entries. Roughly speaking, $\op{od}$ is trying to measure how far $A$ is from diagonal: $\op{od}(A)=0$ if and only if $A$ is diagonal. To measure diagonalization, we define the function
\[\op{attempt}_Q(A):=\op{od}\left(QAQ^\intercal\right)\]
for all matrices $Q\in\RR^{n\times n}.$ For orthogonal matrices $Q,$ we know $Q^\intercal=Q^{-1},$ so the above really is an attempt at diagonalization: if $\op{attempt}_Q(QAQ^{-1})=0$ then $QAQ^{-1}$ is diagonalized. We use $Q^\intercal$ instead of $Q^{-1}$ directly because we will want $\op{attempt}_\bullet(A)$ to be defined over all $\RR^{n\times n}$ for our compactness argument.

For a compactness argument to achieve a minimum, we need a ``do-better'' lemma. This is the meat of the argument.
\begin{lemma}
    Fix a real symmetric matrix $A\in\RR^{n\times n}$ with $\op{od}(A)>0.$ We can construct an orthogonal matrix $Q\in O(n)$ with $\op{attempt}_Q(A)<\op{od}(A).$
\end{lemma}
The fact that $\op{od}(A)>0$ means that we can find a nonzero entry $A_{K,L}$ with $K<L.$ (We can find $A_{K,L}\ne0$ with $K\ne L$; if $K>L,$ then note $A_{L,K}=A_{K,L}.$) The main idea is to apply a ``rotation matrix'' to make $A_{K,L}$ smaller while incrementing the elements on the diagonal. That is, we take some angle $\theta$ to be chosen later and let
\[Q:=\left[\begin{array}{*{20}c}
    1 \\
    & \ddots \\
    & & 1 \\
    & & & \cos\theta & & & & -\sin\theta \\
    & & & & 1 \\
    & & & & & \ddots \\
    & & & & & & 1 \\
    & & & \sin\theta & & & & \cos\theta \\
    & & & & & & & & 1 \\
    & & & & & & & & & \ddots & \\
    & & & & & & & & & & 1 \\
\end{array}\right],\]
where the entries of interest are $Q_{K,K}=\cos\theta,$ $Q_{K,L}=-\sin\theta,$ $Q_{L,K}=\sin\theta,$ and $Q_{L,L}=\cos\theta.$ Otherwise, $Q_{k,\ell}=1_{k=\ell}$ when $\{k,\ell\}\not\subseteq\{K,L\}.$ Very quickly, we show that $Q$ is orthogonal. Indeed,
\[\left(Q^\intercal Q\right)_{k,m}=\sum_{\ell=1}^n\left(Q^\intercal\right)_{k,\ell}Q_{\ell,m}=\sum_{\ell=1}^nQ_{\ell,k}Q_{\ell,m}.\]
We have the following cases.
\begin{itemize}
    \item If $k\not\in\{K,L\},$ then $Q_{\ell,k}=1_{k=\ell}$ always. So the only term of the sum we have to worry about is $Q_{k,k}Q_{k,m}=1\cdot 1_{k,m}=1_{k=m},$ as required.
    \item If $m\not\in\{K,L\},$ then again $Q_{\ell,m}=1_{\ell=m}$ always. So the only term of the sum we have to worry about is $Q_{m,k}Q_{m,m}=1_{m=k}\cdot1=1_{k=m},$ as required.
    \item If both $k,m\in\{K,L\},$ then we still know $Q_{k,\ell}=0$ when $\ell\not\in\{K,L\}.$ So the only terms of the sum we care about are
    \[Q_{K,k}Q_{K,m}+Q_{L,k}Q_{L,m}.\]
    When $k=m=K,$ this reads $\cos^2\theta+\sin^2\theta=1.$ When $k=m=L,$ this reads $(-\sin\theta)^2+\cos^2\theta=1.$ And when $k\ne m,$ this reads $\cos\theta\sin\theta-\sin\theta\cos\theta=0.$ In all cases, we still have $1_{k=m}.$
\end{itemize}
It follows that $Q^\intercal Q=I,$ so $Q$ is indeed orthogonal.

It remains to compare $\op{attempt}_Q(A)=\op{od}(QAQ^\intercal)$ with $\op{od}(A).$ Well, note that
\[\left(QAQ^\intercal\right)_{j,m}=\sum_{k,\ell}Q_{j,k}A_{k,\ell}\left(Q^\intercal\right)_{\ell,m}=\sum_{k,\ell}Q_{j,k}A_{k,\ell}Q_{m,\ell}.\]
Also, observe that $Q^\intercal AQ$ is symmetric: $\left(Q^\intercal AQ\right)^\intercal=Q^\intercal A^\intercal Q^{\intercal\intercal}=Q^\intercal AQ,$ so we only have to compute half of these entries. We proceed, as expected, with more casework.
\begin{itemize}
    \item If both $j,m\not\in\{K,L\},$ then $Q_{j,k}=1_{j=k}$ and $Q_{m,\ell}=1_{m=\ell}$ for any $k$ or $\ell,$ so the only term of the sum we care about is $Q_{j,j}A_{j,m}Q_{m,m}=A_{j,m}.$ So the contribution to $\op{od}$ is the same.
    \item If $j=K$ but $m\not\in\{K,L\},$ then we care about $k\in\{K,L\}$ but require $\ell=m.$ So the sum looks like
    \[Q_{K,K}A_{K,m}+Q_{K,L}A_{L,m}=A_{K,m}\cos\theta-A_{L,m}\sin\theta.\]
    \item If $j=L$ but $m\not\in\{K,L\},$ then we care about $k\in\{K,L\}$ but require $\ell=m.$ So the sum looks like
    \[Q_{L,K}A_{K,m}+Q_{L,L}A_{L,m}=A_{K,m}\sin\theta+A_{L,m}\cos\theta.\]
\end{itemize}
We take a brief interlude to notice that the above two cases accumulate to
\[(A_{K,m}\cos\theta-A_{L,m}\sin\theta)^2+(A_{K,m}\sin\theta+A_{L,m}\cos\theta)^2,\]
which expands into $A_{K,m}^2\left(\cos^2\theta+\sin^2\theta\right)+A_{L,m}^2\left(\cos^2\theta+\sin^2\theta\right),$ which is just $A_{K,m}^2+A_{L,m}^2.$ So the terms with $j\in\{K,L\}$ but $m\not\in\{K,L\}$ still contribute the same amount to $\op{od}.$ We observe that by symmetry, we can swap $j$ and $m$ to say the same for $m\in\{K,L\}$ but $j\not\in\{K,L\}.$

It follows that any change to $\op{od}$ will come from when both $j,m\in\{K,L\}.$ Of course, $\op{od}$ doesn't care about when $j=m,$ so we only have to deal with $j=K$ and $m=L$ or $j=L$ and $m=K.$ But by symmetry, these entries are the same, so we only have one computation left.

In this case, we still only care about $k,\ell\in\{K,L\},$ which is
\[Q_{K,K}A_{K,K}Q_{L,K}+Q_{K,K}A_{K,L}Q_{L,L}+Q_{K,L}A_{L,K}Q_{L,K}+Q_{K,L}A_{L,L}Q_{L,L}.\]
Using what we know about these quantities, this turns into
\[A_{K,K}\cos\theta\sin\theta-A_{L,L}\cos\theta\sin\theta+A_{K,L}\left(\cos^2\theta-\sin^2\theta\right),\]
which is
\[\frac12(A_{K,K}-A_{L,L})\sin2\theta+A_{K,L}\cos2\theta.\]
We need this term to have a smaller contribution than $A_{K,L}^2,$ so we will choose $\theta$ to achieve this. Well, at $\theta=0,$ this term is $A_{K,L}\ne0,$ and when $\theta=\frac\pi2,$ this term is $-A_{K,L}<0,$ so there is some $\theta$ in between which causes this term to vanish completely.

For this chosen $\theta,$ we see that $\op{od}\left(QAQ^\intercal\right)$ is equal to $\op{od}(A)$ with the exception of the $A_{K,L}$ and $A_{L,K}$ terms, where $\op{od}(A)+A_{K,L}^2+A_{L,K}^2=\op{od}\left(QAQ^\intercal\right).$ This finishes the proof of the lemma. $\blacksquare$

With our ``do-better'' lemma in hand, we proceed with the compactness argument. We need something to be compact, and because we hope to only care about the orthogonal matrices $O(n),$ this will be our compact set.
\begin{lemma}
    The set of orthogonal matrices $O(n)$ is compact in $\RR^{n\times n}.$
\end{lemma}
We have to be a bit precise about what we mean by this. For clarity, we imagine vector-izing the matrices in $O(n)$ by reading down each row and turning the matrix from $\RR^{n\times n}$ to $\RR^{n^2},$ and we know what our topology in $\RR^{n^2}$ looks like. Equivalently, we could define the metric
\[d(A,B)=\sum_{1\le k,\ell\le n}(A_{k,\ell}-B_{k,\ell})^2\]
for matrices $A,B\in\RR^{n\times n}$ to induce the topology on $\RR^{n\times n}.$ This is the same as the topology from vector-izing because the vectorizing process essentially takes $A$ to $v$ with $v_{n(k-1)+\ell}=A_{k,\ell},$ so the metric topologies match.

Anyways, to show that $O(n)$ is compact it suffices to show that it is closed and bounded. For bounded, given $Q\in O(n),$ we remarked above that the column vectors of $Q$ form an orthonormal basis, so in particular they all have length $1.$ Thus,
\[d(Q,0)=\sum_{\ell=1}^n\underbrace{\sum_{k=1}^nQ_{k,\ell}^2}_{\text{column vec.}}=\sum_{\ell=1}^n1=n\]
is bounded. And as for closed, we note that we can write the condition $Q^\intercal Q=I$ for $Q\in O(n)$ as
\[\left(Q^\intercal Q\right)_{k,m}=\sum_{\ell=1}\left(Q^\intercal\right)_{k,\ell}Q_{\ell,m}=\sum_{\ell=1}Q_{\ell,k}Q_{\ell,m}=1_{k=m}\]
for any pair of indices $k,m.$ Looking in $\RR^{n^2}$ (say), this is just an intersection of $n^2$ equations, which we can show to be closed. A single one of these equations traces out a closed subset of $\RR^{n^2}$ because the function, for fixed $k$ and $m,$
\[Q\mapsto\sum_{\ell=1}Q_{\ell,k}Q_{\ell,m}\]
from $\RR^{n^2}\to\RR$ is just some large polynomial and hence continuous. Indeed, the preimage of the open set $\RR\setminus\{1_{k=m}\}$ is open, so the preimage of the complement closed set $\{1_{k=m}\}$ must also be closed.

Thus, each equation traces out some closed set, and the intersection of all of $n^2$ them will remain closed. Thus, $O(n)$ is closed and bounded and hence compact. $\blacksquare$

We are almost ready to finish. We quickly pick up a point-set lemma.
\begin{lemma}
    Fix topological spaces $X$ and $Y$ with a continuous map $f:X\to Y.$ The image $f(C)$ of a compact set $C$ is still compact.
\end{lemma}
We show this directly. Fix $\{U_\alpha\}_{\alpha\in\lambda}$ an open cover of $f(C).$ Because $f$ is continuous, we note that each $f^{-1}(U_\alpha)$ is still open, so
\[C\subseteq f^{-1}(f(C))\subseteq f^{-1}\left(\bigcup_{\alpha\in\lambda}U_\alpha\right)=\bigcup_{\alpha\in\lambda}f^{-1}(U_\alpha).\]
Thus, the $f^{-1}(U_\alpha)$ give an open cover of $C.$ Because $C$ is compact, we may extract some finite subcover $\left\{f^{-1}(U_k)\right\}_{k=1}^n\subseteq\left\{f^{-1}(U_\alpha)\right\}_{\alpha\in\lambda}$ around $C.$ Pushing this though $f,$ we see that $\left\{f\left(f^{-1}(U_\alpha)\right)\right\}_{k=1}^n=\left\{U_\alpha\right\}_{k=1}^n$ is a finite subcover of $f(C),$ so we are done here. $\blacksquare$

To finish, we fix $A$ some real symmetric matrix. Observe that the composite function $\op{attempt}_\bullet(A)$ can be thought of as
\[Q\longmapsto QAQ^\intercal\stackrel{\op{od}}\longmapsto\sum_{j\ne m}\left(QAQ^\intercal\right)_{j,m}^2=\sum_{j\ne m}\left(\sum_{k,\ell}Q_{j,k}A_{k,\ell}Q_{m,\ell}\right)^2_{j,m}.\]
This is just some large polynomial in the entries of $Q\in\RR^{n\times n}$ or $\RR^{n^2}$ into $\RR_{\ge0}$ and is therefore a continuous function. In particular, the image of the compact set $O(n)$ remains compact in $\RR_{\ge0},$ so the image is closed and both has and achieves a minimum. (Note $0$ is a lower bound, and then the greatest lower bound is a limit point: an open neightborhood around a lower bound not intersecting the closed set gives a greater lower bound.)

We need to show that the image of $\op{attempt}_\bullet(A)$ contains $0$ to diagonalize, which is where the ``do-better'' lemma comes in. Indeed, the minimum of the image of $O(n)$ must be a $0$: call the minimum $m,$ and certainly $m\in\RR_{\ge0}$ is at least $0.$

On the other hand, for any bound $m_0>0,$ we can find the $Q\in O(n)$ giving $\op{od}(QAQ^\intercal)=m_0>0.$ But $A':=QAQ^\intercal$ is itself symmetric (shown earlier), so the ``do-better'' lemma provides a $Q'\in O(n)$ with
\[\op{od}\left((Q'Q)A(Q'Q)^\intercal\right)=\op{od}\left(Q'(QAQ^\intercal)(Q')^\intercal\right)<\op{od}(QAQ^\intercal)=m,\]
so we can do better than any bound $m_0>0.$ It follows that the minimum $m$ is no greater than $0,$ so we must have $m=0.$

Thus, we can find a $Q\in O(n)$ for which $\op{od}(QAQ^\intercal)=0,$ meaning that $QAQ^\intercal$ is diagonal. This gives the following.
\begin{theorem}
    For any real symmetric matrix $A\in\RR^{n\times n},$ we can find an orthogonal matrix $Q\in O(n)$ such that $QAQ^\intercal$ is diagonal.
\end{theorem}
This follows from the above discussion. $\blacksquare$

We end with a few remarks. The above proof is truly remarkable: the only heavy comptuation we had to do was in the ``do-better'' lemma, and then a quick compactness argument was enough to finish. The clever observation is that the ``do-better'' lemma is even enough to get the result we need, as well as a fairly modern perspective of viewing matrix groups like $O(n)$ topologically.

Indeed, I'm not even convinced that repeated applications of the ``do-better'' lemma will actually approach a diagonal matrix: it's entirely possible that repeated applications will simply approach some small but nonzero number. However, that's fine because compactness has our back.

As an aside, I'll remark that there really isn't an easy way to use the more natural $\op{attempt}_Q(A):=\op{od}\left(QAQ^{-1}\right),$ defined over $Q\in\op{GL}_n(\RR).$ Indeed, $\op{GL}_n(\RR)\subseteq\RR^{n\times n}$ is an open subset because it is defined by the equation
\[M\in\op{GL}_n(\RR)\iff\det M\ne0,\]
and $\det M$ is just some large polynomial and hence continuous. So the compactness of $O(n)\subseteq\RR^{n\times n}$ doesn't easily carry into $\op{GL}_n(\RR).$ This is one roadblock to generalizing this proof past symmetric matrices.

\subsubsection{June 8th}
Today I learned about the semi-failure of function extensionality for polynomials. I think this entry is going to be more story-based than usual. The ``inciting incident'' here is Fermat's little theorem, which asserts that
\[x^p\equiv x\pmod p.\]
More generally, we have that
\[x^q=x\text{ in }\FF_q,\]
for any finite field $\FF_q.$ This is potentially problematic because the literal polynomials $x^q\in\FF_q[x]$ and $x\in\FF_q[x]$ have the exact same behavior over $\FF_q,$ but they are of course different polynomials. In other words, they are the same function (according to function extensionality) while different polynomials.

The resolution to this problem, of course, is to expand the size of the field. Even though $x^p=x$ in $\FF_p,$ if we move up into $\FF_{p^2},$ then for some $a\in\FF_p$ not a quadratic residue, we have
\[\left(\sqrt a\right)^p=a^{(p-1)/2}\sqrt a=\left(\frac ap\right)\sqrt a=-\sqrt a,\]
so we no longer have $x^p=x$ as functions. However, even in this larger field, we still have $x\mapsto x^{p^2}$ equal to the identity map. The only way to resolve these continually popping up problems is to move into the algebraic closure
\[\bigcup_{k>0}\FF_{p^k},\]
where such problems no longer occur. One can try to explain this by saying that being in the algebraic closure implies that all polynomials fully factor, so they have the same set of roots, so they are already pretty much equal.

However, one can make an even simpler argument.
\begin{lemma}
    Fix $K$ a field and two polynomials $a(x),b(x)\in K[x].$ If $a(x)\ne b(x)$ is a nonzero polynomial, then $a(x)$ and $b(x)$ agree on at most $\deg(a-b)(x)$ points.
\end{lemma}
This is an application of Lagrange's theorem on polynomials. The condition that $a(x)\ne b(x)$ is really saying that $(a-b)(x)\ne0$ so that $\deg(a-b)(x)$ is a well-defined nonnegative integer. Now, asserting that $a(x)$ and $b(x)$ agree on $d$ points is saying that
\[a(k)=b(k)\]
for $d$ distinct values of $k\in K.$ This means that $(a-b)(x)$ has a root on each of the $d$ distinct values of $k\in K,$ so Lagrange's theorem on polynomials asserts that the degree of $(a-b)(x)$ must be at least $d.$ This finishes. $\blacksquare$

Looking at our inciting incident $x^q\equiv x$ in $\FF_q,$ we see that this is exactly the equality case of the above: the breaking of function extensionality is saying that $x^q\ne x$ as polynomials while $x^q=x$ for all $x\in\FF_q,$ or equivalently, that these two polynomials agree on $q$ points. The above proposition says that this requires
\[\deg\left(x^q-x\right)\ge q,\]
which is of course true: it is in fact exactly degree $q.$ This kind of thinking lets us reconcile function extensoinality.
\begin{proposition}
    Fix $K$ an infinite field and two polynomials $a(x),b(x)\in K[x].$ If $a(x)=b(x)$ on all elements of $x\in K,$ then we know $a(x)=b(x)$ as polynomials.
\end{proposition}
This is a direct application of the lemma. We show the contrapositive: if $a(x)\ne b(x),$ then $a(x)$ and $b(x)$ must disagree on some value of $x\in K.$ Well, the lemma says that $a(x)$ and $b(x)$ are allowed to agree on at most
\[\deg(a-b)(x)<\infty\]
points of $K,$ so on $K$ minus this finite set of points we have that $a(x)$ and $b(x)$ must disagree. Because $K$ is infinite, we know that $K$ minus this finite set of points is nonempty, so we are done. $\blacksquare$

This helps explain why moving to the algebraic closure of $\FF_p$ earlier fixed the problem: the algebraic closure has infinitely many elements, so we retain function extensionality for polynomials, even when it broke for any component finite fields.

Additionally, we know for free that function extensionality for polynomials over characteristic-$0$ fields will hold because a characteristic-$0$ field must have the infinite field $\QQ$ as a subfield, so the above proposition kicks in immediately.

Of course, the best we can say for finite fields is that $a(x)=b(x)$ on all elements of $x\in\FF_q$ implies that $\deg(a-b)(x)\ge q,$ but if we are careful about the equality case $x^q\equiv x,$ then we can do a little better.
\begin{proposition}
    Fix $K=\FF_q$ a finite field for prime-power $q.$ Then given polynomials $a(x)$ and $b(x)$ which agree on all elements of $\FF_q,$ then $a(x)\equiv b(x)\pmod{x^q-x}.$
\end{proposition}
One way to show this is to use the lemma again. Observe that a full reduction of a polynomial$\pmod{x^q-x}$ will yield a polynomial of degree strictly less than $q.$ Indeed, if we want to reduce
\[a(x)=\sum_{k=0}^\infty a_kx^k\]
where all but finitely many of the $\{a_k\}_{k=0}^\infty$ are nonzero, then we note $x^d\equiv x^{d\pmod q}\pmod{x^q-x}$ (by, say, induction), so we can say
\[a(x)\equiv\sum_{r=1}^{q-1}\left(\sum_{k\equiv r\pmod q}a_k\right)x^r\pmod{x^q-x}.\]
Now, the point of saying this is to note that the two polynomials are still equal on all elements of $\FF_q$ because $x^q-x=0$ for all elements $x\in\FF_q.$

So, after reducing $a(x)$ and $b(x)$ as in the proposition, we note that the lemma implies that $a(x)$ and $b(x)$ agreeing on all $q$ elements of $\FF_q$ requires $a(x)=b(x)$ or $\deg(a-b)(x)\ge q.$ Well, $\deg a,\deg b,$ if defined, are bounded strictly above $q,$ so $\deg(a-b)$ is as well. So we must instead have $a(x)=b(x),$ finishing. $\blacksquare$

An alternate way to view the above proposition is to note that $\FF_q$ is pretty much made of the $x$ such that $x^q-x=0.$ This means that polynomials which are equivalent$\pmod{x^q-x}$ should have the same behavior on $\FF_q$ (and conversely!) because this is rigged by the structure of $\FF_q$ itself.

Anyways, I think that's the furthest we can salvage polynomial functional extensionality. Essentially, the only exceptions are the ones generated by the inciting incident and friends, so we have managed to contain the problem as much as we reasonably can.

\subsubsection{June 9th}
Today I learned a little about group homology and chain complexes, from \href{https://web.archive.org/web/20210506095754/https://venhance.github.io/napkin/Napkin.pdf}{Napkin}. Here is our main definition.
\begin{definition}
    A chain complex is a sequence of (usually abelian) groups $\{A_k\}_{k\in\NN}$ with maps $\varphi_n:A_{n+1}\to A_n$ for which the chain
    \[\cdots\stackrel{\varphi_{n+1}}\longrightarrow A_{n+1}\stackrel{\varphi_n}\longrightarrow A_n\stackrel{\varphi_{n-1}}\longrightarrow A_{n-1}\to\cdots\]
    gives $\varphi_n\circ\varphi_{n+1}$ as the trivial group homomorphism.
\end{definition}
The geometric way to view this is to take $\{A_k\}$ equal to $k$-chains of a topological space $X$ with $\varphi_\bullet$ equal to the boundary operator $\del.$ Then we can say things like ``an $n$-cycle is an element of the kernel of $\del_A:A_n\to A_{n-1}$'' and ``an $n$-boundary is an element of the image of $\del_A:A_{n+1}\to A_n.$'' This perspective leads us to the definition of homology groups.
\begin{definition}
    Given a chain complex $A_\bullet,$ we define the $n$th homology group $H_n(A)$ as the set of $n$-cycles modulo the $n$-boundaries. That is,
    \[H_n(A_\bullet):=\op{ker}\left(A_n\to A_{n-1}\right)/\op{im}\left(A_{n+1}\to A_n\right).\]
\end{definition}

As a more algebraic way to justify homology groups, observe that we can take a long exact sequence
\[\cdots\stackrel{\varphi_{n+1}}\longrightarrow A_{n+1}\stackrel{\varphi_n}\longrightarrow A_n\stackrel{\varphi_{n-1}}\longrightarrow A_{n-1}\to\cdots\]
satisfying $\op{im}(\varphi_{n+1})=\op{ker}(\varphi_n)$ and split this up into a bunch of connected short exact sequences of the form
\[0\to\op{coker}(\varphi_{n+2})\stackrel{\varphi_{n+1}}\to A_{n+1}\stackrel{\varphi_n}\to\op{im}(\varphi_n)\to0.\]
The right half of this sequence is exact because $\varphi_n$ is of course surjective onto $\op{im}(\varphi_n).$ However, the main point is that
\[\op{coker}(\varphi_{n+2}):=A_{n+2}/\op{im}(\varphi_{n+2}).\]
We are not immediately guaranteed that $\varphi_{n+1}:A_{n+2}\to A_{n+1}$ is well-defined over $A_{n+2}/\op{im}(\varphi_{n+2}),$ but it is exactly because $\op{im}(\varphi_{n+2})=\op{ker}(\varphi_{n+1}).$ More explicitly, a coset $a_{n+2}+\op{im}(\varphi_{n+2})$ can simply be taken to $\varphi_{n+1}(a_{n+2})$ because $\varphi_{n+1}(\op{im}(\varphi_{n+2}))=\{e\}.$

To finish, we note that the left half of the sequence is exact because $\varphi_{n+1}$ is injective on $A_{n+2}/\op{im}(\varphi_{n+2})=A_{n+2}/\op{ker}(\varphi_{n+1})$ because it has trivial kernel after modding out by its old kernel. And the image of $\varphi_{n+1}$ doesn't change---we are still more or less pushing all $A_{n+2}$ through---so $\varphi_{n+1}$ continues to surject onto the kernel of $\varphi_n,$ finishing the required exactness.

All of this is to say that long exact sequences are nice: we can study them in terms of short exact sequences, which are simpler because they are shorter. However, our chain complex as defined above does not necessarily have $\op{im}(\varphi_{n+1})=\op{ker}(\varphi_n)$ but merely $\op{im}(\varphi_{n+1})\subseteq\op{ker}(\varphi_n).$ So we define
\[H_n(A_\bullet):=\op{ker}(\varphi_{n-1})/\op{im}(\varphi_n)\]
to measure how badly we failed around $A_n.$

While we're here, we note that we can turn the set of chain complexes into a category. Here are our morphisms.
\begin{definition}
    Given a chain complexes $A_\bullet$ and $B_\bullet,$ we define a morphism $f_\bullet:A_\bullet\to B_\bullet$ to be a set of morphisms $\{f_k\}_{k\in\NN}$ that make the following (infinite) diagram commute.
    \begin{center}
        \begin{tikzcd}
            \cdots \arrow[r, "\varphi_{n+1}"] & A_{n+1} \arrow[r, "\varphi_n"] \arrow[d, "f_{n+1}"] & A_n \arrow[r, "\varphi_{n-1}"] \arrow[d, "f_n"] & A_{n-1} \arrow[r] \arrow[d, "f_{n-1}"] & \cdots \\
            \cdots \arrow[r, "\gamma_{n+1}"]  & B_{n+1} \arrow[r, "\gamma_n"]                       & B_n \arrow[r, "\gamma_{n-1}"]                   & B_{n-1} \arrow[r]                      & \cdots
        \end{tikzcd}
    \end{center}
\end{definition}
While natural, the following convinces us that we really do have the correct definition.
\begin{lemma}
    Given a morphism of chain complexes $f_\bullet:A_\bullet\to B_\bullet,$ we have an induced $H_n(f_\bullet):H_n(A_\bullet)\to H_n(B_\bullet).$
\end{lemma}
The point is that everything commutes, so the kernels and images behave. Starting from the top, note that we have
\[f_n:A_n\to B_n.\]
Now, surely $\op{ker}(\varphi_{n-1})\subseteq\op{ker}(f_{n-1}\circ\varphi_{n-1})$ as well, but $f_{n-1}\circ\varphi_{n-1}=\gamma_{n-1}\circ f_n,$ so we see the downward arrow $f_n$ induces a map from $\op{ker}(\varphi_n)$ to $\op{ker}(\gamma_n)$: $\varphi_{n-1}(a)=e$ implies $f_{n-1}(\varphi_{n-1}(a))=e$ implies $\gamma_{n-1}(f_n(a))=e.$

Similarly,
\[f_n:A_n\to B_n\]
induces a map $f_n:\op{im}(\varphi_n)\to B_n$ by restriction. However, $f_n\circ\varphi_n=\gamma_n\circ\varphi_{n+1},$ so the image of $A_{n+1}\to\op{im}(\varphi_n)\to B_n$ is a subset of $\op{im}(\gamma_n),$ so in fact we may restrict this to a map $f_n:\op{im}(\varphi_n)\to\op{im}(\gamma_n).$

Putting this all together, we saw that $f_n:A_n\to B_n$ restricts to a map
\[f_n:\op{ker}(\varphi_{n-1})\to\op{ker}(\gamma_{n-1}).\]
We still have $\op{im}(\gamma_n)\subseteq\op{ker}(\gamma_{n-1}),$ so we may safely mod out the domain here by $\op{im}(\gamma_n),$ giving an induced map
\[\op{ker}(\varphi_{n-1})\to H_n(B_\bullet).\]
We know that $f_n$ takes elements of $\op{im}(\varphi_n)$ to elements of $\op{im}(\gamma_n),$ so they are all going to the kernel of the above induced map. Thus, we may safely mod out by them in the codomain, giving our desired map $H_n(A_\bullet)\to H_n(B_\bullet).$ This finishes. $\blacksquare$

\subsubsection{June 10th}
Today I learned a proof to Sperner's theorem, from THE BOOK. For background, the problem is to take the subsets of $[n]:=\{1,\ldots,n\}$ for $n\in\NN$ and turn them into a poset with relation $\subseteq.$ Then the question we ask is for the largest possible antichain in the poset, or the largest number of incomparable elements in a single family.

Observe that if all the subsets have the same size, then we are an antichain for free: if $|A|=|B|$ while $A\subseteq B,$ then $A=B,$ so distinct elements must be incomparable. Fixing a size $k,$ the most subsets we can get from this construction is $\binom nk.$ To maximize $\binom nk$ over $k,$ we note that
\[\binom nk=\frac{n!}{k!(n-k)!}=\frac{k+1}{n-k}\cdot\frac{n!}{(k+1)!(n-k-1)!}=\frac{k+1}{n-k}\binom n{k+1}.\]
In particular, $\binom nk$ is increasing as long as $k+1\le n-k,$ or equivalently $k\le(n-1)/2,$ and decreasing as long as $k+1\ge n-k,$ or equivalently $k\ge(n-1)/2.$ The point is that we are maximized at the middle, so $\binom n{\floor{n/2}}$ is the best we can do.

Sperner's theorem says that this is in fact maximal, which is perhaps surprising.
\begin{theorem}
    The largest antichain in the poset of subsets of $[n]$ ordered by $\subseteq$ has size $\binom n{\floor{n/2}}.$
\end{theorem}
The key idea in this proof is that two subsets $A$ and $B$ can belong to an antichain if and only if $A$ and $B$ themselves do not belong to any chain. The way to turn this into combinatorics is to consider the set of chains
\[\emp=A_0\subseteq A_1\subseteq\cdots\subseteq A_n=[n]\]
where $\#A_k=k.$ Call such chain an ``incremental'' chain. As outline for what's about to happen, we will count the number of chains containing each individual subset in an antichain, and all these chains must be distinct because the subsets are in antichain, so we can bound the total number of chains containing any subset in the antichain without tears.

Suppose we require $S\subseteq[n]$ with to live in the chain; we count the number of incremental chains containing $S.$ We claim the following.
\begin{lemma}
    Given a finite set $S,$ there are $(\#S)!$ incremental chains to $S,$ which look like
    \[\emp=A_0\subseteq A_1\subseteq\cdots\subseteq A_{\#S}=S.\]
\end{lemma}
Observe that the only place for $S$ is at $A_{\#S}$ if the size is to match. On one hand, the number of inrecemental chains
\[\emp=A_0\subseteq A_1\subseteq\cdots\subseteq A_{\#S}=S\]
can be bijected with the one-element distinct subsets of $S$ by
\[A_1\setminus A_0,\,A_2\setminus A_1,\,\ldots,\,A_{\#S}\setminus A_{\#S-1}.\]
More specifically, we see that $\sigma(k)\in A_k\setminus A_{k-1}$ (unique because $\#(A_k\setminus A_{k-1})=1$) forms a bijection $\sigma:[\#S]\to S$ because the domain and codomain have the same size, and $\sigma$ is injective because distinct $k+1$ and $\ell+1$ gives $k<\ell$ without loss of generality, implying $A_{k+1}\subseteq A_\ell,$ so $A_{k+1}\setminus A_k\ne A_{\ell+1}\subseteq A_\ell.$

This correspondence between chains to $S$ and bijections $\sigma:[\#S]\to S$ is in fact a one-to-one correspondence, with inverse map
\[A_k=\bigcup_{\ell=1}^k\sigma(\ell).\]
This is actually a chain because because $A_{k+1}=A_k\subseteq\sigma(k+1),$ and it has the right form because $A_{\#S}=\sigma([\#S])=S,$ and $\#A_k=\#\sigma([k])=k.$ We won't be rigorous showing that the forward and backward maps are actually inverses, but we will say that
\[\bigcup_{\ell=1}^k(A_\ell\setminus A_{\ell-1})=A_k\setminus A_0=A_k\]
shows that the forward composed with backward map is the identity, and
\[\{\sigma(k)\}=\bigcup_{\ell=1}^k\sigma(\ell)\bigg\backslash\bigcup_{\ell=1}^{k-1}\sigma(\ell).\]
shows the other direction. All of this to say that there are $(\#S)!$ total incremental chains to $S$ because this is the number of bijections. $\blacksquare$

Continuing, it remains to count the number of incremental chains from $S$ to $[n].$ However, we can biject such chains with incremental chains
\[\emp=A_0\subseteq A_1\subseteq\cdots\subseteq A_{n-\#S}=[n]\setminus S.\]
Indeed, unioning each individual chain with $S$: the difference in size between consecutive elements remains $1,$ we start at $S,$ and we end at $[n].$ We won't be as formal as above showing that this is a bijection, but it is. This implies the number of incremental chains from $S$ to $[n]$ is $(n-\#S)!.$

Thus, for any particular subset $S,$ there are $(\#S)!\cdot(n-\#S)!$ total incremental chains through $S.$ Given an antichain $\mathcal F,$ we note that distinct $S$ and $T$ cannot share any incremental chains. Indeed, if both $S$ and $T$ live in an incremental chain
\[\emp=A_0\subseteq A_1\subseteq\cdots\subseteq A_n=[n],\]
either $S\subseteq T$ or $T\subseteq S$ because this is a chain, implying that $S=T$ because $S$ and $T$ are in the antichain $\mathcal F.$ So the total number of incremental chains through any element of $\mathcal F$ is
\[\sum_{S\in\mathcal F}(\#S)!(n-\#S)!.\]
That was the key step, where we turned the antichain condition into actual numbers. Anyways, to finish, we let $s_k$ be the number of subsets in $\mathcal F$ of size $k$ so that we can write the above as
\[\sum_{k=0}^ns_k\cdot k!(n-k)!.\]
Now, we can bound the number of incremental chains through $\mathcal F$ above by the total number of incremental chains in $[n],$ which is $n!$ by lemma. Doing some rearranging, we see that
\[\sum_{k=0}^n\frac{s_k}{\binom nk}\le1.\]
The theorem statement makes us think that we should only have to care about the $k=\floor{n/2}$ and $k=\ceil{n/2}$ terms, so we may safely note that $\binom nk\le\binom n{\floor{n/2}}$ to get a lower bound
\[\sum_{k=0}^n\frac{s_k}{\binom n{\floor n/2}}\le1.\]
In particular, $\#\mathcal F=\sum_{k=0}^ns_k\le\binom n{\floor{n/2}},$ which is exactly what we wanted. $\blacksquare$

This proof is exceedingly clever. The way to translate the antichain condition into something countable is quite remarkable, and even the inequality manipulation at the end pays very careful attention to the equality case to make sure we get what we need. This is quite nice.

As an aside, we note that the statement of Sperner's theorem doesn't actually tell us what the optimal cases are. Writing everything down, the inequalities we used are
\[\sum_{k=0}^n\frac{s_k}{\binom n{\floor n/2}}\le\sum_{k=0}^n\frac{s_k}{\binom nk}\le1.\]
The second inequality has equality case when our $\mathcal F$ actually uses all incremental chains. But the former inequality will only hold if all the $s_k=0$ except for $s_{\floor{n/2}}$ and $s_{\ceil{n/2}}.$ So we get the following immediately.
\begin{proposition}
    The largest antichain in the poset of subsets of $[n]$ ordered by $\subseteq$ is exactly all the subsets of size $n/2$ for $n$ even.
\end{proposition}
Indeed, in this case, $\floor{n/2}=\ceil{n/2},$ so continuing from the above discussion, all terms except for $k=n/2$ give $s_k=0,$ so $\#\mathcal F=s_{n/2}=\binom n{n/2},$ which is what we wanted. $\blacksquare$

As for $n$ odd, the most we can immediately say is that $s_{\floor{n/2}}+s_{\ceil{n/2}}=n,$ but I am fairly confident that the equality cases are indeed either the set of all subsets of size $\floor{n/2}$ or the set of all subsets of size $\ceil{n/2}.$ This proof is somewhat involved, but intuitively, having a subset of size $\ceil{n/2}$ prevents $\ceil{n/2}$ subsets of size $\floor{n/2}$ from appearing in the antichain, which is a heavy cost. \href{https://www.whitman.edu/mathematics/cgt_online/book/section01.07.html}{Here} is a more complete proof, which I haven't read.

\subsubsection{June 11th}
Today I learned about the P\'olya-Vinogradov inequality, from \href{http://jonismathnotes.blogspot.com/2014/10/character-sums-and-polya-vinogradov.html}{Joni} and \href{https://terrytao.wordpress.com/2009/08/18/the-least-quadratic-nonresidue-and-the-square-root-barrier/#more-2664}{Tao}. Here is our statement.
\begin{theorem}
    Fix $p$ a prime and $\chi$ a nontrivial character$\pmod p.$ Then
    \[\left|\sum_{M\le n\le N}\chi(n)\right|\le\sqrt p\log p.\]
\end{theorem}
There are two steps to this proof: Fourier analysis and bash. Observe that we have a somewhat trivial bound
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le p\]
simply because $\chi$ is periodic, and $\sum_{n=0}^{p-1}\chi(n)=0$ by orthogonality of characters. The way we are going to improve upon this bound is to find that, in fact, the Fourier coefficients of $\chi$ are reasonably small, and then bashing will yield the result.

With that said, here is our Fourier analysis step, finding the (discrete) Fourier coefficients. \begin{lemma}
    Given a nontrivial character $\chi\pmod p$ for $p$ prime and integer $n,$ we have
    \[\chi(n)G(\overline\chi)=\sum_{x\in\FF_p^\times}\overline{\chi(x)}\zeta_p^{nx}.\]
\end{lemma}
This is pretty normal Gauss sums theory. Note that if $p\mid n,$ then $\chi(n)=0,$ and the sum on the right-hand side is
\[\sum_{x\in\FF_p^\times}\overline\chi(x)\cdot1=0\]
by orthogonality of characters. Otherwise, we take $p\nmid n,$ and note that
\[\chi(n)G(\overline\chi)=\chi(n)\sum_{x\in\FF_p^\times}\overline\chi(x)\zeta_p^x=\sum_{x\in\FF_p^\times}\chi(x/n)\zeta_p^x.\]
Now, taking $x\mapsto x/n,$ which is a legal bijection because $n^{-1}\in\FF_p^\times,$ we see that
\[\chi(n)G(\overline\chi)=\sum_{x\in\FF_p^\times}\chi(x)\zeta_p^{nx},\]
which is exactly what we wanted. This finishes. $\blacksquare$

We would like to divide both sides of the above computation by $G(\overline\chi),$ but we need to know that $G(\overline\chi)\ne0$ for this. In fact, we can do better; we have the following.
\begin{lemma}
    Given a nontrivial character $\chi\pmod p$ for $p$ prime, we define the Gauss sum
    \[G(\chi):=\sum_{x\in\FF_p^\times}\chi(x)\zeta_p^x.\]
    Then $|G(\chi)|=\sqrt p.$
\end{lemma}
We evaluate $G(\chi)^2.$ Proceeding by brute force, this is
\[G(\chi)^2=\left(\sum_{x\in\FF_p^\times}\chi(x)\zeta_p^x\right)\left(\sum_{y\in\FF_p^\times}\chi(y)\zeta_p^y\right).\]
Distributing, this turns into
\[G(\chi)^2=\sum_{x,y\in\FF_p^\times}\chi(xy)\zeta_p^{x+y}\]
because $\chi$ is a multiplicative character. At this point we reparamterize the sum to be a sum over $x$ and a sum over $t:=1xy.$ Observe that for fixed $x\in\FF_p^\times,$ $y$ ranging over $\FF_p^\times$ is equivalent to $t=xy$ ranging over $\FF_p^\times$ (because $\FF_p^\times$ is a group), so we can safely write rearrange as
\[G(\chi)^2=\sum_{t\in\FF_p^\times}\chi(t)\sum_{x\in\FF_p^\times}\zeta_p^{x+xt^{-1}},\]
where $t^{-1}$ is taken$\pmod p.$ It will make the argument clearer if we now add in $\sum_t\chi(t)=0$ to the entire right-hand sum, which doesn't change the quantity, but it lets us write
\[G(\chi)^2=\sum_{t\in\FF_p^\times}\chi(t)\left(\sum_{k=0}^{p-1}\left(\zeta_p^{1+t^{-1}}\right)^k\right).\]
Now, whenever $t\ne-1,$ we have that $1+t^{-1}\ne0,$ so the inner sum is a sum of roots of unity around a circle and will vanish. (We can see this directly by summing as a geometric sequence.) Then at $t=-1,$ the summands are all $1,$ so we get out $p.$ This means that
\[G(\chi)^2=\chi(-1)\cdot p.\]
It follows that $|G(\chi)|^2=1p=p,$ and taking square-roots finishes. $\blacksquare$

The above lemma says that the Fourier coefficients of $\chi,$ which look like $\overline\chi/G(\overline\chi),$ are on the order of $1/\sqrt p.$ This is quite small, and we will be able to convert this smallness into the desired inequality. So now we brute force. Using our Fourier step, we see
\[\sum_{M\le n\le N}\chi(n)=\sum_{M\le n\le N}\left(\frac1{G(\overline\chi)}\sum_{x\in\FF_p^\times}\overline\chi(x)\zeta_p^{nx}\right).\]
Rearranging the order of summation and abusing the triangle inequality, we see
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac1{\sqrt p}\sum_{x\in\FF_p^\times}\left(|\overline\chi(x)|\cdot\left|\sum_{M\le n\le N}\zeta_p^{nx}\right|\right).\]
This is the most egregious inequality we will use in this proof. Indeed, we lose a lot of cancellation by separating out the character $\overline\chi$ from the sum of $\overline\chi\zeta_p^\bullet,$ but there isn't an easy way to use this, and hey, maybe we can actually configure $M$ and $N$ to make this all align. Anyways, this collapses into
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac1{\sqrt p}\sum_{x\in\FF_p^\times}\left|\zeta_p^M\cdot\frac{\zeta_p^{N-M+1}-1}{\zeta_p^x-1}\right|.\]
We can't really optimize the numerator of this fraction because maybe it really is $2,$ so we will bound the numerator by $2.$ As for the denominator, we evaluate its magnitude by writing
\[\zeta_p^x-1=e^{2\pi ix/p}-1=2ie^{\pi ix/p}\left(\frac{e^{\pi ix/p}-e^{-\pi ix/p}}{2i}\right),\]
so the magnitude is $2\sin(\pi x/p).$ Thus,
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac1{\sqrt p}\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}.\]
It remains to bound this sum. Observe that nothing we did above really involves number theory in any crucial way, and by now, we can't even see the character $\chi$ anymore, so there will definitely not be any more number theory in what follows.

We start with a heuristic argument to get the inequality.
\begin{lemma}
    Fix $p$ a positive odd integer. We have that
    \[\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}=p\log p+O(1).\]
\end{lemma}
Observe that $\sin(\pi x)\ge2x$ over $x\in(0,1/2),$ which is true because $\sin(\pi x)$ is concave down in this range, and $2x$ is a secant line from $x=0$ to $x=1/2.$ Thus, we may write
\[\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}=2\sum_{k=1}^{(p-1)/2}\frac1{\sin(\pi k/p)}\le2\sum_{k=1}^{(p-1)/2}\frac1{2k/p}.\]
So we can bound the sum by
\[p\sum_{k=1}^{(p-1)/2}\frac1k,\]
and we know the sum is $\log\frac{p-1}2+O(1)=\log p+O(1).$ Thus, the entire sum is $p\log p+O(p),$ which is what we wanted. $\blacksquare$

The above is a quick argument to show that the inequality we want is roughly on the right track, but it turns out that $\sin(\pi k/p)\ge 2k/p$ is pretty expensive in the sum. But now that we understand the type of argument we're looking for, we can skip the harmonic numbers by noticing that the bound comes from an integral, so we can try to directly turn out sum into an integral. In particular,
\[\frac1p\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}\]
looks a lot like a Riemann sum for $\int_0^1\frac1{\sin(x)}\,dx,$ but this integral diverges, so we have to be more careful. We pick up the following.
\begin{lemma}
    Fix $y\in(0,\pi)$ and a small positive $\varepsilon<\min\{y,\pi-y\}$ so that $(y-\varepsilon,y+\varepsilon)\subseteq(0,\pi).$ We have that
    \[\frac1{\sin(y)}\le\frac1{2\varepsilon}\int_{y-\varepsilon}^{y+\varepsilon}\frac1{\sin(x)}\,dx.\]
\end{lemma}
This follows because $1/\sin(x)$ is concave up on $(0,\pi).$ Indeed, let $f(x)=1/\sin(x),$ and we see that $f'(x)=-(\sin x)^{-2}\cos(x)=\frac{-\cos(x)}{\sin^2(x)},$ which gives
\[f''(x)=\frac{\sin^2(x)\cdot\sin(x)+\cos(x)\cdot2\sin(x)\cos(x)}{\left(\sin^2(x))\right)^2},\]
where are all the terms are positive when $x\in(0,\pi).$

Now we prove Jensen's inequality that for any $x\in(0,\pi)$ and any $\varepsilon\ge0$ with $(x-\varepsilon,x+\varepsilon)\subseteq(0,\pi),$ we have
\[\frac{f(x+\varepsilon)+f(x-\varepsilon)}2\ge f(x).\]
This statement is true for $\varepsilon=0,$ so it suffices to show that the function on the left is increasing with respect to $\varepsilon.$ Its derivative is
\[f'(x+\varepsilon)-f'(x-\varepsilon).\]
We need to show this is nonnegative, but $f''(x)>0$ in this region implies that $f'$ is increasing, so $f'(x+\varepsilon)\ge f'(x-\varepsilon),$ which is exactly what we needed.

We are now ready to prove the lemma. Fix $y$ and $\varepsilon$ as needed. From the above we know that, for any large $N,$
\[\sum_{k=-N}^Nf\left(y+\varepsilon\frac kN\right)=f(y)+\sum_{k=1}^N\left(f\left(y+\varepsilon\frac kN\right)+f\left(y-\varepsilon\frac kN\right)\right)\ge(2N+1)f(y).\]
This rearranges into
\[f(y)\le\frac N{2N+1}\cdot\frac1N\sum_{k=-N}^Nf\left(y+\varepsilon\frac kN\right).\]
Sending $N\to\infty$ reveals the right-hand side as a Riemann sum, giving
\[f(y)\le\frac12\int_{-1}^1f(y+\varepsilon x)\,dx=\frac1{2\varepsilon}\int_{y-\varepsilon}^{y+\varepsilon}f(x)\,dx.\]
This is what we wanted, so we are done here. $\blacksquare$

We will use the above lemma with $\varepsilon=\pi/(2p)$ to get
\[\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}\le\sum_{k=1}^{p-1}\frac1{2\cdot\pi/(2p)}\int_{\pi k/p-\pi/(2p)}^{\pi k/p+\pi/(2p)}\frac1{\sin(x)}\,dx=\frac p\pi\int_{\pi/(2p)}^{\pi-\pi/(2p)}\frac1{\sin(x)}\,dx.\]
This turns out to be very sharp. If we want to bee-line to the P\'olya-Vinogradov inequality, Joni has this trick where right now we can write
\[\frac p\pi\int_{\pi/2p}^{\pi(1-1/(2p))}\frac1{\sin(x)}\,dx=p\int_{1/(2p)}^{1-1/(2p)}\frac1{\sin(\pi x)}\,dx=2p\int_{1/(2p)}^{1/2}\frac1{\sin(\pi x)}\,dx\]
and use concavity to note $\sin(\pi x)\ge2x$ over $x\in(0,1/2),$ which gives
\[2p\int_{1/(2p)}^{1/2}\frac1{2x}\,dx=p\left(\log\frac12-\log\frac1{2p}\right)=p\log p.\]
Putting this all together, we see
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac1{\sqrt p}\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}\le\frac1{\sqrt p}\cdot p\log p,\]
which rearranges to $\sqrt p\log p.$ This is the P\'olya-Vinogradov inequality, so we are done. $\blacksquare$

Of course, we can do a little better than the given if we actually contend with $\int1/\sin(x)\,dx.$ To ground ourselves, we quickly start with the following lemma.
\begin{lemma}
    For $\chi\pmod p$ a nontrivial character, we get
    \[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac{\sqrt p}\pi\int_{\pi/(2p)}^{\pi-\pi/(2p)}\frac1{\sin(x)}\,dx.\]
\end{lemma}
This follows from stringing the inequalities established above all together. Explicitly, we see
\[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac1{\sqrt p}\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}\le\frac1{\sqrt p}\cdot\frac p\pi\int_{\pi/(2p)}^{\pi-\pi/(2p)}\frac1{\sin(x),}\]
which rearranges into what we wanted. $\blacksquare$

It remains to deal with the integral. We have the indefinite integral
\[\int\frac1{\sin(x)}\,dx=\log\left|\frac{\sin(x)}{1+\cos(x)}\right|,\]
which can be verified manually. This gives
\[\int_{\pi/(2p)}^{\pi-\pi/(2p)}\frac1{\sin(x)}=\log\left|\frac{\sin\left(\pi-\frac\pi{2p}\right)}{1+\cos\left(\pi-\frac\pi{2p}\right)}\cdot\frac{1+\cos\left(\frac\pi{2p}\right)}{\sin\left(\frac\pi{2p}\right)}\right|.\]
The $\sin$ terms cancel by symmetry, and $\cos\left(\pi-\frac\pi{2p}\right)=-\cos\left(\frac\pi{2p}\right).$ Re-expanding this out, we get
\[\log\left(1+\cos\left(\frac\pi{2p}\right)\right)-\log\left(1-\cos\left(\frac\pi{2p}\right)\right).\]
As $p\to\infty,$ we see $\cos\left(\frac\pi{2p}\right)\to1$ from below, so I have no qualms bounding the positive term directly by $\log2.$ Using the Taylor expansion of $\cos,$ we see $\cos(x)\le1-\frac12x^2+\frac1{24}x^4,$ so we can bound this above by
\[\log2-\log\left(\frac12\left(\frac\pi{2p}\right)^2-\frac1{24}\left(\frac\pi{2p}\right)^4\right).\]
(We wanted to bound $\cos$ above because $-\log(1-x)$ is increasing.) Rearranging a bit, our main term is $2\log p$ from the right $\log,$ leaving behind
\[2\log p+\log2-\log\left(\frac{\pi^2}{2\cdot2^2}-\frac{\pi^4}{24\cdot16p^2}\right).\]
As $p\to\infty,$ the right-hand term will approach its own main term, and we could deal with this, but I don't want to bother. As $p\to\infty,$ the long $\log$ term increases and subtracts more, meaning that to maximize, we take $p=2.$ Now, we see that
\[\frac{\pi^2}{2\cdot2^2}-\frac{\pi^4}{24\cdot16\cdot2^2}\le\frac98-\frac{10^2}{24\cdot16\cdot4}\le\frac98-\frac{192}{24\cdot16\cdot4}=1,\]
so the $\log$ term subtracts at least $\log1=0.$ That is, we may ignore it. Plugging in our estimate of the integral, we get the following.
\begin{proposition}
    For $\chi\pmod p$ a nontrivial character, we get
    \[\left|\sum_{M\le n\le N}\chi(n)\right|\le\frac2\pi\sqrt p\log p+\frac{\log2}\pi\sqrt p.\]
\end{proposition}
The discussion above established
\[\int_{\pi/(2p)}^{\pi-\pi/(2p)}\frac1{\sin(x)}\le2\log p+\log2,\]
which is enough upon plugging into the above lemma. So we are done here. $\blacksquare$

The above improvement isn't drastic, but the constant $\frac2\pi$ is a much better approximation for the constant we get out of the sum $\sum_{k=1}^{p-1}\frac1{\sin(\pi k/p)}.$ I do like the simplicity of the $\sqrt p\log p$ bound for P\'olya-Vinogradov, but I wanted to make the point that we have some more optimizable space.

While we're here, we pick up the following cute result which falls out of our work.
\begin{proposition}
    Give $p$ a prime. The least non-quadratic residue is at most $1+\sqrt p\log p.$
\end{proposition}
Use $\chi=\left(\frac\bullet p\right),$ which is the Legendre symbol. If we let $n_p$ be the least non-quadratic residue$\pmod p,$ then
\[\sum_{k=1}^{n_p-1}\left(\frac kp\right)=n_p-1.\]
However, this character sum is bounded by $\sqrt p\log p,$ so we see $n_p\le1+\sqrt p\log p.$ This is what we wanted. $\blacksquare$

Of course, we probably expect that quadratic and non-quadratic residues should be roughly evenly distributed, so this bound isn't great, but it is still cute. We also remark that we can do a little better than this if we use the multiplicativity of our character $\chi.$ Indeed, if $n_p$ is the least non-quadratic residue, then any $n_p-1$-smooth integer is also a quadratic residue, making for a large character sum. We won't do this here.

\subsubsection{June 12th}
Today I learned about representing numbers as sums of distinct squares, from a bunch of places. We start with a relatively standard result, which comes from \href{https://math.stackexchange.com/a/4021787/869257}{here}.
\begin{proposition}
    Every positive integer $n$ greater than $128$ can be written as the sum of distinct squares.
\end{proposition}
There are a few ways to approach this, but it feels best motivated to just be greedy and pay for the consequences later. The idea is to continually pick up the largest square we can, and these squares should be distinct if the square is on the same scale as what we're subtracting it from. Here's our exact extraction.
\begin{lemma}
    Fix $n$ a positive integer at least $298=13^2+129.$ Then there exists a positive integer $k$ with $n-129\ge k^2>n/2.$
\end{lemma}
Take $k=\floor{\sqrt{n-129}}>0$ as large as possible. Then we note $k^2\le n-129<(k+1)^2$ by construction, so we see
\[\frac n2<\frac{(k+1)^2+129}2.\]
It suffices to show that $k^2$ is greater than or equal to this, which is an inequality only involving $k$ and therefore solvable. Rearranging, we need
\[129\le2k^2-(k+1)^2=k^2-2k-1.\]
So what we need is then equivalent to $(k-1)^2\ge131,$ or $k\ge1+\sqrt{131},$ for which $k\ge13$ is enough. However, our bound $13^2+129$ safely guarantees exactly $\floor{\sqrt{n-129}}\ge13.$ This finishes. $\blacksquare$

So our greedy algorithm is to continually pick the largest square we can but keeping the current total above $128.$ Then eventually we will hit some value below $298,$ but we can just check all numbers manually between $129$ and $297<298.$ For this second justification, a computer program verifies this for us; \href{https://replit.com/@NirElber/Distinct-Squares#main.py}{here} is my code.

Anyways, to rigorize our greedy algorithm, we use a strong induction. Our base cases are the positive integers $n$ from $129$ to $297.$ Now, suppose $n\ge298$ and every positive integer smaller than $n$ but at least $129$ can be written as the sum of distinct squares. The lemma gives us a $k$ such that
\[n>n-k^2>129,\]
so $n-k^2$ can be written as the sum of distinct squares, say $n-k^2=\sum_{\ell=1}^ma_\ell^2.$ However, we now claim that $k^2$ cannot be found among the $\left\{a_\ell^2\right\}_{\ell=1}^m$: indeed, one of those squares $a_\bullet^2$ satisfies
\[a_\bullet^2\le\sum_{\ell=1}^ma_\ell^2=n-k^2<\frac n2\]
because $k^2>n/2$ from the lemma. However, $k^2>n/2,$ so we must have $a_\bullet\ne k^2.$ Thus,
\[n=k^2+\sum_{\ell=1}^ma_\ell^2\]
is a valid decomposition of $n$ into distinct squares. $\blacksquare$

Some remarks in order. First off, it might feel unsatisfying that we have to check the cases from $129$ to $298,$ but I don't really see a way around this. At least this elementary method more or less requires some brute force computation; in other words, I don't think we can push this to ``sufficiently large $n$'' and avoid the casework.

Secondly, the interesting questions are not really about what we can do if given arbitrarily many squares. Indeed, the main trick in the above proof with the greedy algorithm is only possible because we can continually add on more squares to get larger numbers. There is no hope for this approach if we limit the number of squares.

Thirdly, the proof actually gives us a somewhat explicit bound on the number of distinct squares we need to represent $n,$ though it isn't close to sharp. We begin by refining the lemma.
\begin{lemma}
    Fixing an arbitrarily small $\varepsilon>0,$ there exists a large constant $N$ such that for any $n\ge N,$ there exists a positive integer $k$ with $n-129\ge k^2>(1-\varepsilon)n.$
\end{lemma}
The proof of this is identical to the previous lemma. Take $k=\floor{\sqrt{n-129}}$ so that $k^2\le n-129<(k+1)^2.$ It follows
\[(1-\varepsilon)n<(1-\varepsilon)\left((k+1)^2+129\right).\]
To be assured that $k^2>(1-\varepsilon)n,$ it suffices for $k^2$ to exceed the right-hand side here. That is, we want
\[(1-\varepsilon)\left(k^2+2k+130\right)\le k^2,\]
which rearranges into
\[130(1-\varepsilon)\le\varepsilon k^2-(1-\varepsilon)\cdot 2k.\]
The right-hand side will eventually grow arbitrarily large because the $\varepsilon k^2$ term dominates, we so can be sure that this holds for sufficiently large $k.$ But $k=\floor{\sqrt{n-129}}$ means that this will also hold for sufficiently large $n$ giving sufficiently large $k.$ This finishes $\blacksquare$

Observe that the proof above used $\varepsilon=1/2,$ but we had no good reason to prefer this constant other than that it made the proof actually function by having $n-k^2\le n/2.$ Using a smaller $\varepsilon$ will also work just fine, though it increases the number of needed base cases.
\begin{proposition}
    Fixing a small $\varepsilon\in(0,1/2],$ there exists a constant $C$ so that every positive integer $n$ greater than $128$ can be written as the sum of at most $-\log_\varepsilon(n)+C$ distinct squares.
\end{proposition}
Fix some small $\varepsilon>0$ to use with the previous lemma, and extract our $N.$ We essentially do the same greedy algorithm from before but keep track of the number of squares in the induction.

For our base cases, the computer program from earlier can verify that the largest number of distinct squares needed to represent any integer from $129$ to $N$ is some finite number, say $M.$ Setting $C=M+\log_\varepsilon N,$ for example, guarantees the bound holds for our base cases.

The inductive step is also pretty much the same. Fix $m>N$ and suppose that all positive integers $n$ below $m$ but above $129$ can be written as the sum of $-\log_\varepsilon(n)+C$ squares. Again, we can use the lemma to extract some $k$ for which
\[(1-\varepsilon)m<k^2\le m-129.\]
In particular, the number $m-k^2\ge129$ can be written as the sum of at most $-\log_\varepsilon\left(m-k^2\right)+C\le-\log_\varepsilon(\varepsilon m)+C=-\log_\varepsilon(m)+C-1$ distinct squares.

All of these squares are bounded above $m-k^2\le\varepsilon m\le m/2,$ which $k^2$ exceeds, so indeed $k^2$ is distinct from any square in this representation, and we may safely add it back in while keeping all squares distinct. This totals to at most $\log_\varepsilon(m)-C-1+1=\log_\varepsilon(m)+C$ total squares. $\blacksquare$

The statement of the previous proposition is written to be easier to prove, but we can just remark that $-\log_\varepsilon(n)=\frac1{-\log\varepsilon}\cdot\log n.$ As $\varepsilon\to0,$ we have $\log\varepsilon\to-\infty,$ so $\frac1{-\log\varepsilon}\to0,$ so we can refine our statement to the following.
\begin{proposition}
    Fixing a sufficiently small $\varepsilon>0,$ there exists a constant $C$ so that every positive integer $n$ greater than $128$ can be written as the sum of at most $\varepsilon\log(n)+C$ distinct squares.
\end{proposition}
The given $\varepsilon$ can be used to solve for some $\delta$ such that $\frac1{-\log\delta}=\varepsilon.$ Explicitly, we have $\delta:=\exp\left(-\varepsilon^{-1}\right).$ Now, for sufficiently small $\varepsilon,$ we can guarantee $\delta\in(0,1/2]$ because $\varepsilon\to0^+$ sends $\delta\to0^+.$ In this case the previous proposition applies to $\delta,$ and we get the conclusion. $\blacksquare$

Thus, we get that the number of distinct squares required to decompose is something like $o(\log n)$ without having to do any really fancy number theory; after all, the hard part was combinatorics. However, I am obligated to mention that the Hardy-Littlewood circle method can show that five distinct squares is enough to get all but finitely many positive integers. I do not know this proof.

\subsubsection{June 13th}
Today I learned a few ways to evaluate $\zeta$ at even integers, from THE BOOK, and some consequences.
\begin{lemma}
    We have that
    \[\pi\cot(\pi x)=\lim_{N\to\infty}\sum_{k=-N}^N\frac1{x+k}.\]
\end{lemma}
There are a variety of proofs of this, but we will not do this here because I do not know any proof that is not its worth own entry. For completeness, we outline the proof from THE BOOK, which uses the amazingly clever Herglotz trick. Define
\[f(x)=\pi\cot(\pi x)-\lim_{N\to\infty}\sum_{k=-N}^N\frac1{x+k}.\]
Then one can show that $f(x)$ is periodic with period $1$ and continuous at any point in $\RR\setminus\ZZ$ as well as that $f(x)\to0$ as $x\to0.$

Now, the Herglotz trick consists of showing that $f(x)$ satisfies the functional equation
\[f\left(\frac x2\right)+f\left(\frac{x+1}2\right)=2f(x)\]
and then using this to show that $f$ must be constantly $0.$ Indeed, $|f|$ has a maximum value because it is periodic and continuous and so attains a maximum on the compact set $(0,1),$ which is enough. Say that $|f(x_0)|=M$ is maximal. The trick is to note that applying the triangle inequality to
\[f\left(\frac{x_0}2\right)+f\left(\frac{x_0+1}2\right)=2M\]
requires that $\left|f\left(\frac{x_0}2\right)\right|=M$ achieves the maximum also. So $x_0$ achieving maximum magnitude implies $x_0/2$ does as well, which forms a sequence approaching $0$ giving absolute value $M.$ But we know $f(x)\to0$ as $x\to0,$ so $M=0,$ finishing. $\blacksquare$

Anyways, we use this to derive values of $\zeta$ at positive even integers. The main character of our story will be
\[x\cot(x).\]
We will derive a Taylor expansion for this function in two ways and derive the desired result in a generating functions sort of way. We can rearrange the lemma into
\[\pi x\cot(\pi x)=\lim_{N\to\infty}\sum_{k=-N}^N\frac x{x+k}=1+\sum_{k=1}^\infty\left(\frac x{x-k}+\frac x{x+k}\right).\]
This gives
\[\pi x\cot(\pi x)=1+\sum_{k=1}^\infty\frac{2x^2}{x^2-k^2}.\]
The idea now is that we want to expand out $\frac{2x^2}{x^2-k^2}$ as a geometric series to hopefully pick up a series with $\frac1{k^n}$ in it. Because we want $k$ in the denominator later in life, we write
\[\pi x\cot(\pi x)=1-2\sum_{k=1}^\infty\frac{(x/k)^2}{1-(x/k)^2}.\]
Now, for $|x|<1,$ we have $|x/k|<1$ always, so we may expand the inner sum as an infinite geometric series as
\[\pi x\cot(\pi x)=1-2\sum_{k=1}^\infty\sum_{n=1}^\infty\left(\frac xk\right)^{2s}.\]
We know that this converges for $|x|<1,$ and all terms are positive, so in fact it is absolutely converging. Thus, we may swap the order of summation and factor out a $x^{2n}$ to get
\[\pi x\cot(\pi x)=1-2\sum_{n=1}^\infty\left(x^{2n}\sum_{k=1}^\infty\frac1{k^{2n}}\right).\]
That inner sum is $\zeta(2n),$ so we have the following.
\begin{lemma}
    For $|x|<1,$ we have that
    \[\pi x\cot(\pi x)=1-2\sum_{n=1}^\infty\zeta(2n)x^{2n}.\]
\end{lemma}
This follows from the above discussion. $\blacksquare$

It remains to actually evaluate the Taylor expansion of $\pi x\cot(\pi x)$ in a second way. We will not be using the theory developed above. The main hope is that $\exp$ is relatively well-studied, so we write
\[\cot x=\frac{\cos x}{\sin x}=\frac{\left(e^{ix}+e^{-ix}\right)/2}{\left(e^{ix}-e^{-ix}\right)/2i}=i\cdot\frac{e^{2ix}+1}{e^{2ix}-1}.\]
At this point we can almost the Bernoulli numbers appearing, so we introduce their definition.
\begin{definition}
    The Bernoulli numbers $\{B_k\}_{k=0}^\infty$ are defined by the generating function
    \[\frac x{e^x-1}=\sum_{k=0}^\infty B_k\cdot\frac{x^k}{k!},\]
    which has radius convergence $2\pi.$
\end{definition}
One important fact about Bernoulli numbers is that they are rational, which we will outline quickly. We can prove this induction by noting $x\to0$ gives $B_0=1$ as our base case, and then
\[x=\left(e^x-1\right)\left(\sum_{k=0}^\infty B_k\cdot\frac{x^k}{k!}\right)=\left(\sum_{\ell=1}^{\infty}\frac{x^\ell}{\ell!}\right)\left(\sum_{k=0}^\infty B_k\cdot\frac{x^k}{k!}\right).\]
Expanding this equation out gives a series of systems of equations, all of whose terms are rational, and solving this system must preserve being rational, so we conclude the $B_\bullet$ are rational. To be explicit, equating the coefficient for $x^n$ with $n>0,$ we see
\[\sum_{\ell=1}^n\frac{B_{n-\ell}}{\ell!(n-\ell)!}=1_{n=1},\]
which is our equation. Solving for $B_{n-1}$ gives
\[B_{n-1}=(n-1)!\left(1_{n=1}-\sum_{\ell=2}^n\frac{B_{n-\ell}}{\ell!(n-\ell)!}\right)\]
is our explicit recursion, and indeed, it only requires field operations. This is not terribly important, but it's a nice property.

Anyways, we return to talking about $\cot.$ Rearranging what we have to try to employ this definition, we write
\[\cot x=i+\frac{2i}{e^{2ix}-1}.\]
Now, we multiply by $x$ to plug in the definition of Bernoulli numbers with $x\mapsto2ix,$ as
\[x\cot x=ix+\frac{(2ix)}{e^{(2ix)}-1}=ix+\sum_{k=0}^\infty B_k\cdot\frac{(2ix)^k}{k!}.\]
This gives us our other Taylor expansion of $\pi x\cot(\pi x).$
\begin{lemma}
    We have that
    \[\pi x\cot(\pi x)=\pi ix+\sum_{k=0}^\infty\frac{B_k(2\pi i)^k}{k!}x^k.\]
\end{lemma}
This follows from taking $x\mapsto\pi x$ in the above formula. $\blacksquare$

Now, comparing coefficients between our two expansions of $\pi x\cot(\pi x),$ we get a number of nice properties. Indeed, the work above shows that
\[1+\sum_{n=1}^\infty(-2\zeta(2n))x^{2n}=\pi ix+\sum_{k=0}^\infty\frac{B_k(2\pi i)^k}{k!}x^k.\]
These functions define the same function over, say, $|x|\le\min\{1,2\pi\},$ so these must actually be the same power series. To start off, here is our main attraction: an explicit formula for $\zeta(2n),$ showing that it is a rational multiple of $\pi^{2n}.$
\begin{theorem}
    For positive integers $n,$ we have that
    \[\zeta(2n)=\frac{(-1)^{n+1}B_{2n}}{2(2n)!}\cdot(2\pi)^{2n}.\]
\end{theorem}
Comparing the coefficients of $x^{2n}$ and $x^k$ gives
\[-2\zeta(2n)=\frac{B_{2n}(2\pi i)^{2n}}{(2n)!}.\]
This simplifies into
\[\zeta(2n)=(-1)^{n}\cdot\frac{B_{2n}(2\pi)^{2n}}{-2(2n)!},\]
which rearranges into the desired. $\blacksquare$

However, we even get some other properties of the Bernoulli numbers based off of properties of the left-hand side. For example, we have the following.
\begin{proposition}
    We have that $B_1=-1$ and $B_{2n+1}=0$ for positive integers $n.$
\end{proposition}
Indeed, there is no odd power in the right-hand side of the equality
\[1+\sum_{n=1}^\infty(-2\zeta(2n))x^{2n}=\pi ix+\sum_{k=0}^\infty\frac{B_k(2\pi i)^k}{k!}x^k,\]
so we conclude that odd powers must vanish for the right-hand side as well. At $k=1,$ the coefficient is
\[0=\pi i+\frac{B_1(2\pi i)^1}{1!},\]
which implies that $B_1=-\frac12.$ Continuing, when $k=2n+1,$ we have
\[0=\frac{B_{2n+1}(2\pi i)^{2n+1}}{(2n+1)!},\]
from which $B_{2n+1}=0$ immediately follows. $\blacksquare$

Of course, there are more direct proofs of the above fact; for example, $\frac x{e^x-1}+\frac12x=\frac x2\cdot\frac{e^x+1}{e^x-1}$ is the product of two odd functions and is therefore even, implying that its power series must have $B_1=-1$ and $B_{2n+1}=0$ for $n>0.$ However, the fact that we are able to get this result quickly from the theory is remarkable.

We can even talk intelligently about $B_{2n}$ using properties of $\zeta,$ if we're careful.
\begin{proposition}
    We have that $B_{2n}\sim(-1)^{n+1}\cdot\frac{2(2n)!}{(2\pi)^{2n}}.$
\end{proposition}
For this we recall that
\[\zeta(2n)=\frac{(-1)^{n+1}B_{2n}}{2(2n)!}\cdot(2\pi)^{2n}\]
and read this backwards to get
\[B_{2n}=(-1)^{n+1}\cdot\frac{2(2n)!}{(2\pi)^{2n}}\cdot\frac1{\zeta(2n)}.\]
Now, as $n\to\infty,$ we see that
\[1\le\sum_{k=1}^\infty\frac1{k^{2n}}<1+\int_1^\infty\frac1{x^{2n}}\,dx=1+\frac1{2n-1}\]
will cause $\zeta(2n)\to1.$ Thus, we may ignore this term in the limit to conclude
\[B_{2n}\sim(-1)^{n+1}\cdot\frac{2(2n)!}{(2\pi)^{2n}}.\]
This is what we wanted. $\blacksquare$

We could use the Stirling formula approximation to get an estimate which is easier to handle, but the Stirling approximation scares me, so I'll avoid it. Nevertheless, this result, using properties of $\zeta$ to talk about the growth rate of Bernoulli numbers, is cute enough to call it quits here.

\subsubsection{June 14th}
Today I learned more formally about the association between points in $k^n$ and maximal ideals of $k[x_1,\ldots,x_n],$ from Undergraduate Commutative Algebra. We begin with the following motivating statement and proof.
\begin{proposition}
    Fix $K$ a field and $\alpha=(\alpha_1,\ldots,\alpha_n)\in K^n.$ Then the ideal $(x_1-\alpha_1,\ldots,x_n-\alpha_n)$ is maximal in $K[x_1,\ldots,x_n].$
\end{proposition}
The main character in this proof is the evaluation mapping $\op{ev}_\alpha:K[x_1,\ldots,x_n]\to k,$ taking $f\in K[x_1,\ldots,x_n]$ to
\[\op{ev}_\alpha(f)=f(\alpha)=f(\alpha_1,\ldots,\alpha_n).\]
It is a fact that $(f+g)(\alpha)=f(\alpha)+g(\alpha)$ and $(fg)(\alpha)=f(\alpha)g(\alpha),$ so this map preserves structure. We also have $\op{ev}_\alpha(1)=1$ because $1$ is a constant, so the identity is send to the identity. So $\op{ev}_\alpha$ is indeed a ring homomorphism.

We are going to the desired maximality by investigating the isomorphism induced by $\op{ev}_\alpha,$ so we need to compute the kernel and image of $\op{ev}_\alpha.$ Quickly, note that taking $f=k$ a constant function for $k\in K$ guarantees that $\op{ev}_\alpha$ has full image, regardless of $\alpha.$

It remains to compute the kernel of $\op{ev}_\alpha.$ This is somewhat involved, so we separate it out.
\begin{lemma}
    Take everything as defined above. We claim that $\ker(\op{ev}_\alpha)=(x_1-\alpha_1,\ldots,x_n-\alpha_n).$
\end{lemma}
We begin by working in the case where $\alpha=(0,\ldots,0).$ Quickly let
\[f(x)=\sum_{\{d_k\}_{k=1}^n\in\NN^n}a_{d_1,\ldots,d_n}\left(x_1^{d_1}\cdots x_n^{d_n}\right),\]
where all but finitely many of the $a_{d_1,\ldots,d_n}$ are zero. Then $f\in\op{ker}(\op{ev}_\alpha)$ in the case that $\alpha=(0,\ldots,0)$ is equivalent to having
\[0=f(\alpha)=a_{0,\ldots,0}\]
because $a_{d_1,\ldots,d_n}\left(x_1^{d_1}\cdots x_n^{d_n}\right)$ vanishes if any of the $d_\bullet$ are nonzero, implying that we only have to worry about $a_{0,\ldots,0}.$ In other words, $f\in\ker(\op{ev}_{(0,\ldots,0)})$ if and only if $f$ has constant term $0.$

Anyways, we claim that this is the ideal $(x_1,\ldots,x_n).$ Indeed, every element of $(x_1,\ldots,x_n)$ looks like
\[\sum_{k=1}^nf_kx_k\]
for some sequence of $f_\bullet\in K[x_1,\ldots,x_n],$ and this will expand out to have no constant term because each of the summands has no constant term. (Alternatively, plugging in $\alpha=(0,\ldots,0)$ causes the sum to vanish.) On the other hand, if the constant term of $a_{0,\ldots,0}=0,$ then every nonzero term of the sum
\[f(x)=\sum_{\{d_k\}_{k=1}^n\in\NN^n}a_{d_1,\ldots,d_n}\left(x_1^{d_1}\cdots x_n^{d_n}\right)\]
has some nonzero exponent of a $x_\bullet,$ implying that all terms live in $(x_1,\ldots,x_n),$ so $f(x)\in(x_1,\ldots,x_n).$ This finishes the claim.

Working in the general case of $\alpha=(\alpha_1,\ldots,\alpha_n),$ we note that $f\in\op{ker}(\op{ev}_\alpha)$ if and only if
\[g(x_1,\ldots,x_n):=f(x_1+\alpha_1,\ldots,x_n+\alpha_n)\]
vanishes at $(x_1,\ldots,x_n)=(0,\ldots,0).$ In other words, $f$ lives in the kernel if and only if
\[f(x_1+\alpha_1,\ldots,x_n+\alpha_n)\in(x_1,\ldots,x_n).\]
Shifting coordinates by taking $x_\bullet\mapsto x_\bullet-\alpha_\bullet$ (alternatively, fix $y_\bullet=x_\bullet-\alpha_\bullet$ and proceed), we get this equivalent to
\[f(x_1,\ldots,x_n)\in(x_1-\alpha_1,\ldots,x_n-\alpha_n).\]
So indeed, $\ker(\op{ev}_\alpha)=(x_1-\alpha_1,\ldots,x_n-\alpha_n),$ which is what we wanted. $\blacksquare$

Thus, we see that $\op{ev}_\alpha$ is a ring homomorphism with image all of $K$ and kernel exactly $(x_1-\alpha_1,\ldots,x_n-\alpha_n).$ This gives us the isomorphism
\[\frac{K[x_1,\ldots,x_n]}{(x_1-\alpha_1,\ldots,x_n-\alpha_n)}\cong K.\]
But now we're done! The quotient ring of $(x_1-\alpha_1,\ldots,x_n-\alpha_n)$ is a field, so this ideal must be maximal, finishing. $\blacksquare$

This proof has a few potential avenues of generalization. For example, the only part of the proof where we used the fact that $K$ is a field is at the very end where we need $K$ a field to give maximal. However, we can weaken $K$ field to $\mathcal O_K$ an integral domain to remain prime.
\begin{proposition}
    Fix $\mathcal O_K$ an integral domain and $\alpha:=(\alpha_1,\ldots,\alpha_n)\in(\mathcal O_K)^n.$ Then the ideal $(x_1-\alpha_1,\ldots,x_n-\alpha_n)$ is prime in $K[x_1,\ldots,x_n].$
\end{proposition}
The first proof works practically verbatim while replacing $K$ with $\mathcal O_K,$ using the same $\op{ev}_\alpha,$ etc., so we will not repeat it here. However, the very end must change, for now we see
\[\frac{\mathcal O_K[x_1,\ldots,x_n]}{(x_1-\alpha_1,\ldots,x_n-\alpha_n)}\cong\mathcal O_K.\]
Thus, the quotient ring of $(x_1-\alpha_1,\ldots,x_n-\alpha_n)$ is an integral domain, so this ideal must be prime. This is what we wanted. $\blacksquare$

Of course, we can see this result more directly once we know that $\ker(\op{ev}_\alpha)=(x_1-\alpha_1,\ldots,x_n-\alpha_n).$ Indeed, if a product $fg\in\ker(\op{ev}_\alpha),$ then $(fg)(\alpha),$ so
\[f(\alpha)g(\alpha)=0.\]
It follows that $f(\alpha)=0$ or $g(\alpha)=0$ because we are working in an integral domain, so one of $f$ or $g$ lives in $(x_1-\alpha_1,\ldots,x_n-\alpha_n),$ verifying our primality. (And this ideal is not all ideals because it does not have constants.) However, the given proof didn't even have to touch elements once establishing facts about $\op{ev}_\alpha,$ which is nice.

As an aside on this statement, we note that we have somehow created an exact way to detect if a ring $R$ is an integral domain or a field. Indeed, if the ideals $(x-\alpha)$ for $\alpha\in R$ is prime but not maximal in $R[x],$ then $R$ is an integral domain but not a field. Of course, this is just a more elaborate way of saying that $R$ is a field if and only if it has exactly two ideals, but this felt amusing anyways.

Another way to generalize the original proposition is to let $(\alpha_1,\ldots,\alpha_n)$ live in algebraic extension of $K.$ This turns out to work in the best possible way.
\begin{proposition}
    Fix $L/K$ an algebraic extension of fields with $\alpha=(\alpha_1,\ldots,\alpha_n)\in L^n.$ Then the ideal $(x_1-\alpha_1,\ldots,x_n-\alpha_n)\cap K[x_1,\ldots,x_n]$ is maximal in $K[x_1,\ldots,x_n].$
\end{proposition}
Slightly more care is needed for this because neither the image nor kernel are immediately well-behaved. However, we will still be more terse than in our first proof. We still use the original $\op{ev}_\alpha,$ though now $\op{ev}_\alpha:K[x_1,\ldots,x_n]\to L.$

The kernel turns out to be easier to handle. Indeed, fixing $f\in\ker(\op{ev}_\alpha),$ we can imagine embedding $f\in L[x_1,\ldots,x_n]$ in the obvious way, extending $\op{ev}_\alpha$ to domain $L[x_1,\ldots,x_n].$ In $L[x_1,\ldots,x_n],$ the work from earlier shows that our kernel is $(x_1-\alpha_1,\ldots,x_n-\alpha_n),$ so
\[f\in\ker(\op{ev}_\alpha)\iff f\in(x_1-\alpha_1,\ldots,x_n-\alpha_n).\]
However, this equivalence does not give an equality because we fixed $f\in K[x_1,\ldots,x_n]$ as a condition to start. Moving this condition into the equivalence, we get
\[f\in\ker(\op{ev}_\alpha)\iff f\in(x_1-\alpha_1,\ldots,x_n-\alpha_n)\cap K[x_1,\ldots,x_n].\]
Indeed, if $f\in\ker(\op{ev}_\alpha),$ then the argument above gives $f\in(x_1-\alpha_1,\ldots,x_n-\alpha_n),$ and we are in $K[x_1,\ldots,x_n]$ for free. And in the other direction, if $f$ is in $(x_1-\alpha_1,\ldots,x_n-\alpha_n),$ then $f(\alpha)=0,$ and to be in the kernel, we merely need to know $f\in K[x_1,\ldots,x_n]$ as well.

We now deal with the image. We claim that it is $K[\alpha_1,\ldots,\alpha_n].$ Indeed, this comes from moving our definitions around: $K[\alpha_1,\ldots,\alpha_n]$ can be defined as the set of elements of the form
\[\sum_{\{d_k\}_{k=1}^n\in\NN^n}a_{d_1,\ldots,d_n}\left(\alpha_1^{d_1}\cdots\alpha_n^{d_n}\right),\]
which is just $f(\alpha_1,\ldots,\alpha_n)$ where
\[f(x_1,\ldots,x_n)=\sum_{\{d_k\}_{k=1}^n\in\NN^n}a_{d_1,\ldots,d_n}\left(x_1^{d_1}\cdots x_n^{d_n}\right).\]
The annoying part of dealing with the kernel is to have that $K[\alpha_1,\ldots,\alpha_n]$ is actually a field. We show this in somewhat painstaking detail because I can't remember having done this before. Also, in it we read the above trick backwards, which is a bit amusing.
\begin{lemma}
    Fix an algebraic extension of fields $L/K$ and elements $\alpha_1,\ldots,\alpha_n\in L.$ Then $K[\alpha_1,\ldots,\alpha_n]$ is a field.
\end{lemma}
We are not going to be too formal about $K[\alpha_1,\ldots,\alpha_n],$ but we will assert that we can write $K[\alpha_1,\ldots,\alpha_n]=K[\alpha_1,\ldots,\alpha_{n-1}][\alpha_n].$ This can more or less be seen from the definition above by collecting terms with the same $\alpha_n^{d_n}$ term to  view these as polynomials with coefficients in $K[\alpha_1,\ldots,\alpha_{n-1}].$

The point of saying this is so that we can do an induction. Our base case is $n=0,$ for which there is nothing to prove. For the inductive step, We let $K_n:=K[\alpha_1,\ldots,\alpha_n]$ with $\alpha_{n+1}$ algebraic over $K$ (and therefore algebraic over $K_n$). We know that $K_n$ is a field and need to show that $K_n[\alpha_{n+1}]$ is a field.

Well, doing the above trick backwards, observe that we have a ring homomorphism $K_n[x]\to K_n[\alpha_{n+1}]$ by evaluating at $\alpha_{n+1}.$ If we let $\pi(x)$ be the minimal polynomial of $\alpha_{n+1},$ then $(\pi)$ is the kernel of this evaluation mapping. Indeed, certainly $p\in\ker$ if and only if $p(\alpha_{n+1})=0,$ for which the division algorithm dictates
\[p(x)=\pi(x)q(x)+r(x).\]
Evaluating at $\alpha_{n+1}$ requires that $r(\alpha_{n+1})=0,$ but this remainder must be zero or of smaller degree than $\pi$; the latter violates the minimality of $\pi,$ so $r=0,$ giving $\pi\mid p.$ And surely $p\in(\pi)$ gives $0=\pi(\alpha_{n=1})\mid p(\alpha_{n+1}).$ So we know that
\[K_n[\alpha_{n+1}]\cong K_n[x]/(\pi).\]
However, $(\pi)$ is also irreducible, for if $\pi=pq$ for polynomials $p,q\in K_n[x],$ then $p(\alpha_{n+1})q(\alpha_{n+1})=0$ implies $p(\alpha_{n+1})=0$ or $q(\alpha_{n+1})=0,$ so $\pi\mid p$ or $q,$ giving the irreducibility. (Without loss of generality, let $p=\pi p',$ which means $\pi=\pi p'q,$ giving $p'q=1,$ so $q$ is a unit.)

So because $K[x]$ is a unique factorization domain, it follows $(\pi)$ is prime, and because $K[x]$ is a principal ideal domain, it follows $(\pi)$ is maximal, so $K[x]/(\pi)$ is a field. Thus, $K_n[\alpha_{n+1}]$ is a field, completing our inductive step. $\blacksquare$

Returning to the proposition, we now note that our induced isomorphism is
\[\frac{K[x_1,\ldots,x_n]}{(x_1-\alpha_1,\ldots,x_n-\alpha_n)\cap K[x_1,\ldots,x_n]}\cong K[\alpha_1,\ldots,\alpha_n].\]
However, the right-hand side is of a field, so it follows $(x_1-\alpha_1,\ldots,x_n-\alpha_n)\cap K[x_1,\ldots,x_n]$ is maximal. This finishes. $\blacksquare$

We are obligated to remark that we can play the same game as earlier by replacing $K$ with an integral domain $\mathcal O_K$ while keeping $K$ the field of fractions of $\mathcal O_K.$ We won't write out the actual statement because it's not too remarkable, and we've already repeated ourselves enough today.

Anyways, the moral of the story is that individual points $(\alpha_1,\ldots,\alpha_n)$ induce maximal ideals in $K[x_1,\ldots,x_n]$ by being the kernel of the evaluation mapping. It is a bit concerning that we are not guaranteed all maximal ideals in the above propositions, and perhaps this is not surprising.

Indeed, the first proposition seems to generate ``lots'' of maximal ideals---one for each point---but when $K$ is not algebraically closed, then the final proposition gives us even more points to play with and therefore more maximal ideals. We don't really have any good reason to believe that we've found all the maximal ideals yet.

As some evidence that maybe we have found all the maximal ideals, observe that we have at least covered all maximal ideals when $n=1.$ Indeed, $K[x]$ is a principal ideal domain with primes induced by irreducible polynomials (which we won't prove here). Thus, for any maximal ideal $(\pi)$ with $\pi$ irreducible, we can take $L=K[\alpha]$ in our final proposition to retrieve $(\pi)$ as the ideal $(x-\alpha)\cap K[x]$ because $\pi$ is the minimal polynomial of $\alpha.$

\subsubsection{June 15th}
Today I learned the start of some group cohomology. Instead of describing what group cohomology is supposed to do, which is the morally correct path, we're going to describe how to make these groups. So as fair warning, I am well-aware that this will potentially be very boring.

We quickly give the definition of a $G$-module, for completeness.
\begin{definition}
    Fix a finite (but not necessarily abelian!) group $G.$ A $G$-module $M$ is an abelian group equipped with a left action. That is, we have a homomorphism $\sigma_\bullet:G\to\op{Aut}(M),$ meaning that $\sigma_e=\op{id},$ and for each $g\in G$ gives
    \[\sigma_g(a+b)=\sigma_ga+\sigma_gb,\]
    and $\sigma_{gh}=\sigma_g\sigma_h.$ We will suppress the $\sigma$ as much as possible.
\end{definition}
As a quick example, when $G=\langle e\rangle$ is the trivial group, $G$ can act on any, arbitrary abelian group $M$ because the action of $G$ is trivial.

Now that we have defined our algebraic object, we are of course interested in its maps.
\begin{definition}
    Fix a finite group $G.$ Given $G$-modules $M$ and $N,$ a $G$-module homomorphism $\varphi:M\to N$ is a homomorphism of the underlying abelian groups which preserves the group action. In other words, $\varphi$ is a group homomorphism $M\to N$ which also satisfies
    \[g\varphi(m)=\varphi(gm)\]
    for any $g\in G$ and $m\in M.$ (Note $g\varphi(m)$ has $G$ acting on $N.$) We will denote these homomorphisms as $\op{Hom}_G(M,N).$
\end{definition}
I am under the impression that, typically, checking that our group homomorphism preserves the group structure is not going to be too bad of a road block, but it must be done.

Anyways, continuing with our trivial example $G=\langle e\rangle,$ any group homomorphism $\varphi:M\to N$ of abelian groups is a $G$-module homomorphism, for
\[e\varphi(m)=\varphi(m)=\varphi(em)\]
because $G$ acts trivially. So in our trivial case, it's enough to have a group homomorphism.

To define cohomology, we are interested in injective resolutions; namely, quotients of images modulo kernels in injective resolutions will exactly be our group cohomology. But to define an injective resolution, we first have to define an injective $G$-module.
\begin{definition}
    A $G$-module $M$ is injective if and only if for any inclusion $A\subseteq B$ of $G$-modules, then $G$-module homomorphisms $\varphi_A:A\to M$ can always be extended out to maps $\varphi_B:B\to M.$ To be explicit, $\varphi_B(a)=\varphi_A(a)$ for any $a\in A.$
\end{definition}
What this means is that $M$ is somehow ``big enough'' to allow arbitrary refining of $G$-module homomorphisms into it. For our trivial case $G=\langle e\rangle,$ we can actually classify injective $G$-modules.
\begin{proposition}
    Fix $G=\langle e\rangle$ the trivial group. Then a $G$-module $M$ is injective if and only if it is ``divisible,'' meaning that the exponential maps $x\mapsto nx$ for any $n\in\ZZ\setminus\{0\}$ are all surjective.
\end{proposition}
In one direction, suppose that $M$ is injective; we want to show $M$ is injective. That is, for any $m\in M$ and $n\in\ZZ\setminus\{0\},$ there exists some $m_0\in M$ such that $nm_0=m.$

Now, the intuition give above says that we should be thinking about injectivity as permitting refinements, so we let $A=n\ZZ$ and $B=\ZZ$ in the definition of injectivity. The $G$-module homomorphism $A\to M$---which is really just a group homomorphism---has to send $n\in n\ZZ$ somewhere, and we get to choose, so we fix
\[\varphi_A(nk):=km\]
for any $nk\in n\ZZ.$ This is indeed a homomorphism of abelian groups (say, $\varphi(n(k+\ell))=(k+\ell)n,$ but we're really saying that abelian groups are a $(\ZZ,+,\times)$-module), so it is a valid $G$-module homomorphism. But now we need to be able to refine $\varphi_A$ to $\varphi_B:\ZZ\to B.$ This implies that
\[n\varphi_B(1)=\varphi_B(\underbrace{1+\cdots+1}_n)=\varphi_B(n)=\varphi_A(n)=m\]
because $\varphi_B$ is a group homomorphism refining $\varphi_A.$ In particular, we see $m_0:=\varphi_B(1)$ gives $nm_0=m,$ finishing the proof of divisibility.

In the other direction, suppose $M$ is a divisible abelian group, and we need to show that $M$ is an injective $G$-module. So fix an inclusion of $G$-modules (abelian groups) $A\subseteq B$ as well as a $G$-module homomorphism (group homomorphism) $\varphi_A:A\to M.$ We need to expand $\varphi_A$ to $\varphi_B:B\to M.$

This expansion is potentially troubling because we can have, say, $A=\ZZ$ and $B=\RR,$ making an explicit expansion quite difficult due to $\RR$'s transcendentals and the way $\QQ\subseteq\RR$ needs to interact with $\ZZ.$ So we kill this with Zorn's lemma.

Our partially ordered set will consist of pairs $(C,\varphi_C),$ where $C$ is an abelian group $A\subseteq C\subseteq B$ between $A$ and $B,$ and $\varphi_C:C\to M$ is a group homomorphism which restricts to $\varphi_A$ on $A.$ We will compare two pairs as
\[(C_1,\varphi_{C_1})\preceq(C_1,\varphi_{C_2})\]
if and only if $C_1$ is a subgroup of $C_2$ and $\varphi_{C_2}$ restricts to $\varphi_{C_1}$ on $C_1$: $\varphi_{C_2}(c_1)=\varphi_{C_1}(c_1)$ for any $c_1\in C_1.$

Verification of the chain condition is a bit annoying, so we relegate it to a lemma.
\begin{lemma}
    The poset defined above satisfies the chain condition.
\end{lemma}
Showing the chain condition is not that bad. Suppose $\{(C_\alpha,\varphi_{C_\alpha})\}_{\alpha\in\lambda}$ is an ascending chain. We claim that we can define $\varphi_C$ on
\[C:=\bigcup_{\alpha\in\lambda}C_\alpha\]
to be our maximal element. Observe any $c_\alpha\in C$ lives in some $c_\alpha\in C_\alpha,$ so in fact $c\in B.$ Thus, $C\subseteq B.$ Showing that $C$ is actually a subgroup comes down to checking the subgroup test, which we do for completeness.
\begin{itemize}
    \item For any $c_\alpha,c_\beta\in C,$ we can situate $c_\alpha\in C_\alpha$ and $c_\beta\in C_\beta.$ The chain condition means $C_\alpha\subseteq C_\beta$ (without loss of generality), so $c_\alpha c_\beta\in C_\beta$ is well-defined in the subgroup $C_\beta\subseteq B,$ and therefore we have $c_\alpha c_\beta\in C.$
    \item For any $c_\alpha\in C,$ we can again situate $c_\alpha\in C_\alpha.$ Here, $C_\alpha$ is already a subgroup of $B,$ so $c_\alpha^{-1}\in C_\alpha\subseteq C.$
\end{itemize}
We now define $\varphi_C:C\to M.$ Well, for any $c\in C,$ we can find it in some $c\in C_\alpha,$ so we send $\varphi_C(c):=\varphi_{C_\alpha}(c).$ This is well-defined because if we find $c$ in two $c\in C_\alpha$ and $c\in C_\beta,$ then the chain condition asserts $C_\alpha\subseteq C_\beta$ (without loss of generality), giving
\[\varphi_{C_\alpha}(c)=\varphi_{C_\beta}(c)\]
because of the definition of our poset. To verify that $\varphi_C$ is a group homomorphism, we fix $c_\alpha,c_\beta\in C$ and find $c_\alpha\in C_\alpha$ and $c_\beta\in C_\beta.$ Then the chain condition implies $C_\alpha\subseteq C_\beta$ (without loss of generality), so
\[\varphi_C(\underbrace{c_\alpha+c_\beta}_{\in C_\beta})=\varphi_{C_\beta}(c_\alpha+c_\beta)=\varphi_{C_\beta}(c_\alpha)+\varphi_{C_\beta}(c_\beta)=\varphi_C(c_\alpha)+\varphi_C(c_\beta).\]
Finally, to verify that $(C,\varphi_C)$ is indeed maximal to this chain, we note $C_\alpha\subseteq C$ for any $C_\alpha$ because of the way we defined $C$---its group operation restricts to $C_\alpha$ by construction. And we defined $\varphi_C$ exactly to restrict to any of the $\varphi_{C_\alpha}$ (and so also to $\varphi_A$). This finishes. $\blacksquare$

Having satisfied the chain condition, Zorn's lemma lets us pick up a maximal element. It remains to show that we must have $B$ in this maximal element. Notice that we haven't used the divisibility of $M$ yet, so it had better come up now.

Well, now fix any pair $(C,\varphi_C)$ such that $B\ne C,$ and we claim that this pair is not maximal. Indeed, pick up $g\in B\setminus C,$ and we will extend $(C,\varphi_C)$ to $C+\langle g\rangle,$ which is strictly bigger than $C.$ (This sum is well-defined because $C\subseteq B$ is a subgroup, so we can define the arithmetic over $B.$) We still have $A\subseteq C+\langle g\rangle\subseteq B,$ so the task is to extend $\varphi_C$ to some $\varphi_{C+\langle g\rangle}.$

To extend, let $\langle g\rangle\cap C=\langle ng\rangle$ for some $n\in\NN,$ possibly $n=0.$ (Such an $n$ exists as either $n=0$ or the least positive $n$ for which $ng\in C.$) To exhibit our homomorphism, we must say where $g$ goes. If $n=0,$ we set its image to be $m:=0\in M.$ If $n>0,$ then we know where $\varphi(ng)=n\varphi(g)\in M$ goes, so we pick up $m$ using the surjectivity of $x\mapsto nx.$ (Here we used divisibility!) So we set
\[\varphi_{C+\langle g\rangle}(c+kg)=\varphi_C(c)+km\]
for any $c\in C$ and $kg\in\langle g\rangle.$ This is well-defined because $c+kg=c'+k'g$ implies $c'=c+(k-k')g$ as well as $(k-k')g\in C,$ so $(k-k')\in\langle ng\rangle$ and $n\mid k-k'.$ It follows
\[\varphi_C(c')+k'm=\varphi_C(c+(k-k')g)+k'm=\varphi_C(c)+\varphi_C\left(\frac{k-k'}n\cdot ng\right)+k'm=\varphi_C(c)+km.\]
This definition of $\varphi_{C+\langle g\rangle}$ is also a group homomorphism because
\[\varphi_{C+\langle g\rangle}(c+kg+c'+k'g)=\varphi_C(c+c')+(k+k')g=\varphi_C(c+kg)+\varphi_C(c+k'g).\]
And lastly, $\varphi_{C+\langle g\rangle}$ does indeed restrict to $\varphi_C$ (and therefore $\varphi_A$) by setting $k=0$ in its definition. This completes the construction of $(C+\langle g\rangle,\varphi_{C+\langle g\rangle}).$

It follows that any pair $(C,\varphi_C)$ with $C\ne B$ is not maximal, so the maximal element given by Zorn's lemma must be a pair $(B,\varphi_B)$ with $\varphi_B$ restricting to $\varphi_A.$ This completes the proof that the module $M$ is injective. $\blacksquare$

That was a lot of work, but with a classification in hand, we can intelligently talk about embedding modules into injective ones.
\begin{lemma}
    Fix $G=\langle e\rangle$ the trivial group. Then any $G$-module $M$ can be embedded into an injective $G$-module.
\end{lemma}
Translating the statement for our trivial case, this means we need to embed an arbitrary abelian group $M$ into a divisible abelian group. Well, we claim that
\[\QQ[M]:=\left\{\{a_m\}_{m\in M}\subseteq\QQ:\text{all but finitely many }m\text{ give }a_m=0\right\}\]
will do the trick. This is an abelian group, where our operation is pointwise summation; two elements $\{a_m\}_{m\in M}$ and $\{b_m\}_{m\in M}$ are equal if and only if $a_m-b_m\in\ZZ$ for all $m\in M$ and
\[\sum_{m\in M}(a_m-b_m)m=0,\]
where this sum is taken over $M.$ (This sum is finite because all but finitely many of the $\{a_m\}_{m\in M}$ and $\{b_m\}_{m\in M}$ are $0.$) Our embedding (call it $\iota$) takes $m\in M$ to $1m\in\QQ[M],$ meaning $a_m=1$ while $a_{m'}=0$ for $m'\ne m.$ This is a group homomorphism because
\[\iota(m+n)-\iota(m)-\iota(n)\]
is the sequence $\{a_k\}_{k\in M}$ where $a_{m+n}=1,a_m=-1,a_n=-1,$ which is equal to $0\in\QQ[M]$ because it evaluates to $1(m+n)-1m-1n=0\in M.$

It remains to show that $\QQ[M]$ is divisible. Well, fix any $n\in\ZZ\setminus\{0\}$ so that we need to show that the map $x\mapsto nx$ is surjective. This means that, given $\{a_m\}_{m\in M}\subseteq\QQ$ where all but finitely many are $0,$ we need to find an element $\{b_m\}_{m\in M}\subseteq\QQ$ with all but finitely many are $0$ satisfying
\[n\{b_m\}_{m\in M}=\{a_m\}_{m\in M}.\]
With this in mind, we take $b_m=a_m/n$ (note $n\ne0$). Then $\{b_m\}_{m\in M}$ is nonzero only where $\{a_m\}_{m\in M}$ is, so $\{b_m\}_{m\in M}$ is nonzero finitely often. Further, the group operation of pointwise addition implies that $n\{b_m\}_{m\in M}$ is pointwise equal to $\{a_m\}_{m\in M}$ and hence actually equal because $\{a_m\}_{m\in M}-\{b_m\}_{m\in M}$ is identically $0.$ This finishes. $\blacksquare$

It turns out that any $G$-module $M$ (for an arbitrary group $G$) can always be embedded into some injective $G$-module. I do not currently know the full proof of this statement, so we will take it on faith for the time being. The point of all of this is to be able to provide injective resolutions.
\begin{definition}
    Fix a finite group $G$ and $M$ a $G$-module. An injective resolution of $M$ is a sequence of injective $G$-modules $\left\{I^k\right\}_{k\in\NN}$ and $G$-module homomorphisms $\del^k:I^k\to I^{k+1}$ such that
    \[0\to M\to I^0\stackrel{\del^0}\to I^1\stackrel{\del^1}\to I^2\stackrel{\del^2}\to\cdots\]
    is an exact sequence.
\end{definition}
The work we did above (and assumed) tells us that every $G$-module has an injective resolution.
\begin{lemma}
    Fix $G$ a finite group. Then every $G$-module $M$ has an injective resolution.
\end{lemma}
Define $I^0$ to be the injective $G$-module containing $I^{-1}:=M,$ and then after that recursively define $I^{k+1}$ as an injective $G$-module containing $I^k/I^{k-1}.$ Explicitly, for $k\ge0,$ we make $\del^k:I^k\to I^{k+1}$ mod out by $I^{k-1}$ and then embed the new object $I^k/I^{k-1}$ into $I^{k+1}$ so that
\[\op{im}\del^{k-1}=\ker\del^k,\]
where $\del^{-1}:M\to I^0$ is the embedding. Indeed, $\del^k:I^k\to I^{k+1}$ takes $I^k$ to a copy of $I^k/I^{k-1}$ inside of $I^{k+1},$ meaning that the kernel comes exactly from the elements of $I^{k-1}$ that found there way into $I^k$ through $\del^{k-1}.$ This gives the exactness.

We do technically have to check that the $\del^k:I^k\to I^{k+1}$ are $G$-module homomorphisms, which is annoying but not hard. We do this inductively. At $k=-1,$ this is true by hypothesis on our embedding $M\to I^0.$ Now suppose $\del^k$ is a $G$-module homomorphism so that we want to show $\del^{k+1}$ is.

Note $\del^{k+1}:I^{k+1}\to I^{k+2},$ as described, mods out by the image $\del^k\left(I^k\right)$ and then embeds into an injective $G$-module; we show that these two maps are individually $G$-module homomorphisms. The embedding we automatically know is a $G$-module homomorphism because of its construction.

As for the modding, we note that the image of $\del^k$ is itself a $G$-module, inherited from $I^{k+1}.$ This inheritance gives subgroup structure because $\del^k$ is a group homomorphism, and the group action comes from $I^{k+1},$ which $\del^k\left(I^k\right)$ is closed under because, for any $g\in G$ and $a\in I^k,$
\[g\left(\del^k(a)\right):=\del^k(ag),\]
where we have used the inductive hypothesis that $\del^k$ is a $G$-module homomorphism. This means $I^{k+1}/\del^k\left(I^k\right)$ is a $G$-module, where our action is
\[g\left(a+\del^k\left(I^k\right)\right)=ga+\del^k\left(I^k\right),\]
for any $g\in G$ and $a\in I^{k+1}.$ This is well-defined because $a+\del^k\left(I^k\right)=b+\del^k\left(I^k\right)$ implies $a-b\in\del^k\left(I^k\right)$ implies $ga-gb\in\del^k\left(I^k\right)$ implies $ga+\del^k\left(I^k\right)=gb+\del^k\left(I^k\right).$

Now, the modding out $I^{k+1}\to I^{k+1}/\del^k\left(I^k\right)$ is already a group homomorphism, so we have to check this is well-defined as a $G$-module homomorphism. It is already a group homomorphism, so we merely have to show $a\in I^{k+1}$ and $g\in G$ give $g\left(a+\del^k\left(I^k\right)\right)=ga+\del^k\left(I^k\right),$ which is true by definition. So we are done here. $\blacksquare$

We've come all this way, so we will spend a little more time defining group cohomology to make it worthwhile. The following is our motivation.
\begin{definition}
    Fix $G$ a finite group. We define the fixed-point functor $(-)^G$ as taking $G$-modules $M$ to the set of fixed points $M^G=\{x\in M:gx=x\text{ for any }g\in G\}.$
\end{definition}
What's interesting about this functor is that it takes short exact sequences to left exact sequences.
\begin{proposition}
    Fix $G$ a finite group and $G$-modules $A,B,C$ such that
    \[0\to A\to B\to C\to 0\]
    is an exact sequence. Then we have the left exact sequence
    \[0\to A^G\to B^G\to C^G.\]
\end{proposition}
Let $\varphi:A\to B$ and $\psi:B\to C$ be the middle maps. We have to show that $\varphi$ restricts to an injective function $A^G\to B^G$ (making $0\to A^G\to B^G$ exact) whose image is the kernel of $\psi$ (making $A^G\to B^G\to C^G$ exact), and we have to show $\psi$ also restricts to a function $B^G\to C^G.$

We already know that $\varphi$ is injective, so we really just have to show that $a\in A^G$ implies $\varphi(a)\in B^G.$ Well, picking up any $g\in G,$ we see
\[g(\varphi(a))=\varphi(ga)=\varphi(a)\]
because $\varphi$ is a $G$-module homomorphism and so preserves the group action. The same argument, replacing $\varphi$ with $\psi$ and taking $A$ to $B$ and $B$ to $C,$ shows that $\psi$ also restricts to a function $B^G\to C^G.$

Lastly, we already know that the image of $\varphi$ is the kernel of $\psi$ from the original short exact sequence, and this property does not change on restriction; the image of $\varphi$ got smaller, but it was still contained in the domain (and hence kernel) of $\psi.$ So we are done. $\blacksquare$

What we did not show was surjectivity of $\psi:B^G\to C^G,$ and indeed it is possible for surjectivity to fail. This failure gives rise to group cohomology.
\begin{definition}
    Fix $G$ a finite group. Take our injective resolution for a $G$-module $M$
    \[0\to M\to I^0\stackrel{\del^0}\to I^1\stackrel{\del^1}\to I^2\stackrel{\del^2}\to\cdots,\]
    we can apply the fixed-point functor to each element. Because the embedding mappings $\del^k$ are still $G$-module homomorphisms, they will restrict to maps $\left(I^k\right)^G\to\left(I^{k+1}\right)^G.$ The fact that these are restrictions means that the $\op{im}\del^k\subseteq\ker\del^{k+1}$ still, but we lose assurance of equality. We define, then
    \[H^k(G,M):=\ker\del^k/\op{im}\del^{k-1}.\]
\end{definition}
We remark that this definition appears to depend on the injective resolution, but it turns out not to. We will not show this here because I do not know how.

\subsubsection{June 16th}
Today I learned finish to the proof that any $G$-module can be embedded into an injective $G$-module. This turns out to be somewhat involved. We recall from yesterday that any abelian group can be embedded into a divisible abelian group, which is the same thing as an injective $\langle e\rangle$-module. Here is the main attraction for today.
\begin{proposition}
    Fix $G$ a finite group. Then any $G$-module $M$ can be embedded into an injective $G$-module.
\end{proposition}
The key construction is to use the above lemma to embed the abelian group $M$ into a divisible group, which we will call $N,$ and then fix $\op{Hom}_\ZZ(\ZZ[G],N),$ which consist of the group homomorphisms $\ZZ[G]\to N.$ The $\ZZ$ in the subscript means that we require $\varphi:\ZZ[G]\to N$ to be $\ZZ$-linear, meaning that
\[\varphi(ng)=n\varphi(g)\]
for $n\in\ZZ$ and $g\in\ZZ[G].$ However, this follows by, say, induction on the fact $\varphi$ is a group homomorphism already, so the $\ZZ$ is mostly redundant to point out that we are not trying to be $G$-linear.

Anyways, to turn $\op{Hom}_\ZZ(\ZZ[G],N)$ into a right (!) $G$-module, we first note that $\ZZ[G]$ is a $G$-module, where $g$ acts on $\sum_{h\in G}a_hh$ by taking it to $\sum_{h\in G}a_h(hg).$ Indeed, $\times g$ is a group homomorphism because
\[\left(\sum_{h\in G}(a_h+b_h)h\right)g=\sum_{h\in G}(a_h+b_h)(hg)=\sum_{h\in G}a_h(hg)+\sum_{h\in G}b_h(hg).\]
This action is $G$-compatible because
\[\left(\sum_{h\in G}a_h\right)(gg')=\sum_{h\in G}a_hh(gg')=\left(\left(\sum_{h\in H}a_hh\right)g\right)g'.\]
Now, given $\varphi:\ZZ[G]\to N$ and $g\in G,$ we define the action (note the multiplication on the right!)
\[(g\varphi)(a):=\varphi(ag),\]
which works because $\times g$ is a group homomorphism, and the composition of group homomorphisms is still a group homomorphism. The inversion here makes this $G$-compatible: $(gh)\varphi(a)=\varphi\left(a(gh)\right)=\varphi\left((ag)h\right)=g(h\varphi)(a).$ The simplicity of this (not requiring any inversion) is why we made $\ZZ[G]$ a right $G$-module, though of course it is also a left $G$-module.

Anyways, to finish showing that $\op{Hom}_\ZZ(\ZZ[G],N)$ is a $G$-module, we need to know the action is linear, but we can fix $\varphi,\varphi':\ZZ[G]\to N$ and $g\in G$ so that
\[(g(\varphi+\varphi'))(r)=(\varphi+\varphi')(rg)=\varphi(rg)+\varphi'(rg)=(g\varphi+g\varphi')(r).\]
So indeed, $\op{Hom}_\ZZ(\ZZ[G],N)$ is a $G$-module. Note that the proof of this only required knowing that $\ZZ[G]$ was a right $G$-module.

We divide the rest of the proof of the proposition into steps. The first part will focus on $M$ because we need to embed $M$ into $\op{Hom}_\ZZ(\ZZ[G],N).$
\begin{lemma}
    Fixing everything as above, there is a injective $G$-module homomorphism $M\to\op{Hom}_\ZZ(\ZZ[G],N).$
\end{lemma}
We begin by describing our embedding $M\to\op{Hom}_\ZZ(\ZZ[G],N).$ Write $\iota_N:M\to N$ for the embedding of $M$ in $N,$ and then we send $m\in M$ to
\[\varphi_m\left(\sum_{g\in G}a_gg\right):=\sum_{g\in G}a_g\cdot\iota_N(gm),\]
where the final sum is evaluate in $N.$ Of note $gm$ is the group action of $G$ on $M.$ This $\varphi_m$ is a group homorphism because
\[\varphi_m\left(\sum_{g\in G}(a_g+b_g)g\right)=\sum_{g\in G}(a_g+b_g)\cdot\iota_N(gm)=\sum_{g\in G}a_g\cdot\iota_N(gm)+\sum_{g\in G}b_g\cdot\iota_N(gm),\]
which collapses back into $\varphi_m\left(\sum_{g\in G}a_gg\right)+\varphi_m\left(\sum_{g\in G}b_gg\right).$

We have left to show that the map $m\mapsto\varphi_m$ is an injective $G$-module homomorphism. It is a group homomorphism because
\[\varphi_{m+n}\left(\sum_{g\in G}a_gg\right)=\sum_{g\in G}a_g\cdot\iota_N(g(m+n))=\sum_{g\in G}a_g\cdot\iota_N(gm)+\sum_{g\in G}a_g\cdot\iota_N(gn)\]
by using the fact $G$ is an action and $\iota_N$ a group homomorphism. This collapses to $\varphi_m+\varphi_n,$ so we are safe. To finish showing that the embedding is a $G$-module homomorphism, we have to show that it preserves the group action, which means we need $\varphi_{gm}=g\varphi_m,$ for which we note
\[(g\varphi_m)\left(\sum_{h\in G}a_hh\right)=\sum_{h\in G}a_h\iota_N((hg)m)=\varphi_{gm}\left(\sum_{h\in G}a_hh\right).\]
So we have established that our map $M\to\op{Hom}_\ZZ(\ZZ[G],N)$ is a $G$-module homomorphism. Our final promise is that it is an embedding, but note
\[\varphi_m(1e)=1\iota_N(em)=\iota_N(m),\]
so the input $1e$ detects $m.$ Explicitly, if $\varphi_m=\varphi_n,$ then plugging in $1e$ means $\iota_N(m)=\iota_N(n),$ from which $m=n$ follows because $\iota_N$ is an embedding. $\blacksquare$

At this point we can throw out the $G$-module $M.$ Indeed, the second part of the proof will be showing that $\op{Hom}_\ZZ(\ZZ[G],N)$ is an injective $G$-module.
\begin{lemma}
    Fixing everything as above, we have that $\op{Hom}_\ZZ(\ZZ[G],N)$ is an injective $G$-module.
\end{lemma}
The main idea in the proof is currying. Indeed, fix $G$-modules $A\subseteq B$ with a map $\varphi_A:A\to\op{Hom}_\ZZ(\ZZ[G],N)$ so that we need to extend $\varphi_A$ out to $\varphi_B:B\to\op{Hom}_\ZZ(\ZZ[G],N).$ Because $N$ is divisible, we know group homomorphisms into $N$ can be extended, so we try to message such a group homomorphism.

If we ignore all the structure present, then the way to turn a map of the form $A\to\op{Hom}_\ZZ(\ZZ[G],N)$ into a map into $N$ is to think about this like $A\to(\ZZ[G]\to N)$ and then un-curry into $A\times\ZZ[G]\to N.$ However, this will not let us preserve our group structure, for
\[\begin{cases}
    \varphi_A(a_1+a_2)(r_1)=\varphi_A(a_1)(r_1)+\varphi_A(a_2)(r_1), \\
    \varphi_A(a_1)(r_1+r_2)=\varphi_A(a_1)(r_1)+\varphi_A(a_1)(r_2),
\end{cases}\]
for any $a_1,a_2\in A$ and $r_1,r_2\in\ZZ[G].$ The point is that our map $A\times\ZZ[G]\to N$ needs to be linear in both arguments.

To keep track of this algebraic structure, the key observation is to push it into the domain! Namely, we need the tensor product. Simply being linear in both arguments gives $A\otimes_\ZZ\ZZ[G],$ but we note that we also have
\[\varphi_A(ga)(r)=(g\varphi_A(a))(r)=\varphi_A(a)(rg)\]
by definition of the action on $\op{Hom}_\ZZ(\ZZ[G],N),$ so we also want linearity in $G.$

Thus, we define the following object. Define $A\otimes_G\ZZ[G]$ as
\[A\otimes_G\ZZ[G]=(A\times\ZZ[G])/\sim,\]
which is the group generated $\ZZ$-linearly by pairs in $A\times\ZZ[G]$ modulo $\sim$ defined to give all of the following relations: given $a,a_1,a_2\in A$ and $r,r_1,r_2\in\ZZ[G]$ with $g\in G,$
\begin{enumerate}[label=(\alph*)]
    \item $(a,r_1+r_2)\sim(a,r_1)+(a,r_2),$
    \item $(a_1+a_2,r)\sim(a_1,r)+(a_2,r),$
    \item $(ga,r)\sim(a,rg).$
\end{enumerate}
We will denote the equivalence classes of the pair $(a,r)$ under $\sim$ by $a\otimes r$ for $a\in A$ and $r\in\ZZ[G].$ The definition we gave technically defines this group as elements of the form
\[\sum_{a\in A,r\in\ZZ[G]}k_{a,r}(a\otimes r)\]
where all but finitely many of the $k_{a,r}$ are zero. Defining addition pointwise on the coordinates $a\otimes r,$ we need to check that the addition is well-defined and that it creates a group.

Checking that this forms a group is done in the same way as checking $\ZZ[G]$ is a group. Checking that addition is well-defined is annoying, but we will say that if two $\sum_{a,r}k_{a,r}(a\otimes r)$ and $\sum_{a,r}k_{a,r}'(a\otimes r)$ are equal, then there is a finite sequence of steps (a), (b), and (c) proving their equality. So if $\sum_{a,r}\ell_{a,r}(a\otimes r)=\sum_{a,r}\ell_{a,r}'(a\otimes r)$ are also equal by some sequence of steps,
\[\sum_{a,r}(k_{a,r}+\ell_{a,r})(a\otimes r)=\sum_{a,r}(k_{a,r}'+\ell_{a,r}')(a\otimes r)\]
can be established by doing the sequence for $\sum k=\sum k'$ and $\sum\ell=\sum\ell'$ one after the other. We won't dwell on this point.

Anyways, with this construction in hand, we have the following lemma.
\begin{lemma}
    Fix $G$ a finite group and $A$ a $G$-module with $N$ an arbitrary abelian group. A $G$-module homomorphism $\varphi_A:A\to\op{Hom}_\ZZ(\ZZ[G],N)$ contains the same data as a group homomorphism $\psi_A:A\otimes_G\ZZ[G]\to N.$ Namely, having one induces the other.
\end{lemma}
For the forwards direction, suppose we have a $G$-module homomorphism $\varphi_A:A\to\op{Hom}_\ZZ(\ZZ[G],N)$ and want to induce a group homomorphism $\psi_A:A\otimes_G\ZZ[G]\to N.$ We claim that extending
\[\psi_A(a\otimes r):=\varphi_A(a)(r)\]
linearly will work. Most of our time is going to be spent showing that this is a well-defined map. It suffices to show that this definition $\psi_A$ respects the equivalence relation $\sim.$ (If two elements in $A\otimes_G\ZZ[G]$ are equal with $\sim,$ then there should be a finite number of moves of the type given above showing the equality. So respecting the individual moves is enough by, say, induction.) Fix $a,a_1,a_2\in A$ and $r,r_1,r_2\in\ZZ[G]$ with $g\in G,$
\begin{enumerate}[label=(\alph*)]
    \item Note $\varphi_A(a)(r_1+r_2)=\varphi_A(a)(r_1)+\varphi_A(a)(r_2)$ because $\varphi_A(a)$ is a group homomorphism. This gives $\psi_A(a\otimes(r_1+r_2))=\psi_A(a\otimes r_2)+\psi_A(a\otimes r_2).$
    \item Note $\varphi_A(a_1+a_2)(r)=\varphi_A(a_1)(r)+\varphi_A(a_2)(r)$ because $\varphi_A$ is itself a group homomorphism. This gives $\psi_A((a_1+a_2)\otimes r)=\psi_A(a_1\otimes r)+\psi_A(a_2\otimes r).$
    \item Note $\varphi_A(ga)(r)=\big(g\varphi_A(a)\big)(r)=\varphi_A(a)(rg)$ because $\varphi_A$ is a $G$-module homomorphism. This gives $\psi_A(ga\otimes r)=\psi_A(a\otimes rg).$
\end{enumerate}
Having shown all the individual components of $\sim,$ we can be comforted that $\psi_A$ is well-defined. It remains to show that $\psi_A$ is a group homomorphism. Well, we defined $\psi_A$ to extend linearly, which gives this for free. (I am growing tired of being rigorous.) This finishes this direction.

For the backwards direction, suppose we have a $\psi_A:A\otimes_G\ZZ[G]\to N$ and want to induce $\varphi_A:A\to\op{Hom}_\ZZ(\ZZ[G],N).$ Then we simply set
\[\varphi_A(a)(r):=\psi_A(a\otimes r).\]
This time around this is a perfectly good function. We might be afraid of collisions, but at least this function is surely well-defined. To start, this $\varphi_A$ does indeed give homomorphisms $\op{Hom}_\ZZ(\ZZ[G],N)$ because, fixing $a\in A$ and picking up $r_1,r_2\in\ZZ[G],$ we see
\[\varphi_A(a)(r_1+r_2)=\psi_A(a\otimes(r_1+r_2))=\psi_A(a\otimes r_1+a\otimes r_2)=\psi_A(a\otimes r_1)+\psi_A(a\otimes r_2),\]
which is indeed $\varphi_A(a)(r_1)+\varphi(a)(r_2).$ Secondly, $\varphi_A$ is itself a group homomorphism because, picking up $a_1,a_2\in A$ and arbitrary $r\in\ZZ[G],$
\[\varphi_A(a_1+a_2)(r)=\psi_A((a_1+a_2)\otimes r)=\psi_A(a_1\otimes r+a_2\otimes r)=\psi_A(a_1\otimes r)+\psi_A(a_2\otimes r),\]
which again correctly comes out to $\varphi_A(a_1)(r)+\varphi(a_2)(r).$ Lastly, $\varphi_A$ is a $G$-module homomorphism because, fixing $a\in A,$ $r\in\ZZ[G]$ and now $g\in G,$ we have
\[\big(g\varphi_A(a)\big)(r)=\varphi_A(a)(rg)=\psi_A(a\otimes rg)=\psi_A(ga\otimes r)=\varphi_A(ga)(r).\]
This finishes this direction and so the proof of the lemma. $\blacksquare$

With the above lemma in hand, we can quickly get the result. Note that the give $\varphi_A$ induces a group homomorphism $\psi_A:A\otimes_G\ZZ[G]\to N.$ Now, the divisibility of $N$ (we finally use the divisibility!) implies that we can pull this $\psi_A$ back to a $\psi_B:B\otimes_G\ZZ[G]$ because
\[A\otimes_G\ZZ[G]\subseteq B\otimes_G\ZZ[G].\]
(There are literally more pairs in the second.) But then the $\psi_B$ induces a homomorphism of $G$-modules $\varphi_B:B\to\op{Hom}(\ZZ[G],N).$ This is what we wanted. $\blacksquare$

As some final commentary, I am amused by the trick of encoding algebraic information about maps into the structures themselves. From the outset, it is quite reasonable to expect that $G$-module homomorphisms $A\to\op{Hom}_\ZZ(\ZZ[G],N)$ to be harder to talk about than group homomorphisms, but if we make the groups sufficient complicated, it turns out to be the same data!

Anyways, we remark that we have somewhat explicitly given a way to get an injective resolution, with the exception of the use of the Zorn's lemma to retrieve the injectivity maps for a divisible abelian group. Regardless, we do know that we can embed an arbitrary $G$-module $M$ into
\[M\hookrightarrow\op{Hom}_\ZZ(\ZZ[G],\QQ[M]).\]
Now we can write the injective resolution from yesterday as having $I^{-2}=0$ and $I^{-1}=M$ as base cases and then
\[I^{k+2}=\op{Hom}_\ZZ\left(\ZZ[G],\QQ[I^{k+1}/\del^k(I^k)]\right),\]
where $\del^{-2}:0\to M$ and the rest are defined recursively by the composition of
\[I^{k+1}\to I^{k+1}/\del^k\left(I^k\right)\hookrightarrow I^{k+2}.\]
Granted, this construction isn't pretty, but it is still a construction, and we can be guaranteed that it always works.

\subsubsection{June 17th}
Today I learned that group cohomology is well-defined. Recall our definition from a two days ago.
\begin{definition}
    Fix $G$ a finite group. Take our injective resolution for a $G$-module $M$
    \[0\to M\to I^0\stackrel{\del^0}\to I^1\stackrel{\del^1}\to I^2\stackrel{\del^2}\to\cdots,\]
    we can apply the fixed-point functor to each element. Because the embedding mappings $\del^k$ are still $G$-module homomorphisms, they will restrict to maps $\left(I^k\right)^G\to\left(I^{k+1}\right)^G.$ The fact that these are restrictions means that the $\op{im}\del^k\subseteq\ker\del^{k+1}$ still, but we lose assurance of equality. We define, then
    \[H^k(G,M):=\ker\del^k/\op{im}\del^{k-1}.\]
    After taking the fixed-point functor, we will set $\del^{-1}:M^G\to(I^0)^G$ to be the zero mapping so that $H^0(G,M)=M^G/0=M^G.$
\end{definition}
Yesterday we completed the proof that every $G$-module has an injective resolution, which completed the proof of the existence of the cohomology. Today we show that $H(G,M)$ does not depend on our choice of injective resolution.

We will want the homorphism theorem for $G$-modules, so we quickly get it out of the way. This holds in arbitrary abelian categories, but I don't know the definition of an abelian category, so we will just show this.
\begin{lemma}
    Fix $G$ a finite group and $\varphi:M\to N$ a $G$-module homomorphism. Then we have an isomorphism of $G$-modules
    \[\varphi(M)\cong M/\ker\varphi,\]
    and all of these objects are well-defined $G$-modules.
\end{lemma}
We start by showing that $\ker\varphi$ is a $G$-submodule of $M.$ Looking at the underlying abelian groups, it is already a subgroup. To show that it is fixed by the action of $G,$ pick up any $g\in G$ and $m\in\ker\varphi$ and observe
\[\varphi(gm)=g\cdot\varphi(m)=g\cdot 0_N=0_N,\]
so indeed, $gm\in\ker\varphi.$ (I guess $g\cdot0=0$ because $g\cdot0=g\cdot(0+0)=g\cdot0+g\cdot0.$) It follows that $M/\ker\varphi$ is defined as a $G$-module: it is already an abelian group, and its group action is
\[g\left(a+\ker\varphi\right)=ga+\ker\varphi,\]
which is well-defined because $a+\ker\varphi=b+\ker\varphi,$ which implies $a-b\in\ker\varphi,$ which implies $a-b\in\ker\varphi,$ which implies $ga-gb\in\ker\varphi,$ which implies $ga$ and $gb$ are in the same coset.

To get the isomorphism asserted, we send $m+\ker\varphi\in M/\ker\varphi$ to $\varphi(m).$ By group theory, this is already a group isomorphism, so we only have to check that the group action is preserved. Picking up a $g\in G,$ we note
\[g(m+\ker\varphi)=gm+\ker\varphi\mapsto\varphi(gm)=g\cdot\varphi(m),\]
so indeed, this is a $G$-module homomorphism. Finally, we note that $\varphi(M)$ is immediately a $G$-module because it is isomorphic to one. This finishes the proof. $\blacksquare$

Our main tool showing that $H^\bullet(G,M)$ is well-defined will be the fact that a $G$-module homomorphism $f:M\to N$ induces a map between cohomology groups.
\begin{lemma}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Then a $G$-module homomorphism $\varphi:M\to N$ induces maps $\varphi^k:I^k\to J^k$ such that the following diagram commutes.
    \begin{center}
        \begin{tikzcd}
            0 \arrow[r] & M \arrow[r] \arrow[d, "\varphi"] & I^0 \arrow[r, "\del_I^0"] \arrow[d, "\varphi^0"] & I^1 \arrow[r, "\del_I^1"] \arrow[d, "\varphi^1"] & I^2 \arrow[r, "\del_I^2"] \arrow[d, "\varphi^2"] & \cdots \\
            0 \arrow[r] & N \arrow[r]                      & J^0 \arrow[r, "\del_J^0"']                       & J^1 \arrow[r, "\del_J^1"']                       & J^2 \arrow[r, "\del_J^2"']                       & \cdots
        \end{tikzcd}
    \end{center}
\end{lemma}
Note that to prove the diagram commutes, we merely have to prove that each square commutes. We can show this inductively, but is is more directly true that two consecutive arrows in the top or bottom complex give the $0$ mapping, so they will commute just fine. We will not dwell on this.

Anyways, we do induction on $k.$ By convention, we will set $M=I^{-1}$ and $N=K^{-1}$ so that $\varphi^{-1}:=\varphi.$ Additionally, we set $\del_I^{-2}:0\to M$ and $\del_J^{-2}:0\to N$ with $\del_I^{-1}:M\to I^0$ and $\del_J^{-1}:N\to J^0$ to be the embeddings. (Our $\del_\bullet^{-1}$ are not the zero mapping here!) Observe that we can also realize these as first modding out $I^{-1}$ by ``$I^{-2}=0$'' and then embedding into an injective $G$-module, just like the higher $\del^k$s.

For the inductive step, we are allowed to assume the previous square commutes (for $k=-1,$ this square includes $0\to 0$) and that $I^{k+1}$ and $J^{k+1}$ are injective $G$-modules. (For $k=-1,$ $I^k$ and $J^k$ are not injective.) We need to induce $\varphi^{k+1}:I^{k+1}\to J^{k+1}.$ For this, we decompose $\del^k_I$ and $\del^k_J$ into modding followed by embedding.
\begin{center}
    \begin{tikzcd}
        I^k \arrow[d, "\varphi^k"] \arrow[r] & I^k/\del_I^{k-1}\left(I^{k-1}\right) \arrow[r, hook] \arrow[d, "\psi^k", dashed] & I^{k+1} \arrow[d, "\varphi^{k+1}", dashed] \\
        J^k \arrow[r]                         & J^k/\del_J^{k-1}\left(J^{k-1}\right) \arrow[r, hook]                             & J^{k+1}                                   
    \end{tikzcd}
\end{center}
As for why we can do this, we note that $\del^k$ has kernel exactly $\del^{k-1}\left(I^{k-1}\right)$ because $I^{k-1}\to I^k\to I^{k+1}$ is exact, so we are allowed to mod out by this image, and after we do, we have an isomorphism
\[\del^k_I\left(I^k\right)\cong\frac{I^k}{\ker\del^k_I}=\frac{I^k}{\del^{k-1}_I\left(I^{k-1}\right)}\]
induced by the $G$-module homomorphism $\del^k.$ The same holds for $J.$ This means that after the modding out, we have an isomorphism into the image, which is the same thing as an embedding.

Anyways, we now induce $\psi^k$ and $\varphi^{k+1}.$ We can induce $\psi^k$ directly from the previous square.
\begin{center}
    \begin{tikzcd}
        I^{k-1} \arrow[r, "\del_I^{k-1}"] \arrow[d, "\varphi^{k-1}"] & I^k \arrow[d, "\varphi^k"] \arrow[r] & I^k/\del_I^{k-1}\left(I^{k-1}\right) \arrow[d, "\psi^k", dashed] \\
        J^{k-1} \arrow[r, "\del_J^{k-1}"']                           & J^k \arrow[r]                         & J^k/\del_J^{k-1}\left(J^{k-1}\right)                            
    \end{tikzcd}
\end{center}
(Our base case is still valid because $\varphi^{k-1}:I^{k-1}\to J^{k-1}$ is $0\to 0.$) We begin creating $\psi^k$ from the composite
\[I^k\stackrel{\varphi^k}\to J^k\to J^k/\del_J^{k-1}\left(J^{k-1}\right).\]
Now, we want to show that this mapping is defined up to coset of $\del^{k-1}_I\left(I^{k-1}\right),$ so we have to show that the image $\del_I^{k-1}\left(I^{k-1}\right)$ lives in the kernel of the above map. Well, the commutativty of the previous square says
\[\varphi^k\left(\del_I^{k-1}\left(I^{k-1}\right)\right)=\del_J^{k-1}\left(\varphi^{k-1}\left(I^{k-1}\right)\right)\subseteq\del_J^{k-1}\left(J^{k-1}\right),\]
so indeed, we may mod out by $\del_I^{k-1}\left(I^{k-1}\right).$ That is, $\psi^k$ is just $\varphi^k$ after modding out the domain and codomain.

We take a moment to verify that the constructed square commutes.
\begin{center}
    \begin{tikzcd}
        I^k \arrow[d, "\varphi^k"] \arrow[r] & I^k/\del_I^{k-1}\left(I^{k-1}\right) \arrow[d, "\psi^k", dashed] \\
        J^k \arrow[r]                         & J^k/\del_J^{k-1}\left(J^{k-1}\right)                            
    \end{tikzcd}
\end{center}
Indeed, $\psi^k$ is defined to take any coset $a+\del_I^{k-1}\left(I^{k-1}\right)$ directly to $\varphi(a)+\del_J^{k-1}\left(J^{k-1}\right),$ where we just spend the previous paragraph showing well-definedness. So going around the top or bottom of the square sends $a\in I^k$ to $\varphi(a)+\del_J^{k-1}\left(J^{k-1}\right).$

It remains to induce $\varphi^{k+1}.$ For this we focus on the right half of our original diagram, as follows.
\begin{center}
    \begin{tikzcd}
        I^k/\del_I^{k-1}\left(I^{k-1}\right) \arrow[r, hook] \arrow[d, "\psi^k", dashed] & I^{k+1} \arrow[d, "\varphi^{k+1}", dashed] \\
        J^k/\del_J^{k-1}\left(J^{k-1}\right) \arrow[r, hook]                             & J^{k+1}                                   
    \end{tikzcd}
\end{center}
We now induce $\varphi^{k+1}$ from the injectivity of $I^{k+1}$ and $J^{k+1}.$ Indeed, the top embedding lets us write (abusing notation) $I^k/\del_I^{k-1}\left(I^{k-1}\right)\subseteq I^{k+1},$ and we already have a $G$-module homomorphism $I^k/\del_I^{k-1}\left(I^{k-1}\right)\to J^k/\del_J^{k-1}\left(J^{k-1}\right)\to J^{k+1}.$ So we can extend this $G$-module homomorphism to all of $I^{k+1}$ by the injectivity of $J^{k+1},$ which is our $\varphi^{k+1}.$

We now finish by showing that the square we just constructed commutes. Then the inductive step will be done by adjoining the two commuting squares, which will create the desired commuting square. Namely, we show the following commutes.
\begin{center}
    \begin{tikzcd}
        I^k/\del_I^{k-1}\left(I^{k-1}\right) \arrow[r, hook] \arrow[d, "\psi^k", dashed] & I^{k+1} \arrow[d, "\varphi^{k+1}", dashed] \\
        J^k/\del_J^{k-1}\left(J^{k-1}\right) \arrow[r, hook]                             & J^{k+1}                                   
    \end{tikzcd}
\end{center}
Indeed, $\varphi^{k+1}$ is defined as extending the path through the right of the square. Explicitly, $a\in I^k/\del_I^{k-1}\left(I^{k-1}\right)\subseteq I^{k+1}$ is sent to the same place along $I^k/\del_I^{k-1}\left(I^{k-1}\right)\to J^k/\del_J^{k-1}\left(J^{k-1}\right)\to J^{k+1}$ as along $\varphi^{k+1}.$ This finishes the proof. $\blacksquare$

The above construction is so important, it has a name.
\begin{definition}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Then a $G$-module homomorphism $\varphi:M\to N$ induces maps $\varphi^k:I^k\to J^k$ such that the following diagram commutes.
    \begin{center}
        \begin{tikzcd}
            0 \arrow[r] & M \arrow[r] \arrow[d, "\varphi"] & I^0 \arrow[r, "\del_I^0"] \arrow[d, "\varphi^0"] & I^1 \arrow[r, "\del_I^1"] \arrow[d, "\varphi^1"] & I^2 \arrow[r, "\del_I^2"] \arrow[d, "\varphi^2"] & \cdots \\
            0 \arrow[r] & N \arrow[r]                      & J^0 \arrow[r, "\del_J^0"']                       & J^1 \arrow[r, "\del_J^1"']                       & J^2 \arrow[r, "\del_J^2"']                       & \cdots
        \end{tikzcd}
    \end{center}
    We call these induced maps a ``chain morphism lifting $\varphi$'' and label them by $\varphi^\bullet:I^\bullet\to J^\bullet.$
\end{definition}

These induced maps in fact generate a map between cohomology groups, as promised.
\begin{proposition}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, fix a $G$-module homomorphism $\varphi:M\to N$ and the chain morphism $\varphi^\bullet:I^\bullet\to J^\bullet$ lifting $\varphi.$ Then these $\varphi^\bullet$ induce maps $H^\bullet\left(\varphi^\bullet\right):H^\bullet(G,M)\to H^\bullet(G,N).$
\end{proposition}
As discussed a few days ago, a $G$-module homomorphism induces a map between the $G$-fixed points of the domain and codomain. So we have the following commutative diagram created by restricting the hypothesized chain morphism.
\begin{center}
    \begin{tikzcd}
        0 \arrow[r] & M^G \arrow[r] \arrow[d, "\varphi"] & \left(I^0\right)^G \arrow[r, "\del_I^0"] \arrow[d, "\varphi^0"] & \left(I^1\right)^G \arrow[r, "\del_I^1"] \arrow[d, "\varphi^1"] & \left(I^2\right)^G \arrow[r, "\del_I^2"] \arrow[d, "\varphi^2"] & \cdots \\
        0 \arrow[r] & N^G \arrow[r]                      & \left(J^0\right)^G \arrow[r, "\del_J^0"']                       & \left(J^1\right)^G \arrow[r, "\del_J^1"']                       & \left(J^2\right)^G \arrow[r, "\del_J^2"']                       & \cdots
    \end{tikzcd}
\end{center}
Now, we need to induce a map $H^k(G,M)\to H^k(G,N),$ which given these injective resolutions, means we need a map from $\ker\del_I^k/\op{im}\del_I^{k-1}$ to $\ker\del_J^k/\op{im}\del_J^{k-1}.$

To begin, we consider the following commuting square.
\begin{center}
    \begin{tikzcd}
       \left(I^k\right)^G \arrow[r, "\del_I^k"] \arrow[d, "\varphi^k"] & \left(I^{k+1}\right)^G \arrow[d, "\varphi^{k+1}"] \\
       \left(J^k\right)^G \arrow[r, "\del_J^k"']                       & \left(J^{k+1}\right)^G
    \end{tikzcd}
\end{center}
Note that $\ker\del_I^k$ vanishes along $\varphi^{k+1}\circ\del_I^k,$ so it must also vanish along $\del_J^k\circ\varphi^k.$ This means that $\varphi^k$ will restrict to a map $\ker\del_I^k\to\ker\del_J^k,$ so we hope that our $\varphi^k$ will induce a 

Next, we look at the other side.
\begin{center}
    \begin{tikzcd}
       \left(I^{k-1}\right)^G \arrow[r, "\del_I^{k-1}"] \arrow[d, "\varphi^{k-1}"] & \left(I^k\right)^G \arrow[d, "\varphi^k"] \\
       \left(J^{k-1}\right)^G \arrow[r, "\del_J^{k-1}"']                       & \left(J^k\right)^G
    \end{tikzcd}
\end{center}
Now, we notice that the image of $\varphi^k\circ\del_I^{k-1}$ is equal to the image of $\del_J^{k-1}\circ\varphi^{k-1}$ and hence a subset of the image $\del_J^{k-1}.$ Thus, $\varphi^k$ takes the image of $\del_I^{k-1}$ to the image of $\del_J^{k-1}$ on restriction.

Before continuing, we take a second to codify what just happened.
\begin{lemma}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, fix a $G$-module homomorphism $\varphi:M\to N$ and the chain morphism $\varphi^\bullet:I^\bullet\to J^\bullet$ lifting $\varphi.$ Then a $\varphi^k$ restricts to a map $\ker\del_I^k\to\ker\del_J^k$ and $\op{im}\del_I^{k-1}\to\op{im}\del_J^{k-1}.$
\end{lemma}
This follows from the above discussion. $\blacksquare$

Anyways, bringing this together, we take the restricted map
\[\varphi^k:\ker\del_I^k\to\ker\del_J^k\]
and mod out the image by the $G$-(sub)module $\op{im}\del_J^{k-1}.$ Because now $\varphi^k$ (under restriction) takes $\op{im}\del_I^{k-1}$ into the kernel of $\op{im}\del_J^{k-1},$ we are guaranteed a well-defined map
\[\ker\del_I^k/\op{im}\del_I^{k-1}\to\ker\del_J^k/\op{im}\del_J^{k-1}\]
induced by $\varphi^k.$ This is exactly what we wanted. $\blacksquare$

What's amazing about the induced map of cohomology groups is that they are unique up to choice of $\varphi.$ Namely, exactly which $\varphi^\bullet$ we chose is irrelevant.
\begin{lemma}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, fix a $G$-module homomorphism $\varphi:M\to N$ and maps $\varphi_1^\bullet$ and $\varphi_2^\bullet$ to chain morphisms lifting $\varphi.$ In other words, they commute with themselves (but not necessarily each other) as in the following diagram.
    \begin{center}
        \begin{tikzcd}
        	0 & M & {I^0} & {I^1} & {I^2} & \cdots \\
        	0 & N & {J^0} & {J^1} & {J^2} & \cdots
        	\arrow[from=1-1, to=1-2]
        	\arrow[from=2-1, to=2-2]
        	\arrow[from=1-2, to=1-3]
        	\arrow[from=2-2, to=2-3]
        	\arrow["\varphi", from=1-2, to=2-2]
        	\arrow["{\varphi_2^0}", shift left=1, color={rgb,255:red,92;green,92;blue,214}, from=1-3, to=2-3]
        	\arrow["{\varphi_1^0}"', shift right=1, color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=2-3]
        	\arrow["{\varphi_2^1}", shift left=1, color={rgb,255:red,92;green,92;blue,214}, from=1-4, to=2-4]
        	\arrow["{\varphi_1^1}"', shift right=1, color={rgb,255:red,214;green,92;blue,92}, from=1-4, to=2-4]
        	\arrow["{\partial_I^0}", from=1-3, to=1-4]
        	\arrow["{\partial_J^0}"', from=2-3, to=2-4]
        	\arrow["{\partial_J^1}"', from=2-4, to=2-5]
        	\arrow["{\partial_I^1}", from=1-4, to=1-5]
        	\arrow["{\varphi_2^2}", shift left=1, color={rgb,255:red,92;green,92;blue,214}, from=1-5, to=2-5]
        	\arrow["{\varphi_1^2}"', shift right=1, color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-5]
        	\arrow["{\partial_I^2}", from=1-5, to=1-6]
        	\arrow["{\partial_J^2}"', from=2-5, to=2-6]
        \end{tikzcd}
    \end{center}
    (As a warning, the red and blue arrows do not have to commute with each other.) Then the induced maps $H^\bullet(\varphi_1^\bullet)$ and $H^\bullet\left(\varphi_2^\bullet\right)$ on $H^\bullet(G,M)\to H^\bullet(G,N)$ are the same.
\end{lemma}
Before doing anything, we start by saying when $\bullet=0,$ the map of cohomology groups $H^0(G,M)=M^G$ to $H^0(G,N)=N^G$ induced by either $\varphi_1^\bullet$ or $\varphi_2^\bullet$ is the same because this is just the restriction of $\varphi.$ We say this now because we are about to set an unusual convention to $\del_\bullet^{-1}$ (namely, we make it the embedding, not the zero map) in order to make our diagram commute nicer.

Our first goal will be to decrease the number of arrows. Extend all of $I^k,J^k,\varphi^k_\bullet,\del_\bullet^k$ backwards to $k=-1$ and $k=-2,$ where $\varphi_1^{-1}=\varphi_2^{-1}=\varphi$ and $\varphi_1^{-2}=\varphi_2^{-2}:0\to 0.$ The commutative diagram tells us that $\del_J^k\circ\varphi_\bullet^k=\varphi_\bullet^{k+1}\circ\del_I^k$ for any $k\ge-2,$ which implies that, upon subtracting,
\[\del_J^k\circ\left(\varphi_2^k-\varphi_1^k\right)=\left(\varphi_2^{k+1}-\varphi_1^{k+1}\right)\circ\del_I^k.\]
Indeed, the sum/difference of two $G$-module homomorphisms $\varphi_1:A\to B$ and $\varphi_2:A\to B$ is another $G$-module homomorphism because the underlying group homomorphisms work, and for any $g\in G$ and $a\in A,$ $(\varphi_1-\varphi_2)(ga)=\varphi_1(ga)-\varphi_2(ga)=g\cdot(\varphi_1(a)-\varphi_2(a))$ by linearity. Then the fact we have a group homomorphism means, for any $a\in I^k,$
\[\begin{cases}
    \left(\del_J^k\circ\left(\varphi_2^k-\varphi_1^k\right)\right)(a)=\left(\del_J^k\circ\varphi_2^k\right)(a)-\left(\del_J^k\circ\varphi_1^k\right)(a), \\
    \left(\left(\varphi_2^{k+1}-\varphi_1^{k+1}\right)\circ\del_I^k\right)(a)=\left(\varphi_2^{k+1}\circ\del_I^k\right)(a)-\left(\varphi_1^{k+1}\circ\del_I^k\right)(a),
\end{cases}\]
so the commutativity of the diagram gives the assertion. We remark that the above was merely showing that morphisms form a ring of composition over addition, which is checked when we say the category of $G$-modules is abelian. We will not show this explicitly again.

Recollecting our morphisms, the following diagram commutes.
\begin{center}
    \begin{tikzcd}
    	0 & M & {I^0} & {I^1} & {I^2} & \cdots \\
    	0 & N & {J^0} & {J^1} & {J^2} & \cdots
    	\arrow[from=1-1, to=1-2]
    	\arrow[from=2-1, to=2-2]
    	\arrow[from=1-2, to=1-3]
    	\arrow[from=2-2, to=2-3]
    	\arrow["{\varphi-\varphi}", from=1-2, to=2-2]
    	\arrow["{\varphi_2^0-\varphi_1^0}", from=1-3, to=2-3]
    	\arrow["{\varphi_2^1-\varphi_1^1}", from=1-4, to=2-4]
    	\arrow["{\partial_I^0}", from=1-3, to=1-4]
    	\arrow["{\partial_J^0}"', from=2-3, to=2-4]
    	\arrow["{\partial_J^1}"', from=2-4, to=2-5]
    	\arrow["{\partial_I^1}", from=1-4, to=1-5]
    	\arrow["{\varphi_2^2-\varphi_1^2}", from=1-5, to=2-5]
    	\arrow["{\partial_I^2}", from=1-5, to=1-6]
    	\arrow["{\partial_J^2}"', from=2-5, to=2-6]
    \end{tikzcd}
\end{center}
This diagram does indeed have fewer arrows, so we would like to work with it instead. Recall that the maps $\varphi_\bullet^k$ induces of $H^k(G,M)\to H^k(G,N)$ were simply $\varphi_\bullet^k$ restricted and then modded. So if we want to show $\varphi_1^k$ and $\varphi_2^k$ induce the same map, it suffices to show that $\varphi_2^k-\varphi_1^k$ induces (i.e., restricts to) the zero map.

Thus, it suffices to show the following. Note $\varphi-\varphi=0.$
\begin{lemma}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, suppose we have a chain morphism $\varphi^k:I^k\to J^k$ lifting $0:M\to N.$ Then the induced $H^\bullet\left(\varphi^\bullet\right)$ is the $0$ map on $H^\bullet(G,M)\to H^\bullet(G,N)$ for positive exponents.
\end{lemma}
The idea is to create a chain homotopy. Again, extend all of $I^k,J^k,\varphi^k_\bullet,\del_\bullet^k$ backwards to $k=-1$ and $k=-2.$ We claim that we exhibit maps $\psi^k:I^{k+1}\to J^k$ for $k\ge-2$ such that $\varphi^k=\psi^k\circ\del_I^k+\del_J^{k-1}\circ\psi^{k-1}$ for $k\ge-1.$ Here is the diagram.
\begin{center}
    \begin{tikzcd}
    	0 & M & {I^0} & {I^1} & {I^2} & \cdots \\
    	0 & N & {J^0} & {J^1} & {J^2} & \cdots
    	\arrow[from=1-1, to=1-2]
    	\arrow[from=2-1, to=2-2]
    	\arrow[from=1-2, to=1-3]
    	\arrow[from=2-2, to=2-3]
    	\arrow["0", from=1-2, to=2-2]
    	\arrow["{\varphi^0}", from=1-3, to=2-3]
    	\arrow["{\varphi^1}", from=1-4, to=2-4]
    	\arrow["{\partial_I^0}", from=1-3, to=1-4]
    	\arrow["{\partial_J^0}"', from=2-3, to=2-4]
    	\arrow["{\partial_J^1}"', from=2-4, to=2-5]
    	\arrow["{\partial_I^1}", from=1-4, to=1-5]
    	\arrow["{\varphi^2}", from=1-5, to=2-5]
    	\arrow["{\partial_I^2}", from=1-5, to=1-6]
    	\arrow["{\partial_J^2}"', from=2-5, to=2-6]
    	\arrow["{\psi^{-1}}"{description}, dashed, from=1-3, to=2-2]
    	\arrow["{\psi^0}"{description}, dashed, from=1-4, to=2-3]
    	\arrow["{\psi^1}"{description}, dashed, from=1-5, to=2-4]
    	\arrow["{\psi^2}"{description}, dashed, from=1-6, to=2-5]
    	\arrow["{\psi^{-2}}"{description}, dashed, from=1-2, to=2-1]
    \end{tikzcd}
\end{center}
As a warning, the dashed arrows do not commute with the other arrows.

We construct the $\psi^k$ to satisfy $\varphi^k=\psi^k\circ\del_I^k+\del_J^{k-1}\circ\psi^{k-1}$ by induction. We begin with $\psi^{-2}=0$ for free (it is the zero map), and our base case is $\psi^{-1}=0,$ which works because $0=0\circ\del_I^k+\del_J^{k-1}\circ0$ is a true statement. For the inductive step, we suppose we have $\varphi^k=\psi^k\circ\del_I^k+\del_J^{k-1}\circ\psi^{k-1},$ and we construct $\psi^{k+1}.$

For this, we use the long exact sequence to decompose $\del_I^{k+1}:I^{k+1}\to I^{k+2}$ into $I^{k+1}\to I^{k+1}/\op{im}\del_I^k\hookrightarrow I^{k+2}$ in the same lemma establishing the lifting. (Explicitly, $\del_I^{k+1}$ induces the isomorphism $I^{k+1}/\op{im}\del_I^k=I^{k+1}/\ker\del_I^{k+1}\cong\op{im}\del_I^{k+1}.$) Here are the morphisms we care about.
\begin{center}
    \begin{tikzcd}
    	& {I^k} & {I^{k+1}} & {I^{k+1}/\operatorname{im}\partial_I^k} & {I^{k+2}} \\
    	{J^{k-1}} & {J^k} & {J^{k+1}}
    	\arrow["{\partial_I^k}", from=1-2, to=1-3]
    	\arrow["d", from=1-3, to=1-4]
    	\arrow[hook, from=1-4, to=1-5]
    	\arrow["{\psi^{k-1}}"{description}, dashed, from=1-2, to=2-1]
    	\arrow["{\varphi^k}", from=1-2, to=2-2]
    	\arrow["{\partial_J^{k-1}}"', from=2-1, to=2-2]
    	\arrow["{\psi^k}"{description}, dashed, from=1-3, to=2-2]
    	\arrow["{\partial_J^k}"', from=2-2, to=2-3]
    	\arrow["s", dashed, from=1-4, to=2-3]
    	\arrow["{\varphi^{k+1}}", from=1-3, to=2-3]
    \end{tikzcd}
\end{center}
The hard part will be done attempting to induce $s:I^{k+1}/\op{im}\del_I^k$ such that
\[\varphi^{k+1}\stackrel?=s\circ d+\del_J^k\circ\psi^k.\]
Indeed, if we can induce such an $s:I^{k+1}/\op{im}\del_I^k,$ then we note $I^{k+1}/\op{im}\del_I^k\subseteq I^{k+2}$ means injecivity gives us the desired map $\psi^{k+1}:I^{k+2}\to J^{k+1}$ which restricts to $s$ on $I^{k+1}/\op{im}\del_I^k.$ Namely, we will still have
\[\varphi^{k+1}=\psi^{k+1}\circ\del_I^{k+1}+\del_J^k\circ\psi^k\]
because the image of $\del_I^{k+1}$ is the image of $d,$ on which $\psi^k$ restricts to $s.$

So we now attempt to construct $s.$ Rearranging, we see we need
\[s\circ d\stackrel?=\varphi^{k+1}-\del_J^k\circ\psi^k.\]
Because $d$ merely mods out by $\op{im}\del_I^k,$ it suffices to show that $\varphi^{k+1}-\del_J^k\circ\psi^k$ as a map $I^{k+1}\to J^{k+1}$ has $\op{im}\del_I^k$ in its kernel so that we can mod out the domain $I^{k+1}$ by $\op{im}\del_I^k$ safely. That is, we need
\[\left(\varphi^{k+1}-\del_J^k\circ\psi^k\right)\circ\del_I^k\]
is the $0$ mapping. Using the fact our moprhisms form a ring of composition over addition, we see that we need to show
\[\varphi^{k+1}\circ\del_I^k\stackrel?=\del_J^k\circ\psi^k\circ\del_I^k.\]
To finish, we expand out $\psi^k.$ Recall $\varphi^k=\psi^k\circ\del_I^k+\del_J^{k-1}\circ\psi^{k-1},$ so $\psi^k\circ\del_I^k=\varphi^k-\del_J^{k-1}\circ\psi^{k-1}.$ Substituting, we see
\[\del_J^k\circ\psi^k\circ\del_I^k=\del_J^k\left(\varphi^k-\del_J^{k-1}\circ\psi^{k-1}\right)=\del_J^k\circ\varphi^k-\del_J^k\circ\del_J^{k-1}\circ\psi^{k-1}.\]
Note $\del_J^k\circ\varphi^k=\varphi^{k+1}\circ\del_I^k$ by construction of the $\varphi^\bullet$ as a chain morphism, and this is our desired left-hand side. As for $\del_J^k\circ\del_J^{k-1}\circ\psi^{k-1},$ we see $\del_J^k\circ\del_J^{k-1}$ is the $0$ mapping by the exactness of our injective resolution, so it vanishes. This completes the justification of the existence of $s,$ and so the inductive step, and so the construction of the $\psi^\bullet.$

It remains to use the chain homotopy to show that the induced map on the cohomology groups is the zero mapping. We go ahead and apply the fixed-point functor everywhere, but all of our maps are preserved because they were $G$-module homomorphisms (and are now group homomorphisms). Recall that $H^k\left(\varphi^\bullet\right):H^k(G,M)\to H^k(G,N)$ was created by noting $\varphi^k$ restricts to a map
\[\varphi^k:\ker\del_I^k\to\ker\del_J^k\]
and then modding out the image by $\op{im}\del_J^{k-1}$ (and modding out the domain by $\op{im}\del_I^{k-1}$). Explicitly, we have that $H^k\left(\varphi^\bullet\right)\left(a+\op{im}\del_I^{k-1}\right):=\varphi(a)+\op{im}\del_J^{k-1}.$

So to show that our $H^k\left(\varphi^\bullet\right)$ is the $0$ map, it suffices to show that this restriction of $\varphi^k$ maps into $\op{im}\del_J^{k-1}.$ Well, pick up any $a\in\ker\del_I^k,$ and the chain homotopy implies
\[\varphi^k(a)=\left(\psi^k\circ\del_I^k\right)(a)+\left(\del_J^{k-1}\circ\psi^{k-1}\right)(a).\]
Because $a\in\ker\del_I^k,$ the first summand evaporates, and the second summand is $\del_J^{k-1}\left(\psi^{k-1}(a)\right)$ is in the image of $\del_J^{k-1}.$ So indeed, $\varphi^k$ does indeed take $\ker\del_I^k\to\op{im}\del_J^{k-1}.$ All of these maps are well-defined and work as discussed after applying the fixed-point functor as well. $\blacksquare$

As discussed preceding the above lemma, the lemma was enough to complete the proof. $\blacksquare$

In light of the previous lemma, we begin to notate the induced $H^\bullet\left(\varphi^\bullet\right):H(G,M)\to H(G,N)$ from the $G$-module homomorphism $\varphi:M\to N$ by just $H^\bullet(\varphi)$ because the exact choice of lift $\varphi^\bullet$ doesn't matter, established above.

Before continuing, we remark that we are beginning to get the feeling that $H^\bullet$ might actually be a functor from the category of $G$-modules to abelian groups. Indeed, it sends a $G$-module $M$ to $H^\bullet(G,M)$ and a morphism $\varphi$ to $H^\bullet(\varphi).$ To finish establishing this, we need to show the following.
\begin{lemma}
    Fix $G$ a finite group and $G$-modules $M,N,P$ with injective resolutions $I^\bullet,J^\bullet,K^\bullet.$ Additionally, suppose we have $G$-module homomorphisms $\varphi:M\to N$ and $\psi:N\to P.$ Then $H^\bullet(\psi\circ\varphi)=H^\bullet(\psi)\circ H^\bullet(\varphi).$
\end{lemma}
For this, we have to go back to the definition of $H^\bullet(\varphi)$ and $H^\bullet(\psi).$ Fix lifts of $\varphi$ to $\varphi^\bullet:I^\bullet\to J^\bullet$ and $\psi^\bullet:J^\bullet\to K^\bullet.$ For concreteness, here is our diagram.
\begin{center}
    \begin{tikzcd}
    	0 & M & {I^0} & {I^1} & {I^2} & \cdots \\
    	0 & N & {J^1} & {J^2} & {J^2} & \cdots \\
    	0 & P & {K^0} & {K^1} & {K^2} & \cdots
    	\arrow["\varphi", from=1-2, to=2-2]
    	\arrow["\psi", from=2-2, to=3-2]
    	\arrow[from=1-1, to=1-2]
    	\arrow[from=2-1, to=2-2]
    	\arrow[from=3-1, to=3-2]
    	\arrow[from=1-2, to=1-3]
    	\arrow[from=3-2, to=3-3]
    	\arrow["{\partial_I^0}"', from=1-3, to=1-4]
    	\arrow["{\partial_K^0}"', from=3-3, to=3-4]
    	\arrow["{\partial_I^1}"', from=1-4, to=1-5]
    	\arrow["{\partial_K^1}"', from=3-4, to=3-5]
    	\arrow[from=2-2, to=2-3]
    	\arrow["{\partial_J^0}"', from=2-3, to=2-4]
    	\arrow["{\partial_J^1}"', from=2-4, to=2-5]
    	\arrow["{\partial_I^2}"', from=1-5, to=1-6]
    	\arrow["{\partial_K^2}"', from=3-5, to=3-6]
    	\arrow["{\partial_J^2}"', from=2-5, to=2-6]
    	\arrow["{\varphi^0}", from=1-3, to=2-3]
    	\arrow["{\psi^0}", from=2-3, to=3-3]
    	\arrow["{\varphi^1}", from=1-4, to=2-4]
    	\arrow["{\psi^1}", from=2-4, to=3-4]
    	\arrow["{\varphi^2}", from=1-5, to=2-5]
    	\arrow["{\psi^2}", from=2-5, to=3-5]
    \end{tikzcd}
\end{center}
We recall that $H^\bullet(\varphi)$ was defined by applying the fixed-point functor and then restricting $\varphi^k$ to $\ker\del_I^k\to\ker\del_J^k$ followed by modding out the codomain by $\op{im}\del_J^{k-1}$ and then the domain by $\op{im}\del_I^{k-1}.$ Of course, $\psi^k$ has a similar story.

Thus, we note that we have the defining equations (here $\del^{-1}_\bullet$ is the zero map!)
\[\begin{cases}
    H^\bullet(\varphi)\left(a+\op{im}\del_I^k\right):=\varphi(a)+\op{im}\del_J^k, \\
    H^\bullet(\psi)\left(b+\op{im}\del_J^k\right):=\psi(b)+\op{im}\del_K^k, \\
    H^\bullet(\psi\circ\varphi)\left(a+\op{im}\del_I^k\right):=(\varphi\circ\psi)(b)+\op{im}\del_K^k.
\end{cases}\]
From these definitions we see that, indeed, $H^\bullet(\psi)\circ H^\bullet(\varphi)$ and $H^\bullet(\psi\circ\varphi)$ do the same thing to $a+\op{im}\del_I^k,$ so we are done here. $\blacksquare$

We are now ready to show that our cohomology groups are well-defined. The intuition at this point is that our induced map of cohomology groups by a morphism $\varphi$ is ``canonical,'' and this we expect to be enough to show that this map is an isomorphism in special cases. Here is the conversion.
\begin{proposition}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, suppose we have $G$-module homomorphisms $\varphi:M\to N$ and $\psi:N\to M$ such that $\psi\circ\varphi=\op{id}_M.$ Then the induced maps $H^\bullet(\varphi):H(G,M)\to H(G,N)$ and $H^\bullet(\psi):H(G,N)\to H(G,M)$ have $H^\bullet(\psi)\circ H^\bullet(\varphi)=\op{id}_{H^\bullet(G,M)}$ as well.
\end{proposition}
This is reasonably slick. Set up the chain morphisms on top of each other.
\begin{center}
    \begin{tikzcd}
    	0 & M & {I^0} & {I^1} & {I^2} & \cdots \\
    	0 & N & {J^1} & {J^2} & {J^2} & \cdots \\
    	0 & M & {I^0} & {I^1} & {I^2} & \cdots
    	\arrow["\varphi", from=1-2, to=2-2]
    	\arrow["\psi", from=2-2, to=3-2]
    	\arrow[from=1-1, to=1-2]
    	\arrow[from=2-1, to=2-2]
    	\arrow[from=3-1, to=3-2]
    	\arrow[from=1-2, to=1-3]
    	\arrow[from=3-2, to=3-3]
    	\arrow["{\partial_I^0}"', from=1-3, to=1-4]
    	\arrow["{\partial_I^0}"', from=3-3, to=3-4]
    	\arrow["{\partial_I^1}"', from=1-4, to=1-5]
    	\arrow["{\partial_I^1}"', from=3-4, to=3-5]
    	\arrow[from=2-2, to=2-3]
    	\arrow["{\partial_J^0}"', from=2-3, to=2-4]
    	\arrow["{\partial_J^1}"', from=2-4, to=2-5]
    	\arrow["{\partial_I^2}"', from=1-5, to=1-6]
    	\arrow["{\partial_I^2}"', from=3-5, to=3-6]
    	\arrow["{\partial_J^2}"', from=2-5, to=2-6]
    	\arrow["{\varphi^0}", from=1-3, to=2-3]
    	\arrow["{\psi^0}", from=2-3, to=3-3]
    	\arrow["{\varphi^1}", from=1-4, to=2-4]
    	\arrow["{\psi^1}", from=2-4, to=3-4]
    	\arrow["{\varphi^2}", from=1-5, to=2-5]
    	\arrow["{\psi^2}", from=2-5, to=3-5]
    \end{tikzcd}
\end{center}
By construction of the $\varphi^\bullet$ and $\psi^\bullet,$ all of the square commute, so all of the $2\times1$ rectangles. What we see, them, is that $\psi^\bullet\circ\varphi^\bullet$ lifts $\psi\circ\varphi=\op{id}_M:M\to M$ as a chain morphism.

The kicker is that $\op{id}^\bullet:I^k\to I^k$ also lifts $\op{id}_M:M\to M,$ but we know that the induced map $H^\bullet(\op{id}_M)$ takes the coset $a+\op{im}\del_I^\bullet$ directly to $a+\op{im}\del_I^\bullet,$ which is $\op{id}_{H^\bullet(G,M)}.$ Because $\psi^\bullet\circ\varphi^\bullet$ and $\op{id}^\bullet$ both lift $\op{id},$ we conclude
\[H^\bullet(\psi)\circ H^\bullet(\varphi)=H^\bullet(\psi\circ\varphi)=H^\bullet(\op{id})=\op{id}_{H^\bullet(G,M)}\]
using that funtoriality of $H^\bullet$ we just so happened to need. This finishes. $\blacksquare$

And here is the main attraction.
\begin{theorem}
    Fix $G$ a finite group and $G$-modules $M$ and $N$ with injective resolutions $I^\bullet$ and $J^\bullet$ respectively. Additionally, $M\cong N$ are $G$-modules. Then $H^\bullet(G,M)\cong H^\bullet(G,N)$ canonically.
\end{theorem}
Note that $M\cong N$ promises us maps $\varphi:M\to N$ and $\psi:N\to M$ such that $\psi\circ\varphi=\op{id}_M$ and $\varphi\circ\psi=\op{id}_N.$ Then the previous proposition implies
\[H^\bullet(\psi)\circ H^\bullet(\varphi)=\op{id}_{H^\bullet(G,M)},\quad\text{and}\quad H^\bullet(\varphi)\circ H^\bullet(\psi)=\op{id}_{H^\bullet(G,N)}.\]
Thus, $H^\bullet(\varphi)$ and $H^\bullet(\psi)$ witness $H^\bullet(G,M)\cong H^\bullet(G,N),$ and we are done. $\blacksquare$

In particular, if $M=N,$ then $\op{id}_M$ witnesses our isomorphism $M\cong M,$ so it doesn't matter what choice of injective resolutions we have, our group cohomology is always the same. This is perhaps remarkable because we can reliably make an injective module bigger, so the injective resolution is potentially very large. However, the moderating factor is that the definition of cohomology is modding in a long exact sequence, and these mods accommodate for steep size increases.

\subsubsection{June 18th}
Today I learned an algebraic proof of the Cayley-Hamilton theorem. Here is statement.
\begin{theorem}
    Fix $R$ a commutative ring and $M=(m_{k,\ell})_{k,\ell=1}^n$ of a $n\times n$ matrix. Define the characteristic polynomial $p_M(x)\in R[x]$ by $x\mapsto\det(xI_n-M).$ Then plugging in $p_M(M)$ gives the $0$ matrix.
\end{theorem}
The main idea in this proof is to coerce $N$ to live on the same level as normal elements of $R.$ The way we do this is by bringing $N$ down to an endomorphism of $R^n$ and bringing elements of $R$ up to endomorphisms of $R,$ and when everything is an endomorphism, we can interact with them easier.

With this in mind, we note the following bijection does this translation for us.
\begin{lemma}
    Fix $R$ a commutative ring and $n$ a positive integer. Endomorphisms $\mu:R^n\to R^n$ are in bijection with matrices $M=(m_{k,\ell})_{k,\ell=1}^n.$
\end{lemma}
We define the bijection by picking up a matrix $M=(m_{k,\ell})_{k,\ell=1}^n$ and defining $\varphi_M(v):=Mv$ for any vector $v\in R^n.$ To be explicit, this extends linearly to the mapping $\varphi:R^n\to R^n$ which looks like
\[\varphi_M\left(\begin{bmatrix}r_1 \\ \vdots \\ r_n\end{bmatrix}\right):=
\begin{bmatrix}
    m_{1,1} & \cdots & m_{1,n} \\
    \vdots & \ddots & \vdots \\
    m_{n,1} & \cdots & m_{n,n}
\end{bmatrix}\begin{bmatrix}r_1 \\ \vdots \\ r_n\end{bmatrix}=
\begin{bmatrix}
    m_{1,1}r_1 + \cdots + m_{1,n}r_n \\
    \vdots \\
    m_{n,1}r_1 + \cdots + m_{n,n}r_n
\end{bmatrix}.\]
For $a,b\in R$ and vectors $v,w\in R^n,$ we do have
\[\varphi_M(av+bw)=M(av+bw)=M(av)+M(bw)=aMv+bMw=a\varphi_M(v)+b\varphi_M(w),\]
so indeed, $\varphi_M\in\op{End}_R\left(R^n\right).$ This follows because of the linearity of the matrix $M,$ which we will not check explicitly here because it is the same proof verbatim as for vector spaces and merely amounts to writing out the arithmetic.

It remains to show that the correspondence $\varphi_\bullet$ is injective and surjective. For injectivity, fix $M=(m_{k,\ell})_{k,\ell=1}^n$ and $N=(n_{k,\ell})_{k,\ell=1}^n$ giving $\varphi_M=\varphi_N.$ Then, for any basis vector $e_\ell,$ we see
\[\begin{bmatrix}m_{1,\ell} \\ \vdots \\ m_{n,\ell}\end{bmatrix}=Me_\ell=\varphi_M(e_\ell)=\varphi_N(e_\ell)=Ne_\ell=\begin{bmatrix}n_{1,\ell} \\ \vdots \\ n_{n,\ell}\end{bmatrix}.\]
Thus, for any $k$ and $\ell,$ we see $m_{k,\ell}=n_{k,\ell}$ by checking the $k$th entry in the above computation. So we do have $M=N,$ establishing injectivity.

Lastly, we show surjectivity. Fix $\varphi:R^n\to R^n$ any $R$-linear endomorphism, and we define the entries $(m_{k,\ell})_{k,\ell=1}^n$ by
\[\begin{bmatrix}m_{1,\ell} \\ \vdots \\ m_{n,\ell}\end{bmatrix}:=\varphi(e_\ell)\]
for any basis vector $e_\ell.$ These entries do cover all cases $1\le k,\ell\le n,$ so they give us a matrix $M\in R^{n\times n}.$ It remains to show $\varphi(v)=Mv$ for all vectors $v$ so that $\varphi=\varphi_M.$ Well, any $v\in R^n$ can be written as
\[v=\begin{bmatrix}r_1 \\ \vdots \\ r_n\end{bmatrix}=\sum_{\ell=1}^nr_\ell e_\ell.\]
Thus, the $R$-linearity of $\varphi$ tells us
\[\varphi(v)=\sum_{\ell=1}^nr_\ell\varphi(e_\ell),\]
Now, $\varphi(e_\ell)=Me_\ell$ by our row construction, so we get to translate as
\[\varphi(v)=\sum_{\ell=1}^nr_\ell Me_\ell=M\left(\sum_{\ell=1}^nr_\ell e_\ell\right)=Mv\]
using the linearity of matrix-vector multiplication. This completes the verification that $\varphi_\bullet$ is surjective and thus the proof. $\blacksquare$

With the above correspondence in mind, we let $\varphi\in\op{End}_R\left(R^n\right)$ correspond to the matrix $M.$ As promised, next we want lift the elements of $R$ to endomorphisms. Well, we have the $R$-linear endomorphism $(r,v)\mapsto rv$ as just scalar-vector multiplication, and this will do.

To avoid abuse of notation, we let $r\in R$ be the element and $\mu_r$ be the endomorphism. We have the following general statement about $\mu_\bullet.$
\begin{lemma}
    Fix $R$ a commutative ring and $M$ an $R$-module. Then multiplication by $r\in R$ is an $R$-linear endomorphism, and the corresponding map $r\mapsto\mu_r$ is a homomorphism of commutative rings.
\end{lemma}
The idea is that this follows from the ring action of $R$ on $M.$ We begin by showing that $\mu_\bullet(R)$ is itself a commutative subring of $\op{End}_R(M).$ Here, addition is pointwise and multiplication is composition.

To start, we have to actually show $\mu_r\in\op{End}_R(M)$ for any $r\in R.$ Well, it suffices to pick up any $r_1,r_2\in R$ and $m_1,m_2\in M$ and note that
\[\mu_r(r_1m_1+r_2m_2)=r(r_1m_1+r_2m_2)=rr_1m_1+rr_2m_2=r_1rm_1+r_2rm_2=r_1\mu_r(m_1)+r_2\mu_r(m_2)\]
by repeatedly using properties of the ring action. Next we show closure of $\mu_\bullet(R)$ under $+$ and $\circ.$ For any two $\mu_r,\mu_s\in\mu_\bullet(R)$ and $m\in M,$ we note
\[(\mu_r+\mu_s)(m)=\mu_r(m)+\mu_s(m)=rm+sm=(r+s)m=\mu_{r+s}(m),\]
so we get closure under addition and a homomorphism of abelian groups both at once. Continuing,
\[(\mu_r\circ\mu_s)(m)=\mu_r\big(\mu_s(m)\big)=\mu_r(sm)=r(sm)=(rs)m=\mu_{rs}(m),\]
so we get closuder under $\circ$ and a homomorphism of rings. As for commutativity, we note
\[(\mu_r\circ\mu_s)=(rs)m=(sr)m=(\mu_s\circ\mu_s)(m),\]
which finishes showing that this is a commutative ring. Thus, we are done. $\blacksquare$

The point of doing that is to make the lemma less painful to prove.
\begin{lemma}
    Fix a commutative ring $R$ and $\varphi\in\op{End}_R\left(R^n\right).$ The ring $\mu_\bullet(R)[\varphi]$ is a commutative subring of $\op{End}\left(R^n\right).$
\end{lemma}
Addition is done pointwise as with morphisms as usual, and multiplication is composition. To show that this is a subring, we starting by noting that any element $\sum_{k=1}^m\mu_{r_k}\varphi^k$ is still an endomorphism because $\op{End}_R\left(R^n\right)$ is itself a ring.

To finish showing this is a subring, we have to show that $\mu_\bullet(R)[\varphi]$ is closed under these operations, and indeed, this is essentially a polynomial ring in $\varphi$ with coefficients in $\mu_\bullet.$ So the polynomial addition
\[\sum_{k=1}^\infty\mu_{r_k}\varphi^k+\sum_{k=1}^\infty\mu_{s_k}\varphi^k=\sum_{k=1}^\infty\underbrace{(\mu_{r_k}+\mu_{s_k})}_{\in\mu_\bullet(R)}\varphi^k\]
verifies closure under addition, where all but finitely many of the $\mu_{r_\bullet}$ and $\mu_{s_\bullet}$ are zero. For closure under multiplication, we do a similar computation to find
\[\left(\sum_{k=1}^\infty\mu_{r_k}\varphi^k\right)\left(\sum_{\ell=1}^\infty\mu_{s_\ell}\varphi^\ell\right)=\sum_{m=1}^\infty\underbrace{\left(\sum_{k+\ell=m}\mu_{r_k}\mu_{s_\ell}\right)}_{\in\mu_\bullet(R)}\varphi^m\]
verifies closure under multiplication, where all but finitely many of the $\mu_{r_\bullet}$ and $\mu_{s_\bullet}$ are zero.

Finally, the commutativity of multiplication comes from writing
\[\sum_{m=1}^\infty\left(\sum_{k+\ell=m}\mu_{r_k}\mu_{s_\ell}\right)\varphi^m=\sum_{m=1}^\infty\left(\sum_{k+\ell=m}\mu_{s_\ell}\mu_{r_k}\right)\varphi^m,\]
which holds because of the commutativity of multiplication in $R.$ Thus,
\[\left(\sum_{k=1}^\infty\mu_{r_k}\varphi^k\right)\left(\sum_{\ell=1}^\infty\mu_{s_\ell}\varphi^\ell\right)=\left(\sum_{\ell=1}^\infty\mu_{s_\ell}\varphi^\ell\right)\left(\sum_{k=1}^\infty\mu_{r_k}\varphi^k\right).\]
This completes the proof. $\blacksquare$

Our next goal will be to create an actual matrix which can interpret $p_M(M)$ more like a determinant, so we aim to create something like $M\cdot I_n-M.$ With our correspondence, we create $\varphi I_n-M.$ So we let $\Delta$ have coefficients
\[(\Delta)_{k,\ell=1}^n=\varphi1_{k=\ell}-\mu_{m_{k,\ell}},\]
where everything here is an element of $\op{End}_R\left(R^n\right).$ Namely, $1_{k=\ell}$ is the identity morphism when $k=\ell$ and the zero morphism otherwise so that $I_n:=(1_{k=\ell})_{k,\ell=1}^n\in\left(\mu_\bullet(R)[\varphi]\right)^{n\times n}.$ Similarly, we let $\mu_\bullet(M)\in\left(\mu_\bullet(R)[\varphi]\right)^{n\times n}$ by the matrix with entries $(\mu_{m_{k,\ell}})_{k,\ell=1}^n.$ So the above definition translate into
\[\Delta=\varphi I_n-\mu_\bullet(M)\in\left(\mu_\bullet(R)[\varphi]\right)^{n\times n},\]
which is close enough to what we want $M\cdot I_n-M.$

It remains to compute the determinant. The key translator is note we have a ring homomorphism $R[x]\to\mu_\bullet(R)[\varphi)$ defined by taking $r\mapsto\mu_r$ and $x\mapsto\varphi.$ Indeed, we already showed $R\to\mu_\bullet(R)$ is a ring homomorphism, so $R[x]\to\mu_\bullet(R)[x]$ is a ring homomorphism, and we get to $\mu_\bullet(R)[\varphi]$ by evaluating.

The reason this is helpful is because we can apply this homomorphism before or after applying $\det$ to the matrix $xI_n-M$ defining $p_M(x).$ Because we have a homomorphism of rings, and determinants only require a large number of additions and multiplications to evaluate, we will get the same answer out. More explicitly,
\[\det\left(\begin{bmatrix}
    x-m_{1,1} & -m_{1,2} & \cdots & -m_{1,n} \\
    -m_{2,1} & x-m_{2,2} & \cdots & -m_{2,n} \\
    \vdots & \vdots & \ddots & \vdots \\
    -m_{n,1} & -m_{n,2} & \cdots & x-m_{n,n}
\end{bmatrix}\right)\mapsto\det\left(\begin{bmatrix}
    \varphi-\mu_{m_{1,1}} & -\mu_{m_{1,2}} & \cdots & -\mu_{m_{1,n}} \\
    -\mu_{m_{2,1}} & \varphi-\mu_{m_{2,2}} & \cdots & -\mu_{m_{2,n}} \\
    \vdots & \vdots & \ddots & \vdots \\
    -\mu_{m_{n,1}} & -\mu_{m_{n,2}} & \cdots & \varphi-\mu_{m_{n,n}}
\end{bmatrix}\right),\]
and we see the right-hand side is $\det\Delta.$ The point is that $\mu_\bullet(p_M)(\varphi)=\det\Delta$ (here $\mu_\bullet(p_M)$ is an abuse of notation with $\mu_\bullet:x\mapsto x$), and it will suffice to show $\det\Delta=0\in\mu_\bullet(R)[\varphi].$ The punchline is the following lemma.
\begin{lemma}
    Fix everything as above. Then $\det\Delta=0$ as an element of $\mu_\bullet(R)[\varphi]\subseteq\op{End}_R\left(R^n\right)$ if and only if $p_M(M)=0$ as an $n\times n$ matrix.
\end{lemma}
Working backwards a bit, we note that $p_M(M)$ is the $0$ matrix if and only if $p_M(M)$ evaluates to $0$ on every vector, for this would correspond to the zero endomorphism, and the bijection between matrices and endomorphisms would force $p_M(M)=0.$

Anyways, suppose we can write out the characteristic polynomial as
\[p_M(x)=\sum_{k=0}^na_kx^k\]
Then plugging in $M$ gives
\[p_M(M)=\sum_{k=0}^na_kM^k.\]
Plugging in an arbitrary vector $v,$ we use the fact $\varphi(v)=Mv$ to note
\[p_M(M)(v)=\left(\sum_{k=0}^na_kM^k\right)v=\sum_{k=0}^n\left(a_kM^k\right)(v)=\sum_{k=0}^n\left(\mu_{a_k}\varphi^k\right)v\]
by making everything into an endomorphism. The point is that the right-hand side is $p_M(x)$ under the homomorphism $R[x]\to\mu_\bullet(R)[\varphi]$ discussed above, which we called $\mu_\bullet(p_M)(\varphi)$ by abuse of notation.

Continuing, the corresponding movement of matrices $xI_n-M\mapsto\varphi I_n-\mu_\bullet(M)=\Delta$ we found in the discussion gave $\mu_\bullet(p_M)(\varphi)=\det\Delta.$ So $p_M(M)(v)=0$ for arbitrary vectors $v$ if and only if $(\det\Delta)(v)=0$ for arbitrary vectors $v,$ which is equivalent to $\det\Delta$ being the $0$ endomorphism. $\blacksquare$

It remains to compute $\det\Delta.$ We start by picking up some evidence that this should be $0.$ Well, pick up any basis vector $e_\ell\in R^n,$ and we see
\[\varphi e_\ell=Me_\ell=\begin{bmatrix}m_{1,\ell} \\ \vdots \\ m_{n,\ell}\end{bmatrix}=\sum_{k=1}^nm_{k,\ell}e_k.\]
Attempting to move everything to one side, we write $m_{k,\ell}e_k=\mu_{m_{k,\ell}}e_k$ and $\varphi e_\ell=\sum_{k=1}^n\varphi1_{k=\ell}e_k$ so that this rearranges to
\[\sum_{k=1}^n(\varphi1_{k=\ell}-\mu_{m_{k,\ell}})e_k=0.\]
The inner entry is $\Delta_{k,\ell},$ so our evidence that $\det\Delta=0$ will really be
\[\sum_{k=1}^n\Delta_{k,\ell}e_k=0\]
for any $\ell.$ Of course, it's not yet obvious how to convert this, but this is really the only relation we have, so we hope that it is enough. To finish the conversion, we introduce the adjugate matrix.
\begin{definition}
    Fix $R$ a commutative ring and $M\in R^{n\times n}$ a matrix. We define the adjugate matrix $\op{adj}M$ as the transpose of the cofactor matrix. To be explicit,
    \[\op{adj}M:=\begin{bmatrix}
        (-1)^{1+1}\overline M_{1,1} & \cdots & (-1)^{1+n}\overline M_{1,n} \\
        \vdots & \ddots & \vdots \\
        (-1)^{n+1}\overline M_{n,1} & \cdots & (-1)^{n+n}\overline M_{n,n}
    \end{bmatrix}^\intercal,\]
    where $\overline M_{k,\ell}$ is the determinant of the matrix $M$ with the $k$th row and $\ell$th column removed.
\end{definition}
The reason we care about the adjugate matrix is the following property.
\begin{proposition}
    Fix $R$ a commutatieve ring and $M=(m_{k,\ell})_{k,\ell=1}^n\in R^{n\times n}$ a matrix. Then we have
    \[M\cdot(\op{adj}M)=\det M\cdot I_n.\]
\end{proposition}
This more or less reduces to computing the determinant by expansion by minors. Fixing indices $j$ and $\ell,$ we see that
\[(M\cdot\op{adj}M)_{j,\ell}=\sum_{k=1}^nm_{j,k}(\op{adj}M)_{k,\ell}=\sum_{k=1}^n(-1)^{k+\ell}m_{j,k}\cdot\overline M_{\ell,k}.\]
Note $(\op{adj}M)_{\ell,k}$ became $(-1)^{\ell+k}\overline M_{\ell,k}$ because $\op{adj}M$ is the transpose of the cofactor matrix. Anyways, this sum is exactly computing the determinant of
\[\begin{bmatrix}
    m_{1,1} & m_{1,2} & \cdots & m_{1,n} \\
    \vdots & \vdots & \ddots & \vdots \\
    m_{\ell-1,1} & m_{\ell-1,2} & \cdots & m_{\ell-1,n} \\
    m_{j,1} & m_{j,2} & \cdots & m_{j,n} \\
    m_{\ell+1,1} & m_{\ell+1,2} & \cdots & m_{\ell+1,n} \\
    \vdots & \vdots & \ddots & \vdots \\
    m_{n,1} & m_{n,2} & \cdots & m_{n,n}
\end{bmatrix}\]
by expanding along the $\ell$th row, a rule which works in arbitrary commutative rings roughly because it is a purely symbolic fact. Formally, we cold define this matrix with indices and whatnot, but we aren't going to be rigorous about the expansion by minors either, so we won't bother.

If $j\ne\ell,$ then two rows are equal, so the entire determinant vanishes; this is also true in arbitrary commutative rings again roughly because it is a purely symbolic fact. When $j=k,$ this is the matrix $M$ exactly, so we get $\det M.$ Thus,
\[(M\cdot\op{adj}M)_{j,\ell}=\det M\cdot1_{j=\ell}.\]
Removing the indices, this gives $M\cdot\op{adj}M=\det M\cdot I_n,$ which is what we wanted. $\blacksquare$

From here, converting our evidence to proof that $\det\Delta=0$ is not terribly hard. Observing that our evidence already almost looks the start of a matrix multiplication, we fix an index $j$ and sum over $\ell$ so that
\[0=\sum_{\ell=1}^n\left((\op{adj}\Delta)_{\ell,j}\sum_{k=1}^n\Delta_{k,\ell}e_k\right)=\sum_{k=1}^n\left(\sum_{\ell=1}^n\Delta_{k,\ell}(\op{adj}\Delta)_{\ell,j}\right)e_k.\]
Now the inner sum is $(\Delta\cdot\op{adj}\Delta)_{k,j}=\det\Delta1_{k=j}$ from the above proposition. Thus, summing over $k,$ all terms vanish except for $(\det\Delta)e_j=0.$ Because $j$ was arbitrary, we see that $\det\Delta$ vanishes on all basis vectors.

To turn this into having $\det\Delta$ vanishing on all vectors, fix any vector $v=(v_1,\ldots,v_n)=\sum_{k=1}^nv_ke_k\in R^n$ so that
\[(\det\Delta)v=(\det\Delta)\left(\sum_{k=1}^nv_ke_k\right)=\sum_{k=1}^nv_k(\det\Delta)(e_k)=0,\]
so indeed, $\det\Delta$ is the zero endomorphism. As promised, this completes the proof. $\blacksquare$

As a final remark, I do know another proof of the Cayley-Hamilton theorem by arguing in $\CC$ and then moving into arbitrary commutative rings, but I like this proof due to its directness and almost blatantness. It is not correct to directly evaluate $p_M(M)$ by writing $\det(MI_n-M)$ because we need to evaluate $MI_n$ as a matrix with matrix coefficients. Thus, the object we really need is a matrix with endomorphism entries, and the rest is conversion.

\subsubsection{June 19th}
Today I learned a more thorough version of the Snake Lemma to set up for the long exact sequence of cohomology. The following is the main attraction.
\begin{theorem}
    Fix $G$ a finite group (or, more generally, a commutative ring), and suppose we have the following commutative diagram with exact rows.
    \begin{center}
        \begin{tikzcd}
        	& {M_0} & {M_1} & {M_2} & 0 \\
        	0 & {N_0} & {N_1} & {N_2}
        	\arrow["{\partial_M^0}", from=1-2, to=1-3]
        	\arrow["{\partial_M^1}", two heads, from=1-3, to=1-4]
        	\arrow[from=1-4, to=1-5]
        	\arrow[from=2-1, to=2-2]
        	\arrow["{\partial_N^0}"', hook, from=2-2, to=2-3]
        	\arrow["{\partial_N^1}"', from=2-3, to=2-4]
        	\arrow["{\varphi_0}", from=1-2, to=2-2]
        	\arrow["{\varphi_1}", from=1-3, to=2-3]
        	\arrow["{\varphi_2}", from=1-4, to=2-4]
        \end{tikzcd}
    \end{center}
    Then we have an exact sequence
    \[\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2\stackrel\delta\to\op{coker}\varphi_0\to\op{coker}\varphi_1\to\op{coker}\varphi_2,\]
    where all maps are canonical.
\end{theorem}
The maps on the left and right of the sequence are fairly standard. We cover these quickly. Here are the $\ker$ maps.
\begin{lemma}
    Fix everything as above. We can exhibit canonical maps so that the sequence $\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2$ is exact.
\end{lemma}
We begin by defining our maps.
\begin{itemize}
    \item The map $\ker\varphi_0\to\ker\varphi_1$ is $\del_M^0$ on restriction. Indeed, any $a\in\ker\varphi_0$ has $\del_M^0\big(\varphi_0(a)\big)=0,$ so $\varphi_1\big(\del_M^0(a)\big)=0$ because of the square. This gives $\del_M^0(a)\in\ker\varphi_1.$
    \item The map $\ker\varphi_1\to\ker\varphi_1$ is $\del_M^1$ on restriction (again). To repeat, any $a\in\ker\varphi_1$ has $\varphi_1\big(\del_M^0(a)\big)=\del_M^0\big(\varphi_0(a)\big)=0,$ because of the square.
\end{itemize}
To check the exactness on the left, we need to know $\del_M^0:\ker\varphi_0\to\ker\varphi_1$ surjects onto the kernel of $\del_M^1\ker\varphi_1\to\ker\varphi_2.$ Well, the exactness of the original diagram means
\[\ker\del_M^1=\op{im}\del_M^1\]
without the restriction. Beginning to introduce the restriction, we note $m_1\in\del_M^0(\ker\varphi_0)$ gives $m_1\in\op{im}\del_M^0=\ker\del_M^1$ for free, so $m_1\in\ker\del_M^1\cap\ker\varphi_1$ from the work above. In the other direction, if $m_1\in\ker\del_M^1\cap\ker\varphi_1=\op{im}\del_M^0\cap\ker\varphi_1,$ we are granted $m_1$ such that
\[\del_M^0(m_0)=m_1.\]
Using the commuting square, $(\del_N^0\circ\varphi_0)(m_0)=(\varphi_1\circ\del_M^0)(m_0)=\varphi_1(m_1)=0.$ The key step, now, is that $\del_N^1$ is injective in the diagram, so we must have $\varphi_0(m_0)=0,$ implying that, indeed, $m_1\in\del_M^0(\ker\varphi_0).$ Thus, the left side is exact. $\blacksquare$

And here are the $\op{coker}$ maps.
\begin{lemma}
    Fix everything as above. We can exhibit canonical maps so that $\op{coker}\varphi_0\to\op{coker}\varphi_1\to\op{coker}\varphi_2$ is exact.
\end{lemma}
We begin by defining our maps.
\begin{itemize}
    \item The map $\op{coker}\varphi_0\to\op{coker}\varphi_1$ is induced by $\del_N^0.$ Indeed, $\del_N^0$ takes $N_0\to N_1,$ so we can mod out the image by $\op{im}\varphi_1$ to map into $\op{coker}\varphi_1.$ Now, any $n_0\in\op{im}\varphi_0$ promises $n_0=\varphi_0(m_0)$ so that $(\del_N^0\circ\varphi_0)(m_0)=(\varphi_1\circ\del_M^0)(m_0)\in\op{im}\varphi_1.$ That is, $\del_N^0:\op{im}\varphi_0\to\op{im}\varphi_1,$ meaning we may safely mod the domain by $\op{im}\varphi_0$ to induce $\op{coker}\varphi_0\to\op{coker}\varphi_1.$
    \item The map $\op{coker}\varphi_0\to\op{coker}\varphi_1$ is induced by $\del_N^0$ in the same way. Again, $\del_N^1\circ\varphi_1=\varphi_2\circ\del_M^1$ implies $\del_N^1:\op{im}\varphi_1\to\op{im}\varphi_2,$ so we safely take $\del_N^1:N_1\to N_2$ and mod out the image by $\op{im}\varphi_2$ and then the domain by $\op{im}\varphi_1.$
\end{itemize}
We now check the exactness on the right. Again, we start off by knowing that
\[\op{im}\del_N^0=\ker\del_N^1,\]
so we just have to show that this is maintained after we've done our modding. On one hand, any coset $n_1+\op{im}\varphi_1\in\op{im}\del_N^0$ lands in $\del_N^1(n_1)+\op{im}\varphi_2=0+\op{im}\varphi_2$ after applying $\del_N^1,$ for the bottom row of the diagram is exact.

On the other hand, pick up any coset $n_1+\op{im}\varphi_1\in\ker\del_N^1,$ which we need to show lives in the image $\del_N^0(\op{coker}\varphi_0).$ We have $\del_N^1(n_1)\in\op{im}\varphi_2,$ so we are promised $m_2$ such that $\del_N^1(n_1)=\varphi_2(m_2).$ But $\del_M^1$ is surjective, so we also have $m_1$ such that $\del_N^1(n_1)=(\varphi_2\circ\del_M^1)(m_1).$ But using the square, we see
\[\del_N^1(n_1)=\del_N^1\big(\varphi_1(m_1)\big),\]
so $n_1-\varphi_1(m_1)\in\ker\del_N^1=\op{im}\del_N^0.$ So we are promised $n_0\in N_0$ such that $\del_N^0(n_0)\in n_1+\op{im}\varphi_1,$ which is what we needed to finish exactness. $\blacksquare$

It remains to talk about $\delta:\ker\varphi_2\to\op{coker}\varphi_0.$ The idea, now, is to pick up an element $m_2\in\ker\varphi_2$ and push it backwards along the diagram to reach $N_0.$ Note we cannot go left twice consecutively in a meaningful way because $\del_\bullet^1\circ\del_\bullet^0=0,$ so we travel $M_2\leftarrow M_1\to N_1\leftarrow N_0.$

Indeed, $\del_M^1$ is surjective, so we are promised an $m_1\in M_1$ such that $\del_M^1(m_1)=m_2.$ This choice is not canonical, but we do have the isomorphism
\[\frac{M_1}{\ker\del_M^1}\cong\del_M^1(M_1)=M_2,\]
so the choice of $m_1$ is unique up to the (now canonical) coset $m_1+\ker\del_M^1.$ By exactness, we are granted the canonical coset $m_1+\op{im}\del_M^0.$

Continuing, we go through $\varphi_1.$ Note that $\varphi_2(m_2)=0$ means $(\del_N^1\circ\varphi_1)(m_1)=(\varphi_2\circ\del_M^1)(m_1)=0,$ so $\varphi_1(m_1)\in\ker\del_N^1.$ By exactness, we also know $\varphi_1(m_1)\in\op{im}\del_N^0,$ so indeed, we are promised $n_0$ such that $\del_N^0(n_0)=\varphi_1(m_1).$ Note this $n_0$ is unique by the injectivity of $\del_N^0.$

The hope is that, with $m_1$ unique up to choice of coset $m_1+\op{im}\del_M^0,$ we have $n_0$ unique up to $n_0+\op{im}\varphi_0.$ So suppose $m_1'\in m_1+\op{im}\del_M^0$ and pick up $n_0'=(\del_N)^{-1}(\varphi_1(m_1')).$ The point is that
\[\del_N^0(n_0-n_0')=\varphi_1(m_1-m_1')\in\op{im}\varphi_1\circ\del_M^0=\op{im}\del_N^0\circ\varphi_0.\]
By injectivity, we conclude $n_0-n_0'\in\op{im}\varphi_0$ are indeed unique up to coset of $\op{im}\varphi_0.$ This completes the construction of $\delta.$ To review, our map is
\[\left(\del_N^0\right)^{-1}\circ\varphi_1\circ\left(\del_M^1\right)^{-1},\]
where we showed the answer is well-defined up to choice of coset $\op{im}\varphi_0$ at the very end.

Now we have to talk about exactness. As a warning, this is about to get messy. Here is the left.
\begin{lemma}
    Fix everything as above. Then $\ker\varphi_1\to\ker\varphi_1\stackrel\delta\to\op{coker}\varphi_0$ is exact.
\end{lemma}
We mostly focus on the kernel of $\delta.$ Elements in the kernel of $\delta$ are those that have made their way into $\op{im}\varphi_0$ by the end, so we manually track $\op{im}\varphi_0$ backwards along
\[N_0\stackrel{\del_N^0}\to N_1\stackrel{\varphi_1}\leftarrow M_1\stackrel{\del_M^1}\to M_0.\]
We start with $\op{im}\varphi_0\subseteq N_0.$ Moving along $\del_N^0,$ we have $\del_N^0(\op{im}\varphi_0)=\op{im}\del_N^0\circ\varphi_0.$ Using the commutative diagram, this is $\op{im}\varphi_1\circ\del_M^0=\varphi_1(\op{im}\del_M^0),$ and using the exactness of the top row, this is $\varphi_1(\ker\del_M^1).$

Now, we need to move $\varphi_1(\ker\del_M^1)\subseteq M_1$ backwards along $\varphi_1.$ Every single element $\varphi_1(m_1)\in\varphi_1(\ker\del_M^1)$ is going to induce an entire fiber $m_1+\ker\varphi_0$ back in $M_1,$ and unioning over all fibers gives us
\[\bigcup_{m_1\in\ker\del_M^1}m_1+\ker\varphi_0=\left\{m_1+m_1':m_1\in\ker\del_M^1\text{ and }m_1'\in\ker\varphi_0\right\}.\]
However, when we push this through $\del_M^1,$ every single pair sum $m_1+m_1'$ will have the $m_1\in\ker\del_M^1$ term vanish, leaving us with $\del_M^1(\ker\varphi_1).$

Amazingly, we can check now that the image of $\ker\varphi_1\to\ker\varphi_2,$ induced by $\del_M^1$ by restriction, is exactly $\del_M^1(\ker\varphi_1),$ so the exactness follows. $\blacksquare$

And here is the right.
\begin{lemma}
    Fix everything as above. Then $\ker\varphi_2\stackrel\delta\to\op{coker}\varphi_0\to\op{coker}\varphi_1$ is exact.
\end{lemma}
Again, we begin by focusing on the image of $\delta,$ which we do by tracking forwards along
\[M_2\stackrel{\del_M^1}\leftarrow M_1\stackrel{\varphi_1}\to N_1\stackrel{\del_N^0}\leftarrow N_0.\]
Because we want the image, we begin with $M_2$ at the left. Because $\del_M^1$ is surjective, the movement along $\del_M^1$ is (boringly) all of $M_1.$ Then we push $M_1$ through $\varphi_1$ to get $\op{im}\varphi_1.$ If we go ahead and push all to $N_0,$ we get $\left(\del_N^0\right)^{-1}(\op{im}\varphi_1).$ We remark that we are actually interested in the cosets $n_0+\op{im}\varphi_0$ for $n_0\in\left(\del_N^0\right)^{-1}(\op{im}\varphi_1).$

Now we talk about the kernel of $\op{coker}\varphi_0\to\op{coker}\varphi_1,$ which is induced by modding out of $\del_N^0.$ Thus, our kernel is the set of cosets $n_0+\op{im}\varphi_0$ such that $\del_N^0(n_0)\in\op{im}\varphi_1.$ This latter condition is equivalent (by the injectivity of $\del_N^0$) to requiring $n_0\in\left(\del_N^0\right)^{-1}(\op{im}\varphi_1),$ so this kernel is the image of $\delta.$ The exactness follows. $\blacksquare$

Putting all of this together, we can splice our exact sequences into a long sequence
\[\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2\stackrel\delta\to\op{coker}\varphi_0\to\op{coker}\varphi_1\to\op{coker}\varphi_2.\]
This is exactly what we wanted. $\blacksquare$

With the proof out of the way, I remark that this is not really a proof that should be done in public. There are a very large number of small steps we did above that are true but really annoying to justify, and they're not very instructive anyways.

Regardless, we mention a few related statements, for completeness. Suppose we make the top row of the snake lemma diagram a short exact sequence so that we have the following commutative diagram.
\begin{center}
    \begin{tikzcd}
    	0 & {M_0} & {M_1} & {M_2} & 0 \\
    	0 & {N_0} & {N_1} & {N_2}
    	\arrow["{\partial_M^0}", hook, from=1-2, to=1-3]
    	\arrow["{\partial_M^1}", two heads, from=1-3, to=1-4]
    	\arrow[from=1-4, to=1-5]
    	\arrow[from=2-1, to=2-2]
    	\arrow["{\partial_N^0}"', hook, from=2-2, to=2-3]
    	\arrow["{\partial_N^1}"', from=2-3, to=2-4]
    	\arrow["{\varphi_0}", from=1-2, to=2-2]
    	\arrow["{\varphi_1}", from=1-3, to=2-3]
    	\arrow["{\varphi_2}", from=1-4, to=2-4]
    	\arrow[from=1-1, to=1-2]
    \end{tikzcd}
\end{center}
In this case, $\del_M^0$ is injective, so $\ker\varphi_0\to\ker\varphi_1,$ induced by $\del_M^0$ on restriction, is also injective. It follows that we can extend the long exact sequence as
\[0\to\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2\stackrel\delta\to\op{coker}\varphi_0\to\op{coker}\varphi_1\to\op{coker}\varphi_2\]
due to this injectivity.

Similarly, suppose we make the bottom of the snake lemma diagram a short exact sequence so that we have the following commutative diagram.
\begin{center}
    \begin{tikzcd}
    	& {M_0} & {M_1} & {M_2} & 0 \\
    	0 & {N_0} & {N_1} & {N_2} & 0
    	\arrow["{\partial_M^0}", hook, from=1-2, to=1-3]
    	\arrow["{\partial_M^1}", two heads, from=1-3, to=1-4]
    	\arrow[from=1-4, to=1-5]
    	\arrow[from=2-1, to=2-2]
    	\arrow["{\partial_N^0}"', hook, from=2-2, to=2-3]
    	\arrow["{\partial_N^1}"', two heads, from=2-3, to=2-4]
    	\arrow["{\varphi_0}", from=1-2, to=2-2]
    	\arrow["{\varphi_1}", from=1-3, to=2-3]
    	\arrow["{\varphi_2}", from=1-4, to=2-4]
    	\arrow[from=2-4, to=2-5]
    \end{tikzcd}
\end{center}
In this case, $\del_N^1$ is surjective, so the map $\op{coker}\varphi_1\to\op{coker}\varphi_2$ (induced by modding $\del_N^1:N_1\to N_2$ by modding) is also surjective. It follows that we can extend the long exact sequence as
\[\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2\stackrel\delta\to\op{coker}\varphi_0\to\op{coker}\varphi_1\to\op{coker}\varphi_2\to0\]
due to this surjectivity.

Anyways, while we're here, we mention how to use the Snake lemma to construct the long exact sequence of cohomology.
\begin{theorem}
    Fix $G$ a finite group and $0\to M\to N\to P\to 0$ a short exact sequence of $G$-modules. Further suppose that we have injective resolutions $I^\bullet,J^\bullet,K^\bullet$ of $M,N,P$ respectively such that $0\to(I^\bullet)^G\to(J^\bullet)^G\to(K^\bullet)^G\to0$ is an exact sequence of complexes. Then we have the following canonical long exact sequence.
    \begin{center}
        \begin{tikzcd}
        	0 & {H^0(G,M)} & {H^0(G,N)} & {H^0(G,P)} \\
        	& {H^1(G,M)} & {H^1(G,N)} & {H^1(G,P)} \\
        	& {H^2(G,M)} & {H^2(G,N)} & \cdots
        	\arrow[from=1-1, to=1-2]
        	\arrow[from=1-2, to=1-3]
        	\arrow[from=1-3, to=1-4]
        	\arrow[from=1-4, to=2-2]
        	\arrow[from=2-2, to=2-3]
        	\arrow[from=2-3, to=2-4]
        	\arrow[from=2-4, to=3-2]
        	\arrow[from=3-2, to=3-3]
        	\arrow[from=3-3, to=3-4]
        \end{tikzcd}
    \end{center}
\end{theorem}
We begin by saying the map $H^0(G,M)\to H^0(G,N)$ is injective because it is just the map $M^G\to N^G$ induced by $\varphi$ itself on restriction, which is injective by the original short exact sequence. We spend the rest of our time ignoring the $0$ at the start of the long exact sequence.

The strength of the short exact sequence of injective resolutions $0\to I^\bullet\to J^\bullet\to K^\bullet\to0$ is that the following commutative diagram has exact columns; note we have already taken the fixed-point functor.
\begin{center}
    \begin{tikzcd}
    	& 0 & 0 & 0 & 0 \\
    	0 & {M^G} & {(I^0)^G} & {(I^1)^G} & {(I^2)^G} & \cdots \\
    	0 & {N^G} & {(J^0)^G} & {(J^1)^G} & {(J^2)^G} & \cdots \\
    	0 & {P^G} & {(K^0)^G} & {(K^1)^G} & {(K^2)^G} & \cdots \\
    	&& 0 & 0 & 0
    	\arrow["\varphi", from=2-2, to=3-2]
    	\arrow["\psi", from=3-2, to=4-2]
    	\arrow[from=1-2, to=2-2]
    	\arrow[from=1-3, to=2-3]
    	\arrow["{\varphi_0}", from=2-3, to=3-3]
    	\arrow["{\psi_0}", from=3-3, to=4-3]
    	\arrow[from=4-3, to=5-3]
    	\arrow[from=1-4, to=2-4]
    	\arrow["{\varphi_1}", from=2-4, to=3-4]
    	\arrow["{\psi_1}", from=3-4, to=4-4]
    	\arrow[from=4-4, to=5-4]
    	\arrow[from=1-5, to=2-5]
    	\arrow["{\varphi_2}", from=2-5, to=3-5]
    	\arrow["{\psi_2}", from=3-5, to=4-5]
    	\arrow[from=4-5, to=5-5]
    	\arrow[from=2-1, to=2-2]
    	\arrow[from=2-2, to=2-3]
    	\arrow["{\partial_I^0}", from=2-3, to=2-4]
    	\arrow["{\partial_I^1}", from=2-4, to=2-5]
    	\arrow[from=3-2, to=3-3]
    	\arrow["{\partial_J^1}", from=3-4, to=3-5]
    	\arrow["{\partial_I^2}", from=2-5, to=2-6]
    	\arrow["{\partial_J^2}", from=3-5, to=3-6]
    	\arrow[from=3-1, to=3-2]
    	\arrow[from=4-1, to=4-2]
    	\arrow[from=4-2, to=4-3]
    	\arrow["{\partial_K^1}", from=4-4, to=4-5]
    	\arrow["{\partial_K^2}", from=4-5, to=4-6]
    	\arrow["{\partial_K^0}", from=4-3, to=4-4]
    	\arrow["{\partial_J^0}", from=3-3, to=3-4]
    \end{tikzcd}
\end{center}
To extend backwards a bit, we set $\del_\bullet^{-1}$ to be the zero map, and we set $\varphi^{-1}=\varphi$ and $\psi^{-1}=\psi.$

The proof of this statement consists of applying the Snake lemma twice. The first application will give us the long exact sequence, and the second will show that the first is actually valid. Indeed, our first snake diagram is as follows; fix $k\ge0$ a nonnegative integer.
\begin{center}
    \begin{tikzcd}
    	& {(I^k)^G/\operatorname{im}\partial_I^{k-1}} & {(J^k)^G/\operatorname{im}\partial_J^{k-1}} & {(K^k)^G/\operatorname{im}\partial_K^{k-1}} & 0 \\
    	0 & {\ker\partial_I^{k+1}} & {\ker\partial_J^{k+1}} & {\ker\partial_K^{k+1}}
    	\arrow["{\varphi^k}", from=1-2, to=1-3]
    	\arrow["{\psi^k}", from=1-3, to=1-4]
    	\arrow[from=1-4, to=1-5]
    	\arrow["{\partial_I^k}", from=1-2, to=2-2]
    	\arrow[from=2-1, to=2-2]
    	\arrow["{\varphi^{k+1}}", from=2-2, to=2-3]
    	\arrow["{\psi^{k+1}}", from=2-3, to=2-4]
    	\arrow["{\partial_J^k}", from=1-3, to=2-3]
    	\arrow["{\partial_K^k}", from=1-4, to=2-4]
    \end{tikzcd}
\end{center}
We don't know if the rows are exact yet, so to build suspens for a bit, we will show this in a lemma. Note quickly that the vertical maps are well-defined because $\op{im}\del_\bullet^{k-1}\subseteq\ker\del_\bullet^k$ due to the complex, so these maps are safe.
\begin{lemma}
    Fix everything as above. Then, for $k\ge0$ a nonnegative integer, the following commutative diagram has exact rows.
    \begin{center}
        \begin{tikzcd}
        	& {(I^k)^G/\operatorname{im}\partial_I^{k-1}} & {(J^k)^G/\operatorname{im}\partial_J^{k-1}} & {(K^k)^G/\operatorname{im}\partial_K^{k-1}} & 0 \\
        	0 & {\ker\partial_I^{k+1}} & {\ker\partial_J^{k+1}} & {\ker\partial_K^{k+1}}
        	\arrow["{\varphi^k}", from=1-2, to=1-3]
        	\arrow["{\psi^k}", from=1-3, to=1-4]
        	\arrow[from=1-4, to=1-5]
        	\arrow["{\partial_I^k}", from=1-2, to=2-2]
        	\arrow[from=2-1, to=2-2]
        	\arrow["{\varphi^{k+1}}", from=2-2, to=2-3]
        	\arrow["{\psi^{k+1}}", from=2-3, to=2-4]
        	\arrow["{\partial_J^k}", from=1-3, to=2-3]
        	\arrow["{\partial_K^k}", from=1-4, to=2-4]
        \end{tikzcd}
    \end{center}
\end{lemma}
Note that $(I^k)^G/\op{im}\del_I^{k-1}$ is exactly $\op{coker}\del_I^{k-1},$ and similar for $J$ and $K,$ so we really want the exactness of the sequences
\[\begin{cases}
    \op{coker}\del_I^{k-1}\to\op{coker}\del_J^{k-1}\to\op{coker}\del_K^{k-1}\to0, \\
    0\to\ker\del_I^k\to\ker\del_J^k\to\ker\del_K^k.
\end{cases}\]
When written this way, this is begging for an application of the Snake lemma. In particular, we have the following diagram, for $\ell\ge0.$
\begin{center}
    \begin{tikzcd}
    	0 & {(I^\ell)^G} & {(J^\ell)^G} & {(K^\ell)^G} & 0 \\
    	0 & {(I^{\ell+1})^G} & {(J^{\ell+1})^G} & {(K^{\ell+1})^G} & 0
    	\arrow[from=1-1, to=1-2]
    	\arrow["{\varphi^\ell}", from=1-2, to=1-3]
    	\arrow["{\psi^\ell}", from=1-3, to=1-4]
    	\arrow[from=1-4, to=1-5]
    	\arrow[from=2-1, to=2-2]
    	\arrow["{\varphi^{\ell+1}}", from=2-2, to=2-3]
    	\arrow["{\psi^{\ell+1}}", from=2-3, to=2-4]
    	\arrow[from=2-4, to=2-5]
    	\arrow["{\partial_I^\ell}", from=1-2, to=2-2]
    	\arrow["{\partial_J^\ell}", from=1-3, to=2-3]
    	\arrow["{\partial_K^\ell}", from=1-4, to=2-4]
    \end{tikzcd}
\end{center}
This diagram is just a piece of the short exact sequence of chain complexes, so we exact rows for free. Here the snake lemma (this is the second application of the snake lemma) gives a long exact sequence
\[0\to\ker\del_I^\ell\to\ker\del_J^\ell\to\ker\del_K^\ell\to\op{coker}\del_I^\ell\to\op{coker}\del_J^\ell\to\op{coker}\del_K^\ell\to0.\]
Setting $\ell=k$ gives the needed $0\to\ker\del_I^k\to\ker\del_J^k\to\ker\del_K^k$ for any $k\ge0.$ Setting $\ell=k-1$ gives the needed $\op{coker}\del_I^{k-1}\to\op{coker}\del_J^{k-1}\to\op{coker}\del_K^{k-1}\to0$ for any $k\ge1.$

It remains to show that the top row is exact for $k=0.$ In this case, $\op{im}\del_\bullet^{k-1}=\op{im}\del_\bullet^{-1}$ is the zero map, so we are really asking for the exactness of $I^0\to J^0\to K^0\to0.$ However, this comes for free from the short exact sequence of complexes, so we are safe. $\blacksquare$

Now that we know our rows are exact, we see the snake lemma gives the induced long exact sequence
\[\ker\del_I^k\to\ker\del_J^k\to\ker\del_K^k\to\op{coker}\del_I^k\to\op{coker}\del_J^k\to\op{coker}\del_K^k.\]
Here, $\ker\del_\bullet^k$ consists of the cosets $a+\op{im}\del_\bullet^{k-1}$ such that $\del_I^k(a)=0.$ In particular, this kernel is $\ker\del_I^k/\op{im}\del_\bullet^{k-1}=H^k(G,\bullet).$

Further, $\op{coker}\del_\bullet^k$ is $\ker\del_\bullet^{k+1}$ modded by the image of $\del_\bullet^k$ under $(\bullet^k)^G.$ (The modding by $\op{im}\del_\bullet$ doesn't matter because all these elements go to $0.$) Thus, the cokernel is $\ker\del_\bullet^{k+1}/\op{im}\del_\bullet^k=H^{k+1}(G,\bullet).$ This cokernel computation is why our bottom row consits of kernels.

Bringing this all together, we have the long exact sequence
\[H^k(G,M)\to H^k(G,N)\to H^k(G,P)\to H^{k+1}(G,M)\to H^{k+1}(G,N)\to H^{k+1}(G,N).\]
We also note the first two maps $H^k(G,N)\to H^k(G,N)\to H^k(G,P)$ are induced by $\varphi^k$ and $\psi^k$ by restriction, so they are really the map $\varphi^k:(I^k)^G\to(J^k)^G$ and $\psi^k:(J^k)^G\to (K^k)^G$ upon modding out in the codomain and domain. This is to say that the maps on the left are the good ones $H^k(\varphi)$ and $H^k(\psi).$

Similarly, the last two maps $H^{k+1}(G,N)\to H^{k+1}(G,N)\to H^{k+1}(G,N)$ are induced by the restrictions of $\varphi^{k+1}:\ker\del_I^{k+1}\to\ker\del_J^{k+1}$ and $\psi^{k+1}:\ker\del_J^{k+1}\to\ker\del_K^{k+1}$ after modding out the codomain and domain by some images. Again, this tells us the maps on the right are the good ones $H^{k+1}(\varphi)$ and $H^{k+1}(\psi)$ by construction.

Thus, we have a series of long exact sequences for $k\ge0,$ as follows; here, $\delta^k$ is given by the Snake lemma.
\begin{center}
    \begin{tikzcd}
    	{H^k(G,M)} & {H^k(G,N)} & {H^k(G,P)} \\
    	{H^{k+1}(G,M)} & {H^{k+1}(G,N)} & {H^{k+1}(G,P)}
    	\arrow["\delta^k"{description}, from=1-3, to=2-1]
    	\arrow["{H^k(\varphi)}", from=1-1, to=1-2]
    	\arrow["{H^k(\psi)}", from=1-2, to=1-3]
    	\arrow["{H^{k+1}(\varphi)}"', from=2-1, to=2-2]
    	\arrow["{H^{k+1}(\psi)}"', from=2-2, to=2-3]
    \end{tikzcd}
\end{center}
Because of each these sequences are exact on their own, and the sequence for $k$ and $k+1$ share the bottom and top row respectively (that is, the maps are the same), we may splice all of these sequences together into the desired long exact sequence. $\blacksquare$

As some commentary, it might feel very strong that we required the short exact sequence of complexes, and indeed, this is not easy to prove, and I do not yet know its proof. However, it is always possible. We say (but do not prove) that merely fixing the injective resolution $J^\bullet$ of $N$ to be $I^\bullet\oplus K^\bullet$ will do the trick.

\subsubsection{June 20th}
Today I learned some stuff around the Cauchy-Schwartz inequality, from THE BOOK. We begin with a particularly simple proof of the Cauchy-Schwartz inequality.
\begin{theorem}
    Fix $V$ a real vector space with $\langle\bullet,\bullet\rangle$ the usual inner product. Then for any $v,w\in V,$ we have $\lVert v\rVert\cdot\lVert w\rVert\ge\langle v,w\rangle,$ with equality if and only if $v$ and $w$ are linearly dependent.
\end{theorem}
If $v=0,$ then the equality reads $0\ge0,$ so inequality holds, and indeed, $w$ and $0=0w$ are linearly dependent. Otherwise, $v$ and $w$ linearly dependent guarantees $\lambda\in\RR$ such that $w=\lambda v.$ In this case, the inequality reads
\[\lVert v\rVert\cdot\lVert\lambda w\rVert\ge\langle v,\lambda v\rangle.\]
Factoring out the $\lambda$ gives this equivalent to $\lambda\lVert v\rVert^2=\lambda\langle v,v\rangle,$ which is in fact an inequality.

It remains to show the inequality for $v$ and $w$ not linearly dependent. Let $x\in\RR$ be a real number. The main character in our story is
\[\lVert xv+w\rVert^2=\langle xv+w,xv+w\rangle=\langle xv,xv\rangle+\langle xv,w\rangle+\langle w,xv\rangle+\langle w,\rangle.\]
This collapses to $x^2\lVert v\rVert^2+\lVert w\rVert^2+2x\langle v,w\rangle.$ Now, as $x$ varies, the linear independence of $v$ and $w$ requires $\lVert xv+w\rVert>0$ always (!), so the discriminant of the created quadratic is negative. That is,
\[(2\langle v,w\rangle)^2-4\cdot\lVert v\rVert^2\cdot\lVert w\rVert^2<0.\]
This rearranges into the desired inequality. $\blacksquare$

In particular, we will be using a particular case of the Cauchy-Schwartz, which is Titu's lemma.
\begin{lemma}
    Fix $\{x_k\}_{k=1}^n$ a sequence of real numbers. Then
    \[\left(\sum_{k=1}^nx_k\right)^2\le n\sum_{k=1}^nx_k^2,\]
    with equality if and only if all the $x_\bullet$ are equal.
\end{lemma}
Apply the Cauchy-Schwartz to $v=\langle x_1,\ldots,x_n\rangle$ and $w=\langle 1,\ldots,1\rangle.$ In particular, note that $w\mapsto-w$ in the inequality $\lVert v\rVert\cdot\lVert w\rVert\ge\langle v,w\rangle$ flips the sign of the dot product while not changing the left-hand side. Thus,
\[|\langle v,w\rangle|=\max\{\langle v,w\rangle,-\langle v,w\rangle\}\le\lVert v\rVert\cdot\lVert w\rVert.\]
Squaring, we see that $\langle v,w\rangle^2\le\lVert v\rVert^2\cdot\lVert w\rVert^2.$ Plugging in for $v$ and $w$ gives
\[\left(\sum_{k=1}^nx_k\cdot1\right)^2\le\left(\sum_{k=1}^nx_k^2\right)\left(\sum_{k=1}^n1^2\right).\]
This rearranges into the desired inequality.

As for the equality case, we know Cauchy-Schwartz has equality if and only if $v$ and $w$ are linearly independent, but here $w\ne0,$ so equality holds if and only if $v=\lambda w$ for some real $\lambda.$ But this means
\[\langle x_1,\ldots,x_n\rangle=v=\lambda\langle1,\ldots,1\rangle=\langle\lambda,\ldots,\lambda\rangle.\]
Such a $\lambda$ exists if and only if all the $x_\bullet$ are equal, in which case $\lambda=x_1$ will do. $\blacksquare$

Anyways, here is our main attraction.
\begin{theorem}
    Suppose $f(x)=x^n+\sum_{k=0}^{n-1}a_kx^k$ is a polynomial with real coefficients and exclusively real roots. Then all the roots are contained in the interval with endpoints
    \[-\frac{a_{n-1}}n\pm\frac{n-1}n\sqrt{a_{n-1}^2-\frac{2n}{n-1}a_{n-2}}.\]
\end{theorem}
This statement comes out of applying Titu's lemma to Vieta's formulae. Let the roots of the polynomial be $r_1,\ldots,r_n,$ and we will show that $r_n$ lives in the desired interval. Relabeling the roots shows that all of them will live in the desired interval because the endpoints of the interval are independent of the $r_\bullet.$

By Vieta's formulae, we have that
\[-a_{n-1}=\sum_{k=1}^nr_k\qquad\text{and}\qquad a_{n-2}=\sum_{1\le k<\ell\le n}r_kr_\ell.\]
In order to make Titu's lemma more viable, as well as to relate these two equations, we square the first to get
\[a_{n-1}^2=\sum_{k,\ell=1}^nr_kr_\ell=\left(\sum_{k=1}^nr_k^2\right)+2\left(\sum_{1\le k,\ell\le n}r_kr_\ell\right)=\left(\sum_{k=1}^nr_k^2\right)+2a_{n-2}.\]
We have a sum of squares, so we now anticipate an application of Titu's lemma. However, to isolate $r_n,$ we remove it from the sum and note
\[\sum_{k=1}^{n-1}r_k^2\ge\frac1{n-1}\left(\sum_{k=1}^{n-1}r_k\right)^2.\]
The work above shows that the left-hand side is $a_{n-1}^2-2a_{n-2}-r_n^2,$ and the right-hand side is $\frac1{n-1}(-a_{n-1}-r_n)^2$ by Vieta's formulae.

Thus, we have the quadratic inequality
\[a_{n-1}^2-2a_{n-2}-r_n^2\ge\frac1{n-1}(a_{n-1}+y_n)^2.\]
This will give the desired statement after a computation. Alias $r=r_n.$ Multiplying by $n-1$ and moving everything to the right, we see
\[0\ge nr^2+2a_{n-1}r-(n-1)\left(a_{n-1}^2-2a_{n-2}\right)+a_{n-1}^2.\]
Completing the square, we see
\[0\ge n\left(r+\frac{a_{n-1}}n\right)^2-n\cdot\frac{a_{n-1}^2}{n^2}-(n-1)a_{n-1}^2+2(n-1)a_{n-2}+a_{n-1}^2,\]
which is
\[\left(r+\frac{a_{n-1}}n\right)^2\le\frac{a_{n-1}^2+n(n-2)a_{n-1}^2}{n^2}-\frac{2n(n-1)a_{n-2}}{n^2}.\]
Quickly, note that $1+n(n-2)=(n-1)^2.$ Anyways, taking the square root, we see
\[\left|r+\frac{a_{n-1}}n\right|\le\sqrt{\frac{(n-1)^2}{n^2}a_{n-1}^2-\frac{2n(n-1)^2}{n^2(n-1)}a_{n-2}}.\]
After factoring out $n/(n-1),$ we see that $r$ does indeed live in the desired interval. $\blacksquare$

As an aside, our application of Titu's lemma means that equality here holds if and only if all roots except for the one being examined are equal. So, for example, equality holds for any quadratic or if all roots are equal. However, this isn't sharp. For example, $f(x)=x^3+x^2-x-1=(x-1)(x+1)^2$ gives the interval
\[\left|r+\frac13\right|\le\frac23\sqrt{1^2-\frac62\cdot-1}=\frac23\cdot2=\frac43,\]
for which the $r=1$ provides the equality case. This is legal because the Titu's lemma application when examining $r=1$ does have all the other roots, which are $-1$ and $-1,$ equal.

What amuses me most about this statement is that it can act like a detector for when a polynomial with real coefficients must have a non-real root. For example, here is a quick application.
\begin{proposition}
    A polynomial of the form $f(x)=x^n+\sum_{k=0}^{n-3}a_kx^k\in\RR[x]$ (note the $n-3$) has exclusively real roots if and only if $f(x)=x^n.$
\end{proposition}
In one direction, $f(x)=x^n$ has all of its roots real because $x^n=0$ if and only if $x=0.$

The other direction is harder. Fix $f(x)$ of the desired form with exclusively real roots. The point is that we have the coefficients $a_{n-1}x^{n-1}=0x^{n-1}$ and $a_{n-2}x^{n-2}=0x^{n-2}.$ So plugging in $a_{n-1}=a_{n-2}=0$ into the interval given above, we see that our roots live in the interval with endpoints
\[-\frac0n\pm\frac{n-1}n\sqrt{0^2-\frac{2n}{n-1}\cdot0}=0\pm0.\]
In other words, all roots of $f(x)$ must be $0,$ which means $f(x)$ has the root $0$ with multiplicity $n,$ or $f(x)=x^n.$ This completes the proof. $\blacksquare$

As an example, we see that $x^n-10x+1$ has a non-real root for any $n\ge4,$ immediately. I guess the way to read the above statement is that $a_{n-1}=a_{n-2}=0$ forces $f(x)$ to grow ``too much'' like $x^n,$ and if there are deviations, then there's no way to introduce enough ripples into the graph of $f(x)$ to give $n$ real roots.

Here is another application.
\begin{proposition}
    A polynomial of the form $f(x)=x^n+\sum_{k=0}^{n-1}a_kx^k\in\RR[x]$ with exclusively real roots must have
    \[2na_{n-2}\le(n-1)a_{n-1}^2,\]
    with equality if and only if all the roots of $f(x)$ are equal.
\end{proposition}
The trick here is to examine
\[xf(x)=x^{n+1}+\sum_{k=1}^{n-1}a_{k-1}x^k.\]
The point is that now $0$ a root because the constant term has vanished, and $xf(x)$ still has exclusively real roots. Thus, we may plug in $0$ to the interval for $xf(x),$ we see that
\[\left|0+\frac{a_{n-1}}{n+1}\right|\le\frac n{n+1}\sqrt{a_{n-1}^2-\frac{2n+2}na_{n-2}}.\]
Working backwards a bit, we can make this criterion a bit more palatable. For example, squaring gives
\[\frac{a_{n-1}^2}{(n+1)^2}\le\frac{n^2}{(n+1)^2}a_{n-1}^2-\frac{2n}{n+1}a_{n-2}.\]
This rearranges into
\[2n(n+1)a_{n-2}\le\left(n^2-1\right)a_{n-1}^2,\]
or $2na_{n-2}\le(n+1)a_{n-1}^2.$ This is exactly the equality case what we wanted.

Our equality case comes from the equality case of the applied theorem. Indeed, equality holds if and only if all roots of $xf(x)$ other than the chosen $0$ are equal, which is equivalent to all roots of $f(x)$ being equal. $\blacksquare$

We remark that we could have looked at the proof of the theorem statement and simply used Titu's lemma directly. Indeed, if we let the roots be $r_1,\ldots,r_n,$ then Vieta's formulae implies that the given inequality is equivalent to
\[2n\sum_{1\le k<\ell\le n}r_kr_\ell\le(n-1)\left(\sum_{k=1}^nr_k\right)^2.\]
To make the left side more palatable, we note as we did in the proof that it is $n\left(\left(\sum_kr_k\right)^2-\sum_kr_k^2\right).$ Everything rearranges into
\[\left(\sum_{k=1}^nr_k\right)^2\le n\sum_{k=1}^nr_k^2,\]
which is exactly Titu's lemma. And now it is more apparent that the equality case is when all roots are equal, as this is the equality case for Titu's lemma. However, it is impressive that our theorem was powerful enough to retrieve this result without requiring another appeal to Titu's lemma.

Anyways, it's kind of magic to see the above proposition work.
\begin{example}
    Take $f(x)=(x+1)(x+2)(x+3)=x^3+6x^2+11x+6,$ which yields
    \[2\cdot3\cdot11\le2\cdot6^2,\]
    which is $66\le72.$ So indeed, all is well in the world.
\end{example}

\subsubsection{June 21st}
Today I learned a proof of Nakayama's lemma as a corollary of the determinant trick. Here is the determinant trick.
\begin{proposition}
    Fix $R$ a ring and $M$ an $R$-module, where $M$ is generated by $n\in\NN$ elements. Now suppose $\varphi:M\to M$ is an $R$-module homomorphism such that $\varphi(M)\subseteq IM$ for an ideal $I\subseteq R.$ Then we can write
    \[\varphi^n+\sum_{k=0}^{n-1}\mu_{r_k}\varphi^k=0,\]
    where $r_k\in I.$
\end{proposition}
Very quickly, we recall that $\mu_{r_k}:M\to M$ is the $R$-module homomoprhism induced by $\mu_{r_k}(m):=r_km.$ When we proved the Cayley-Hamtilton theorem, we showed that $\mu_\bullet(R)[\varphi]$ is a commutative ring.

Anyways, this is proved in pretty much the same way we proved the Cayley-Hamilton theorem. Instead of using the basis vectors $\{e_k\}_{k=1}^n$ of $R^n,$ we pick up generators $\{m_k\}_{k=1}^n$ of $M.$ Our next task is defining $\Delta$ as in Cayley-Hamtilton theorem, then we'll show $\det\Delta=0,$ and this determinant expansion will expand in the desired identity.

Running things in the fast lane a bit, we don't have a matrix to define a $\Delta,$ but we can use the evidence that $\det\Delta=0$ to define $\Delta,$ a bit backwards. Indeed, we write, for any generator $m_\ell,$
\[\varphi(m_\ell)=\sum_{k=1}^na_{k,\ell}m_k\]
for some $a_{k,\ell}\in R$; in fact, we may restrict $a_{k,\ell}\in I$ because $\varphi(m_\ell)\in IM.$ Attempting to move everything over to the left-hand side as in our proof of Cayley-Hamilton, we write $\varphi m_\ell=\sum_{k=1}^n\varphi1_{k=\ell}m_k$ and $a_{k,\ell}m_k=\mu_{a_{k,\ell}}m_k$ so that
\[\sum_{k=1}^n\left(\varphi1_{k=\ell}-\mu_{a_{k,\ell}}\right)m_k=0.\]
Last time, we found that $\Delta_{k,\ell}:=\varphi1_{k=\ell}-\mu_{a_{k,\ell}},$ but here we will just define it this way. Here, $\Delta\in(\mu_\bullet(R)[\varphi])^{n\times n},$ and in fact $\Delta\in(\mu_\bullet(I)[\varphi])^{n\times n}.$

From here, the conversion from evidence to statement is the same.
\begin{lemma}
    Fix everything as above. Then $\det\Delta\in\mu_\bullet(R)[\varphi]$ is the zero endomorphism.
\end{lemma}
Fixing an index $j,$ we show that each generator $m_j$ vanishes under $\det\Delta.$ Indeed, let $\op{adj}\Delta$ be the adjugate matrix so that $\Delta\cdot\op{adj}\Delta=(\det\Delta)I_n.$ Continuing, we sum our evidence over $\ell$ like a matrix multiplication as
\[\sum_{\ell=1}^n\Bigg((\op{adj}\Delta)_{\ell,j}\underbrace{\sum_{k=1}^n\Delta_{k,\ell}m_k}_0\Bigg)=0.\]
Actually making this a matrix multiplication turns this into
\[0=\sum_{k=1}^n\left(\sum_{\ell=1}^n\Delta_{k,\ell}(\op{adj}\Delta)_{\ell,j}\right)m_k=\sum_{k=1}^n(\Delta\cdot\op{adj}\Delta)_{k,j}m_k.\]
Now, $\Delta\cdot\op{adj}\Delta=(\det\Delta)I_n$ means that $(\Delta\cdot\op{adj}\Delta)_{k,j}=(\det\Delta)1_{k=j}.$ In particular, the above summands vanish except when $k=j,$ collapsing this to $(\det\Delta)m_j=0.$

To finish off showing $\det\Delta$ is the zero endomorphism, we note that any $m\in M$ can be written as $\sum_{k=1}^nr_km_k$ for some $\{r_k\}_{k=1}^n\subseteq R.$ Then
\[(\det\Delta)(m)=\sum_{k=1}^nr_k(\det\Delta)m_k=0\]
by linearity. This finishes the proof. $\blacksquare$

It remains to expand $\det\Delta.$ Doing this by brute force, we write
\[\det\Delta=\sum_{\sigma\in S_n}\op{sgn}\sigma\prod_{k=1}^n\Delta_{k,\sigma(k)}.\]
Because each of the entries $\Delta_{k,\sigma(k)}$ lives in $\mu_\bullet(I)[\varphi],$ and $\mu_\bullet(I)[\varphi]$ is a commutative ring---indeed, $I\subseteq R$ is a subring and induces its own action on $M$---the above expansion shows provides a polynomial in $\mu_\bullet(I)[\varphi]$ which vanishes.

Further, a term of maximum degree in the expansion $\det\Delta$ will feature a $\varphi$ in each term, and because $\varphi$ only appears when $k=\sigma(k),$ a term of maximum degree must come from $\sigma=\op{id}$ and is so in fact unique. This term is
\[\op{sgn}\sigma\prod_{k=1}^n\Delta_{k,\sigma(k)}=+\prod_{k=1}^n(\varphi-\mu_{a_{k,\ell}}).\]
We have written this out entirely to say that the leading term of our polynomial $\det\Delta$ is $1\varphi^n.$ Thus, we have a monic polynomial in $\mu_\bullet(I)[\varphi]$ which vanishes, which is exactly what we wanted. $\blacksquare$

We remark quickly that we can strengthen this statement some with little effort. Indeed, in our expansion of $\det\Delta,$ a term with degree $k$ will come from choosing a $\sigma\in S_n$ with at least $k$ fixed points (these are the only way to have a $\varphi$) and then choosing exactly $k$ of the available $\varphi$s from the resulting product
\[\op{sgn}\sigma\prod_{k=1}^n(\varphi1_{k=\sigma(k)}-\mu_{a_{k,\sigma(k)}}).\]
The point of saying this is that choosing $k$ of the $\varphi$s means that the coefficient is a product of $n-k$ elements in $I.$ That is, this coefficient lies in the ideal $I^{n-k}.$ Adding up all of the $\varphi^k$ terms means that its coefficient lives in $I^{n-k}.$ So we have the following.
\begin{proposition}
    Fix $R$ a ring and $M$ an $R$-module, where $M$ is generated by $n\in\NN$ elements. Now suppose $\varphi:M\to M$ is an $R$-module homomorphism such that $\varphi(M)\subseteq IM$ for an ideal $I\subseteq R.$ Then we can write
    \[\varphi^n+\sum_{k=0}^{n-1}\mu_{r_k}\varphi^k=0,\]
    where $r_k\in I^{n-k}.$
\end{proposition}
This follows from the above (albeit terse) discussion. $\blacksquare$

Anyways, we will not actually use this strengthening in our discussion of Nakamaya's lemma. Here is our main attraction.
\begin{theorem}
    Fix $R$ a ring and $M$ a finitely generated $R$-module. Further suppose there is an ideal $I\subseteq A$ such that $1+I\subseteq A^\times$ and $IM=M.$ Then actually $M=0.$
\end{theorem}
We begin by using the $IM=M$ condition. Here, the determinant trick with $\varphi=\op{id}$ gives us some expression
\[\op{id}^n+\sum_{k=0}^{n-1}\mu_{r_k}\op{id}^k=0\]
for $r_k\in I.$ Because $\op{id}^\bullet=\op{id}$ is the identity endomorphism, we can rearrange the above into
\[\op{id}+\sum_{k=0}^{n-1}\mu_{r_k}=0.\]
In particular, because each $r_k\in I$ and because $\mu_\bullet(I)$ is a commutative ring (again, $M$ is also an $I$-module), we find that $\mu_r:=\sum_{k=0}^{n-1}\mu_{r_k}\in\mu_\bullet(I).$ Noticing that $\op{id}=\mu_1,$ the above reads $\mu_{1+r}=\mu_1+\mu_r=0.$

However, now we use the fact $1+I\subseteq A^\times,$ implying $1+r$ is a unit; let $(1+r)^{-1}$ be its inverse. To finish, we fix any $m\in M$ so that
\[m=(1+r)^{-1}\cdot(1+r)m=(1+r)^{-1}\cdot0=0\]
because $\mu_{1+r}=0.$ Thus, $M=0,$ and we are done here. $\blacksquare$

I am under the impression Nakamaya's lemma is mostly important in the case where $R$ is a local ring with $I=\mf m$ its maximal ideal. We have to show that $1+\mf m\subseteq R^\times$ in this case, but this is not hard.
\begin{lemma}
    Suppose $R$ is a local ring with maximal ideal $\mf m.$ Then $1+m\subseteq R^\times.$
\end{lemma}
Indeed, pick up any $m\in\mf m,$ and we examine $1+m.$ Well, $1+m\in\mf m$ implies that $1=(1+m)-m\in\mf m,$ which implies $\mf m=R,$ violating that $\mf m$ is a maximal ideal. Thus, $1+m$ is not in $\mf m.$

On the other hand, $1+m$ not a unit implies the ideal $(1+m)$ can be lifted to some maximal ideal by Zorn's lemma, so the uniqueness of the maximal ideal implies $1+m\in\mf m,$ which we just showed cannot be the case. Thus, $1+m\in R^\times,$ which is what we wanted. $\blacksquare$

In fact, more is true: if $R$ is a ring with $\mf m$ a maximal ideal with $1+\mf m\subseteq R^\times,$ then we can actually conclude that $R$ is a local ring. We outline because this is cute. Very roughly, this is true because a ideal $\mf m'\not\subseteq\mf m$ would have $\mf m'+\mf m=R,$ which promises $m'\in\mf m'$ and $m\in\mf m$ such that $m'+m=1.$ But then
\[m'=1+(-m)\in(1+\mf m)\subseteq R^\times.\]
Thus, $\mf m'$ is actually all of $R.$ This verifies that any maximal ideal must be $\mf m,$ for otherwise $\mf m'$ would contain elements not in $\mf m,$ requiring $\mf m'=R.$

The point of all of this is to say that if we want easy access to an ideal $I$ such that $1+I\subseteq R^\times,$ as is required by Nakamaya's lemma, in particular a maximal ideal $I,$ then we really want $R$ to be a local ring. In particular, having $I$ be maximal means that $R/I$ is a field, which is nice.

\subsubsection{June 22nd}
Today I learned some stuff about short exact sequences. I am under the impression that a short exact sequences more or less acts like a size limitation on the middle term: the left term lower-bounds the size, and the right-term upper-bounds the size. Here is an example of this.
\begin{proposition}
    Fix $R$ a ring. Suppose we have a short exact sequence $0\to M\to N\to P\to 0$ of $R$-modules. If $M$ and $P$ are finitely generated, than $N$ is also finitely generated.
\end{proposition}
For concreteness, we let $f:M\to N$ and $g:N\to P$ be the maps from the short exact sequence. The intuitive idea here is that we can ``pull back'' any element from $P$ to a coset of $N,$ and the exact element of the coset can be chosen from a particular push forward from $M.$

Let $M=(m_1,\ldots,m_a)$ and $P=(p_1,\ldots,p_c)$ generate $M$ and $P$ respectively. As promised, we pull back each generator $p_k$ to some $n_k\in N$ such that $g(n_k)=p_k$; this is legal because $g$ is surjective. We also $n'_k=f(m_k)$ to be our push forward of $M,$ and we claim that
\[N\stackrel?=(n'_1,\ldots,n'_a,n_1,\ldots,n_c).\]
This will finish because then $N$ is generated by the above $a+c$ elements. On one hand, $n'_\bullet,n_\bullet\in N$ implies that $(n'_1,\ldots,n'_a,n_1,\ldots,n_c)\subseteq N$ by taking $R$-linear combinations.

On the other hand, we fix any $n\in N,$ and we attempt to find its representation. We begin by finding which coset we are in with respect to $P.$ Well, $g(n)\in P,$ so we can write
\[g(n)=\sum_{k=1}^cr_kp_k\]
for some sequence $\{r_k\}_{k=1}^a\subseteq R$ because $P=(p_1,\ldots,p_k).$ Because $g(n_k)=p_k,$ the above rearranges into
\[0=g(n)-\sum_{k=1}^cr_kg(n_k)=g\left(n-\sum_{k=1}^cr_kn_k\right)\]
using the $R$-linearity of $g.$ The point is that we have found an element of $\ker g=\im f$ by exactness. So we can find some $m\in M$ for which
\[f(m)=n-\sum_{k=1}^cr_kn_k.\]
Now, $m\in M=(m_1,\ldots,m_a),$ so we can write $m=\sum_{k=1}^ar'_km_k$ for some sequence $\{r'_k\}_{k=1}^a\subseteq R.$ Pushing this through $f,$ we see
\[n-\sum_{k=1}^cr_kn_k=f\left(\sum_{k=1}^ar'_km_k\right)=\sum_{k=1}^ar'_kf(m_k),\]
where we have again used the $R$-linearity of $f.$ To finish, we recall $f(m_k)=n'_k,$ so the above rearranges into
\[n=\sum_{k=1}^ar'_kn'_k+\sum_{k=1}^cr_kn_k\in(n'_1,\ldots,n'_c,n_1,\ldots,n_c).\]
Thus, $N\subseteq(n'_1,\ldots,n'_a,n_1,\ldots,n_c),$ finishing the proof. $\blacksquare$

It is not exactly true that the left-hand element $M$ in the above short exact sequence is lower-bounding the size of $N$ insofar as everything is finitely generated, but we do see that the number of generators we needed for $N$ is lower-bounded by those needed for $M.$ This is in fact a common case.
\begin{definition}
    Fix $R$ a ring. A split short exact sequence
    \[0\to M\stackrel f\to N\stackrel g\to P\to 0\]
    is a split short exact sequence if and only if there is an isomorphism $\varphi:N\to M\oplus P$ such that $(\varphi\circ f)(m)=(m,0)$ and $(g\circ\varphi^{-1})((m,p))=p.$
\end{definition}
The requirement that $\varphi$ is well-behaved with respect to $f$ and $g$ is important to this definition, albeit for reasons which aren't clear to me. I am told that for Noetherian rings this behavior is free, but for non-Noetherian rings counterexamples can be given where these do not behave. I am not thinking too hard about this.

It feels like the intuition from a short exact sequence is that, given $m\in M$ and $p\in P,$ we ``should'' be able to determine an element in the middle: pull $p$ back up to $N$ to determine our coset of $\ker g,$ and then push forward from $m$ to determine the precise element in the coset of $\im f=\ker g.$ Split exact sequences give the best way for this to happen.

Anyways, it turns out that we can describe split exact sequences in a nicer way: the functions $f$ and $g$ have reasonable semi-inverses. Here is one of the directions.
\begin{proposition}
    Fix $R$ a ring and a short exact sequence $0\to M\stackrel f\to N\stackrel g\to P\to0.$ This short exact sequence is split if and only if there is an $R$-module homomorphism $\gamma:P\to N$ such that $g\circ\gamma=\op{id}_P.$
\end{proposition}
As a quick remark, we note that $\gamma$ is not a full-power inverse of $g$ because $g$ is not injective. But $g$ is surjective, so we can hope that after pulling back along $\gamma,$ maybe $g$ will take us where we started.

The forward direction is fairly easy. Fix $\varphi:N\to M\oplus P$ the isomorphism promised by the split short exact sequence, and we construct $\gamma:P\to N$ by
\[\gamma(p):=\varphi^{-1}((0,p)).\]
This is an $R$-module homomorphism because 
\[\varphi^{-1}((0,a_1p_1+a_2p_2))=\varphi^{-1}((0,a_1p_1))+\varphi^{-1}((0,a_2p_2))=a_1\varphi^{-1}((0,p_1))+a_2\varphi^{-1}((0,p_2))\]
for any $p_1,p_2\in P$ and $a_1,a_2\in R.$ And we see
\[(g\circ\gamma)(p)=g\left(\varphi^{-1}((0,p))\right)=\left(g\circ\varphi^{-1}\right)((0,p))=p\]
by construction of $\varphi.$ This finishes this direction.

In the other direction, suppose we have $\gamma:P\to N$ such that $g\circ\gamma=\op{id}_P,$ and we need to construct our $\varphi^{-1}:M\oplus P\to N.$ We work backwards: pick up $n\in N.$ We would like to make the $N$-coordinate $g(n),$ but we need a way to afterwards retrieve an element of $M.$ The key observation is that, pulling back along $\gamma,$
\[g\big(n-(\gamma\circ g)(n)\big)=g(n)-(g\circ\gamma)(g(n))=g(n)-g(n)=0.\]
Thus, $n-(\gamma\circ g)(n)\in\ker g=\im f$ by exactness, allowing us to find $m\in M$ such that
\[n=f(m)+\gamma(g(n)).\]
This tells us that we want $\varphi^{-1}$ to take the pair $(m,p)\in M\oplus P$ to $f(m)+\gamma(p).$ The above argument gives surjectivity: we just constructed $m$ and $p:=g(n)$ such that $(m,p)$ maps to $f(m)+\gamma(p)=n.$

This $\varphi^{-1}$ is an $R$-module homomorphism because $(a_1m_1+a_1m_2,a_1p_1+a_2p_2)$ goes to
\[f(a_1m_1+a_2m_2)+\gamma(a_1p_1+a_2p_2)=a_1(f(m_1)+\gamma(p_1))+a_2(f(m_2)+\gamma(p_2)).\]
Lastly, we need to check the kernel of $\varphi^{-1}.$ Well, if $(m,p)$ has $f(m)+\gamma(p)=0,$ then applying $g$ gives
\[(g\circ f)(m)+(g\circ\gamma)(p)=g(0)=0.\]
The first summand vanishes because $\im f=\ker g,$ and the second term is just $\op{id}_P(p)=p.$ Thus, $p=0.$ As for $m,$ we see $\gamma(p)=0$ implies $f(m)=0,$ which requires $m=0$ by the injectivity of $f.$ It follows that $(m,p)=(0,0),$ making our kernel trivial.

Thus, we have constructed an $A$-module isomorphism $\varphi^{-1}:M\oplus P\to N.$ It remains to check that $(\varphi\circ f)(m)=(m,0)$ and $(g\circ\varphi^{-1})((m,p))=p.$ For the first condition, we have to check that $f(m)=\varphi^{-1}((m,0)),$ but
\[\varphi^{-1}((m,0))=f(m)+\gamma(0)=f(m)\]
is exactly what we need. For the second condition, we simply note
\[(g\circ\varphi^{-1})((m,p))=g\big(f(m)+\gamma(p)\big)=(g\circ f)(m)+(g\circ\gamma)(p).\]
Now, $g\circ f$ is the zero map by exactness, and $g\circ\gamma=\op{id}_P$ by definition of $\gamma.$ Thus, this is indeed $p.$ This finishes the proof. $\blacksquare$

Of course, the same thing happens if we try to invert $f.$
\begin{proposition}
    Fix $R$ a ring and a short exact sequence $0\to M\stackrel f\to N\stackrel g\to P\to0.$ This short exact sequence is split if and only if there is an $R$-module homomorphism $\gamma:N\to M$ such that $\gamma\circ f=\op{id}_M.$
\end{proposition}
Again, we quickly remark that $\gamma$ is not a full-power inverse of $f$ because $f$ is not necessarily surjective. However, we can hope that $\gamma$ is an inverse on $\im f.$

Again, the forward direction is not that bad. Fix $\varphi:N\to M\oplus P$ the isomorphism promised by the split short exact sequence, and we construct $\gamma:N\to M$ by
\[\gamma(n):=\op{pr}_1(\varphi(n)),\]
the left coordinate of $\varphi(n).$ This is an $R$-module homomorphism because $\varphi$ and $\op{pr}_1$ are, so they're composition is as well. And we see
\[(\gamma\circ f)(m)=\op{pr}_1\big((\varphi\circ f)(m)\big)=\op{pr}_1((m,0))=m\]
by construction of $\varphi.$ This finishes this direction.

In the other direction, suppose we have $\gamma:N\to M$ such that $\gamma\circ f=\op{id}_M,$ and we need to construct our $\varphi:N\to M\oplus P.$ Looking around, we already have functions $\gamma:N\to M$ and $g:N\to P,$ so we hope that
\[\varphi(n):=(\gamma(n),g(n))\]
will do the trick. This is an $R$-module homomorphism because, for any $a_1,a_2\in R$ and $n_1,n_2\in N,$ we see $\varphi(a_1n_1+a_2n_2)$ is
\[(\gamma(a_1n_1+a_2n_2),g(a_1n_1+a_2n_2))=(a_1\gamma(n_1)+a_2\gamma(n_2),a_1g(n_1)+a_2g(n_2)),\]
which does indeed split into $a_1\varphi(n_1)+a_2\varphi(n_2).$

Next we show surjectivity and injectivity of $\varphi.$ For injectivity, we check the kernel: $\varphi(n)=(0,0)$ would require
\[\gamma(n)=0\qquad\text{and}\qquad g(n)=0.\]
The fact that $g(n)=0$ implies $n\in\ker g=\im f$ by exactness. Then $n\in\im f$ promises $m\in M$ such that $n=f(m),$ but now we see $\gamma(n)=(\gamma\circ f)(m)=m$ by definition of $\gamma.$ Thus, $\gamma(n)=0$ requires $m=0,$ and $n=f(m)=0$ as well. Thus, our kernel is trivial.

As for surjectivity, fix a pair $(m,p)\in M\oplus P$ so that we need to find $n\in N$ such that $\gamma(n)=m$ and $g(n)=p.$ Here we use the intuition we built a while ago: $g(n)=p$ is equivalent to $n\in n_0+\ker g$ for any $n_0\in N$ such that $g(n_0)=p.$ (Such an $n_0$ exists because $g$ is surjective.) Indeed,
\[g(n-n_0)=g(n)-g(n_0)=p-p=0,\]
so $n-n_0\in\ker g=\im f.$ Now that we have an element in $\im f,$ we can get a bearing on the behavior of $\gamma.$ Pick up $m_0$ such that $f(m_0)=n-n_0,$ and it follows
\[m_0=(\gamma\circ f)(m_0)=\gamma(n)-\gamma(n_0).\]
Because we want $\gamma(n)=m,$ we had better have $m_0=m-\gamma(n_0).$ So in total, we have
\[n:=n_0+f(m-\gamma(n_0))\]
for any $n_0$ such that $g(n_0)=p.$ We can quickly verify that applying $g$ to this gives $g(n)=g(n_0)=p$ because $f(m-\gamma(n_0))\in\im f=\ker g.$ Further, applying $\gamma$ to this gives $\gamma(n_0)+m-\gamma(n_0)=m$ because $\gamma\circ f=\id_M.$

Thus, we have verified that $\varphi:N\to M\oplus P$ is an $R$-module isomorphism. It remains to check the conditions on $\varphi.$ Well, for any $m\in M,$
\[(\varphi\circ f)(m)=\big(\gamma(f(m)),g(f(m))\big)=(m,0),\]
where $\gamma\circ f=\id_M$ by construction of $\gamma$ and $g\circ f=0$ by exactness. As for $g\circ\varphi^{-1},$ we remark that we have constructed our $\varphi^{-1}$ above in the proof of surjectivity. So fix some $(m,p)$ and pick up the $n$ for which $\varphi(n)=(m,p)$ so that $\varphi^{-1}((m,p))=n.$ Then
\[g(n)=\op{pr}_2\big(\gamma(n),g(n)\big)=\op{pr}_2(\varphi(n))=p.\]
This is the last condition on $\varphi$ we had to check, so this finishes the proof. $\blacksquare$

In essence, the above two propositions say that a short exact sequence where there is hope of reversing the maps turn out to have pretty strong ability to reversetheir maps because they're just $0\to M\to M\oplus P\to P\to 0.$ Perhaps this is disappointing as this short exact sequence is somewhat boring, but it is also quite surprising that we have this power-up.

Anyways, we close out our narrative that short exact sequences give size conditions on the middle term by proving the short five lemma.
\begin{proposition}
    Fix $R$ a ring, and suppose we have the following commutative diagram of $R$-modules with exact rows.
    \begin{center}
        \begin{tikzcd}
        	& {M_0} & {M_1} & {M_2} & 0 \\
        	0 & {N_0} & {N_1} & {N_2}
        	\arrow["{\partial_M^0}", from=1-2, to=1-3]
        	\arrow["{\partial_M^1}", two heads, from=1-3, to=1-4]
        	\arrow[from=1-4, to=1-5]
        	\arrow[from=2-1, to=2-2]
        	\arrow["{\partial_N^0}"', hook, from=2-2, to=2-3]
        	\arrow["{\partial_N^1}"', from=2-3, to=2-4]
        	\arrow["{\varphi_0}", from=1-2, to=2-2]
        	\arrow["{\varphi_1}", from=1-3, to=2-3]
        	\arrow["{\varphi_2}", from=1-4, to=2-4]
        \end{tikzcd}
    \end{center}
    If $\varphi_0$ and $\varphi_2$ are isomorphisms, then $\varphi_1$ is also an isomorphism.
\end{proposition}
We can do this as a direct application of the Snake lemma, and we will do so momentarily, but for now, it feels significantly more concrete to do this directly. I feel fuzzy about the maps in the Snake lemma anyways, so we can view this proof as ``evidence'' that the snake lemma is really correct.

We merely have to prove that $\varphi_1$ has trivial kernel and is surjective. This comes down to a diagram chase. Suppose we have $m_1\in\ker\varphi_1.$ Then, using the commutativity of the diagram,
\[\left(\varphi_2\circ\del_M^1\right)(m_1)=\left(\del_N^1\circ\varphi_1\right)(m_1)=0,\]
so $\del_M^1(m_1)\in\ker\varphi_2.$ However, $\varphi_2$ has trivial kernel, so $\del_M^1(m_1)=0,$ so $m_1\in\ker\del_M^1=\im\del_M^0$ by exactness, which promises $m_0\in M_0$ such that $m_1=\del_M^0(m_0).$ Using the commutativity of the diagram again,
\[\left(\del_N^0\circ\varphi_0\right)(m_0)=\left(\varphi_1\circ\del_M^0\right)(m_0)=\varphi_1(m_1)=0.\]
However, $\del_N^0$ is injective, so we must have $\varphi_0(m_0)=0,$ and $\varphi_0$ is injective as well, so $m_0=0.$ It follows $m_1=\del_M^0(m_0)=0,$ so indeed, $\varphi_1$ has trivial kernel. Note that we have used the injectivity of $\del_N^0$ here but not the surjectivity of $\del_M^1.$

Now we show the surjectivity of $\varphi_1$ and hope to see the surjectivity of $\del_M^1$ appear. Fix any $n_1\in N_1$ which we need to be in the image of $\varphi_1.$ Well, note $\varphi_2$ and $\del_M^1$ are both surjective, so we can pull $\del_N^1(n_1)\in N_2$ back to $M_2$ through $\varphi_2$ and then back to $M_1$ through $\del_M^1.$ So we have $m_1\in M_1$ such that
\[\left(\del_N^1\circ\varphi_1\right)(m_1)=\left(\varphi_2\circ\del_M^1\right)(m_1)=\del_N^1(n_1).\]
This $m_1$ is not the $m_1$ we want, however. After all, $m_1$ is not even canonically chosen because the pull back along $\del_M^1$ only gives a canonical coset $m_1+\ker\del_M^1.$ Anyways, what we do know is that
\[\del_N^1\big(n_1-\varphi_1(m_1)\big)=0,\]
so $n_1-\varphi_1(m_1)\in\ker\del_N^1=\im\del_N^0$ by exactness. Thus, we can pull $n_1-\varphi_1(m_1)$ backwards along $\del_N^0$ to $N_0$ (we no longer care that this is injective, but it is), and then the surjectivity of $\varphi_0$ lets us pull back to $M_0.$ So we get $m_0\in M_0$ such that
\[\left(\varphi_1\circ\del_M^0\right)(m_0)=\left(\del_N^0\circ\varphi_0\right)(m_0)=n_1-\varphi_1(m_1).\]
The above rearranges to $n_1=\varphi_1\left(\del_N^0(m_0)+m_1\right)\in\im\varphi_1.$ This completes the proof of surjectivity and hence the proof. $\blacksquare$

As promised, we now prove the short five lemma again by the Snake lemma. The Snake lemma provides us a long exact sequence
\[\ker\varphi_0\to\ker\varphi_1\to\ker\varphi_2\to\coker\varphi_0\to\coker\varphi_1\to\coker\varphi_2.\]
We remark that $\varphi:M\to N$ is an isomorphism if and only if it has trivial kernel so that $\ker\varphi=0$ and full image so that $\coker\varphi=N/\varphi(M)=0.$ So the fact that $\varphi_0$ and $\varphi_2$ are isomorphisms lets us fill this in as
\[0\to\ker\varphi_1\to0\to0\to\coker\varphi_1\to0.\]
Now, we state the following ``fill in the blank'' lemma.
\begin{lemma}
    Fix a $R$ a ring. If $0\to M\to 0$ is an exact sequence of $R$-modules, them $M=\{0\}$ only has its zero element.
\end{lemma}
The image of $0\to M$ must be equal to the kernel of $M\to 0$ for this to be exact. Well, $0\to M$ has image $0\in M$ and $M\to 0$ takes everything to $0$ and hence has kernel $M.$ It follows that $M=\{0\}.$ This is what we wanted. $\blacksquare$

In light of the above, $0\to\ker\varphi_1\to0$ forces $\ker\varphi_1=0,$ and $0\to\coker\varphi_1\to0$ forces $\coker\varphi_1=0.$ Thus, $\varphi_1$ is indeed an isomorphism, and we are done here. $\blacksquare$

We could be more careful about exactly what conditions are needed in the above proof. For example, injectivity of $\varphi_1$ only required $\varphi_0,\varphi_2,\del_N^0$ to be injective, and surjectivity of $\varphi_1$ only required $\varphi_0,\varphi_2,\del_M^1$ to be surjective. This doesn't really matter but should be remarked.

Anyways, the point of bringing up the short five lemma at all is that our exact sequences essentially forced an isomorphism in the middle term. The rows are not necessarily short exact sequences, but even still, we can feel the lower-bound of injectivity and the upper-bound of surjectivity along the left and right forcing the middle column to behave in a particular way.

\subsubsection{June 23rd}
Today I learned (and reviewed) some proofs of the Sylow theorems that I like. These all come from group actions, and many from the $p$-group fixed-point theorem. To not be too bogged down in details (as I usually am), we often won't show that these group actions are actually actions. That kind of rigor is not the point today.

To avoid digressions in what follows, we review the necessary theory at the start, knowingly sacrificing motivation. Here is the orbit-stabilizer theorem.
\begin{proposition}
    Let $G$ be a group acting on a set $S.$ Given $s\in S,$ elements of the orbit $Gs$ are in bijection with cosets in $G/\op{Stab}(s),$ where $\op{Stab}(s)$ is the stabilizer.
\end{proposition}
We send a coset $g\op{Stab}(s)$ to the element $gs$ in the orbit. This mapping is well-defined because $g\op{Stab}(s)=h\op{Stab}(s)$ if and only if $h^{-1}g\in\op{Stab}(s)$ if and only if $\left(h^{-1}g\right)(s)=s$ if and only if $gs=hs.$

To show the bijection, we note the map is surjective because we can hit $gs\in Gs$ with the coset $g\op{Stab}(s).$ And the map is injective because we showed $gs=hs$ if and only if $g\op{Stab}(s)=h\op{Stab}(s)$ above. This finishes the proof. $\blacksquare$

This gives us the $p$-group fixed-point theorem.
\begin{theorem}
    Let $G$ be a $p$-group acting on a set $S.$ Then the number of fixed points $\op{Fix}(S)$ satisfies
    \[\#\op{Fix}(S)\equiv\#S\pmod p.\]
\end{theorem}
This is done by writing down the class equation. Note that orbits partition our set $S$ because being in the same orbit is an equivalence relation. This lets us write
\[\#S=\sum_{Gs\text{ orbit}}\#(Gs).\]
If the size of the orbit is $1,$ then $Gs=\{s\}$ is a fixed point, so we can remove all these points and write
\[\#S=\op{Fix}(S)+\sum_{\substack{Gs\text{ orbit}\\\#(Gs)>1}}\#(Gs).\]
It remains to show that the sum on the right-hand side is divisible by $p$; we show that all terms in the sum are divisible by $p.$

Indeed, the size of the orbit $Gs$ is equal to the number of cosets in $G/\op{Stab}(s),$ which is equal to $[G:\op{Stab}(s)]>1.$ This index divides the size of the total group $G,$ so it is a power of $p.$ Powers of $p$ larger than $1$ are divisible by $p,$ so we are done. $\blacksquare$

A few of our group actions will make the normalizer of a subgroup appear, so it is probably safest to define it at the outset so that we aren't surprised when it comes up.
\begin{definition}
    Fix $G$ a group and $H$ a subgroup. We define the normalizer $N(H)$ as
    \[N(H):=\left\{g\in G:gHg^{-1}=H\right\}\]
\end{definition}
It is not difficult to show that $N(H)$ is in fact a subgroup. Indeed, it is closed under inversion because $gHg^{-1}=H$ impies $H=g^{-1}Hg,$ and it is closed under multiplication because $g_1Hg_1^{-1}=g_2Hg_2^{-1}=H$ implies
\[(g_1g_2)H(g_1g_2)^{-1}=g_1g_2Hg_2^{-1}g_1^{-1}=g_1Hg_1^{-1}=H.\]
Anyways, what we like about the normalizer is that $H$ is a subgroup of $N(H)$ ($H$ is closed), and in fact a normal subgroup. Indeed, for any $g\in N(H),$ we see $gHg^{-1}=H,$ which is what we need for $H$ to be normal in $N(H).$

Before we continue with the proofs of the Sylow theorems, we codify the following intuition.
\begin{idea}
    Our purely group-theoretic actions are mostly one of the following. (I.e., these are the ``best'' ways for a group to act in group theory.)
    \begin{itemize}
        \item A group or subgroup acting by multiplication on cosets.
        \item A group or subgroup acting by conjugation on some subgroups.
        \item A group or subgroup acting naturally on a more exotic set (such as tuples) by multiplication.
    \end{itemize}
    We say explicitly that we don't typically have a group or subgroup acting naturally on exotic sets by conjugation because this tends to make our action useless for abelian groups.
\end{idea}

We are now ready to talk about the Sylow theorems. We will prove them in sequence; here is the first Sylow theorem, which asserts existence.
\begin{theorem}
    Let $G$ be a finite group and $p$ a prime so that $\#G=p^\nu m$ with $p\nmid m.$ Then $G$ has a subgroup of order $p^\nu.$ Such a subgroup is called a Sylow $p$-subgroup.
\end{theorem}
It is not easy, but it is possible, to prove this directly using group actions. We will start with a relatively unmotivated approach and then return to prove this in a more motivated way. It might be productive to not read this proof at all.

Let $S$ be the set of all subsets of $G$ of order $p^\nu.$ The idea here is that we would like to use a group action to ``detect'' a subgroup out of $S.$ That is not exactly what is about to happen, but it explains why we might consider such a thing. Because this set $S$ is somewhat exotic, we will have $G$ act on it by left multiplication.

Very quickly, this action is well-defined because multiplication by $g\in G$ is a bijection. Namely, we need to know $\#gA=\#A$ for $A\in S$ to have $gA\in S,$ and indeed, the maps $a\mapsto ga$ and $a\mapsto g^{-1}a$ are inverses of each other and therefore each bijections. We will not say more than this.

Our action is not by a $p$-group, so we cannot use the $p$-group fixed-point theorem, but we can say something similar. Note that
\[\#S=\binom{p^\nu m}{p^\nu}\equiv\binom m1=m\not\equiv0\pmod p\]
by Lucas's theorem on binomial coefficients; see \href{https://en.wikipedia.org/wiki/Lucas's_theorem}{here} or even \href{https://dfoiler.github.io/TIL/2020/9/17/}{here} for a reference, but we won't prove it now because it would take us too far afield. The point of saying is that when we partition $S$ into orbits
\[\#S=\sum_{Gs\text{ orbit}}\#(Gs),\]
we cannot have all of the orbits have size divisible by $p.$ So in a style akin to the $p$-group fixed-point theorem, we get to pick up a set $A\in S$ whose orbit $G\cdot A$ has size not divisible by $p.$

We said above that we would like to detect a Sylow $p$-subgroup directly from our action, but we cannot quite do that because $A$ might, for example, be the coset of a Sylow $p$-subgroup by merely living in the orbit of a bona fide Sylow $p$-subgroup. There are coarse ways of dealing with this, but to use $p\nmid\#(G\cdot A),$ we note
\[\#(G\cdot A)\cdot\#\op{Stab}(A)=\#G=p^\nu m\]
by the orbit-stabilizer theorem. Because $\#A$ has lots of $p$s, it is reasonable (a la Lagrange) to expect $\#(G\cdot A)$ to have few $p$s, and indeed it has no $p$s by construction of $A.$ This forces $\op{Stab}(A)$ to have lots of $p$s, in fact all of them, so $p^\nu\mid\#\op{Stab}(A).$ This indicates that we want to show $\op{Stab}(A)$ is Sylow $p$-subgroup!

So far we have used the size of the group $G$ as well as some number theory to show that $\#\op{Stab}(A)$ is divisible by $p^\nu.$ To show that $\#\op{Stab}(A)$ is actually equal to $p^\nu,$ it remains to use the group action in a meaningful way. We only know what $\op{Stab}(A)$ does on $A,$ namely it will take any $a\in A$ somewhere else in $A.$ This tells us
\[\op{Stab}(A)\cdot a\subseteq A\]
for a given $a\in A.$ We would like to say that this means $\#\op{Stab}(A)\le\#A,$ and indeed, if $g,h\in\op{Stab}(A)$ with $ga=ha,$ then right multiplication by $a^{-1}$ implies $g=h.$ So the multiplication by $\op{Stab}(A)$ sends $a$ to a different element each time, so we get
\[\#\op{Stab}(A)=\#\big(\op{Stab}(A)\cdot a\big)\le\#A=p^\nu.\]
Because $p^\nu\mid\#\op{Stab}(A)$ already, it follows $\op{Stab}(A)$ has size $p^\nu$ and is our Sylow $p$-subgroup. $\blacksquare$

Above we said around the middle that a barrier to showing $A$ is the desired Sylow $p$-subgroup is that $A$ might be a coset of a Sylow $p$-subgroup. We say now that the end of our proof shows that this is exactly the barrier: we showed
\[\op{Stab}(A)\cdot a\subseteq A,\]
and because we established both sides have size $p^\nu,$ it follows $\op{Stab}(A)\cdot a=A.$ Because $\op{Stab}(A)$ was our Sylow $p$-subgroup, we see that, indeed, $A$ was a (right) coset of a Sylow $p$-subgroup all along.

The above proof was overly clever, and though I like it, it is not using any of the more standard group actions I am trying to get used to. So we present a second proof, which proves the result through the following lemma.
\begin{lemma}
    Let $G$ be a finite group and $p$ a prime so that $\#G=p^\nu$ with $p\nmid m.$ If $H$ is a $p$-subgroup with $\#H=p^\alpha$ and $\alpha<\nu,$ then $H$ is a subgroup of a larger $p$-subgroup $H'\subseteq G$ of order $p^{\alpha+1}.$
\end{lemma}
This time we want to use the $p$-group fixed-point theorem, and the only $p$-group immediately at our disposal is $H.$ It needs to act on something, and for this proof, it will act by multiplication on its cosets $G/H.$ Now, the $p$-group fixed-point theorem gives
\[\#\op{Fix}(G/H)\equiv[G:H]\pmod p.\]
The next step is to talk about what it means to be a fixed point under this action. Well, a coset $gH$ is a fixed point under multiplication by $H$ if and only if
\[hgH=gH\]
for any $h\in H.$ This is equivalent to $g^{-1}hgH=H,$ or $g^{-1}hg\in H.$ But this simply means that $g^{-1}Hg\subseteq H,$ and because $x\mapsto g^{-1}x\mapsto g^{-1}xg$ is a composite of bijections, we see $\#\left(g^{-1}Hg\right)=\#H,$ so in fact this is equivalent to
\[g^{-1}Hg=H.\]
(Of course $g^{-1}Hg=H$ implies $g^{-1}Hg\subseteq H.$) So this is equivalent to $g^{-1}\in N(H)$ and $g\in N(H).$ (Here is our normalizer, as promised.)

This classifies our fixed points. Namely, our fixed points are cosets $gH$ where $g\in N(H),$ which is just the cosets in $N(H)/H.$ So the fixed-point theorem actually asserts
\[\#(N(H)/H)\equiv0\pmod p.\]
Because $N(H)/H$ has order divisible by $p,$ we may pick up a coset $gH$ of order $p$ by Cauchy's theorem. (We will prove Cauchy's theorem after this, but we will not pause to do so.)

It remains to construct the needed $H'.$ We take $H$ and simply extend it by $g$ to get $\langle g\rangle H$; this can be viewed set-theoretically or as a union of cosets. We get $H\subseteq H'$ for free. Further, $H'$ is a subgroup because, for any $g^ah_1,g^bh_2\in H,$ we see
\[g^ah_1\cdot\left(g^bh_2\right)^{-1}\in g^aHg^{-b}=g^{a-b}H\subseteq H',\]
where $g^bHg^{-b}=H$ because $g^b\in N(H).$ (This is a manifestation of $H$ being normal in $N(H).$) Lastly, $H'$ has the desired order because all the cosets $g^\bullet H$ are disjoint, and there are $p$ of them by construction of $g,$ implying $\#H'=p\cdot\#H.$ This finishes the proof. $\blacksquare$

We promised to show Cauchy's theorem, so we quickly do that now. We present a group actions proof, which aligns somewhat with our intuition.
\begin{theorem}
    Fix $G$ a finite group and $p$ a prime so that $p\nmid\#G.$ Then the number of elements of $G$ of order $p$ is $-1\pmod p.$ In particular, it is nonzero.
\end{theorem}
The equivalence in the statement makes us want to use the $p$-group fixed-point theorem. The action, which comes pretty much out of nowhere, is to have $\ZZ/p\ZZ$ act on the tuples
\[S:=\left\{(g_k)_{k=0}^{p-1}\in G^p:g_0\cdot\ldots\cdot g_{p-1}=e\right\}\]
by cycling. Namely, we define
\[k\cdot(g_0,\ldots,g_{p-1}):=(g_k,\ldots,g_{p-1+k}),\]
where indices are taken$\pmod p.$ (Our intuition said that we should act by multiplication in a natural way, and this is close to that.) Now that we have this action, the proof follows quickly.

The $p$-group fixed-point theorem gives
\[\#\op{Fix}(S)\equiv\#S=(\#G)^p\equiv0\pmod p.\]
The next step, as before, is to talk about what a fixed point under this action is. Well, such a tuple $(g_0,\ldots,g_{p-1})$ must have, for any $k\in\ZZ/p\ZZ,$
\[(g_0,\ldots,g_{p-1})=k\cdot(g_0,\ldots,g_{p-1})=(g_k,\ldots,g_{p-1+k}),\]
so all entries are equal to $g_0.$ And indeed, if all entries are equal, then, $g_\ell=g_{\ell+k}$ for any $k$ or $\ell,$ implying the above holds, and we have a fixed point. Noting the condition on $S,$ we see that we have a fixed point $(g,\ldots,g)$ if and only if
\[g^p=e.\]
Taking a pause for a second, we see that
\[\#\left\{g\in G:g^p=e\right\}\equiv0\pmod p\]
from the work we've done so far. Now, a $g$ lives in the left-hand set if and only if its order divides $p,$ which means that the order is either $1$ or $p.$ Order $1$ only occurs for $g=e,$ so the remaining elements all have order $p.$ Subtracting the $g=e$ case from both sides of the above gives the desired statement. $\blacksquare$

Anyways, we now use the lemma to prove the first Sylow theorem.
\begin{theorem}
    Let $G$ be a finite group and $p$ a prime so that $\#G=p^\nu m$ with $p\nmid m.$ Then $G$ has a subgroup of order $p^\nu.$
\end{theorem}
We show this by induction, where the lemma is the inductive step. Intuitively, the lemma lets us start with the $p$-subgroup $H=\langle e\rangle,$ apply the given algorithm $\nu$ times, and end up with a subgroup of the needed order $p^\nu.$

To codify this, we show that we can construct a $p$-subgroup of every order $p^\alpha$ from $\alpha=0$ to $\alpha=\nu,$ by induction. For $\alpha=0,$ we choose $H=\langle e\rangle$ as our base case. For the inductive step, we may assume that we have a $p$-subgroup $H$ of order $p^\alpha$ with $\alpha+1\le\nu,$ and we want to construct $H'$ of order $p^{\alpha+1}.$ But this is exactly what the lemma does, so we are done. $\blacksquare$

While we are here, we remark that the lemma can actually show that Sylow $p$-subgroups are maximal, which is nice.
\begin{proposition}
    Let $G$ be a finite group and $p$ a prime so that $\#G=p^\nu m$ with $p\nmid m.$ Then any $p$-subgroup $H$ is contained in a Sylow $p$-subgroup.
\end{proposition}
Again, we do induction, where the intuition is that the lemma lets us successively build up a Sylow $p$-subgroup from any $p$-subgroup base. Letting $\#H=p^{\nu-k},$ we prove this by induction on this $k$ as long as $k\le\nu.$ Our base case is $k=0,$ for which $H$ is immediately a Sylow $p$-subgroup.

As for the inductive step, we suppose that all $p$-subgroups of order $p^{\nu-k}$ are contained in a Sylow $p$-subgroup, and we show that all subgroups of order $p^{\nu-(k+1)}$ are contained in a Sylow $p$-subgroup, given $k+1\le\nu.$ Indeed, pick up any $H$ with $\#H=p^{\nu-(k+1)},$ and the lemma provides a $p$-subgroup $H'$ with $H\subseteq H'$ and
\[\#H'=p^{\nu-(k+1)+1}=p^{\nu-k}.\]
Now, $H'$ is contained in a Sylow $p$-subgroup by the inductive hypothesis, so $H$ is also. This completes the induction $\blacksquare$

With a vaguely more standard proof of the first Sylow theorem in hand, we now turn to the second and third Sylow theorems. Here is the second.
\begin{theorem}
    Fix $G$ a finite group and $p$ a prime. Then all Sylow $p$-subgroups of $G$ are conjugate.
\end{theorem}
It might be tempting to let $G$ act on the Sylow $p$-subgroups by conjugation, but that will be an action helpful later. For now, we are still attached to the $p$-group fixed-point theorem, so we want a $p$-group to act.

Instead, our action will be one of the other named ones. We recall from the proof of the lemma that having a subgroup $H$ acting on cosets of $H$ made the condition $g^{-1}Hg=H$ appear as our fixed point condition, which is similar to what we want here: it shows two subgroups are conjugate (though $H$ and $H$ are conjugate trivially, of course).

Regardless, we charge ahead. Pick up two Sylow $p$-subgroups of $G$ named $P$ and $Q,$ and we will have $P$ act on the cosets $G/Q$ by multiplication. Now, $P$ is a $p$-group, so the $p$-group fixed-point theorem asserts
\[\#\op{Fix}(G/Q)\equiv[G:Q]\pmod p.\]
Because $[G:Q]=|G|/|Q|$ is not divisible by $p$ ($Q$ is Sylow), the above equivalence gives us a fixed point. As we saw in Cauchy's theorem, this is a more common use of the $p$-group fixed-point theorem: nonzero$\pmod p$ implies nonzero.

Anyways, we let $gQ$ be the fixed point under the action of $P.$ Arguing as we did in the lemma, this means that
\[pgQ=gQ\]
for any $p\in P,$ which is equivalent to $g^{-1}pg\in Q$ for any $p\in P,$ which is equivalent to $g^{-1}Pg\subseteq Q.$ But again, $x\mapsto g^{-1}xg$ is a bijection, so $g^{-1}Pg$ has the same size as $P,$ which has the same size as $Q$ because all Sylow $p$-subgroups have the same size. So in fact
\[g^{-1}Pg=Q.\]
It follows $P=gQg^{-1},$ so $P$ and $Q$ are indeed conjugate. This is what we wanted. $\blacksquare$

We can see the second Sylow theorem as a ``power-up'' for our group actions. In particular, it is now much more meaningful to have a group or subgroup act on the group's Sylow $p$-subgroups by conjugation. Not only is this well-defined, but when the full group does it, this action is transitive by the second Sylow.

Anyways, this thinking suggests two kinds of actions we might use.
\begin{itemize}
    \item We let a $p$-subgroup (more specifically, a Sylow $p$-subgroup) act on the Sylow $p$-subgroups by conjugations and hope to use the $p$-group fixed-point theorem.
    \item We have all $G$ act on the Sylow $p$-subgroups by conjugation and hope to use the transitivity.
\end{itemize}
Each of these will give a part of the third Sylow theorem. We take these in sequence.
\begin{proposition}
    Fix $G$ a finite group and $p$ a prime, and let $n_p$ be the number of Sylow $p$-subgroups. Then $n_p\equiv1\pmod p.$
\end{proposition}
The$\pmod p$ in the statement suggests we want to use the $p$-group fixed-point theorem, so we hope to use the first of the suggested actions. Pick up a Sylow $p$-subgroup $P$ and have it act on the set of Sylow $p$-subgroups $S_p$ by conjugation. We established in the proof of the second Sylow theorem that conjugation does not change the size of the subgroup, so this action is safe.

Now, the $p$-group fixed-point theorem gives
\[\#\op{Fix}(S_p)\equiv\#S_p=n_p\pmod p.\]
Thus, it remains to show that $\#\op{Fix}(S_p)\equiv1\pmod p$; in fact, we show that $\#\op{Fix}(S_p)=1.$ As usual, our next step is to examine the fixed points.

For this, we have to examine what it means for a particular Sylow $p$-subgroup $Q\in S_p$ to be a fixed point. Our action is conjugation, which means that
\[pQp^{-1}=Q\]
for all $p\in P.$ This is equivalent to having $p\in N(Q)$ for each $p\in P,$ or simply $P\subseteq N(Q).$ Of course, $P\subseteq N(P),$ so $P$ is one option for a fixed point, so we now have to show that $P$ is the only option.

This requires some cleverness. Note that $\#N(Q)\mid\#G$ by Lagrange's theorem, so the largest power of $p$ dividing $N(Q)$ is upper-bounded by the largest power of $p$ dividing $G.$ But $Q\subseteq N(Q)$ provides a lower bound as the largest power of $p$ dividing $G,$ so in fact $Q$ is a Sylow $p$-subgroup of $N(Q),$ and in fact so is $P\subseteq N(Q).$

The kicker is that $Q$ is normal in $N(Q),$ and by the second Sylow theorem, all Sylow $p$-subgroups are conjugate, so $P$ is conjugate to $Q,$ giving $P=Q.$ Thus, $P$ is the only fixed point of our action, and we are done. $\blacksquare$

And the other action is featured as follows.
\begin{proposition}
    Fix $G$ a finite group and $p$ a prime so that $G=p^\nu m$ with $p\nmid m.$ Further, let $n_p$ be the number of Sylow $p$-subgroups. Then $n=[G:N(P)]$ for any Sylow $p$-subgroup $P,$ and $n_p\mid m.$
\end{proposition}
This time around we let $G$ act on the set $S_p$ of all Sylow $p$-subgroups by conjugation. Again, this is well-defined because conjugation does not change order.

This action is transitive by the second Sylow theorem. Indeed, if $P$ and $Q$ are Sylow $p$-subgroups, then $P=gQg^{-1}$ for some $g\in G$ because $P$ and $Q$ are conjugate. This means that the orbit of $Q$ includes $P,$ giving our transitivity.

Now, we don't have access to the $p$-group fixed-point theorem because $G$ isn't a $p$-group, so we instead use the orbit-stabilizer theorem for our counting. Fix a Sylow $p$-subgroup $P,$ and we know
\[\#(G\cdot P)=[G:\op{Stab}(P)].\]
We just established that the action is transitive, so $\#(G\cdot P)=\#S_p=n_p.$ As for $\op{Stab}(P),$ this is the set of $g$ such that
\[gPg^{-1}=P,\]
which is exactly what we need for $g\in N(P).$ Thus, $\op{Stab}(P)=N(P),$ and we do indeed have $n_p=[G:N(P)].$

It remains to show that $n_p\mid m.$ Well, $P\subseteq N(P),$ so $\#P\mid\#N(P),$ and dividing from $\#G$ gives
\[n_p=[G:N(P)]\mid[G:P].\]
But now $[G:P]=|G|/|P|=p^\nu m/p^\nu=m.$ This gives $n_p\mid m,$ completing the proof. $\blacksquare$

We quickly remark that a lot of the above proof doesn't need to be about Sylow $p$-subgroups. If we fix $P$ as any subgroup and then let $S_p$ be the set of $P$'s conjugate subgroups, then we still get $\#S_p=[G:N(P)]$ following the same logic. We will not flesh this out here.

The above two propositions together make the third Sylow theorem, so we have proved all three Sylow theorems. I think that's enough for today.

\subsubsection{June 24th}
Today I learned about the Wronskian. To not bury the lead, here is the definition.
\begin{definition}
    Suppose the functions $f_0,\ldots,f_{n-1}$ have $n-1$st derivatives over a domain $D.$ Then we define the Wronskian by
    \[W(f_0,\ldots,f_{n-1}):=\det\begin{bmatrix}
        f_0^{(0)} & \cdots & f_{n-1}^{(0)} \\
        \vdots & \ddots & \vdots \\
        f_0^{(n-1)} & \cdots & f_{n-1}^{(n-1)}
    \end{bmatrix}.\]
\end{definition}
The idea here is to detect linear independence of some functions $f_0,\ldots,f_{n-1}.$ For the sake of concreteness, we will make $f_\bullet:D\to\RR$ for fixed $D\subseteq\RR,$ though this is not entirely necessary for the discussion. Observe that if the $f_\bullet$ are linearly dependent, then we may resurrect constants $\{c_k\}_{k=0}^{n-1}$ not all zero so that
\[c_0f_0(t)+\cdots+c_{n-1}f_{n-1}(t)=0\]
for any suitable $t\in D.$

Now, supposing that the $f_\bullet$ have $n-1$ derivatives on the domain $D,$ the key observation is that we may take the $k$th derivative of both sides with respect to $t,$ for $k\le n-1.$ So in fact we get a full system of equations
\[c_0f_0^{(k)}(t)+\cdots+c_{n-1}f_{n-1}^{(k)}(t)=0\]
for any $t\in D.$ Collecting all of these equations for $0\le k\le n-1$ into a matrix, we get
\[\begin{bmatrix}
    f_0^{(0)}(t) & \cdots & f_{n-1}^{(0)}(t) \\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)}(t) & \cdots & f_{n-1}^{(n-1)}(t) \\
\end{bmatrix}\begin{bmatrix}c_0 \\ \vdots \\ c_{n-1}\end{bmatrix}=\begin{bmatrix}0 \\ \vdots \\ 0\end{bmatrix},\]
still for any $t\in D.$ Because the $c_\bullet$ are not all $0,$ we see that this provides a nontrivial vector in the null space, implying that
\[W(f_0,\ldots,f_{n-1})=\det\begin{bmatrix}
    f_0^{(0)}(t) & \cdots & f_{n-1}^{(0)}(t) \\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)}(t) & \cdots & f_{n-1}^{(n-1)}(t) \\
\end{bmatrix}=0\]
because now the matrix is singular. Writing these ideas down, we have the following.
\begin{proposition}
    Suppose the functions $f_0,\ldots,f_{n-1}:D\to\RR$ for some $D\subseteq\RR$ have $n-1$st derivatives on $D.$ If they are linearly dependent, then
    \[W(f_0,\ldots,f_{n-1})=0\]
    over all $D.$
\end{proposition}
This follows from the above discussion. $\blacksquare$

The converse of the above statement is not necessarily true. The classic example is as follows.
\begin{example}
    Take $f_0(x)=x^2$ and $f_1(x)=x|x|.$ Then we only care about first derivatives, which both of these functions have as $f_0'(x)=2x$ and $f_1'(x)=x\op{sgn}(x)+|x|=2|x|.$ Then
    \[W\left(x^2,x|x|\right)=\det\begin{bmatrix}
        x^2 & x|x| \\
        2x & 2|x|
    \end{bmatrix}=2x^2|x|-2x^2|x|=0.\]
\end{example}
However, for any domain $D$ containing a neighborhood around $0,$ the functions $f_0$ and $f_1$ are not linearly dependent. Indeed, we have
\[\begin{cases}
    f_0(t) + f_1(t) = 0 & t\le0, \\
    f_0(t) - f_1(t) = 0 & t\ge0,
\end{cases}\]
and these cannot be stitched together to a single equation for all $t.$ Explicitly, if $t\in D$ is negative, then $c_0f_0(t)+c_1f_1(t)=0$ with $c_0\ne0$ (without loss of generality) implies
\[\frac{c_1}{c_0}=-\frac{f_0(t)}{f_1(t)}=-\frac{t^2}{-t^2}=1,\]
so $c_1=c_0.$ However, if $t\in D$ is positive, then $c_0f_0(t)+c_1f_1(t)=c_0t^2$ is nonzero. So it is impossible to give constants $c_0$ and $c_1$ such that
\[c_0f_0+c_1f_1=0\]
that holds for the entire domain $D.$

It is not clear exactly what the problem is. Speaking in more generality, having
\[W(f_0,\ldots,f_{n-1})=\det\begin{bmatrix}
    f_0^{(0)} & \cdots & f_{n-1}^{(0)}\\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)} & \cdots & f_{n-1}^{(n-1)}
\end{bmatrix}=0\]
over a domain $D$ does not immediately give a vector $\langle c_0,\ldots,c_{n-1}\rangle$ such that
\[\begin{bmatrix}
    f_0^{(0)} & \cdots & f_{n-1}^{(0)}\\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)} & \cdots & f_{n-1}^{(n-1)}
\end{bmatrix}\begin{bmatrix} c_0 \\ \vdots \\ c_{n-1} \end{bmatrix}=\begin{bmatrix} 0 \\ \vdots \\ 0\end{bmatrix}\]
still over the entire domain $D.$ The most immediate block here is that functions with $n-1$ derivatives over $D$ are not a field, so we cannot do the Gaussian elimination we would like. This is to say that the obvious the most obvious argument for the converse fails.

One way to view this block is via quantifiers. One possible fix is to take a particular $t_0\in D$ and then
\[\begin{bmatrix}
    f_0^{(0)}(t_0) & \cdots & f_{n-1}^{(0)}(t_0)\\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)}(t_0) & \cdots & f_{n-1}^{(n-1)}(t_0)
\end{bmatrix}\begin{bmatrix} c_0 \\ \vdots \\ c_{n-1} \end{bmatrix}=\begin{bmatrix} 0 \\ \vdots \\ 0\end{bmatrix}\]
does indeed have a solution vector. However, as $t_0$ varies, our vector span might change, as is seen in the given example when $t$ cross over $0.$

Here is a clean way to lift the vector from a singele $t_0$ to all of the domain $D,$ taken from \href{https://math.stackexchange.com/a/1680662/869257}{here}.
\begin{proposition}
    Suppose the functions $f_0,\ldots,f_{n-1}:D\to\RR$ for some open interval $D\subseteq\RR$ have $n-1$st derivatives on $D.$ Further add the condition that the $f_\bullet$ are solutions to the differential equation
    \[y^{(n)}(t)+\sum_{k=0}^{n-1}p_k(t)y^{(k)}=0\]
    for continuous functions $p_\bullet.$ Then if the Wronskian $W(f_0,\ldots,f_{(n-1)})=0$ for any particular $t_0\in D,$ then the $f_\bullet$ are linearly dependent. In particular, $W(f_0,\ldots,f_{n-1})=0$ for all $t\in D.$
\end{proposition}
The idea is to use uniqueness of solutions to the differential equation under initial conditions. As alluded to above, we note $W(f_0,\ldots,f_{(n-1)})=0$ at $t_0$ at a provides a nonzero vector $\langle c_0,\ldots,c_{n-1}\rangle$ such that
\[\begin{bmatrix}
    f_0^{(0)}(t_0) & \cdots & f_{n-1}^{(0)}(t_0)\\
    \vdots & \ddots & \vdots \\
    f_0^{(n-1)}(t_0) & \cdots & f_{n-1}^{(n-1)}(t_0)
\end{bmatrix}\begin{bmatrix} c_0 \\ \vdots \\ c_{n-1} \end{bmatrix}=\begin{bmatrix} 0 \\ \vdots \\ 0\end{bmatrix}.\]
Working backwards from our intuition for the Wronskian, we get constants $c_\bullet$ not all zero such that
\[c_0f_0^{(k)}(t_0)+\cdots+c_{n-1}f_{n-1}^{(k)}(t_0)=0\]
for $k\le n-1.$ We hope that this is enough information to force $c_0f_0+\cdots c_{n-1}f_{n-1}=0$ over all $D.$

The way to use the differential equation is to note that its solutions form a vector space: the derivative is a linear operator as is multiplying by a continuous function, so we are essentially asking for the kernel of some linear operator over the space of functions with $n-1$ derivatives on $D.$ Thus,
\[f(t):=c_0f_0(t)+\cdots+c_{n-1}f_{n-1}(t)\]
is also a solution to the given differential equation. The information from the Wronskian tells us
\[f(t_0)=f'(t_0)=\cdots=f^{(n-1)}(t_0)=0.\]
Viewing this as an initial-value problem for the differential equation, this $f$ is unique the solution to the differential equation with these derivatives vanishing. However, the zero function is one such function that satisfies the linear equation and has all of its first $n-1$ derivatives equal to $0,$ so we conclude $f\equiv 0.$

Thus,
\[c_0f_0+\cdots+c_{n-1}f_{n-1}=0\]
over all of $D,$ so we do indeed have linear dependence. Now, the previous proposition implies $W(f_0,\ldots,f_{n-1})=0$ over all $D,$ completing the proof. $\blacksquare$

Amusingly, we can read this proposition backwards to conclude that $f_0(t)=t^2$ and $f_1(t)=t|t|$ are not solutions to any single second-order linear differential equation on any $D$ containing an open neighborhood around $0,$ for their Wronskian vanishes despite their linear independence.

There is actually a cleverer to retrieve the end of the above proposition, via Abel's identity. We do this for second-order differential equations, though a similar statement holds in an arbitrary orders.
\begin{theorem}
    Suppose the functions $f_0,f_1:D\to\RR$ for some open interval $D\subseteq\RR$ have second derivatives. Further, add the condition that $f_1$ and $f_2$ are solutions to the differential equation
    \[y''+p(t)y'+q(t)y=0\]
    for continuous functions $p,q:D\to\RR.$ Then, for any fixed $t_0\in D,$ we have
    \[W(f_0,f_1)(t)=W(f_0,f_1)(t_0)\exp\left(-\int_{t_0}^tp(t)\,dt\right).\]
\end{theorem}
The idea is to create a differential equation that $W:=W(f_0,f_1)=f_0f_1'-f_0'f_1$ satisfies. To this end, observe
\[W'=\frac d{dt}(f_0f_1'-f_0'f_1)=(f_0f_1''+f_0'f_1')-(f_0'f_1'+f_0''f_1)=f_0f_1''-f_0''f_1.\]
Now, we know $f_0''=-pf_0'-qf_0$ and $f_1''=-pf_1'-qf_1,$ so we can write this as
\[W'=f_0(-pf_1'-qf_1)-(-pf_0'-qf_0)f_1=-pf_0f_1'+pf_0'f_1.\]
So, magically, we have that $W'=-pW.$

Fix the $t_0\in D$ as promised. We could solve $y'=-py$ by separation of variables, but we have potential problems if $W$ vanishes somewhere. Instead, we claim by uniqueness that
\[W(t)\stackrel?=f(t):=W(t_0)\exp\left(-\int_{t_0}^tp(t)\,dt\right).\]
Indeed, this gives $f(t_0)=W(t_0)\exp(-0)=f(t_0)$ as our first condition. Further, $f'$ is
\[W(t_0)\exp\left(-\int_{t_0}^tp(t)\,dt\right)\cdot-\frac d{dt}\int_{t_0}^tp(t)\,dt=-p(t)W(t_0)\exp\left(-\int_{t_0}^tp(t)\,dt\right),\]
which is indeed $f'=-pf.$ Thus, $f$ does indeed satisfy the differential equation $y'=-py,$ and we see $f'(t_0)=-p(t_0)f(t_0)=-p(t_0)W(t_0)=W'(t_0).$ So $f$ and $W$ both satisfy the same differential equation and have the same initial conditions, so we conclude $W\equiv f.$ This completes the proof. $\blacksquare$

What we see from this is that if $W(f_0,f_1)(t_0)\ne0$ for some particular $t_0\in D,$ then in fact $W(f_0,f_1)(t)\ne0$ for all $t\in D.$ Of course, vanishing Wronskian on the entire interval is still not enough to conclude linear independence, as seen by our example. However, Abel's identity is nice because we can compute the Wronskian without really computing the determinant in full generality.

\subsubsection{June 25th}
Today I learned a little more about polynomial statistics over finite fields. I did two things with these today. For the first, we recall the following result which we proved a while ago.
\begin{theorem}
    Fix $f(x)\in\ZZ[x]$ an irreducible polynomial. As $p$ varies over all primes, the expected number of roots of $f(x)\pmod p$ is $1.$ In other words,
    \[1=\lim_{N\to\infty}\frac1{\pi(N)}\sum_{p\le N}\#\{k\in\FF_p:f(k)\equiv0\pmod p\}.\]
\end{theorem}
This is an application of Chebotarev's theorem. As an outline, apply Dedekind-Kummer to turn factorization of $f(x)\pmod p$ into factorization of $(p)$ in $\QQ(\alpha)$ where $\alpha$ is a root of $f.$ Then we extend $\QQ(\alpha)$ to a Galois closure $M$ and provide $p$ a Frobenius $\varphi_p,$ where $\varphi_p$ is equidistributed by Chebotarev. Investigating roots of $f(x)\pmod p$ turns into counting certain double cosets in $\op{Gal}(M/\QQ),$ which is an algebra problem. $\blacksquare$

It is a valid question to ask what happens if, instead of fixing $f(x)$ and letting $p$ vary, we fix $p$ and let $f(x)$ vary. This will be one of our motivating questions. We still cannot look at all polynomials at once, so we have to bound the degree to $d$ and send $d\to\infty$ later.

Additionally, some care must be taken because there are still infinitely many polynomials in $\ZZ[x]$ of bounded degree, so we count polynomials in $\FF_p[x]$ instead. And if we're in $\FF_p,$ we might as well be in $\FF_q$ for some prime-power $q.$ So the well-defined version of our question is asking for
\[\lim_{d\to\infty}\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le d}}\#\{k\in\FF_q:f(k)=0\}\Bigg/\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le D}}1.\]
We let the denominator here be $N_d,$ the number of polynomials in $\FF_q[x]$ of degree less than or equal to $d.$ We can compute $N_d$ without too much effort, but we won't actually need to; we place this computation in a lemma.
\begin{lemma}
    Fix everything as above. Then $N_d=q^{d+1}.$
\end{lemma}
Note that a $f(x)\in\FF_q[x]$ of degree less than or equal to $d$ takes the form
\[f(x)=\sum_{k=0}^da_kx^k,\]
where each of the coordinates $a_\bullet\in\FF_q$ has $q$ options, with no restrictions. So the total number of possible polynomials is $q^{d+1}.$ $\blacksquare$

For the purposes of the computation, we will fix a degree $d\in\NN.$ Taking a second to verify that the limit is really what we want, we note
\[\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le d}}\#\{k\in\FF_q:f(k)=0\}\cdot\frac1{N_d}\]
is asking for the expected value of $\#\{k\in\FF_q:f(k)=0\}$ as $f(x)$ is chosen with probability $1/N_d$ from all $N_d$ options. So yes, we really are computing an expected value $\mathbb E_f(\#\{k\in\FF_q:f(k)=0\}).$

The key step to compute this expected value is linearity of expectation. The earlier proposition does not get to use linearity of expectation because looping over all the various $\FF_p$ is somewhat poorly behaved. But now we get to say
\[\mathbb E_f[\#\{k\in\FF_q:f(k)=0\}]=\sum_{k\in\FF_q}\mathbb E_f\left[1_{f(k)=0}\right].\]
Indeed, for a fixed $f,$ we have that
\[\#\{k\in\FF_q:f(k)=0\}=\sum_{k\in\FF_q}1_{f(k)=0},\]
so linearity of expectation lets us slap a $\mathbb E_f$ on both sides.

However, we can compute $\mathbb E_f\left[1_{f(k)=0}\right]$ for any $k\in\FF_q$ without too much effort.
\begin{lemma}
    Fix everything as above and some $k\in\FF_q.$ Then $\mathbb E_f\left[1_{f(k)=0}\right]=\mathbb E_f\left[1_{f(0)=0}\right].$
\end{lemma}
The idea here is to apply a shift in coordinates. Let $S_d$ be the set of polynomials in $\FF_q[x]$ with degree bounded by $d.$ We claim that $f(x)\mapsto f(x+k)$ provides a bijection of $S_d.$

Technically this is a purely symbolic bijection, but we will merely assert that this is injective because $f(x+k)=g(x+k)$ under $x\mapsto x-k$ implies
\[f((x-k)+k)=g((x-k)+k).\]
Then this is a bijection because it is a map of finite sets $S_d\to S_d,$ so injectivity is enough.

The point of this is that, for fixed $k\in\FF_p,$
\[\mathbb E_f\left[1_{f(k)=0}\right]=\sum_{f(x)\in S_d}1_{f(k)=0}\cdot\frac1{N_d}=\frac{\#\{f(x)\in S_d:f(k)=0\}}{N_d}.\]
However, if we shift coordinates by setting $g(x):=f(x+k),$ then $f(k)=0$ if and only if $g(0)=0.$ Because the map $f\mapsto g$ is a bijection as given above, it suffices to loop over $g$ instead of $f$ so that
\[\mathbb E_f\left[1_{f(k)=0}\right]=\frac{\#\{g(x)\in S_d:g(0)=0\}}{N_d}.\]
This right-hand side is $\mathbb E_f\left[1_{f(0)=0}\right]$ by running this argument with $k=0.$ Thus, $\mathbb E_f\left[1_{f(k)=0}\right]=\mathbb E_f\left[1_{f(0)=0}\right],$ as desired. $\blacksquare$

Continuing with our computation, we see
\[\mathbb E_f[\#\{k\in\FF_q:f(k)=0\}]=\sum_{k\in\FF_q}\mathbb E_f\left[1_{f(k)=0}\right]=q\cdot\mathbb E_f\left[1_{f(0)=0}\right].\]
We now finish by computing $\mathbb E_f\left[1_{f(0)=0}\right].$
\begin{lemma}
    Fix everything as above. Then $\mathbb E_f\left[1_{f(0)=0}\right]=1/q.$
\end{lemma}
The idea is that $f(0)=0$ if and only if its constant coefficient is $0.$ Indeed, again letting $S_d$ be the set of polynomials in $\FF_q[x]$ of degree less than or equal to $d,$ we have that
\[\mathbb E_f\left[1_{f(k)=0}\right]=\frac{\#\{f(x)\in S_d:f(0)=0\}}{N_d}.\]
Now, any polynomial $f(x)\in S_d$ takes the form
\[f(x)=\sum_{k=0}^da_kx^k.\]
Here, $f(0)=a_0,$ so $f(0)=0$ if and only if $a_0=0.$ To compute the desired ratio, we split up the task of fixing a sequence of coefficients $\{a_k\}_{k=0}^d$ into choosing $\{a_k\}_{k=1}^d$ and then choosing $a_0.$

To make an element of $S_d,$ there are some nonzero number of ways, say $n_d$ ways, to choose $\{a_k\}_{k=1}^d,$ and then $q$ ways to choose to $a_0.$ However, to make an element $f(x)\in S_d$ with $f(0)=0,$ there are $n_d$ ways to choose $\{a_k\}_{k=1}^d$ and then only one way to choose $a_0=0.$ Thus,
\[\mathbb E_f\left[1_{f(k)=0}\right]=\frac{\#\{f(x)\in S_d:f(0)=0\}}{N_d}=\frac{n_d\cdot1}{n_d\cdot q}=\frac1q.\]
This is what we wanted. $\blacksquare$

It follows that
\[\mathbb E_f[\#\{k\in\FF_q\}]=q\cdot\mathbb E_f\left[1_{f(0)=0}\right]=q\cdot\frac1q=1.\]
This gives the following statement.
\begin{theorem}
    Fix $q$ a prime power. Then, as $f(x)\in\FF_q[x]$ varies, the expected number of roots of $f(x)$ in $\FF_q$ is $1.$ In other words,
    \[1=\lim_{d\to\infty}\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le d}}\#\{k\in\FF_q:f(k)=0\}\Bigg/\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le D}}1.\]
\end{theorem}
We just showed that, for any fixed degree $d\in\NN,$ we know
\[\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le d}}\#\{k\in\FF_q:f(k)=0\}\Bigg/\sum_{\substack{f(x)\in\FF_q[x]\\\deg f\le D}}1\]
is equal to $q\mathbb E\left[1_{f(0)=0}\right]=1.$ So the limit is actually
\[\lim_{d\to\infty}1=1.\]
This competes the proof. $\blacksquare$

I find it somewhat amusing that either fixing the prime $p$ or fixing the polynomial $f(x)$ both give out $1$ as our answer. I still don't really feel like I understand these results, as much as I do like their proofs. The various conversion steps feel too deep to be natural, I suppose.

Anyways, while looking for this result, I ran across a combinatorial way to compute the number of sqaurefree polynomials of fixed degree in $\FF_q[x],$ from \href{https://math.uchicago.edu/~may/REU2015/REUPapers/AlegreJuarezPajela.pdf}{here}. For this we define the $\zeta$ generating function
\[\zeta_q(T):=\sum_{\text{monic }f\in\FF_q[x]}T^{\deg f}.\]
Perhaps the morally correct way to view this function is to take $T\mapsto q^{-s}$ to make this the Dedekind $\zeta$ function for finite fields: indeed, in $K=\FF_q(t),$ any ideal $I$ of $\mathcal O_K=\FF_q[x]$ is principal and hence generated by a monic $f(x)\in\FF_q[x].$ Its norm comes out to $q^{\deg f},$ so
\[\zeta_{\mathcal O_K}(s)=\sum_{I\subseteq\FF_q[x]}\frac1{\op{Norm}(I)}.\]
We will not be interfacing directly with this Dedekind $\zeta$ function because the generating function will be suitable for our purposes.

We begin with a few basic results about $\zeta_q(T).$
\begin{lemma}
    We have that $\zeta_q(T)=\frac1{1-qT}.$
\end{lemma}
We proceed the same way we proceeded with the Dedekind $\zeta$ function for $\FF_q[x].$ Indeed, splice the sum for $\zeta_q(T)$ up by degree $d\in\NN$ so that
\[\zeta_q(T)=\sum_{d=0}^\infty\#\{\text{monic }f(x)\in\FF_q[x]:\deg f=d\}T^{\deg f}.\]
To compute the number of monic $f(x)\in\FF_q$ of fixed degree $d\in\NN,$ we as usual note that every such $f(x)$ takes the form
\[f(x)=\sum_{k=0}^da_kx^k\]
where $a_d=1.$ Now, there are $q$ options for each $a_k$ with $k<d,$ which totals to $q^d$ polynomials. Thus,
\[\zeta_q(T)=\sum_{d=0}^\infty q^dT^d=\frac1{1-qT}\]
by the geometric series formula. This completes the proof. $\blacksquare$

The above is more or less our rationality result for the Dedekind $\zeta$ function for $\FF_q[x].$ We also have an Euler product.
\begin{lemma}
    We have that
    \[\zeta_q(T)=\prod_{(\pi)\text{ irred.}}\frac1{1-T^{\deg\pi}}.\]
\end{lemma}
We use unique prime factorization into irreducible polynomials. Namely,
\[\zeta_q(T)=\sum_{\text{monic }f(x)\in\FF_q[x]}T^{\deg f}=\sum_{\text{monic }f(x)\in\FF_q[x]}\prod_{(\pi)\text{ irred.}}T^{\nu_\pi(f)\deg\pi}.\]
Here we have used the fact that $\deg(fg)=\deg f+\deg g.$ Now each, monic polynomial $f(x)\in\FF_q[x]$ has a unique infinite tuple $\{\nu_\pi(f)\}_{(\pi)\text{ irred.}}$ of nonnegative integers by unique prime factorization where all but finitely many are zero, and the reverse is true by direct multiplication. So we can sum over these sequences instead of over monic polynomials, giving
\[\zeta_q(T)=\sum_{\{\nu_\pi\}_{(\pi)\text{ irred.}}}\prod_{(\pi)\text{ irred.}}T^{\nu_\pi\deg\pi}.\]
We can now undistribute this massive sum (formally!) into
\[\zeta_q(T)=\prod_{(\pi)\text{ irred.}}\sum_{\nu_\pi=0}^\infty T^{\nu_\pi\deg\pi}.\]
There is some technicality here because we only want to consider tuples $\{\nu_\pi\}_{(\pi)\text{ irred.}}$ where all but finitely many are zero, but if infinitely many are nonzero, then the power of $T$ is an infinite sum of positive integers, which is infinite, and so it doesn't appear in the generating function.

Anyways, using the geometric series formula, we do indeed get
\[\zeta_q(T)=\prod_{(\pi)\text{ irred.}}\frac1{1-T^{\deg\pi}},\]
which is what we wanted. $\blacksquare$

We now turn to squarefree polynomials. Our exposition will reasonably closely follow showing that the proportion of squarefree integers is $\zeta(2)^{-1}.$ We start off as we did with $\zeta_q(T)$ by noting
\[\sum_{\substack{\text{monic }f(x)\in\FF_q[x]\\f(x)\text{ squarefree}}}T^{\deg f}=\sum_{d=0}^\infty\#\{\text{monic }f(x)\in\FF_q[x]:\deg f\le d,f(x)\text{ squarefree}\}T^{\deg f}.\]
For brevity, we will let this coefficient by $s_d$ and this function $S_q(T).$ The way we are going to use the squarefree condition is by an Euler product. Namely, we can again factor
\[\sum_{\substack{\text{monic }f(x)\in\FF_q[x]\\f(x)\text{ squarefree}}}T^{\deg f}=\sum_{\substack{\text{monic }f(x)\in\FF_q[x]\\f(x)\text{ squarefree}}}\prod_{(\pi)\text{ irred.}}T^{\nu_\pi(f)\deg\pi}\]
using unique prime factorization. However, now when we biject monic polynomials $f(x)\in\FF_q[x]$ to infinite tuples $\{\nu_\pi\}_{(\pi)\text{ irred.}},$ we have $\nu_\pi\in\{0,1\}$ because squarefree is equivalent to $\nu_\pi(f)\in\{0,1\}$ for each $\pi.$ 

So, summing over all infinite tuples $\{\nu_\pi\}_{(\pi)\text{ irred.}}$ where all but finitely many are zero and the rest are one, we see
\[S_q(T)=\sum_{\{\nu_\pi\}_{(\pi)\text{ irred.}}}\prod_{(\pi)\text{ irred.}}T^{\nu_\pi(f)\deg\pi}\]
Again, we remark that we can actually drop the condition that all but finitely many of the $\nu_\pi$ are zero because if infinitely many of the $\nu_\pi$ are $1,$ we get a term of whose degree is an infinite sum of $1$s, which won't show up in the generating function. Thus, we may safely allow $\{\nu_\pi\}_{(\pi)\text{ irred.}}$ to merely require $\nu_\pi\in\{0,1\}$ and undistribute as
\[S_q(T)=\prod_{(\pi)\text{ irred.}}\sum_{\nu_\pi\in\{0,1\}}T^{\nu_\pi(f)\deg\pi}.\]
This is our Euler product.
\begin{lemma}
    We have that
    \[S_q(T)=\prod_{(\pi)\text{ irred.}}1+T^{\deg\pi}.\]
\end{lemma}
This follows from the above discussion. $\blacksquare$

Now, we simplify this Euler product into something usable. The key idea is that $1+T=\frac{1-T^2}{1-T},$ so
\[S_q(T)=\prod_{(\pi)\text{ irred.}}\frac{1-(T^2)^{\deg\pi}}{1-T^{\deg\pi}}.\]
Then this is
\[S_q(T)=\prod_{(\pi)\text{ irred.}}\frac1{1-T^{\deg\pi}}\Bigg/\prod_{(\pi)\text{ irred.}}\frac1{1-(T^2)^{\deg\pi}}=\frac{\zeta_q(T)}{\zeta_q\left(T^2\right)}.\]
Now we can use the fact $\zeta_q(T)=\frac1{1-qT}$ to collapse this completely into
\[S_q(T)=\frac{1-qT^2}{1-qT}.\]
To actually extract the coefficients, we expand $\frac1{1-qT}$ out to get
\[S_q(T)=\left(1-qT^2\right)\sum_{d=0}^\infty q^dT^d,\]
which after shifting indices is
\[S_q(T)=\sum_{d=0}^\infty q^dT^d-\sum_{d=2}^\infty q^{d-1}T^d=1+qT+\sum_{d=2}^\infty\left(q^d-q^{d-1}\right)T^d.\]
This gives the following.
\begin{theorem}
    Let $s_d$ be the number of monic, squarefree polynomials in $\FF_q[x]$ of degree $d.$ Then $s_d=q^d$ if $d<2$ and $s_d=q^d-q^{d-1}$ if $d\ge2.$
\end{theorem}
This follows from comparing the coefficients of
\[\sum_{d=0}^\infty s_dT^d=S_q(T)=1+qT+\sum_{d=2}^\infty\left(q^d-q^{d-1}\right)T^d.\]
This completes the proof. $\blacksquare$

I suppose I am not too surprised that this generating functions approach is good enough to compute this, but I am somewhat surprised that it came out so nicely. Comparing this with the case in $\ZZ,$ we essentially just proved the analog of
\[\sum_{n\text{ squarefree}}\frac1{n^s}=\frac{\zeta(s)}{\zeta(2s)},\]
which can be seen by taking $T=q^{-s}$ in the above proofs. If we squint really hard and pretend we have Dirichlet density over $\ZZ,$ then we feel justified in writing
\[\sum_{n\text{ squarefree}}\frac1{n^s}\Bigg/\sum_{n=1}^\infty\frac1{n^s}=\frac1{\zeta(2s)},\]
and we can send $s\to1^+$ to get the desired $\zeta(2)^{-1}$ ratio for squarefree integers. This heuristic computation works over both $\QQ$ and $\FF_q(t)$ as seen above, but getting the actual final answer requires more care, care which is different in both cases.

Indeed, if we must be formal about the computation of the proportion of squarefree monic polynomials in $\FF_q[x],$ then note the number of monic polynomials of degree less than or equal to $D>2$ is
\[\sum_{d=0}^Dq^d=\frac{q^{D+1}-1}{q-1},\]
and the number of squarefree monic polynomials of degree less than or equal to $D$ is
\[1+q+\sum_{d=2}^D\left(q^d-q^{d-1}\right)\]
from the above. However, this sum telescopes into $q^D-q,$ so our answer here is $q^D+1.$ Thus, the proportion of monic polynomials of degree bounded by $D$ which are squarefree is
\[\frac{q^D+1}{\frac{q^{D+1}-1}{q-1}}=\frac{(q-1)\left(q^D+1\right)}{q^{D+1}-1}=\left(1-\frac1q\right)\left(1+\frac1{q^D}\right)\left(1-\frac1{q^{D+1}}\right)^{-1}.\]
Taking $D\to\infty$ gives the following statement.
\begin{theorem}
    A randomly chosen monic polynomial in $\FF_q[x]$ has probability $1-\frac1q$ of being squarefree. More precisely,
    \[\lim_{D\to\infty}\frac{\#\{\text{monic }f(x)\in\FF_q[x]:\deg f\le D,f(x)\text{ squarefree}\}}{\#\{\text{monic }f(x)\in\FF_q[x]:\deg f\le D\}}\]
    is equal to $1-\frac1q.$
\end{theorem}
This follows from the above discussion. $\blacksquare$

And just to verify that we aren't crazy, we see that
\[\zeta_{\FF_q[x]}(2)^{-1}=\zeta_q\left(q^{-2}\right)^{-1}=\left(\frac1{1-qq^{-2}}\right)^{-1}=1-\frac1q,\]
just as we expected.

\subsubsection{June 26th}
Today I learned a proof of the Hilbert basis theorem. Here is the main attraction.
\begin{theorem}
    Fix $R$ a Noetherian ring. Then the polynomial ring $R[x]$ is also Noetherian.
\end{theorem}
At a high level, this proof attempts to emulate the proof that $K[x]$ is a principal ideal domain when $K$ is a field. Namely, we are going to attempt to do Euclidean division to show that $R[x]$ is ``small enough'' to be Noetherian.\todo{maybe grobner; maybe R[[x]]}

\subsubsection{June 27th}
Today I learned about Artinian rings. These are rings which satisfy the ``descending chain condition''; here is our definition.
\begin{definition}
    A ring $R$ is called Artinian if and only if, for any sequence of ideals $\{I_k\}_{k=0}^\infty$ satisfying
    \[I_0\supseteq I_1\supseteq I_2\supseteq\cdots,\]
    there exists a nonnegative $N$ for which $I_k=I_N$ for each $k\ge N.$ Such a sequence of ideals is called a descending chian.
\end{definition}
In general, I feel like the Artinian condition makes rings very small. More precisely, the Artinian condition forces us to have some bottom bound on our ideals, making our ring feel inherently ``discrete,'' where we imagine our ideals as a topology.

To solidify the intuition that Artinian rings are small, or at least discrete, we note that $\ZZ$ itself is not even Artinian despite being essentially the smallest characteristic-$0$ ring. Namely, we see
\[\left(2^0\right)\supseteq\left(2^1\right)\supseteq\left(2^2\right)\supseteq\cdots\]
is a descending chain of ideals, but for each nonnnegative integer $N,$ we have $\left(2^N\right)\supsetneq\left(2^{N+1}\right).$ Indeed, $2^N$ is not in $\left(2^{N+1}\right)$ by unique prime factorization. Thus, this descending chain has no bottom, so $\ZZ$ is not Artinian.

This construction leads to the following statement.
\begin{proposition}
    Any Artinian integral domain is a field.
\end{proposition}
Fix $R$ an Artinian integral domain. We already have the necessary ring structure, so we have to show that all nonzero elements of $R$ are units. Well, fix $r\in R,$ and the main idea is to use the above construction. In particular, let $I_k:=\left(r^k\right)$ so that
\[\left(r^0\right)\supseteq\left(r^1\right)\supseteq\left(r^2\right)\supseteq\left(r^3\right)\supseteq\cdots.\]
Formally, we see $I_k\supseteq I_{k+1}$ because $r^{k+1}=r\cdot r^k\in\left(r^k\right),$ implying $\left(r^{k+1}\right)\subseteq\left(r^k\right).$

The punchline is that, because $R$ is Artinian, this descending chain of ideals must bottom out. So extract our $N$ for which $I_k=I_N$ for any $k\ge N,$ and then we see
\[\left(r^N\right)=\left(r^{N+1}\right).\]
The interesting side of this is that $r^N\in\left(r^{N+1}\right),$ so $r^N=ar^{N+1}$ for some $a\in R.$ This rearranges into
\[r^N\left(1-ar\right)=0.\]
Using the fact we live in an integral domain, we see that either $r^N=0$ or $1-ar=0.$ In the former case, we can write the product out as $r\cdot r\cdot\ldots\cdot r=0$ and use the integral domain condition to force $r=0.$ In the latter case, we see $ar=1,$ so $r$ is a unit. Thus, all elements of $R$ are either $0$ or a unit, which is what we wanted. $\blacksquare$

We are a bit obligated to say that $\ZZ$ is in fact not Artinian for multiple compelling reasons. For example, we don't even have to mess with powers and can just write
\[(2)\supseteq(2\cdot3)\supseteq(2\cdot3\cdot5)\supseteq(2\cdot3\cdot5\cdot7)\supseteq\cdots.\]
Namely, we define $I_k=(p_1p_2\cdots p_{k+1})$ to be generated by the product of the first $k+1$ primes. We won't bother formally seeing that this is an infinite descending chain of ideals, but it is.

Ultimately, the barrier here is that we have ``too many'' maximal ideals to choose from. Artinian rings cannot have this.
\begin{proposition}
    Any Artinian ring has finitely many maximal ideals.
\end{proposition}
It is more natural to show the contrapositive (though we could do this more directly if we wanted). Fix $R$ a commutative ring with infinitely many maximal ideals, and we will show that it fails the descending chain condition.

Well, pick up some countable sequence of pairwise distinct maximal ideals $\{m_k\}_{k=0}^\infty.$ The correct way to generalize the above construction with $\ZZ$ is to write
\[m_0\supseteq m_0\cap m_1\supseteq m_0\cap m_1\cap m_2\supseteq\cdots.\]
We won't bother to show that the intersection of two ideals is an ideal, but it is. So explicitly, we choose
\[I_k=\bigcup_{\ell\le k}m_\ell\]
for each $k\in\NN.$ Because $I_{k+1}=I_k\cap m_{k+1},$ we do indeed have $I_{k+1}\subseteq I_k.$ In fact, this chain is strictly decreasing, which holds by the Chinese remainder theorem, roughly speaking.
\begin{lemma}
    Suppose $I$ and $J$ are proper $R$-ideals such that $I+J=R$ (i.e., relatively prime). Then there exists $a\in R$ such that $a\equiv0\pmod I$ and $a\equiv1\pmod J.$
\end{lemma}
The fact that $I+J=R$ means we get $a\in I$ and $b\in J$ such that $a+b=1.$ We claim that $a\in I$ is the element we are looking for. Indeed, if $a\in J$ as well, then $a+b=1\in J,$ implying that $(1)=R\subseteq J,$ forcing $J$ to be non-proper. So by contrapositive, forcing $J$ to be a proper ideal forces $a\not\in J.$ $\blacksquare$

The point of picking up this lemma is the following.
\begin{lemma}
    Suppose $\{I_k\}_{k=1}^n$ is a sequence of pairwise relatively prime proper $R$-ideals. Then there exists $a\in R$ such that $a\equiv0\pmod{I_k}$ for $k<n$ and $a\equiv1\pmod{I_n}.$
\end{lemma}
This is by induction on $n.$ The case of $n=2$ is settled by the above lemma. Now we suppose the statement is true for all sequences of $n$ pairwise relatively prime proper $R$-ideals, and we will show it for $\{I_k\}_{k=1}^{n+1}.$ The key trick is to extract auxiliary $a_\ell\in R$ which satisfy
\[\begin{cases}
    a_\ell\equiv1\pmod{I_{n+1}}, \\
    a_\ell\equiv0\pmod{I_k} & k\ne\ell,
\end{cases}\]
for each $\ell\le n.$ Indeed, this is legal because $\{I_k\}_{k=1,k\ne\ell}^{n+1}$ is a sequence of $n$ pairwise relatively prime proper $R$-ideals by hypothesis on $\{I_k\}_{k=1}^{n+1}.$ But now modular arithmetic tells us that
\[\prod_{\ell=1}^na_\ell\equiv\begin{cases}
    1\pmod{I_{n+1}}, \\
    0\pmod{I_k} & k\le n.
\end{cases}\]
This finishes the inductive step and therefore the proof. $\blacksquare$

Thus, it suffices to note that maximal ideals are always relatively prime proper ideals: if $m$ and $m'$ are distinct and maximal, then the ideal $m+m'$ is strictly larger than at least one of the two by distinctness, forcing $m+m'=R$ by maximality. Fixing $k\in\NN,$ the above lemmas let us pick up $a\in R$ such that
\[a\equiv\begin{cases}
    1\pmod{m_{k+1}}, \\
    0\pmod{m_\ell} & \ell\le k.
\end{cases}\]
Namely, $a\in I_k$ by the bottom condition, but $a\not\in m_{k+1}$ by the top condition, implying that $a\not\in I_{k+1},$ so $I_{k+1}\subsetneq I_k.$ Thus, our $\{I_k\}_{k=0}^\infty$ is indeed a strictly decreasing chain of ideals, implying that $R$ is not Artinian. $\blacksquare$

The above two results establish some basic ``smallness'' of Artinian rings. We take a moment to acknowledge that this inituition is not foolproof, however. For example, Noetherian is another ``smallness'' condition, but these definitions are somewaht compatible.

Namely, it turns out that all Artinian rings (we are assuming commutative anyways) are in fact Noetherian, but this need not be the case of modules. We already showed $\ZZ$ is not Artinian (even though it is Noetherian), and we also have the following.
\begin{proposition}
    There exists Artinian modules $M$ which are not Notherian.
\end{proposition}
The construction we choose is $\ZZ[1/p]/\ZZ$ for some prime $p,$ which we view as a $\ZZ$-module (i.e., a commutative ring without unity). (The primality here is for convenience and isn't totally necessary.) We quickly note that our submodules are ideals, so we are still looking at ascending and descending chains of ideals here.

Showing that this is non-Noetherian is not difficult: we have the ascending chain of ideals
\[\left(1/p^0\right)\subseteq\left(1/p^1\right)\subseteq\left(1/p^2\right)\subseteq\left(1/p^3\right)\subseteq\cdots.\]
Explicitly, we set $I_k:=\left(1/p^k\right),$ and we note $1/p^k=p/p^{k+1}\in\left(1/p^{k+1}\right),$ so $I_k\subseteq I_{k+1}.$ Furthermore, $1/p^{k+1}$ is not in $\left(1/p^k\right)$ because this is equivalent to there being some $a\in\ZZ$ such that
\[\frac1{p^{k+1}}\equiv\frac a{p^k}\pmod\ZZ.\]
However, this requires $ap\equiv1\pmod{p^{k+1}\ZZ},$ which is impossible after reducing$\pmod p.$ Thus, $I_{k+1}$ is in fact strictly larger than $I_k,$ so the give ascending chain of ideals never stabilizers.

It remains to show that $\ZZ[1/p]/\ZZ$ is Artinian. The main claim is as follows.
\begin{lemma}
    All ideals are either finite (and of the form $\left(1/p^\bullet\right)$) or all of $\ZZ[1/p]/\ZZ.$
\end{lemma}
Fix an ideal $I,$ and look at
\[N=\{-\nu_p(a):a\in I\setminus\{0\}\}.\]
In other words, $N$ is made of the set of powers in the denominator of the elements $a/p^\nu\in I.$ We have two cases.
\begin{itemize}
    \item If $N$ is unbounded, then for any representative $a/p^\nu\in\ZZ[1/p]/\ZZ$ with $p\nmid a,$ there exists some integer larger than $\nu$ in $N.$ We will show $I$ is everything; pick up any $a/p^\nu\in\ZZ[1/p]/\ZZ\setminus\{0\}.$
    
    Because $N$ is unbounded, we have some $a_0/p^{\nu_0}$ where $\nu_0>\nu$ and $p\nmid a_0.$ We claim that $a/p^\nu\in\left(a_0/p^{\nu_0}\right)\subseteq I.$ Indeed, after some rearranging, this is equivalent to requiring
    \[ap^{\nu_0-\nu}\equiv ka_0\pmod{p^{\nu_0}\ZZ}\]
    for some $k\in\ZZ.$ Well, $p\nmid a_0,$ so $a_0^{-1}\pmod{p^{\nu_0}}$ exists, so we can solve for $k$ in the above expression. It follows that $\ZZ[1/p]/\ZZ\subseteq I,$ which finishes this case.
    \item If $N$ is bounded, then there exists some largest element $n\in N$ by well-order. We claim that $I=\left(1/p^n\right).$ By similar argument to above, we are given $a/p^n\in I$ with $p\nmid a.$ Then for any $b/p^n\in\left(1/p^n\right),$ we claim $b/p^n\subseteq\left(a/p^n\right)\subseteq I.$ Indeed, this is equivalent to requiring
    \[b\equiv ak\pmod{p^n\ZZ}\]
    for some $k\in\ZZ.$ But again, $p\nmid a$ means that we can solve for $k$ here. Thus, $\left(1/p^n\right)\subseteq I.$
    
    It remains to show that $I\subseteq\left(1/p^n\right).$ Well, for any $a_0/p^{\nu_0}\in I\setminus\{0\}$ chosen with $p\nmid a_0,$ we know that $\nu_0\le n$ by maximality of $n.$ Thus,
    \[\frac{a_0}{p^{\nu_0}}=\frac{a_0p^{n-n_0}}{p^n}\in\left(\frac1{p^n}\right).\]
    Thus, $I\subseteq\left(1/p^n\right),$ finishing the proof that $I=\left(1/p^n\right).$
\end{itemize}
Thus, we have shown that all ideals are either everything or of the form $\left(1/p^n\right)$ for some positive integer $n.$ We close the proof of the main claim by saying $\left(1/p^n\right)$ only has finitely many elements because any $a/p^n\in\left(1/p^n\right)$ can be represented as
\[\frac a{p^n}\equiv\frac{a\pmod{p^n}}{p^n}\pmod\ZZ,\]
and there are only finitely many options of $a\pmod{p^n}.$ $\blacksquare$

We now use the main claim to show that $\ZZ[1/p]/\ZZ$ is an Artinian $\ZZ$-module. Suppose we have a descending chain of ideals
\[I_0\supseteq I_1\supseteq I_2\supseteq\cdots.\]
If $I_k=\ZZ[1/p]/\ZZ$ for arbitrarily large $k,$ then for any particular $\ell,$ there is a larger $k$ such that
\[I_\ell\supseteq I_k=\ZZ[1/p]/\ZZ.\]
Thus, the entire sequence is $\ZZ[1/p]/\ZZ,$ and this stabilizes for free.

Otherwise, there exists a bound $N$ for which $k\ge N$ implies $I_k\ne\ZZ[1/p]/\ZZ.$ By the above lemma, we know that $I_k=\left(1/p^{a_k}\right)$ for some positive integer $a_k.$ We claim the $a_\bullet$ are decreasing. Indeed, we note that $I_k\supseteq I_{k+1}$ requires $1/p^{a_{k+1}}\in\left(1/p^{a_k}\right),$ which is equivalent to
\[\frac a{p^{a_k}}\equiv\frac1{p^{a_{k+1}}}\pmod\ZZ\]
for some $a\in\ZZ.$ This condition is equivalent to $a\equiv p^{a_k-a_{k+1}}\pmod{p^{a_k}},$ and as both sides of this had better be integral, we see $a_{k+1}\le a_k.$

However, by well-order, the decreasing sequence of positive integers $a_\bullet$ must stabilize. That is, we are granted a positive integer $M$ such that $a_k=a_M$ for any $k>M.$ Translating back to the ideal world, this tells us
\[I_k=\left(\frac1{p^{a_k}}\right)=\left(\frac1{p^{a_M}}\right)=I_M\]
for any $k>M.$ In other words, the descending sequence $I_\bullet$ does indeed stabilize. $\blacksquare$

I am under the impression that $\ZZ[1/p]/\ZZ$ being Artinian is more or less the dual of $\ZZ$ being Noetherian. Namely, the reason why ascending chains of ideals in $\ZZ$ stabilize is, ultimately, because of the well-order on $\ZZ,$ which is the same property (albeit in exponents) that forces descending chains in $\ZZ[1/p]/\ZZ$ to stabilize.

The last thing we will say about Artinian rings today is a property of Artinian local rings.
\begin{proposition}
    Suppose $R$ is an Artinian local ring with maximal ideal $m.$ Then the ideal $m$ is nilpotent: there exists an integer $k$ such that $m^k=0.$
\end{proposition}
We are interested in powers, so without much better to do, we write out the descending sequence
\[m^0\supseteq m^1\supseteq m^2\supseteq\cdots.\]
Because $R$ is Artinian, this must stabilizer eventually, so pick up a positive integer $N$ such that $m^k=m^N$ for any $k\ge N.$ In particular, $m^{N+1}=m^N.$ We hope that $m^N=0$; for example, certainly if $m^N=0,$ then we have stabilized.

The trick to finish the proof is quite clever. Suppose for the sake of contradiction that $m^N\ne0,$ and to measure how far we are from $0,$ extract an ideal $I$ minimal to the condition that
\[Im^N\ne0.\]
We can do this by Zorn's lemma, or we can simply say that a lack of such minimal ideals would allow us to build an infinite descending chain, violating the Artinian condition on $R.$

To put this funny step into perspective, we will say that we hope to use Nakayma's lemma in the near future. To begin, we claim that $I$ is principal. Indeed, $Im^N\ne0$ implies that there is some product of $r\in I$ and $m_0\in m^N$ such that $rm_0\ne0,$ implying $rm^N\ne0.$ But then $rm_0\in(r)m^N$ as well, so
\[(r)m^N\ne0.\]
The minimality of $I$ thus requires $I\subseteq(r),$ but with $r\in I$ implying $(r)\subseteq I$ already, we conclude $I=(r)$ is principal.

We now directly go after Nakayama's lemma. We show that $(r)m=(r).$ The motivation for what follows is that $(r)m\subseteq(r)$ already, but we might be able to use the minimality of $(r)$ to save us. Indeed, we can check that
\[\big((r)m\big)m^N=(r)m^{N+1}=(r)m^N,\]
where the last equality holds because $m^{N+1}=m^N.$ However, $(r)m^N$ is nonzero by construction of $(r),$ so indeed, $\big((r)m\big)m^N\ne0,$ and we get $(r)\subseteq (r)m$ by minimality.

Thus, $(r)m=(r),$ and Nakayama's lemma kicks in to tell us $(r)=0.$ However, this means $(r)m^N=0,$ which is a contradiction to the construction of $(r).$ This completes the proof. $\blacksquare$

We remark that a fake proof of the above statement would apply Nakayama's lemma directly to $m^Nm=m^N.$ This doesn't work because we don't know if $m^N$ is finitely generated, which is a crucial condition here. Instead, we have to go through all those gymnastics with $Im^N\ne0.$

The contradiction in the above proof is somewhat subtle, and I am not convinced it is easy to remove it. This is one of the very few instances of a contradiction proof where we have used the assumption we are contradicting in a nontrivial way that cannot be separated from the rest of the logic. Anyways, it is clever.

\subsubsection{June 28th}
Today I learned a toy model for schemes, as more or less the opposite category of commutative algebras. The idea is that for a ring map $R\to S,$ we induce a map of $\op{Spec}S\to\op{Spec}R$ by pre-imaging. Explicitly, we have the following.
\begin{proposition}
    Fix rings $R$ and $S$ with a ring homomoprhism $\varphi:R\to S.$ This induces a canonical map $\varphi^{-1}:\op{Spec}S\to\op{Spec}R.$ In fact, this is functorial: if $\varphi:R\to S$ and $\psi:S\to T,$ then the induced $(\psi\circ\varphi)^{-1}$ is equal to $\varphi^{-1}\circ\psi^{-1}.$
\end{proposition}
The map is by pre-image. Take any prime $\mf q\in\op{Spec}S,$ and we show that $\varphi^{-1}(\mf q)\in\op{Spec}(R).$ Indeed, if $r_1r_2\in\varphi^{-1}(\mf q),$ then
\[\varphi(r_1)\varphi(r_2)=\varphi(r_1r_2)\in\mf q\]
because $\varphi$ is a ring homomorphism. The primality of $\mf q$ implies $\varphi(r_1)\in\mf q$ or $\varphi(r_2)\in\mf q,$ so $r_1\in\varphi^{-1}(\mf q)$ or $r_2\in\varphi^{-1}(\mf q).$ So indeed, $\varphi^{-1}(\mf q)$ is a prime, so $\varphi^{-1}:\op{Spec}S\to\op{Spec}R$ is well-defined.

To show that this functorial is the same thing as showing that taking pre-images is functorial. Namely, given $\mf p\in\op{Spec}T,$ we see that
\[(\psi\circ\varphi)^{-1}(\mf p)=\{r\in R:\psi(\varphi(r))\in\mf p\}\]
while
\[\varphi^{-1}\left(\psi^{-1}(\mf p)\right)=\{r\in R:\varphi(r)\in\psi^{-1}(\mf p)\},\]
and these are indeed equal. $\blacksquare$

We can think of $\op{Spec}R$ as a topological space, but I don't really care yet, so we'll think of it like as a set for now. Anyways, today the goal is to provide a different language to solving Diophantine equations, so we go after that now. Fix
\[f\in\ZZ[t_1,\ldots,t_n]\]
some irreducible polynomial. We are interested in the solutions of $f$ over $\ZZ,$ but it will be interesting to look at solutions to $f$ over various commutative $\ZZ$-algebras, so we define
\[V_{(f)}(A):=\left\{a\in A^n:f(a)=0\right\},\]
where $A$ is a commutative $\ZZ$-algebra. The way we talk about solutions, or elements of $V_{(f)}(A)$ more generally, is through the following.
\begin{lemma}
    Given a commutative $\ZZ$-algebra $A,$ solutions in $V_{(f)}(A)$ are in natural bijection to maps
    \[\op{Hom}_{\op{CAlg}_\ZZ}\left(\frac{\ZZ[t_1,\ldots,t_n]}{(f)},A\right).\]
\end{lemma}
We work backwards a bit. We notice that maps
\[\varphi:\frac{\ZZ[t_1,\ldots,t_n]}{(f)}\to A\]
are in natural bijection with maps
\[\ZZ[t_1,\ldots,t_n]\to A\]
with kernel $(f).$ However, maps $\varphi:\ZZ[t_1,\ldots,t_n]\to A$ are in natural bijection with $n$-tuples $(a_1,\ldots,a_n)\in A^n$ which dictate $\varphi(t_1),\ldots,\varphi(t_n).$

Thus, we claim that maps $\varphi:\ZZ[t_1,\ldots,t_n]\to A$ with kernel $(f)$ are in natural bijection with solutions in $V_{(f)}(A)$ under the bijection mentioned above. Indeed, if the kernel is $f,$ then
\[\varphi(f(t))=f\big(\varphi(t_1),\ldots,\varphi(t_2)\big)\]
because $\varphi$ is a homomorphism of commutative $\ZZ$-algebras. So the tuple provided by the map $\varphi$ does indeed live in $V_{(f)}(A).$ Conversely, any solution $a=(a_k)_{k=1}^n\in V_{(f)}(A)$ bijects to $\varphi$ with $\varphi(t_\bullet)=a_\bullet$ which satisfies
\[\varphi(f(t))=f\big(\varphi(t_1),\ldots,\varphi(t_2)\big)=f(a_1,\ldots,a_n)=0.\]
Thus, the produced $\varphi$ does indeed have $f$ and therefore $(f)$ in its kernel. This completes the proof. $\blacksquare$

In light of the above, we write $R:=\ZZ[t_1,\ldots,t_n]/(f)$ to be the quotient space associated to $f.$ Note that $R$ a commutative $\ZZ$-algebra itself and in fact a ring, so the point is that any solution $V_S(A)$ for a commutative $\ZZ$-algebra $A$ naturally induces $X\to A,$ which naturally induces
\[\op{Spec}A\to\op{Spec}R.\]
This means that we can think about solutions to $(f)$ as maps into $\op{Spec}R,$ which has the advantage of not having to touch elements anywhere. This is our geometric way to view solving Diophantine equations, and it is essentially the slogan for today.
\begin{idea}
    Fix $A$ a commutative $\ZZ$-algebra. A solution $a\in V_{(f)}(A)$ naturally induces a map $\op{Spec}A\to\op{Spec}R.$
\end{idea}
We remark that the converse is not exactly true in full generality because a mapping on spectrums does not induce a map on the rings.

Let's get some practice with this idea. I think the following is true in the case of affine things, though I do not know its proof.
\begin{lemma}
    Fix a ring homomorphism $\varphi:R\to S$ of affine varieties with induced $\varphi^{-1}:\op{Spec}S\to\op{Spec}R.$ If $\varphi$ is monic, then $\varphi^{-1}$ is epic; if $\varphi$ is epic, then $\varphi^{-1}$ is monic.
\end{lemma}
I do not know the proof of this. $\blacksquare$

We are going to use the above to begin to build some geometric intuition, so aside from the notation on our arrows, we will actually prove our claims without the above. For example, we have the canonical surjection $\ZZ\twoheadrightarrow\FF_p$ for $p$ prime, so we have induced an embedding
\[\op{Spec}\FF_p\hookrightarrow\op{Spec}\ZZ.\]
In fact, $\FF_p$ has only one prime, namely $(0),$ and if we track everything back through, its inverse image is $(p)\in\op{Spec}\ZZ.$ So we are really taking $(0)\mapsto(p)$ here: we have $\op{Spec}\FF_p$ embedding as a point in $\ZZ.$

Similarly, we have an embedding $\ZZ\hookrightarrow R$ because $R$ is a commutative $\ZZ$-algebra. This means that we induce a surjection/covering
\[\op{Spec}R\twoheadrightarrow\op{Spec}\ZZ.\]
This is a bit more annoying to describe explicitly, so we won't. The point is that we have the following diagram, virtually for free.
\begin{center}
    \begin{tikzcd}
    	& {\operatorname{Spec}R} \\
    	{\operatorname{Spec}\mathbb F_p} & {\operatorname{Spec}\mathbb Z}
    	\arrow[two heads, from=1-2, to=2-2]
    	\arrow[hook, from=2-1, to=2-2]
    \end{tikzcd}
\end{center}
Geometrically, $\op{Spec}R$ covers $\op{Spec}\ZZ,$ and $\op{Spec}\ZZ$ has a point $\op{Spec}\FF_p.$

Now, a solution in $a\in V_{(f)}(\ZZ)$ induce a map $\varphi_a:R\to\ZZ,$ and because this is a map of $\ZZ$-algebras, the composite $\ZZ\hookrightarrow R\to\ZZ$ should be the identity. Explicitly, the map $\ZZ\hookrightarrow R$ cares nothing for the polynomial structure of $R,$ but $R\to\ZZ$ holds the embedded $\ZZ$ in place and only uses that polynomial structure.

The point of saying this is that the induced map $\varphi_a:R\to\ZZ$ from a solution $a\in V_{(f)}(\ZZ)$ is making a section of $\ZZ\hookrightarrow R.$ Thus, when we look to the spectrum, functoriality says
\[\op{Spec}\ZZ\twoheadleftarrow\op{Spec}R\stackrel{\varphi_a^{-1}}\leftarrow\op{Spec}\ZZ\]
should also be the identity. That is, $\varphi_a^{-1}$ will be a section of the covering $\op{Spec}R\twoheadrightarrow\op{Spec}\ZZ.$ Adding to the diagram, we see the following.
\begin{center}
    \begin{tikzcd}
    	& {\operatorname{Spec}R} \\
    	{\operatorname{Spec}\mathbb F_p} & {\operatorname{Spec}\mathbb Z}
    	\arrow[two heads, from=1-2, to=2-2]
    	\arrow[hook, from=2-1, to=2-2]
    	\arrow["{\varphi_a^{-1}}"', shift right=2, dashed, from=2-2, to=1-2]
    \end{tikzcd}
\end{center}
So, at a high level, we are searching for sections of $\op{Spec}R\to\op{Spec}\ZZ.$

Geometrically, one barrier to constructing a full upwards map $\op{Spec}\ZZ\to\op{Spec}R$ is that we need to know where we can send the point $\op{Spec}\FF_p$ in a reasonable way. In other words, we need a map $\op{Spec}\FF_p\to\op{Spec}R.$ The way we know how to do this is by looking for maps $R\to\FF_p,$ which means we need a solution in $V_{(f)}(\FF_p).$

So at a high level, we have geometrically discovered that a ``local'' barrier to the Diophantine equation $f=0$ over $\ZZ$ is solving the Diophantine equation $\FF_p.$ In other words, we need to be able to solve $f\pmod p$ before we can solve over $\ZZ.$ We have more or less recovered the global-to-local principle.

We take a second to be as explicit as we can with out current technology. We work in a little more generality than $\FF_p.$
\begin{lemma}
    Let $n_0$ be a positive integer, and fix $a\in V_{(f)}(\ZZ/n_0\ZZ)$ a solution inducing $\varphi_a:R\to\ZZ/n_0\ZZ.$ We claim that the following diagram commutes.
    \begin{center}
        \begin{tikzcd}
        	& R \\
        	{\ZZ/n_0\ZZ} & {\mathbb Z}
        	\arrow["{\varphi_a}"', dashed, from=1-2, to=2-1]
        	\arrow[two heads, from=2-2, to=2-1]
        	\arrow[hook, from=2-2, to=1-2]
        \end{tikzcd}
    \end{center}
\end{lemma}
This is just a matter of unravelling all of our maps. Fixing $a=(a_k)_{k=1}^n,$ we have that $\varphi_a:R\to\ZZ/n_0\ZZ$ is the canonical map $\ZZ\to\ZZ/n_0\ZZ$ with $t_\bullet\mapsto a_\bullet.$ However, when $\ZZ\hookrightarrow R,$ then $\varphi_a$ will only act like $\ZZ\to\ZZ/n_0\ZZ$ on these inputs. So indeed, the diagram commutes by restriction. $\blacksquare$

Moving back to the spectrum world, we get that the following diagram also commutes by functoriality.
\begin{center}
    \begin{tikzcd}
    	& {\operatorname{Spec}R} \\
    	{\operatorname{Spec}\mathbb Z/n_0\mathbb Z} & {\operatorname{Spec}\mathbb Z}
    	\arrow[two heads, from=1-2, to=2-2]
    	\arrow[hook, from=2-1, to=2-2]
    	\arrow["{\varphi_a^{-1}}", dashed, from=2-1, to=1-2]
    \end{tikzcd}
\end{center}
We quickly note that even though $\op{Spec}\ZZ/n_0\ZZ$ is embedding into $\op{Spec}\ZZ,$ it is no longer embedding as a single point because $\op{Spec}\ZZ/n_0\ZZ$ is allowed to have more than one prime ideal (and does when $n_0$ is not prime).

It is currently unclear what we gained from moving from $\FF_p$ to $\ZZ/n_0\ZZ,$ but we note that the real global-to-local principle is about trying to lift solutions from $\ZZ_p$ over all $p$ to $\ZZ.$ Thus, we want to expand our view from the point $\op{Spec}\FF_p$ to $\op{Spec}\ZZ/p^\bullet\ZZ.$

In particular, we have that finding a solution modulo $p$ corresponds in some weak way with finding a map $\op{Spec}\ZZ/p\ZZ\to\op{Spec}R.$ Because finding solutions in $\ZZ_p$ corresponds with finding a whole bunch of compatible solutions in $\ZZ/p^\bullet\ZZ,$ we would like to know that the induced $\varphi_a$ maps commute when the solutions do. Indeed, they do.
\begin{lemma}
    Suppose $n_1$ and $n_2$ are positive integers with $n_1\mid n_2.$ Additionally, suppose we are given solutions $a_1\in V_{(f)}(\ZZ/n_1\ZZ)$ and $a_2\in V_{(f)}(\ZZ/n_2\ZZ)$ such that $a_2\equiv a_1\pmod{n_1}.$ Then the induced maps $\varphi_1:R\to\ZZ/n_1\ZZ$ and $\varphi_2:R\to\ZZ/n_2\ZZ$ cause the following diagram to commute.
    \begin{center}
        \begin{tikzcd}
        	{\mathbb Z/n_1\mathbb Z} & R \\
        	{\mathbb Z/n_2\mathbb Z}
        	\arrow["{\varphi_2}", dashed, from=1-2, to=2-1]
        	\arrow["{\varphi_1}"', dashed, from=1-2, to=1-1]
        	\arrow[two heads, from=2-1, to=1-1]
        \end{tikzcd}
    \end{center}
\end{lemma}
This is again a matter of unravelling our maps. The main point is that the following diagram commutes because $n_1\mid n_2.$
\begin{center}
    \begin{tikzcd}
    	{\mathbb Z/n_1\mathbb Z} \\
    	{\mathbb Z/n_2\mathbb Z} & {\mathbb Z}
    	\arrow[two heads, from=2-1, to=1-1]
    	\arrow[two heads, from=2-2, to=1-1]
    	\arrow[two heads, from=2-2, to=2-1]
    \end{tikzcd}
\end{center}
Setting $a=(\alpha_k)_{k=1}^n,$ the map $R\to\ZZ/n\ZZ$ is the map which does $\ZZ\to\ZZ/n\ZZ$ and sends $t_\bullet\mapsto\alpha_\bullet.$ In particular, we set $a_1=(a_{1,k})_{k=1}^n$ and $a_2=(a_{2,k})_{k=1}^n$ so that
\[a_{1,\bullet}\equiv a_{2,\bullet}\pmod{n_1}.\]
In particular, when we send $R\to\ZZ/n_2\ZZ$ and then to $\ZZ/n_1\ZZ,$ a $t_\bullet$ gets sent to $a_{2,\bullet}\equiv a_{1,\bullet}$ in $\ZZ/n_1\ZZ.$ So the $t_\bullet$ are all being sent to the same place as by $R\to\ZZ/n_1\ZZ.$ As for our coefficients in $\ZZ,$ we are safe because the diagram at the beginning commutes. $\blacksquare$

The punchline is that we can use the above lemma to say that a compatible sequence of solutions $a_\bullet\in V_{(f)}(\ZZ/p^\bullet\ZZ)$ (i.e., with $a_k\equiv a_\ell\pmod{p^{\min\{k,\ell\}}}$) induces maps $\varphi_\bullet:R\to\ZZ/p^\bullet\ZZ$ which make the following diagram commute.
\begin{center}
    \begin{tikzcd}
    	{\mathbb Z/p^1\mathbb Z} & R \\
    	{\mathbb Z/p^2\mathbb Z} \\
    	{\mathbb Z/p^3\mathbb Z} \\
    	\vdots
    	\arrow[two heads, from=2-1, to=1-1]
    	\arrow[two heads, from=3-1, to=2-1]
    	\arrow[two heads, from=4-1, to=3-1]
    	\arrow["{\varphi_1}"{description}, dashed, from=1-2, to=1-1]
    	\arrow["{\varphi_2}"{description}, dashed, from=1-2, to=2-1]
    	\arrow["{\varphi_3}"{description}, dashed, from=1-2, to=3-1]
    \end{tikzcd}
\end{center}
In fact, we can add in $\ZZ$ and all of its maps $\ZZ\twoheadrightarrow\ZZ/p^\bullet\ZZ,$ and the diagram still commutes by a prior lemma. We won't do this because it gets messy, but the point is that we get a unique map $\varphi_p:R\to\ZZ_p$ by universal property which commutes with all of the $\varphi_\bullet.$

We say now that we would like to make the following diagram commute.
\begin{center}
    \begin{tikzcd}
    	& R \\
    	{\mathbb Z_p} & {\mathbb Z}
    	\arrow["{\varphi_p}"', dashed, from=1-2, to=2-1]
    	\arrow[hook, from=2-2, to=1-2]
    	\arrow[hook, from=2-2, to=2-1]
    \end{tikzcd}
\end{center}
Well, the map $\ZZ\hookrightarrow\ZZ_p$ is uniquely defined to commute with the projections $\ZZ\twoheadrightarrow\ZZ/p^\bullet\ZZ,$ so it suffices to show that the composite $\ZZ\hookrightarrow R\to\ZZ_p$ also commutes with these projections. For this, $R\to\ZZ_p$ is uniquely defined to commute with the $\varphi_\bullet:R\to\ZZ/p^\bullet\ZZ,$ so
\[\ZZ\hookrightarrow R\to\ZZ_p\to\ZZ/p^\bullet\ZZ=\ZZ\hookrightarrow R\to\ZZ/p^\bullet\ZZ.\]
However, the composite on the right-hand side is simply the projection $\ZZ\to\ZZ/p^\bullet\ZZ$ by a previous lemma. So indeed, $\ZZ\hookrightarrow R\to\ZZ_p$ commutes with the projections $\ZZ\to\ZZ/p^\bullet\ZZ,$ making the diagram commute.

After all of that, we have the following.
\begin{lemma}
    A compatible sequence of solutions $a_\bullet\in V_{(f)}(\ZZ/p^\bullet\ZZ)$ yields a canonical map $R\to\ZZ_p$ making the following diagram commute.
    \begin{center}
        \begin{tikzcd}
        	& R \\
        	{\mathbb Z_p} & {\mathbb Z}
        	\arrow["{\varphi_p}"', dashed, from=1-2, to=2-1]
        	\arrow[hook, from=2-2, to=1-2]
        	\arrow[hook, from=2-2, to=2-1]
        \end{tikzcd}
    \end{center}
\end{lemma}
This follows from the above discussion. $\blacksquare$

Now, moving into spectrum land, we see that a compatible sequence of solutions really induces the following commutative diagram.
\begin{center}
    \begin{tikzcd}
    	& {\operatorname{Spec}R} \\
    	{\operatorname{Spec}\mathbb Z_p} & {\operatorname{Spec}\mathbb Z}
    	\arrow["{\varphi_p^{-1}}", dashed, from=2-1, to=1-2]
    	\arrow[two heads, from=1-2, to=2-2]
    	\arrow[from=2-1, to=2-2]
    \end{tikzcd}
\end{center}
Observe here that $\op{Spec}\ZZ_p$ does not surject onto $\op{Spec}\ZZ$ even though $\ZZ$ embeds into $\ZZ_p.$ So our intuition lemma does not work in full generality.

The geometric way to view $\op{Spec}\ZZ_p$ is as a small disk/neighborhood around the point $(p)\in\op{Spec}\ZZ.$ It still holds that $\ZZ_p$ is a local obstruction to creating the desired solution $\op{Spec}\ZZ\to\op{Spec}R,$ and in fact $\ZZ_p$ is essentially the local obstruction at $(p).$ Anyways, the point is that we managed to recover the idea ``to solve in $\ZZ,$ solve in $\ZZ_p$ first'' somewhat geometrically.

\subsubsection{June 29th}
Today I learned more about the Zariski topology. To review, here is the definition in the specialized case of interest.
\begin{definition}
    Fix $k$ be an integral domain and $\AA_k^n$ the set of $n$-tuples. Then we define the Zariski topology of $\AA_k^n$ as consisting of closed sets
    \[V(S):=\left\{a\in\AA_k^n:f(a)=0\text{ for each }f\in S\right\}\]
    for any $S\subseteq k[t_1,\ldots,t_n].$ We call the sets $V(-)$ ``algebraic'' sets.
\end{definition}
We remark that, for any $\{f_\ell\}_{\ell=1}^m\subseteq S$ and $\{a_\ell\}_{\ell=1}^m\subseteq k,$ we have that $a\in V(S)$ implies
\[\sum_{\ell=1}^ma_\ell f_\ell(a)=\sum_{\ell=1}^ma_\ell\cdot0=0\]
for free. In other words, adding in $k$-linear combinations of elements of $S$ does not limit the vanishing set any more $V(S).$ Because adding in more functions certainly cannot make the vanishing set larger (the condition is ``for all''), we conclude that $V(S)=V((S)),$ where $(S)$ is the ideal of $k[t_1,\ldots,t_n]$ generated by $S.$

Anyways, it is not difficult to verify that this actually makes a topology, which we do now.
\begin{proposition}
    When $k$ is an integral domain, the Zariski closed sets actually define a topology.
\end{proposition}
To show that some family of closed sets makes a topology, we have to show the following.
\begin{itemize}
    \item The empty set is an algebraic set. For example, the constant polynomial $1\in k[t_1,\ldots,t_n]$ doesn't vanish anywhere, so $V(\{1\})=\emp.$
    \item The entire set $\AA_k^n$ is an algebraic set. Indeed, the constant polynomial $0\in k[t_1,\ldots,t_n]$ vanishes everywhere, so $V(\{0\})=\AA_k^n.$
    \item Algebraic sets are closed under finite union. More precisely, we need to show that, for any two algebraic sets $V(S_1)$ and $V(S_2),$ then $V(S_1)\cup V(S_2)$ is also algebraic. We claim that
    \[V(S_1)\cup V(S_2)\stackrel?V(S_1S_2),\]
    where $S_1S_2:=\{f_1f_2:f_1\in S_1,f_2\in S_2\}.$ In one direction, note that $a\in V(S_1)\cup V(S_2)$ lives in $V(S_1)$ without loss of generality. Thus, for any $f_1f_2\in S_1S_2$ with $f_1\in S_1$ and $f_2\in S_2,$ we see
    \[(f_1f_2)(a)=f_1(a)f_2(a)=0f_2(a)=0\]
    because $a\in V(S_1)$ implies $f_1(a)=0.$
    
    In the other direction, fix $a\in V(S_1S_2).$ If $a\in V(S_1),$ then we are done already. Otherwise, $a\notin V(S_1),$ so there is some $f_1\in S_1$ such that $f_1(a)\ne0.$ We want to show $a\in V(S_2).$ Well, for any $f_2\in V(S_2),$ we know $f_1f_2\in S_1S_2,$ so it follows
    \[f_1(a)f_2(a)=(f_1f_2)(a)=0\]
    because $a\in V(S_1S_2).$ However, $f_1(a)\ne0,$ so $f_2(a)=0$ because we live in an integral domain (!). Thus, $a\in V(S_2).$
    
    \item Algebraic sets are closed under arbitrary intersection. For this, we claim that, given a collection $\{S_\alpha\}_{\alpha\in\lambda}$ with $S_\alpha\subseteq k[t_1,\ldots,t_n],$ we have
    \[\bigcap_{\alpha\in\lambda}V(S_\alpha)=V\left(\bigcup_{\alpha\in\lambda}S_\alpha\right).\]
    Showing this is a matter of unravelling the definitions. The set on the left expands as
    \[\bigcap_{\alpha\in\lambda}\left\{a\in\AA_k^n:f(a)=0\text{ for each }f\in S_\alpha\right\}.\]
    The set on the right expands as
    \[\left\{a\in\AA_k^n:f(a)=0\text{ for each }f\in\bigcup_{\alpha\in\lambda}S_\alpha\right\}.\]
    These set constructions are saying the same thing: we want $a\in\AA_k^n$ such that $a$ vanishes on any $f\in S_\alpha$ for any $S_\alpha.$ So these are equal.
\end{itemize}
Having verified all properties needed to make a topology on closed sets, we are done here. $\blacksquare$

Today we are interested in more topological properties of $\AA_k^n.$ However, we understand $\AA_k^n$ best as some algebraic object, so we will slowly try to turn algebraic definitions into statements about the topology. This means that we will want a way to turn our closed sets into algebraic objects, namely ideals. We define
\[I(V):=\{f\in k[t_1,\ldots,t_n]:f(a)=0\text{ for each }a\in V\}\]
for any algebraic set $V.$ Because algebraic sets $V$ are defined as $V(S)$ for some $S\subseteq k[t_1,\ldots,t_n],$ this definition is somewhat natural as an attempt to define an inverse.

Quickly, it is true that $I(V)$ is an ideal. Indeed, $0\in I(V),$ and if $a_1,a_2\in k[t_1,\ldots,t_n]$ and $f_1,f_2\in I(V)$ has $a_1f_1+a_2f_2\in I(V)$: for any $a\in V,$ we know
\[(a_1f_1+a_2f_2)(a)=a_1f_1(a)+a_2f_2(a)=0.\]
So $I(V)$ is closed under $k[t_1,\ldots,t_n]$-linear combination and is an ideal.

It turns out that $I(-)$ is not actually an inverse of $V(-).$ For example, if $S$ is not an ideal, we already remarked that $S$ and $(S)$ generated the same vanishing set, so it is impossible for $I(-)$ to distinguish between these. We can say a bit more than this.
\begin{proposition}
    Again, $k$ is an integral domain. Fix $S$ a subset of $k[t_1,\ldots,t_n].$ Then
    \[\sqrt{(S)}\subseteq I(V(S)),\]
    where $\sqrt{(S)}$ is the radical of the ideal $(S).$
\end{proposition}
Fix $f\in\sqrt{(S)}$ so that we want to show that $f\in I(V(S)).$ By definition,
\[I(V(S))=\left\{g\in k[t_1,\ldots,t_n]:g(a)=0\text{ for each }a\in V(S)\right\},\]
so we need to show that $f$ satisfies $f(a)=0$ for each $a\in V(S).$

Well, $f\in\sqrt{(S)}$ promises a positive integer $N$ such that $f^N\in(S),$ which promises finite sequences $\{a_\ell\}_{\ell=1}^m\subseteq k$ and $\{f_\ell\}_{\ell=1}^m$ such that
\[f^N=\sum_{\ell=1}^ma_\ell f_\ell.\]
Finishing up now, we can pick up any $a\in V(S),$ and we note that
\[f^N(a)=\sum_{\ell=1}^ma_\ell f_\ell(a)=0\]
because the $f_\bullet\in S.$ However, $f(a)^N=0$ forces $f(a)=0$ because we live in an integral domain. Thus, we do indeed have $f\in I(V(S)),$ which finishes the proof. $\blacksquare$

In fact, the above proposition can be strengthened into an equality in most cases we care about, but this requires quite a bit more effort; it is Hilbert's Nullstellensatz. We will not dwell on this further because we have geometry to get back to.

Quickly, we pick up the following.
\begin{lemma}
    We have $S_1\subseteq S_2\subseteq k[t_1,\ldots,t_n]$ implies $V(S_1)\supseteq V(S_2).$ Similarly, we have $V_1\subseteq V_2$ as algebraic sets if and only if $I(V_1)\supseteq I(V_2).$ In short, $I$ and $V$ are inclusion-reversing.
\end{lemma}
This is a matter of expanding out our definitions. To start, if $S_1\subseteq S_2\subseteq k[t_1,\ldots,t_n],$ pick up $a\in V(S_2)$ so that we need to show $V(S_1).$ Well, for any $f\in S_1,$ we know $f\in S_2,$ so it follows
\[f(a)=0.\]
It follows $a\in V(S_1),$ so $V(S_2)\subseteq V(S_1)$ follows. We note that we don't have the converse here because $S_1$ and $S_2$ are not necessarily behaved as sets, or even as ideals; e.g., $V((t))=V\left(\left(t^2\right)\right)$.

Similarly, fix algebraic sets $V_1\subseteq V_2$ so that we want $I(V_2)\subseteq I(V_1).$ Well, for any $f\in I(V_2),$ we know that any $a\in V_1$ also has $a\in V_2,$ so
\[f(a)=0\]
because $f\in I(V_2).$ Thus, $f\in I(V_1),$ so $I(V_2)\subseteq I(V_1)$ indeed.

As for the converse, suppose $I(V_2)\subseteq I(V_1)$ so that we want $V_1\subseteq V_2.$ Well, we may write $V_2=V(S_2)$ for some $S_2\subseteq k[t_1,\ldots,t_n],$ and because $V_2$ vanishes on $S_2,$ we see
\[S_2\subseteq I(V_2)\subseteq I(V_1).\]
Thus, any $a_1\in V_1,$ which vanishes on all of $I(V_1),$ will vanish on $S_2,$ which gives the needed $a_1\in V_2.$ This finishes the proof.
$\blacksquare$

Now, primality is a condition we care about, so it would be nice to have a geometric condition on the closed set $V$ which gives $I(V)$ prime. Well, a proper ideal $\mf p$ is prime if and only if $ab\in\mf p$ implies $a\in\mf p$ or $b\in\mf p.$ A more ideal-theoretic way to say this is that $IJ\subseteq\mf p$ implies $I\subseteq\mf p$ or $J\subseteq\mf p.$

Translating this into vanishing sets means we want to make all ideals into closed sets and reverse the inclusions. This gives the following definition.
\begin{definition}
    A topological space $X$ is called irreducible if and only if $X=X_1\cup X_2$ for closed sets $X_1$ and $X_2$ implies $X_1=X$ or $X_2=X.$
\end{definition}
We might be tempted to say $X\subseteq X_1$ or $X\subseteq X_2$ is a more natural conclusion, but these closed sets would be induced by the subspace topology, so we get $X_1,X_2\subseteq X$ for free.

Intuitively, irreducible is essentially a very strong version of connected; after all, the definition of irreducible is the definition of connected where $X_1\sqcup X_2$ is replaced with $X_1\cup X_2.$ Explicitly, we have the following.
\begin{proposition}
    An irreducible topological space $X$ is connected.
\end{proposition}
Well, to verify connectivity, suppose $X=X_1\sqcup X_2$ for disjoint closed sets $X_1$ or $X_2.$ Then $X=X_1\cup X_2,$ so $X_1=X$ or $X_2=X.$ Without loss of generality, $X_1=X,$ so $X_2\subseteq X\setminus X_1=\emp$ because $X_1$ and $X_2$ are disjoint, giving $X_2=\emp.$ It follows $X$ is connected. $\blacksquare$

However, it turns out that being irreducible is a rather strong condition.
\begin{example}
    The topological space $\RR$ is not irreducible. For example, $(-\infty,0]$ and $[0,\infty)$ union together to make $\RR,$ but neither of these sets is actually $\RR.$
\end{example}
In fact, we have the following.
\begin{proposition}
    A Hausdorff irreducible space $X$ has a single point.
\end{proposition}
We show that a Hausdorff space $X$ with at least two points is not irreducible. We have to use the Hausdorff condition somehow, so pick up our two distinct points $x_1,x_2\in X.$ Then because $X$ is Hausdorff, we are granted two disjoint open sets $U_1$ and $U_2$ such that $x_1\in U_1$ and $x_2\in U_2.$

Now, set
\[X_1=X\setminus U_1\qquad\text{and}\qquad X_2=X\setminus U_2.\]
We know that $X_1$ and $X_2$ are closed because they are the complement of open sets, and $U_1\cap U_2=\emp$ implies
\[X=X\setminus\emp=X\setminus(U_1\cap U_2)=(X\setminus U_1)\cup(X\setminus U_2),\]
so $X=X_1\cup X_2.$ However, we know $X\ne X_1$ and $X\ne X_2$ because $X\setminus X_1=U_1$ and $X\setminus X_2=U_2$ are both nonempty. It follows that $X$ is not irreducible. $\blacksquare$

It follows that, if irreducibility is a condition we care about, most of the spaces we are looking at aren't going to be Hausdorff. Granted, our topology wasn't made to be pretty; it was made to work.

Anyways, we introduced irreducible spaces to be the topological version of prime. Here is the codification of this.
\begin{proposition}
    We live inside of $\AA_k^n.$ Fix $V\subseteq\AA_k^n$ an algebraic set. Then $V$ is irreducible if and only if $I(V)$ is a prime ideal.
\end{proposition}
The closed set $V$ is irreducible if and only if $V$ is irreducible under the subspace topology. Because our closed sets look like $V_\bullet\cap V$ for some other closed $V_\bullet,$ we can get rid of this intersection and say that our irreducible condition is asserting
\[V\subseteq V_1\cup V_2\implies V\subseteq V_1\text{ or }V\subseteq V_2,\]
where $V_1,V_2$ are algebraic sets themselves. Anyways, we need to show $I(V)$ is prime. Well, working element-wise, suppose $f_1f_2\in I(V)$ so that we need $f_1\in I(V)$ or $f_2\in I(V).$ To use irreducibility, we fix the algebraic sets
\[V_1:=V(\{f_1\})\qquad\text{and}\qquad V_2=V(\{f_2\}).\]
In our proof that vanishing sets form a topology, we the union of vanishing sets is the vanishing set on the product, so $V_1\cup V_2=V(\{f_1f_2\}).$

Now, we want $V\subseteq V_1\cup V_2$ to use irreducibility. Indeed, any $a\in V$ vanishes on $f_1f_2$ because $f_1f_2\in I(V),$ implying $a\in V(\{f_1f_2\}=V_1\cup V_2.$ Thus, irreducibility of $V$ now kicks in to tell us $V\subseteq V_1$ or $V\subseteq V_2.$ Without loss of generality, we take $V\subseteq V_1.$

We hope that this corresponds to $f_1\in I(V),$ which will finish this direction. Indeed,
\[V\subseteq V_1=V(\{f_1\})\]
implies that, for each $a\in V,$ we know $a\in V_1,$ so $f_1(a)=0.$ Thus, all of $V$ vanishes under $f_1,$ so $f_1\in I(V).$ This completes the proof of this direction.

In the other direction, suppose that $I(V)$ is a prime ideal, and we need to show that $V$ is irreducible. Well, suppose we can decompose $V=V_1\cup V_2$ where $V_1$ and $V_2$ are closed subsets of $V.$ Expanding our view as before, we are really supposing $V\subseteq V_1\cup V_2$ for $V_1$ and $V_2$ algebraic subsets of $\AA_k^n,$ and we need $V\subseteq V_1$ or $V\subseteq V_2.$

Taking $I(-),$ we know
\[I(V)\supseteq I(V_1\cup V_2).\]
We want to turn the right-hand side into a product of ideals to use the primality condition, so we pick up the following lemma.
\begin{lemma}
    Fix everything as above. Then $I(V_1\cup V_2)\supseteq I(V_1)I(V_2).$
\end{lemma}
It suffices to show that any generator $f_1f_2$ of $I(V_1)I(V_2)$ where $f_1\in I(V_1)$ and $f_2\in I(V_2)$ lives in $I(V_1\cup V_2).$ Well, for any $a\in V_1\cup V_2,$ we know $a\in V_1$ or $a\in V_2,$ so
\[f_1(a)=0\text{ or }f_2(a)=0.\]
Thus, $(f_1f_2)(a)=0,$ so indeed, $f_1f_2\in I(V_1\cup V_2).$ This finishes. $\blacksquare$

So we see $I(V)\supseteq I(V_1)I(V_2).$ Looking ideal-theoretically at the primality of $I(V),$ it follows $I(V_1)\subseteq I(V)$ or $I(V_2)\subseteq I(V).$ Thus, we get $V\subseteq V_1$ or $V\subseteq V_2$ by inclusion-reversing again, which completes the proof that $V$ is irreducible. $\blacksquare$

To reiterate what just happened, we codify the intuition as follows.
\begin{idea}
    Irreducible closed sets correspond to prime ideals.
\end{idea}
This is so important that irreducible algebraic sets is called a variety, though we will not use this terminology.

We will now move on from primality. Noetherian is another nice algebraic condition, so we translate it into geometry. Well, Noetherian asserts that any ascending chain of ideals
\[I_1\subseteq I_2\subseteq I_3\subseteq\cdots\]
eventually stabilizes with $I_\ell=I_N$ for each $\ell\ge N$ for some positive integer $N.$ When we want to talk about geometric, we turn ideals into closed sets and reverse the inclusions. So we have the following definition.
\begin{definition}
    A topological space $X$ is Noetherian if and only if any descending chain of closed sets eventually stabilizes. Explicitly, given a sequence of closed sets $\{V_\ell\}_{\ell=1}^\infty$ such that
    \[V_1\supseteq V_2\supseteq V_3\supseteq\cdots,\]
    then there exists a positive integer $N$ for which $V_\ell=V_N$ for all $\ell\ge N.$
\end{definition}
If we wanted to define Artinian in a similar way, I think we could, but we won't have as much use for those.

Noetherian is a pretty strong condition on topological spaces. For example, we have the following.
\begin{example}
    We claim that $\RR$ is not Noetherian. We have to exhibit a strictly decreasing chain of closed sets, for which we choose $V_\ell=\{j\in\ZZ:j\ge\ell\}$ for positive integers $\ell.$ The $V_\ell$ are closed because their complement is
    \[\RR\setminus V_\ell=(-\infty,\ell)\cup\bigcup_{j=\ell}^\infty(j,j+1).\]
    As a union of intervals, this is open. To finish, $V_{\ell+1}\subsetneq V_\ell$ because $j\ge\ell+1$ implies $j\ge\ell,$ and $\ell\in V_\ell$ while $\ell\notin V_{\ell+1}.$ So indeed, $\{V_\ell\}_{\ell=1}^\infty$ is an infinite strictly descending chain of closed sets.
\end{example}
Luckily, the space we are most interested in is typically Noetherian.
\begin{proposition}
    The space $\AA_k^n$ with the Zariski topology is Noetherian if $k[t_1,\ldots,t_n]$ is Noetherian.
\end{proposition}
Without much else to do, fix a descending chain of closed (i.e., algebraic) sets
\[V_1\supseteq V_2\supseteq V_3\supseteq\cdots.\]
We want to turn this into a chain of ideals in order to use the Noetherian condition. Because each of the $V_\bullet$ is algebraic, we are promised subsets $S_\bullet\subseteq k[t_1,\ldots,t_n]$ such that $V_\bullet=V(S_\bullet).$ Because $V(S_\bullet)=V((S_\bullet)),$ we may replace each $S_\bullet$ with the ideal generated by $S_\bullet.$

So we have a sequence of ideals $\{S_\ell\}_{\ell=1}^\infty$ generated from the $V_\bullet.$ However, this sequence is not necessarily ascending because, for example, we can have $V((t))\subseteq V\left(\left(t^2\right)\right),$ but $(t)$ is not a subset of $\left(t^2\right).$ Instead, we define
\[I_\ell=\sum_{j=1}^\ell S_\ell\]
to ensure we are ascending. Indeed, this definition gives $I_{\ell+1}=I_\ell+S_{\ell+1},$ so $I_\ell\subseteq I_{\ell+1}.$ 

This might make us nervous, but we do still have $V_\bullet=V(I_\bullet).$ Indeed, $S_\ell\subseteq I_\ell$ gives $V_\ell=V_\bullet(S_\ell)\supseteq V(I_\ell)$ already. As for the other direction, pick up $a\in V_\ell.$ Then $a\in V_j\supseteq V_\ell$ for each $j\le\ell$ because our chain is descending so
\[a\in V(S_j)=V_j.\]
The point is that, for any $f_j\in S_j$ for any $j\le\ell,$ we still get $f_j(a)=0.$ Converting this, for any $\{f_j\}_{j=1}^\ell$ with $f_j\in S_j,$ we see
\[\sum_{j=1}^\ell f_j(a)=\sum_{j=1}^\ell0=0.\]
Thus, $a\in V(S_1+S_2+\cdots+S_\ell)=V(I_\ell),$ completing the proof that $V_\bullet=V(I_\bullet).$

We are now ready to finish. Finally using the Noetherian condition on $k[t_1,\ldots,t_n],$ we see that the chain of ideals
\[I_1\subseteq I_2\subseteq I_3\subseteq\cdots\]
must stabilize. That is, there is a positive integer $N$ for which $I_k=I_N$ for each $k\ge N.$ But then $V_k=V(I_k)=V(I_N)=V_N$ for each $k\ge N$ as well, so we see that the chain of closed sets also stabilizes. This finishes the proof. $\blacksquare$

In particular, when $k$ is a field, we get that $k[t_1,\ldots,t_n]$ is Noetherian by Hilbert's basis theorem. So in this case $\AA_k^n$ is indeed a Noetherian space. I think I can imagine a converse, where Noetherian spaces induce Noetherian rings, but we would need to have a process to turn (say, Zariski) spaces back into rings, which we don't have yet.

Anyways, the proof of the proposition dealt in more generality than $k$ fields in order to make clear how the Noetherian condition on our ring gave the Noetherian space. So here is the idea.
\begin{idea}
    Noetherian rings yield Noetherian spaces.
\end{idea}
There is of course other stuff we can do, but I think that's enough for today.

\subsubsection{June 30th}
Today I learned more formally about integral elements and closures. Here is the definition we are focusing on today.
\begin{definition}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra, which means that $\mathcal O_L$ is a ring with a homomorphism $\varphi:\mathcal O_K\to\mathcal O_L$ inducing an $\mathcal O_K$-action on $\mathcal O_L.$ Then an element $y\in\mathcal O_L$ is integral if and only if $y$ is the root of a monic polynomial in $\varphi(\mathcal O_K)[x].$
\end{definition}
Very quickly, we explicitly say that the $\mathcal O_K$-action on $\mathcal O_L$ granted by $\varphi:\mathcal O_K\to\mathcal O_K$ is given by
\[a\cdot b:=\varphi(a)b\]
for $a\in\mathcal O_K$ and $b\in\mathcal O_L.$ In this way we can think of $\mathcal O_L$ as basically a $\mathcal O_K$-module with a multiplication.

There is some subtlety going on here because, importantly, $\mathcal O_L$ does not have all roots of all monic polynomials over $\mathcal O_K.$ Instead, an element is integral in $\mathcal O_L$ when it already lives in $\mathcal O_L$ and happens to be the root of a monic polynomial. Elements are not integral over $A\mathcal O_K$ in isolation: they have to come from a particular $\mathcal O_K$-algebra $\mathcal O_L.$

Anyways, the intuition here is to generalize the situation of $\ZZ$ living inside of $\QQ$ in a purely algebraic way. A first approximation of this would be to take a ring $R$ living inside of its field of fractions $K$; then we can try to imitate $\ZZ\subseteq\QQ$ by picking up all the elements with no denominator.

For a variety of reasons, this is not actually what we want to do. For example, the field of fractions of $\ZZ[\sqrt{-3}]$ is $\QQ(\zeta_3),$ but the situation of $\ZZ[\sqrt{-3}]\subseteq\QQ(\zeta_3)$ is not similar to $\ZZ$: we don't have unique prime factorization of prime ideals here, for example.

It turns out that being integral is the correct algebraic generalization of $\ZZ\subseteq\QQ.$ It is true that the elements of $\QQ$ that are integral over $\ZZ$ make up exactly $\ZZ.$ Here is the general fact.
\begin{proposition}
    Fix $\mathcal O_K$ a unique factorization domain with field of fractions $K$ and interpret $K$ as a $\mathcal O_K$-algebra with the canonical $\varphi:\mathcal O_K\hookrightarrow K.$ Then $\mathcal O_K$ is exactly the elements of $K$ integral over $\mathcal O_K.$
\end{proposition}
For one of these inclusions, note that any $a\in\mathcal O_K$ is the root of the monic polynomial $x-a\in\mathcal O_K[x].$ It follows that all elements of $\mathcal O_K$ are integral over $\mathcal O_K.$

It remains to show that these are the only elements of $K$ integral over $\mathcal O_K.$ For the purpose of the proof, it is useful to read $\mathcal O_K=\ZZ$ and $K=\QQ,$ but of course we will work in generality Fix $a/b\in K$ with $a,b\in\mathcal O_K$ and $b\ne0$ such that $a/b$ is the root of a monic polynomial in $\mathcal O_K[x]$ which looks like
\[x^n+\sum_{k=0}^{n-1}c_kx^k\]
for some $\{c_k\}_{k=1}^{n-1}\subseteq\mathcal O_K.$ Without much better to do, we plug in $a/b$ and get
\[a^n+\sum_{k=0}c_ka^kb^{n-k}=0\]
after clearing fractions by multiplying by $b^n.$

Now, we want to show that $a/b\in\mathcal O_K,$ so it suffices to show that $b\mid a.$ Because we live in a unique factorization domain, we show that $\nu_p(a)\ge\nu_p(b)$ for any fixed irreducible $p.$ Because $p^\bullet\mid x,y$ implies $p^\bullet x+y$ for any $x,y\in\mathcal O_K,$ we see $\nu_p(x+y)\ge\min\{\nu_p(x),\nu_p(y)\}.$ In particular,
\[\nu_p\left(a^n\right)=\nu_p\left(-\sum_{k=0}^{n-1}c_ka^kb^{n-k}\right)\ge\min_{0\le k\le n-1}\left\{\nu_p\left(-c_ka^kb^{n-k}\right)\right\}.\]
To continue, we note that $\nu_p(xy)=\nu_p(x)+\nu_p(y)$ by expanding the prime factorization of $xy$ and of $x\cdot y.$ It follows that
\[\nu_p\left(c_ka^kb^{n-k}\right)=\nu_p(c_k)+k\nu_p(a)+(n-k)\nu_p(b)\ge k\nu_p(a)+(n-k)\nu_p(b),\]
where we have intentionally killed the $c_\bullet$ term. At this point, we are forgetting that the polynomial ever existed. To review, we know that
\[n\nu_p(a)\ge\min_{0\le k\le n-1}\{k\nu_p(a)+(n-k)\nu_p(b)\}\]
and need to show $\nu_p(a)\ge\nu_p(b).$

Well, when $\nu_p(a)<\nu_p(b),$ we have that
\[k\nu_p(a)+(n-k)\nu_p(b)>n\nu_p(a)\]
for any $k,$ so the minimum over all $k$ of $k\nu_p(a)+(n-k)\nu_p(a)$ is also greater than $n\nu_p(a).$ This is the exact opposite of what we know, so we conclude that we must have $\nu_p(a)\ge\nu_p(b)$ by contraposition or contradiction. So indeed, $b\mid a,$ and $a/b\in\mathcal O_K,$ completing the proof. $\blacksquare$

As a quick side remark, the notation of $\mathcal O_K$ we are using is intended to be reminiscent of algebraic number theory with number fields, but the above proof shows that the elements of $k(t)$ integral over $k[t]$ are exactly $k[t]$ as well, where things are potentially harder to think directly about. This is the benefit we get from working in generality.

Our main attraction today will be showing that the integral closure is integrally closed, but we have to build some theory to get there. We say now that this will be similar to the proof that the set of algebraic numbers is algebraically closed, but we could also say that that proof is similar to this one.

We start with a more abstract definition of integral; the purpose of a more abstract definition in terms of extensions is that it becomes easier to show that products and sums of integral elements are integral because extensions are.
\begin{lemma}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ An element $y\in\mathcal O_L$ is integral over $\mathcal O_K$ if and only if $\varphi(\mathcal O_K)[y]\subseteq\mathcal O_L$ is finitely generated (as a module) over $\mathcal O_K.$
\end{lemma}
In one direction, suppose $y\in\mathcal O_L$ is integral over $\mathcal O_K,$ and we want to show $\varphi(\mathcal O_K)[y]$ is a finitely generated $\mathcal O_K$-module. We won't show that $\varphi(\mathcal O_K)[y]$ is a commutative ring (we would have to show closure under $+$ and $\times$ from $\mathcal O_L$). It is a $\mathcal O_K$-submodule of $\mathcal O_L$ because, for any $a\in\mathcal O_K$ and $\sum_k\varphi(a_k)y^k\in\varphi(\mathcal O_K)[y],$
\[a\cdot\sum_{k=0}^n\varphi(a_k)y^k=\sum_{k=0}^n\varphi(aa_k)y^k\in\varphi(\mathcal O_K)[y]\]
gives closure under the action.

The interesting part is showing that this is finitely generated. Taking inspiration from field theory, we use a ``power basis'' of $\left\{1,y,y^2,\ldots\right\}.$ To be explicit, $y$ being integral over $\mathcal O_K$ gives us a monic polynomial in $\varphi(\mathcal O_K)[x]$ making $y$ vanish. Giving this polynomial a name, we see
\[y^d+\sum_{k=0}^{d-1}\varphi(c_k)y^k=0.\]
The important part is this $d.$ We claim that $\left\{1,y,y^2,\ldots,y^d\right\}$ generates $\varphi(\mathcal O_K)[y].$ For this, we proceed by induction on the degree $n$ of an arbitrary element
\[\sum_{k=0}^n\varphi(a_k)y^k\]
in $\varphi(\mathcal O_K)[y].$ When $n\le d,$ then the above counts as a representation with our power basis.

For the inductive step, we suppose everything with degree $n$ is achievable, and we show degree $n+1.$ Well,
\[y^{n+1}=y^{n+1-d}\cdot y^d=-\sum_{k=0}^{d-1}\varphi(c_k)y^{n+1-d+k}.\]
Thus, for any element with degree $n+1,$ we can write
\[\sum_{k=0}^{n+1}\varphi(a_k)y^k=-\varphi(a_{n+1})\sum_{k=0}^{d-1}\varphi(c_k)y^{n+1-d+k}+\sum_{k=0}^n\varphi(a_k)y^k.\]
All powers of $y$ are now bounded by $n+1-d+d-1=n,$ so this element is representable under our power basis by hypothesis. This completes the inductive step and therefore this direction.

As for the other direction, suppose that $\varphi(\mathcal O_K)[y]$ is a finitely generated $\mathcal O_K$-module, and we need to make $y$ the root of some monic polynomial in $\varphi(\mathcal O_K)[x].$ Well, fix its generators $\{p_1(y),\ldots,p_n(y)\}\subseteq\varphi(\mathcal O_K)[y]$ where each $p_\bullet$ comes from $\varphi(\mathcal O_K)[x].$ (Explicitly, each generator can be pulled back from $\varphi(\mathcal O_K)[y]$ into $\varphi(\mathcal O_K)[x]$ by $y^\bullet\mapsto x^\bullet.$)

The idea is that $\mathcal O_K$-lienar combinations of the $p_\bullet$ can only have so large a degree. Namely, each $p_\bullet\in\varphi(\mathcal O_K)[x]$ has some finite degree, and because there are finitely many $p_\bullet,$ we can let $d$ be the largest of these degrees. But now there must be some way to represent $y^{d+1}\in\varphi(\mathcal O_K)[y]$ from these generators, namely
\[y^{d+1}=\sum_{k=1}^n\varphi(a_k)p_k(y).\]
for some $a_\bullet$ in $\mathcal O_K.$ By construction of $d,$ the largest power of $y$ on the right-hand side is less than $d+1,$ so fully expanding
\[x^{d+1}-\sum_{k=1}\varphi(a_k)p_k(x)\]
gives a monic polynomial in $\varphi(\mathcal O_K)[x]$ (of degree $d+1$) causing $y$ to vanish. Thus, $y$ is indeed integral over $\mathcal O_K,$ completing this direction. $\blacksquare$

We will actually need a slightly more versatile condition than ``$\varphi(\mathcal O_K)[y]$ is finitely generated'' for some of where we are going. This is as follows.
\begin{lemma}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ An element $y\in\mathcal O_L$ is integral if and only if there exists an intermediate $\mathcal O_K$-subalgebra $M\subseteq\mathcal O_L$ containing $\varphi(\mathcal O_K)[y]$ and is finitely generated as an $\mathcal O_K$-module.
\end{lemma}
In the forward direction, note that $y$ integral gives $\varphi(\mathcal O_K)[y]$ finitely generated as a $\mathcal O_K$-module. It is also a submodule of $\mathcal O_L$ and hence a subalgebra (i.e., it's closed under the $\mathcal O_K$-action), so $M:=\varphi(\mathcal O_K)[y]$ will do the trick.

The backwards direction is trickier. Essentially all we have is a finitely generated module $M$ which we know contains what the action from $\mathcal O_K$ and $y$ requires. Somehow this needs to turn into a monic linear combination of the $y^\bullet.$ The key idea is to use the determinant trick, whose one job is to generate monic relations among endomorphisms.
\begin{proposition}
    Fix $R$ a ring and $M$ an $R$-module, where $M$ is generated by $n\in\NN$ elements. Now suppose $\varphi:M\to M$ is an $R$-module homomorphism such that $\varphi(M)\subseteq IM$ for an ideal $I\subseteq R.$ Then we can write
    \[\varphi^n+\sum_{k=0}^{n-1}\mu_{r_k}\varphi^k=0,\]
    where $r_k\in I.$
\end{proposition}
We showed this a while ago. It is the determinant trick. $\blacksquare$

Now, because $M$ actually has some ring structure (it's an $\mathcal O_K$-algebra), $\mu_y:m\mapsto ym$ is an $\mathcal O_K$-module endomorphism of $M.$ Indeed, for $a_1,a_2\in\mathcal O_K$ and $m_1,m_2\in M,$ we have
\[\mu_y(\varphi(a_1)m_1+\varphi(a_2)m_2)=y\varphi(a_1)m_1+y\varphi(a_2)m_2=a_1\mu_y(m_1)+a_2\mu_y(m_2)\]
because of our ring structure. Thus, $\mu_y$ is a valid $\mathcal O_K$-linear endomoprhism, and we can use the determinant trick on it, with $I=\mathcal O_K.$ Letting $n$ be the number of generators, we get
\[\mu_y^n+\sum_{k=0}^{n-1}\mu_{\varphi(a_k)}\mu_y^k=0\]
for some $a_\bullet\in\mathcal O_K.$ Now we note that $\mu_a\mu_b=\mu_{ab}$ because both send $m$ to $abm,$ so applying this multiplicativity to the above gives
\[y^n+\sum_{k=0}^{n-1}\varphi(a_k)y^k=0\]
upon plugging in $1$ to the relation of endomorphisms. This gives our monic polynomial in $\varphi(\mathcal O_K)[x]$ causing $y$ to vanish, which proves $y$ integral. This finishes the proof. $\blacksquare$

It might feel a bit unclear what we gained from the above redefinition of integral. What's nice is that we want to say something like ``all elements of $\varphi(\mathcal O_K)[y]$ are integral when $y$ is because the generated extension is finite,'' but this does not immediately follow. After all, $y^2\in\varphi(\mathcal O_K)[y]$ potentially gives a different extension $\varphi(\mathcal O_K)\left[y^2\right],$ which we don't immediately know is finite. But
\[\varphi(\mathcal O_K)\left[y^2\right]\subseteq\varphi(\mathcal O_K)[y],\]
and the latter is finite, so the above redefinition tells us that we do actually get $y^2$ integral.

Now, to prove that the integral closure is integrally closed, we need to define what it means for rings to be integrally closed.
\begin{definition}
    Fix $\mathcal O_L$ a $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ If $y\in\mathcal O_L$ integral over $\mathcal O_K$ implies $y\in\varphi(\mathcal O_K),$ then $\mathcal O_K$ is integrally closed under $\mathcal O_L.$
\end{definition}
As an example, we showed above that $\mathcal O_K$ is integrally closed under its field of fractions when $\mathcal O_K$ is a unique factorization domain.

This is also inherently a finiteness condition. Namely, suppose $\mathcal O_K$ is integrally closed under $\mathcal O_L,$ and if we pick up $y\in\mathcal O_L$ such that $\varphi(\mathcal O_K)[y]$ is finite, then $y$ is integral, so we had $\varphi(\mathcal O_K)[y]=\varphi(\mathcal O_K)$ all along. I think we can modify this a bit to say that $\varphi(\mathcal O_K)$ has no finite extensions.

To not bury the lead, we go ahead and define the integral closure now.
\begin{definition}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ We define the integral closure of $\mathcal O_K$ as
    \[\overline{\mathcal O_K}:=\{y\in\mathcal O_L:y\text{ is integral over }\mathcal O_K\}.\]
\end{definition}
In light of what we said earlier, to show that $\overline{\mathcal O_K}$ is integrally closed over $\mathcal O_K,$ we have to carefully control our finite extensions.

With this in mind, the following is the main step towards our goal.
\begin{proposition}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ Further, pick up $\{y_k\}_{k=1}^m$ elements of $\mathcal O_L$ integral over $\mathcal O_K.$ Then the giant ring
    \[\varphi(\mathcal O_K)[y_1,\ldots,y_m]\]
    is still finitely generated as a $\mathcal O_K$-module.
\end{proposition}
This is by induction on $m.$ When $m=1,$ this is asserting that $\varphi(\mathcal O_K)[y_1]$ is finitely generated when $y_1$ is integral, which we know to be true by one of the equivalent properties of integral.

Now we can assume the statement is true for $m$ and show $m+1.$ That is, we are given $\{y_k\}_{k=1}^{m+1}$ elements of $\mathcal O_L$ integral over $\mathcal O_K$ and need to show $\varphi(\mathcal O_K)[y_1,\ldots,y_{m+1}]$ is a finitely generated $\mathcal O_K$-module. Well, we already know
\[\varphi(\mathcal O_K)[y_1,\ldots,y_m]\]
is finitely generated as a $\mathcal O_K$-module. Because our action is induced by $\varphi,$ we are saying that $\varphi(\mathcal O_K)[y_1,\ldots,y_m]$ is finitely generated over the ring $\varphi(\mathcal O_K).$

Additionally, $y_{m+1}$ being integral over $\mathcal O_K$ means that there is a monic polyomial in $\varphi(\mathcal O_K)[x]$ for which $y_{m+1}$ vanishes. But we can lift this polynomial to $\varphi(\mathcal O_K)[y_1,\ldots,y_m][x],$ meaning that $y_{m+1}$ is also integral over $\varphi(\mathcal O_K)[y_1,\ldots,y_m],$ which $\mathcal O_L$ is an algebra over by embedding. The point is that
\[\varphi(\mathcal O_K)[y_1,\ldots,y_m,y_{m+1}]\]
is finitely generated as a ring over $\varphi(\mathcal O_K)[y_1,\ldots,y_m].$

Thus, the situation is that we have the chain of rings
\[\varphi(\mathcal O_K)\subseteq\varphi(\mathcal O_K)[y_1,\ldots,y_m]\subseteq\varphi(\mathcal O_K)[y_1,\ldots,y_{m+1}],\]
where each step is finite generation as rings. Our end goal is to show that $\varphi(\mathcal O_K)[y_1,\ldots,y_{m+1}]$ is finitely generated over $\varphi(\mathcal O_K),$ which is covered by the following lemma.
\begin{lemma}
    Suppose $R\subseteq S\subseteq T$ are rings such that $T$ is finitely generated over $S$ and $S$ is finitely generated over $R.$ Then $T$ is finitely generated over $T.$
\end{lemma}
This is the same proof as for field extensions. Let $\{s_k\}_{k=1}^n\subseteq R$ generate $S$ over $R$ and $\{t_\ell\}_{\ell=1}^m$ generate $T$ over $S.$ We claim that the set of products $s_kt_\ell$ generate $T$ over $R.$ Indeed, any $t\in T$ can be written as
\[t=\sum_{\ell=1}^ma_\ell t_\ell\]
for $a_\ell\in S$ because the $t_\bullet$ generate $T$ over $S.$ Now, each coefficient $a_\ell$ can be further expanded as
\[a_\ell=\sum_{k=1}^na_{\ell,k}s_k\]
for $a_{\ell,k}\in R$ because the $s_\bullet$ generate $S$ over $R.$ Thus,
\[t=\sum_{\ell=1}^m\left(\sum_{k=1}^na_{\ell,k}s_k\right)t_{\ell}=\sum_{\ell=1}^m\sum_{k=1}^na_{\ell,k}(s_kt_\ell).\]
So indeed, the $s_kt_\ell$ generate $T$ over $R.$ $\blacksquare$

Setting $R=\varphi(\mathcal O_K),$ $S=\varphi(\mathcal O_K)[y_1,\ldots,y_m],$ and $T=\varphi(\mathcal O_K)[y_1,\ldots,y_{m+1}]$ and plugging into the above lemma finishes the inductive step and hence the proof. $\blacksquare$

This proposition will serve a dual purpose. To start off, we quickly get the following.
\begin{lemma}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ Then the integral closure $\overline{\mathcal O_K}$ is a subalgebra of $\mathcal O_L.$
\end{lemma}
Most of the work is showing $\overline{\mathcal O_K}$ is a subring. Note that $\varphi(\mathcal O_K)\subseteq\overline{\mathcal O_K}$ because any $\varphi(a)\in\varphi(\mathcal O_K)$ is the root of the monic polynomial $x-\varphi(a).$ So $\overline{\mathcal O_K}$ does have $0$ and $1.$

To show that $\overline{\mathcal O_K}$ is a subring, it now suffices to show it is closed under $\pm$ and $\times.$ Well, picking up any $y,z\in\overline{\mathcal O_K},$ we note that $y\pm z$ and $yz$ all live in
\[\varphi(\mathcal O_K)[y,z],\]
which is finitely generated over $\varphi(\mathcal O_K)$ by the above proposition. We know that this is a subring of $\mathcal O_L$ and contains $\varphi(\mathcal O_K),$ so it is also a subalgebra of $\mathcal O_L,$ where we are closed under the action because we are closed under multiplication.

So we see that $\varphi(\mathcal O_K)[y,z]$ is a finitely generated intermediate sub-algebra of $\mathcal O_L$ containing $\varphi(\mathcal O_K)[y\pm z]$ and $\varphi(\mathcal O_K)[yz].$ Using our more versatile definition of integral, it follows that $y\pm z$ and $yz$ are all integral, so they still live in $\overline{\mathcal O_K}.$

To finish showing that $\overline{\mathcal O_K}$ is a subalgebra of $\mathcal O_L,$ we have to show that it is closed under the action. Well, the action is multiplication by elements of $\varphi(\mathcal O_K),$ which $\overline{\mathcal O_K}$ contains, so $\overline{\mathcal O_K}$ is indeed closed under the action because it is a subring. $\blacksquare$

And lastly, we finally arrive at the main attraction.
\begin{theorem}
    Fix $\mathcal O_L$ an $\mathcal O_K$-algebra with $\varphi:\mathcal O_K\to\mathcal O_L.$ Then the integral closure $\overline{\mathcal O_K}$ is a subalgebra and integrally closed in $\mathcal O_L.$
\end{theorem}
We have already showed that $\overline{\mathcal O_K}$ is a subalgebra, so our work will be to show that it is integrally closed. Pick up some $y\in\mathcal O_L$ integral over $\overline{\mathcal O_K},$ and we need to show $y\in\overline{\mathcal O_K},$ or equivalently, $y$ is integral over $\mathcal O_K.$ Being integral gives us a vanishing monic polynomial
\[y^n+\sum_{k=0}^{n-1}c_ky^k=0\]
for $c_\bullet\in\overline{\mathcal O_K}.$

To show that $y$ is integral over $\mathcal O_K,$ we reuse the trick from the previous lemma. Namely, we know that
\[\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}]\]
is finitely generated over $\varphi(\mathcal O_K)$ because the $c_\bullet$ are integral over $\mathcal O_K.$ And the given monic polynomial for $y$ has coefficients in $\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}],$ so we know $y$ is integral over $\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}],$ meaning that
\[\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1},y]\]
is finitely generated over $\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}].$

To finish, we note that we have a chain of rings
\[\varphi(\mathcal O_K)\subseteq\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}]\subseteq\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1},y],\]
where each ring is finitely generated over the previous. It follows that $\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}][y]$ is finitely generated over $\varphi(\mathcal O_K).$ So $\varphi(\mathcal O_K)[c_0,\ldots,c_{n-1}][y]$ is a finitely generated $\mathcal O_K$-algebra (it's closed under the action because it contains $\varphi(\mathcal O_K)$) containing $\varphi(\mathcal O_K)[y].$ This means $y$ is integral over $\mathcal O_K,$ or $y\in\overline{\mathcal O_K},$ finishing the proof. $\blacksquare$

This was the main attraction, so I think that's enough for today.